---
title: "RQ1 zip code"
author: "ESINCLAIR"
date: "2026-08-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# LIBRARIES
library(dplyr)
library(igraph)
library(randomcoloR)
library(ggplot2)
library(tidyr)
library(gt)
library(patchwork)
```

https://link.springer.com/content/pdf/10.1007/s11274-025-04755-3.pdf

# Network construction and overall structure

Construct 12 period-specific networks

```{r}
# copy of graph pipeline function - modified to make one panel of 12 graphs
node_colours <- c(
  "Cow-Calf" = "green",
  "Starter Feedlot" = "purple",
  "Finishing Feedlot" = "orange"
)

# margins
par(mar = c(0,0,0,0), oma = c(2,0,0,0)) # hash this is doing 12 plot layout

plot_movement_network <- function(df, month_label = "XX") {

  # convert IDs to character before building nodes and edges
  df$org_id  <- as.character(df$org_id)
  df$dest_id <- as.character(df$dest_id)

  # 1. Node table
  org_nodes  <- df[, c("org_id", "org_type")]
  colnames(org_nodes) <- c("name", "type")

  dest_nodes <- df[, c("dest_id", "dest_type")]
  colnames(dest_nodes) <- c("name", "type")

  nodes <- unique(rbind(org_nodes, dest_nodes))

  # 2. Edge list
  edges <- df[, c("org_id", "dest_id")]

  # 3. Graph
  g <- graph_from_data_frame(
    d = edges,
    vertices = nodes,
    directed = TRUE
  )

  # Add weights
  
  E(g)$weight <- df$total_cattle # 'E(g) is igraphs built-in edge sequence, meaning 'all edges in graph' - so this adds a new attribute called 'weight' to each edge
  
  # 4. Colours
  cols <- node_colours[V(g)$type]
  # margins
  #par(mar = c(0,0,0,0), oma = c(4,0,0,0)) # hash this is doing 12 plot layout

  # 5. Layers
  layer_order <- c("Cow-Calf",
                   "Starter Feedlot",
                   "Finishing Feedlot")

  V(g)$layer <- as.numeric(factor(V(g)$type, levels = layer_order))

  # Layout
  lay <- layout_with_sugiyama(g, layers = V(g)$layer)$layout

  # Plot
  plot(
    g,
    layout = lay,
    vertex.color = cols,
    vertex.label = NA,# V(g)$name - change to this if want ID numbers printed
    vertex.size = 25,
    vertex.label.cex = 0.9,
    edge.arrow.size = 0.5
  )

  # Title
  mtext(
    paste("Period", month_label),
    side = 3, line = 0.2, cex = 0.8
  )

  # Legend - comment out for combined panel
  #legend(
   # "topright",
    #legend = names(node_colours),
    #col = node_colours,
    #pch = 19,
    #pt.cex = 1.5,
    #bty = "n"
  #)
  # Return the graph object for metrics
  return(g)
}
```

Import data 

```{r}
Q12_M1_RQ1 <- read.csv("Q12(D1_RQ1).csv", check.names = TRUE) # tells R to sanitise column names
#head(Q12_M1)
Q12_M2_RQ1 <- read.csv("Q12(D2_RQ1).csv", check.names = TRUE) 
#head(Q12_M2)
Q12_M3_RQ1 <- read.csv("Q12(D3_RQ1).csv", check.names = TRUE) 
#head(Q12_M3)
Q12_M4_RQ1 <- read.csv("Q12(D4_RQ1).csv", check.names = TRUE) 
#head(Q12_M4)
Q12_M5_RQ1 <- read.csv("Q12(D5_RQ1).csv", check.names = TRUE) 
#head(Q12_M5)
Q12_M6_RQ1 <- read.csv("Q12(D6_RQ1).csv", check.names = TRUE) 
#head(Q12_M6)
Q12_M7_RQ1 <- read.csv("Q12(D7_RQ1).csv", check.names = TRUE) 
#head(Q12_M7)
Q12_M8_RQ1 <- read.csv("Q12(D8_RQ1).csv", check.names = TRUE) 
#head(Q12_M18)
Q12_M9_RQ1 <- read.csv("Q12(D9_RQ1).csv", check.names = TRUE) 
#head(Q12_M9)
Q12_M10_RQ1 <- read.csv("Q12(D10_RQ1).csv", check.names = TRUE) 
#head(Q12_M10)
Q12_M11_RQ1 <- read.csv("Q12(D11_RQ1).csv", check.names = TRUE) 
#head(Q12_M11)
Q12_M12_RQ1 <- read.csv("Q12(D12_RQ1).csv", check.names = TRUE) 
#head(Q12_M12)
```

```{r}
# call function for all 12 periods and plot as panel with legend

par(mfrow = c(4, 4), mar = c(0,0,1.5,0), oma = c(0,0,0,0)) # 3 rows, 4 columns, no margins

# plot all 12 networks

g_M1_RQ1 <- plot_movement_network(Q12_M1_RQ1, "1")
g_M2_RQ1 <- plot_movement_network(Q12_M2_RQ1, "2")
g_M3_RQ1 <- plot_movement_network(Q12_M3_RQ1, "3")
g_M4_RQ1 <- plot_movement_network(Q12_M4_RQ1, "4")
g_M5_RQ1 <- plot_movement_network(Q12_M5_RQ1, "5")
g_M6_RQ1 <- plot_movement_network(Q12_M6_RQ1, "6")
g_M7_RQ1 <- plot_movement_network(Q12_M7_RQ1, "7")
g_M8_RQ1 <- plot_movement_network(Q12_M8_RQ1, "8")
g_M9_RQ1 <- plot_movement_network(Q12_M9_RQ1, "9")
g_M10_RQ1 <- plot_movement_network(Q12_M10_RQ1, "10")
g_M11_RQ1 <- plot_movement_network(Q12_M11_RQ1, "11")
g_M12_RQ1 <- plot_movement_network(Q12_M12_RQ1, "12")

# Shared legend
par(fig = c(0, 1, 0, 0.08), new = TRUE, mar = c(0, 0, 0, 0))

plot.new()

legend(
  "center",
  legend = names(node_colours),
  col = node_colours,
  pch = 19,
  pt.cex = 1.5,
  horiz = TRUE,
  bty = "n"
)

par(mfrow = c(1, 1))   # reset afterwards
```
Calculate network level metrics

```{r}
# put all months into one list
monthly_data <- list(
  Q12_M1_RQ1,
  Q12_M2_RQ1,
  Q12_M3_RQ1,
  Q12_M4_RQ1,
  Q12_M5_RQ1,
  Q12_M6_RQ1,
  Q12_M7_RQ1,
  Q12_M8_RQ1,
  Q12_M9_RQ1,
  Q12_M10_RQ1,
  Q12_M11_RQ1,
  Q12_M12_RQ1
)
```
```{r}
# name the list so know which month is which
names(monthly_data) <- paste0("Period_", 1:12)
```
```{r}
# build 12 networks using lapply

monthly_graphs <- lapply(monthly_data, function(d) {
  g <- graph_from_data_frame(
    d[, c("org_id", "dest_id")], # N.B. does not contain node type - but this is not used in analysis
    directed = TRUE
  )
  E(g)$weight <- d$total_cattle
  g
})

```
```{r}
# compute network level metrics
monthly_network_stats <- lapply(monthly_graphs, function(g) {
  list(
    total_nodes = vcount(g),
    total_edges = ecount(g),
    density = edge_density(g, loops = FALSE),
    largest_component = max(components(g, mode = "weak")$csize), # weakly connected - direction ingnored
    diameter = diameter(g, directed = FALSE, weights = NA) # N.B. this isnt weighted - measures the shortest topological path, not the cattle-volume-weigthed path
  )
})
```
```{r}
# turn into a table

monthly_network_summary <- do.call(
  rbind,
  lapply(seq_along(monthly_network_stats), function(i) {
    data.frame(
      Period = paste0("Period_", i),
      total_edges = monthly_network_stats[[i]]$total_edges,
      total_nodes = monthly_network_stats[[i]]$total_nodes,
      largest_component = monthly_network_stats[[i]]$largest_component,
      density = monthly_network_stats[[i]]$density,
      diameter = monthly_network_stats[[i]]$diameter
    )
  })
)
```
```{r}
monthly_network_summary
```
```{r}
# AI how do i turn this table into a png/pdf to export?

network_table <- monthly_network_summary |>
  gt() |>
  cols_label(
    Period = "Observation period",
    total_edges = "Total edges",
    total_nodes = "Total nodes",
    largest_component = "Largest weakly connected component",
    density = "Network density",
    diameter = "Network diameter"
  ) |>
  fmt_number(
    columns = density,
    decimals = 3
  )

gtsave(
  network_table,
  "network_characteristics.html"
)
```

```{r}
# AI prompt - how to turn this into 5 separate line plots in one panel with matching headers/axis labels to metric
# plot variation over 12 periods
png(
  "network_metrics_12_periods.png",
  width = 3000,
  height = 2400,
  res = 300
)

par(
  mfrow = c(3, 3),
  mar = c(4, 4, 3, 1)
)

# Plot 1: Total nodes
plot(
  1:12,
  monthly_network_summary$total_nodes,
  type = "o",
  pch = 16,
  xlab = "Period",
  ylab = "Number of nodes",
  main = "Total nodes",
  xaxt = "n"
)
axis(1, at = 1:12, labels = 1:12)

# Plot 2: Total edges
plot(
  1:12,
  monthly_network_summary$total_edges,
  type = "o",
  pch = 16,
  xlab = "Period",
  ylab = "Number of edges",
  main = "Total edges",
  xaxt = "n"
)
axis(1, at = 1:12, labels = 1:12)

# Plot 3: Largest weakly connected component
plot(
  1:12,
  monthly_network_summary$largest_component,
  type = "o",
  pch = 16,
  xlab = "Period",
  ylab = "Number of nodes",
  main = "Largest weakly connected component",
  xaxt = "n"
)
axis(1, at = 1:12, labels = 1:12)

# Plot 4: Density
plot(
  1:12,
  monthly_network_summary$density,
  type = "o",
  pch = 16,
  xlab = "Period",
  ylab = "Density",
  main = "Network density",
  xaxt = "n"
)
axis(1, at = 1:12, labels = 1:12)

# Plot 5: Diameter
plot(
  1:12,
  monthly_network_summary$diameter,
  type = "o",
  pch = 16,
  xlab = "Period",
  ylab = "Diameter",
  main = "Network diameter",
  xaxt = "n"
)
axis(1, at = 1:12, labels = 1:12)

dev.off()

par(mfrow = c(1, 1))
```

REF - network analysis descriptive review paper

1. Network Density (0-1) = proportion of possible edges that actually exist (equation for this). high = many connected farms = infection spreads easily
2. component size - size of the largest weakly connected component (ignores direction) - large - infection can reach many farms, network not fragmented
3. diameter - longest shortest pat between any two nodes - use weakly connected diameter - large = longmovement chains -> bridging farms matter/small diameter - compact network - outdegree dominates

# Calculate node level metrics

```{r}
# compute node metrics for each month
monthly_metrics_nodes <- lapply(monthly_graphs, function(g) {
  data.frame(
    node = V(g)$name,
    outdegree = degree(g, mode = "out"),
    strength_out = strength(g, mode = "out", weights = E(g)$weight),
    betweenness = betweenness(g, weights = E(g)$weight)
  )
})

```
```{r}
# create summary table with monthly node-level metrics

monthly_node_summary <- do.call(
  rbind,
  lapply(seq_along(monthly_metrics_nodes), function(i) {
    
    m <- monthly_metrics_nodes[[i]]
    
    data.frame(
      Period = paste0("Period_", i),
      median_outdegree = median(m$outdegree),
      max_outdegree = max(m$outdegree),
      outdegree_95th = quantile(m$outdegree, 0.95),
      
      median_strength_out = median(m$strength_out),
      max_strength_out = max(m$strength_out),
      strength_out_95th = quantile(m$strength_out, 0.95),
      
      median_betweenness = median(m$betweenness),
      max_betweenness = max(m$betweenness),
      betweenness_95th = quantile(m$betweenness_out, 0.95)
    )
  })
)
```
```{r}
monthly_node_summary
```
```{r}
# AI prompt - i want to turn the monthly node summary into 3 panels of 12 box plots to show the distribution of node metrics. I want clear labeling, legends, and titles

# this uses the raw values

# combine 12 periods into one dataframe
monthly_node_data <- do.call(
  rbind,
  lapply(seq_along(monthly_metrics_nodes), function(i) {
    
    m <- monthly_metrics_nodes[[i]]
    
    data.frame(
      Period = i,
      node = m$node,
      outdegree = m$outdegree,
      strength_out = m$strength_out,
      betweenness = m$betweenness
    )
  })
)
```
```{r}
# plot, using log of y-axis because highly right skewed and compresses boxes
par(mfrow = c(1, 3))


# Out-degree
boxplot(
  outdegree ~ Period,
  data = monthly_node_data,
  xlab = "Period",
  ylab = "Out-degree",
  main = "Out-degree",
  col = "forestgreen"
)

# Out-strength
boxplot(
  strength_out ~ Period,
  data = monthly_node_data,
  xlab = "Period",
  ylab = "Out-strength",
  main = "Out-strength",
  col = "forestgreen"
)

# Betweenness
boxplot(
  betweenness ~ Period,
  data = monthly_node_data,
  xlab = "Period",
  ylab = "Betweenness centrality",
  main = "Betweenness centrality",
  col = "forestgreen"
)

par(mfrow = c(1, 1))
```

restrick y-axis to 95th percentile - not helpful as majority of boxes are 

```{r}
# jittered plot including median
ggplot(monthly_node_data, aes(
  x = factor(Period),
  y = outdegree
)) +
  geom_jitter(
    width = 0.2,
    alpha = 0.12,
    size = 1
  ) +
  stat_summary(
    fun = median,
    geom = "point",
    size = 3,
    colour = "red"
  ) +
  labs(
    x = "Period",
    y = "Out-degree"
  ) +
  theme_classic()
```
```{r}
# how many farms have zero/very low values
monthly_node_data %>%
  group_by(Period) %>%
  summarise(
    n = n(),
    zero_outdegree = sum(outdegree == 0),
    median_outdegree = median(outdegree),
    max_outdegree = max(outdegree)
  )
```

3 panel figure showing distributional summaries - plot median/75th& 95th percentile, maximum over 12 periods. 95th percentile will show if the upper tail changes over time

```{r}
node_distribution_summary <- monthly_node_data %>%
  group_by(Period) %>%
  summarise(
    outdegree_median = median(outdegree),
    outdegree_75 = quantile(outdegree, 0.75),
    outdegree_95 = quantile(outdegree, 0.95),
    outdegree_max = max(outdegree),

    strength_median = median(strength_out),
    strength_75 = quantile(strength_out, 0.75),
    strength_95 = quantile(strength_out, 0.95),
    strength_max = max(strength_out),

    betweenness_median = median(betweenness),
    betweenness_75 = quantile(betweenness, 0.75),
    betweenness_95 = quantile(betweenness, 0.95),
    betweenness_max = max(betweenness)
  )
```
```{r}
node_distribution_summary
```
```{r}
# reshape summary data
# Out-degree
outdegree_plot_data <- node_distribution_summary %>%
  select(
    Period,
    median = outdegree_median,
    `75th percentile` = outdegree_75,
    `95th percentile` = outdegree_95,
    maximum = outdegree_max
  ) %>%
  pivot_longer(
    cols = -Period,
    names_to = "Statistic",
    values_to = "Value"
  )
```
```{r}
# plot outdegree - shows typical farm (median), upper middle distribution (75th) and upper tail (95th)
ggplot(
  outdegree_plot_data,
  aes(x = Period, y = Value, colour = Statistic)
) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  scale_x_continuous(breaks = 1:12) +
  labs(
    x = "Period",
    y = "Out-degree",
    colour = NULL
  ) +
  theme_classic()
```

?remove maximum - could be highly skewed by one farm in the period

```{r}
# Reshape the three metrics into long format
node_distribution_long <- node_distribution_summary %>%
  select(
    Period,
    
    # Out-degree
    outdegree_median,
    outdegree_75,
    outdegree_95,
    
    # Out-strength
    strength_median,
    strength_75,
    strength_95,
    
    # Betweenness
    betweenness_median,
    betweenness_75,
    betweenness_95
  ) %>%
  pivot_longer(
    cols = -Period,
    names_to = c("Metric", "Statistic"),
    names_pattern = "(.*)_(median|75|95)$",
    values_to = "Value"
  )
```
```{r}
ggplot(
  node_distribution_long,
  aes(
    x = Period,
    y = Value,
    colour = Statistic,
    linetype = Statistic,
    group = Statistic
  )
) +
  geom_line(linewidth = 0.8) +
  geom_point(size = 2) +
  
  facet_wrap(
    ~ Metric,
    ncol = 3,
    scales = "free_y",
    labeller = labeller(
      Metric = c(
        outdegree = "Out-degree",
        strength = "Out-strength",
        betweenness = "Betweenness centrality"
      )
    )
  ) +
  
  scale_x_continuous(
    breaks = 1:12
  ) +
  
  scale_colour_manual(
    values = c(
      median = "red",
      `75` = "forestgreen",
      `95` = "blue"
    ),
    labels = c(
      median = "Median",
      `75` = "75th percentile",
      `95` = "95th percentile"
    )
  ) +
  
  scale_linetype_manual(
    values = c(
      median = "solid",
      `75` = "dashed",
      `95` = "solid"
    ),
    labels = c(
      median = "Median",
      `75` = "75th percentile",
      `95` = "95th percentile"
    )
  ) +
  
  labs(
    x = "Period",
    y = NULL,
    colour = NULL,
    linetype = NULL
  ) +
  
  theme_classic() +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  )
```
```{r}
# check in degree to see why do many farms without outward movement included
monthly_inout <- lapply(monthly_graphs, function(g) {
  data.frame(
    node = V(g)$name,
    indegree = degree(g, mode = "in"),
    outdegree = degree(g, mode = "out")
  )
})
```
```{r}
monthly_inout
```
```{r}
sapply(monthly_inout, function(m) {
  sum(m$indegree >= 1 & m$outdegree == 0)
})
```
```{r}
# identical in terms of which farms are zero
monthly_node_data %>%
  summarise(
    same = all(
      (outdegree == 0) == (strength_out == 0)
    )
  )
```
# Identify structurally important farms

OUTDEGREE

```{r}
# define what counts as structuraly important - justify top 5% threshold using ranked centrality distributions and visible elbow/cut off
# ranked out degree plot to look at 5% threshold
par(mfrow = c(3,4), mar = c(4,4,3,1)) # plotting window for grid

for(i in seq_along(monthly_metrics)){ # loops through each months network metrics stored in monthly_metrics

  m <- monthly_metrics[[i]]

  deg <- sort(m$outdegree, decreasing = TRUE) # sorts farms descending by OD

  plot(deg,
       type = "l",
       lwd = 2,
       xlab = "Farm rank",
       ylab = "Out-degree",
       main = paste("Period", i))
  
 # cutoff <- quantile(deg, 0.95) # this makes top 5% by quantile not count - DIFFERENT be sure to include in methodolgy
  
  #cutoff_rank <- max(which(deg >= cutoff))

  cutoff_rank <- ceiling(0.05 * length(deg)) # calculates the rank corresponding to the top 5% of farms - by count as per rest of analysis

  abline(v = cutoff_rank, col = "red", lty = 2, lwd = 2) # vertical dashed line a 5% threshold
}
```

```{r}
# colour look-up
all_nodes <- unique(unlist(lapply(monthly_metrics, function(m) m$node)))
node_cols <- setNames(distinctColorPalette(length(all_nodes)), all_nodes)
```
Identify structurally important farms through time
which farms are structurally important at each point in time
```{r}
# identify top 5% outdegree rank each month + plot - selects the top 5% of farms numerically by rank
barplots_outdeg <- lapply(seq_along(monthly_metrics), function(i) {

  m <- monthly_metrics[[i]]

  n_top <- ceiling(0.05 * nrow(m))

  top_DO <- m[order(-m$outdegree), ][1:n_top, ]
  top_DO <- top_DO[order(-top_DO$outdegree), ]

  barplot(
    top_DO$outdegree,
    names.arg = top_DO$node,
    las = 2,
    col = node_cols[top_DO$node],
    main = paste("Top 5% Farms by Outdegree - Period", i),
    ylab = "Number of Farms Sent Cattle To",
    xlab = "Farm ID"
  )
})
```
```{r}
# AI prompt - these are my top 5% plots for outdegree - give me code to display in a panel of (3,4), maintaining the axis labels per plot

png( # gets around plot to large issue - save directly
  "top5_outdegree_12_periods.png",
  width = 3600,
  height = 2400,
  res = 300
)

par(mfrow = c(3, 4),
    mar = c(5, 4.1, 3, 0.5)
    #oma = c(0,0,1,0)
)

for (i in seq_along(monthly_metrics)) {
  
  m <- monthly_metrics[[i]]
  
  n_top <- ceiling(0.05 * nrow(m))
  
  top_DO <- m[order(-m$outdegree), ][1:n_top, ]
  top_DO <- top_DO[order(-top_DO$outdegree), ]
  
  barplot(
    top_DO$outdegree,
    names.arg = top_DO$node,
    las = 2,
    col = node_cols[top_DO$node],
    main = paste("Period", i),
    ylab = "Outdegree",
    xlab = "Farm ID",
    cex.names = 0.7, # farm id labels
    cex.axis = 0.8, # y-axis numbers
    cex.lab = 1.0, # x/y axis labels
    cex.main = 1.2 # period titles
  )
}
dev.off()
# reset plotting layout afterwards
par(mfrow = c(1, 1))
```
As a table:
```{r}
# create peiods top 5% table
monthly_top5_outdeg <- lapply(monthly_metrics, function(m) {
  n_top <- ceiling(0.05 * nrow(m))
  top <- m[order(-m$outdegree), ][1:n_top, ]
  top$node
})
```
```{r}
# create table
max_len <- max(sapply(monthly_top5_outdeg, length))

monthly_top5_padded <- lapply(monthly_top5_outdeg, function(x) {
  c(x, rep(NA, max_len - length(x)))
})

top5_table <- do.call(cbind, monthly_top5_padded)
colnames(top5_table) <- paste0("Period_", seq_along(monthly_top5_padded))
top5_table
```
```{r}
# persistence analysis
# convert monthly rankings into longitudinal measure
top5_long <- data.frame(
  month = rep(1:12, sapply(monthly_top5_outdeg, length)),
  node = unlist(monthly_top5_outdeg)
)

counts <- table(top5_long$node)
counts_sorted <- sort(counts, decreasing = TRUE)
persistence_distribution <- table(counts)
counts_sorted
persistence_distribution
```
```{r}
# persistence plot - in how many of the 12 periods does each farm appear in the top 5%

png(
  "outdegree_persistence.png",
  width = 3000,
  height = 2000,
  res = 300
)

barplot(
  counts_sorted,
  las = 2,
  cex.names = 1.0, # farm IDs
  col = "forestgreen",
  main = "Outdegree Persistence",
  ylab = "Periods in Top 5%",
  xlab = "Farm ID",
  cex.lab = 1.4,# x & y axis labels
  cex.main = 3, # title
  cex.axis = 1 # numbers on y axis
)
```
e.g. a farm appearing in 1 period = highly transient/volatile
many months = persistent structural importance etc
```{r}
# persistence distribution plot - shows distribution of temporal persistence across the farm population - from thic can describe whether structural importance is predominantly persistent or transient

persistence_df <- data.frame(
  months = as.numeric(names(persistence_distribution)),
  farms = as.numeric(persistence_distribution)
)

ggplot(persistence_df, aes(x = months, y = farms)) +
  geom_col() +
  labs(
    x = "Number of months appearing in top 5% outdegree",
    y = "Number of farms",
    title = "Persistence distribution of high outdegree farms"
  ) +
  theme_classic()
```
^^^ do you want this in final writeup?

Create lists for scenarios in RQ2
```{r}
OD_persistence <- data.frame(
  farm = names(counts), # farm IDs
  months_top5 = as.numeric(counts) # numebr of months each farm appeared in the top 5% outdegree list
)
```
```{r}
# farms in top 5% OD 8 or more months
OD_high_persistence <- OD_persistence %>%
  filter(months_top5 >= 8) %>%
  pull(farm)
# saved in current working directory
saveRDS(
  OD_high_persistence,
  "outdegree_high_persistent.rds"
)

OD_high_persistence
length(OD_high_persistence) # number removed
```
```{r}
# Farms in top 5% OD & appear in 6 or more months
OD_med_persistence <- OD_persistence %>%
  filter(months_top5 >= 6) %>%
  pull(farm)

saveRDS(
  OD_med_persistence, "Outdegree_med_persistent.rds"
)

OD_med_persistence
length(OD_med_persistence) # number removed
```
```{r}
# Farms in top 5% OD & appear in any months
OD_all_candidates <- OD_persistence %>%
  filter(months_top5 >= 1) %>%
  pull(farm)

saveRDS(
  OD_all_candidates, "outdegree_all_candidates.rds"
)

OD_all_candidates
length(OD_all_candidates) # number removed
```

BETWEENNESS

repeat above pipeline (limitation of months 1-5 have betweenness = 0 for all farms).
Remember thresholds are not statistically validated thresholds = analytical persistence thresholds used to create candidate groups for RQ2 analysis. 

```{r}
# ranked BETWEENNESS plot to look at 5% threshold
# red line = the rank corresponding to the top 5% of farms

par(mfrow = c(3,4), mar = c(4,4,3,1)) # plotting window for grid

for(i in seq_along(monthly_metrics)){ # loops through each months network metrics stored in monthly_metrics

  m_BT <- monthly_metrics[[i]]

  deg_BT <- sort(m_BT$betweenness, decreasing = TRUE) # sorts farms descending by BT

  plot(deg_BT,
       type = "l",
       lwd = 2,
       xlab = "Farm rank",
       ylab = "Betweenness",
       main = paste("Month", i))
  
 # cutoff <- quantile(deg, 0.95) # this makes top 5% by quantile not count - DIFFERENT be sure to include in methodolgy
  
  #cutoff_rank <- max(which(deg >= cutoff))

  cutoff_rank_BT <- ceiling(0.05 * length(deg_BT)) # calculates the rank corresponding to the top 5% of farms - by count as per rest of analysis

  abline(v = cutoff_rank_BT, col = "red", lty = 2, lwd = 2) # vertical dashed line a 5% threshold
}
```
```{r}
# Create colour lookup for all nodes across all months
all_nodes <- unique(unlist(lapply(monthly_metrics, function(m) m$node)))
node_cols <- setNames(distinctColorPalette(length(all_nodes)), all_nodes)
```

```{r}
# Barplots for top 5% by betweenness
barplots_betw <- lapply(seq_along(monthly_metrics), function(i) {

  m <- monthly_metrics[[i]]

  n_top <- ceiling(0.05 * nrow(m))

  # Order by betweenness (descending)
  top_BW <- m[order(-m$betweenness), ][1:n_top, ]
  top_BW <- top_BW[order(-top_BW$betweenness), ]

  barplot(
    top_BW$betweenness,
    names.arg = top_BW$node,
    las = 2,
    col = node_cols[top_BW$node],
    main = paste("Betweenness Top 5% Farms - Period", i),
    ylab = "Betweenness Centrality",
    xlab = "Farm ID"
  )
})
```
```{r}
png(
  "top5_betweenness_periods6-12.png",
  width = 3600,
  height = 2400,
  res = 300
)

par(
  mfrow = c(2, 4),
  mar = c(5, 4.1, 3, 0.5)
)


for (i in 6:12) {

  m <- monthly_metrics[[i]]

  n_top <- ceiling(0.05 * nrow(m))

  # Order by betweenness (descending)
  top_BW <- m[order(-m$betweenness), ][1:n_top, ]
  top_BW <- top_BW[order(-top_BW$betweenness), ]

  barplot(
    top_BW$betweenness,
    names.arg = top_BW$node,
    las = 2,
    col = node_cols[top_BW$node],
    main = paste("Period", i),
    ylab = "Betweenness",
    xlab = "Farm ID",
    cex.names = 0.7,
    cex.axis = 0.8,
    cex.lab = 1.0,
    cex.main = 1.2
  )
}

dev.off()

par(mfrow = c(1, 1))
```

Code here corrected to remove false top 5% farms that were generated in first 5 months when there is no betweenness - for original code & debugs see RQ1 notebook
```{r}
# as table
# top 5% for betweeness
monthly_top5_betw2 <- lapply(monthly_metrics, function(m) {
  
  if(max(m$betweenness) == 0) {
    return(character(0))
  }
  
  n_top <- ceiling(0.05 * nrow(m))
  top <- m[order(-m$betweenness), ][1:n_top, ]
  top$node
})
```
```{r}
# pad months to equal length
max_len <- max(sapply(monthly_top5_betw2, length))

# add NAs to shorter monthly lists so every month has same number rows
monthly_top5_padded2 <- lapply(monthly_top5_betw2, function(x) {
  c(x, rep(NA, max_len - length(x)))
})
```
```{r}
# build comparison table
top5_table2 <- do.call(cbind, monthly_top5_padded2)
colnames(top5_table2) <- paste0("Period_", seq_along(monthly_top5_padded2))
top5_table2
```
```{r}
top5_long_BT2 <- data.frame(
  month = rep(1:12, sapply(monthly_top5_betw2, length)),
  node  = unlist(monthly_top5_betw2)
)

counts_BT2 <- table(top5_long_BT2$node)
counts_BT_sorted2 <- sort(counts_BT2, decreasing = TRUE)
persistence_distribution_BT2 <- table(counts_BT2)
counts_BT2
persistence_distribution_BT2
```
```{r}
# persistence plot
barplot(
  counts_BT_sorted2,
  las = 2,
  cex.names = 0.4,
  col = "forestgreen",
  main = "Farm Persistence in Top 5% Betweeness",
  ylab = "Number of Periods Appearing in Top 5%",
  xlab = "Farm ID"
)

```
```{r}
# persistence plot - in how many of the 12 periods does each farm appear in the top 5% for months 6-12 when BT present

png(
  "betweenness_persistence.png",
  width = 3000,
  height = 2000,
  res = 300
)

barplot(
  counts_BT_sorted2,
  las = 2,
  cex.names = 1.0, # farm IDs
  col = "forestgreen",
  main = "Betweenness Persistence",
  ylab = "Periods in Top 5%",
  xlab = "Farm ID",
  cex.lab = 1.4,# x & y axis labels
  cex.main = 3, # title
  cex.axis = 1 # numbers on y axis
)
```
```{r}
# Persistence distribution of betweenness farms
persistence_df_BT2 <- data.frame(
  months = as.numeric(names(persistence_distribution_BT2)),
  farms = as.numeric(persistence_distribution_BT2)
)

ggplot(persistence_df_BT2, aes(x = months, y = farms)) +
  geom_col() +
  labs(
    x = "Number of periods appearing in top 5% betweeness",
    y = "Number of farms",
    title = "Persistence distribution of high betweeness farms"
  ) +
  theme_classic()
```
# Create data for RQ2

```{r}
BT_persistence2 <- data.frame(
  farm = names(counts_BT2), # farm IDs
  months_top5 = as.numeric(counts_BT2) # numebr of months each farm appeared in the top 5% betweeness list
)
```

```{r}
# farms in top 5% BT >= 6 months
BT_high_persistence2 <- BT_persistence2 %>%
  filter(months_top5 >= 6) %>%
  pull(farm)


BT_high_persistence2
length(BT_high_persistence2) # numebr removed
```
```{r}
# Farms in top 5% BT & appear in 4 or more months
BT_med_persistence2 <- BT_persistence2 %>%
  filter(months_top5 >= 4) %>%
  pull(farm)

BT_med_persistence2
length(BT_med_persistence2) # number removed
```
```{r}
# Farms in top 5% BT & appear in any months
BT_any_persistence2 <- BT_persistence2 %>%
  filter(months_top5 >= 1) %>%
  pull(farm)

BT_any_persistence2
length(BT_any_persistence2) # number removed
```
```{r}
# saved in current working directory
saveRDS(BT_high_persistence2, "BT_high_persistent_corrected.rds")
saveRDS(BT_med_persistence2, "BT_med_persistent_corrected.rds")
saveRDS(BT_any_persistence2, "BT_any_persistent_corrected.rds")
```

What these analyses show:

  threshold analysis- why was top 5% selected as the definitionof structural importance
  
  period-specific analysis - which farms were structurally important in each period
  
  persistence analysis - to what extent was structural importance persistent or transient across the study period
  
# Multidimensional 

standardise OD/strength/BT
parallel-coordinate plots
examine how farms identified by each metric behave across the other metrics
demonstrate that structural importannce is multidimensional

https://search.r-project.org/CRAN/refmans/GGally/html/ggparcoord.html
https://search.r-project.org/CRAN/refmans/lehuynh/html/MinMaxScaling.html
file:///C:/Users/errai/Downloads/Visual_analysis_of_network_centralities.pdf & https://www.researchgate.net/publication/221536310_Visual_analysis_of_network_centralities (webpage)

```{r}
# standardise the three node-level metrics - min-max scaling within each obs period
monthly_metrics_scaled <- lapply(monthly_metrics_nodes, function(m) {

  m$outdeg_scaled <- (m$outdegree - min(m$outdegree)) /
    (max(m$outdegree) - min(m$outdegree))

  m$strength_scaled <- (m$strength_out - min(m$strength_out)) /
    (max(m$strength_out) - min(m$strength_out))

  # Betweenness can be zero for all farms in early periods
  if (max(m$betweenness) == min(m$betweenness)) {
    m$betweenness_scaled <- 0
  } else {
    m$betweenness_scaled <- (m$betweenness - min(m$betweenness)) /
      (max(m$betweenness) - min(m$betweenness))
  }

  m
})
```
```{r}
# make the general parallel coordinate plots
monthly_plots <- lapply(monthly_metrics_scaled, function(m) {

  df_long <- m |>
    pivot_longer(
      cols = c(outdeg_scaled, strength_scaled, betweenness_scaled),
      names_to = "metric",
      values_to = "value"
    )

  ggplot(
    df_long,
    aes(x = metric, y = value, group = node)
  ) +
    geom_line(alpha = 0.4) +
    theme_minimal()
})
```
```{r}
# example
monthly_plots[[8]]
```
```{r}
# highlight the top 5% oudegree farms by farm rank/count

monthly_top5_OD_multi <- lapply(monthly_metrics_scaled, function(m) {

  n_top <- ceiling(0.05 * nrow(m))

  top_nodes <- m$node[
    order(m$outdegree, decreasing = TRUE)[1:n_top]
  ]

  m$top5_OD <- m$node %in% top_nodes

  m
})
```
```{r}
# make highlighted parallel plots - OUTDEGREE
monthly_plots_ODTOP5 <- lapply(monthly_top5_OD_multi, function(m) {

  df_long <- m |>
    pivot_longer(
      cols = c(outdeg_scaled, strength_scaled, betweenness_scaled),
      names_to = "metric",
      values_to = "value"
    )

  ggplot(
    df_long,
    aes(
      x = metric,
      y = value,
      group = node,
      colour = top5_OD
    )
  ) +
    geom_line(alpha = 0.6) +
    scale_colour_manual(
      values = c("FALSE" = "grey70", "TRUE" = "red")
    ) +
    theme_minimal() +
    ggtitle("Top 5% Farms Highlighted — Outdegree")
})
```
```{r}
# example
monthly_plots_ODTOP5[[8]]
```
```{r}
# AI prompt: this is my code to make the outdegree upper 5% highlighted plots - give me the r code to put periods 6-12 in one panel
# make panel of periods 6-12

png(
  "parallel_outdegree_top5_periods6-12.png",
  width = 4200,
  height = 2600,
  res = 300
)

OD_top5_panel <- wrap_plots(
  monthly_plots_ODTOP5[6:12],
  ncol = 4
)

OD_top5_panel

dev.off()
```
```{r}
# Highlighted parallel plots - OUTDEGREE - recreate for panel plot

monthly_plots_ODTOP5 <- lapply(
  seq_along(monthly_top5_OD_multi),
  function(i) {
    
    m <- monthly_top5_OD_multi[[i]]
    
    df_long <- m |>
      pivot_longer(
        cols = c(outdeg_scaled, strength_scaled, betweenness_scaled),
        names_to = "metric",
        values_to = "value"
      ) |>
      mutate(
        metric = recode(
          metric,
          outdeg_scaled = "Outdegree",
          strength_scaled = "Out-strength",
          betweenness_scaled = "Betweenness"
        )
      )
    
    ggplot(
      df_long,
      aes(
        x = metric,
        y = value,
        group = node,
        colour = top5_OD
      )
    ) +
      geom_line(alpha = 0.6, linewidth = 0.5) +
      
      scale_colour_manual(
        values = c(
          "FALSE" = "grey75",
          "TRUE" = "forestgreen"
        ),
        guide = "none"
      ) +
      
      scale_y_continuous(
        limits = c(0, 1),
        expand = c(0, 0.03)
      ) +
      
      labs(
        title = paste("Period", i),
        x = NULL,
        y = "Scaled value"
      ) +
      
      theme_minimal() +
      theme(
        plot.title = element_text(
          size = 11,
          hjust = 0.5
        ),
        axis.text.x = element_text(
          size = 12
        ),
        axis.text.y = element_text(
          size = 12
        ),
        axis.title.y = element_text(
          size = 12
        ),
        panel.grid.minor = element_blank(),
        legend.position = "none"
      )
  }
)
```
```{r}
# save just 6:12
png(
  "parallel_outdegree_top5_periods6-12.png",
  width = 3600,
  height = 4200,
  res = 300
)

OD_top5_panel <- wrap_plots(
  monthly_plots_ODTOP5[6:12],
  ncol = 2
)

OD_top5_panel

dev.off()
```

```{r}
# make highlighted parallel plots - BETWEENNESS (with corrected zero handling)

monthly_top5_BT_multi <- lapply(monthly_metrics_scaled, function(m) {

  if (max(m$betweenness) == 0) {
    m$top5_BT <- FALSE
    return(m)
  }

  n_top <- ceiling(0.05 * nrow(m))

  top_nodes <- m$node[
    order(m$betweenness, decreasing = TRUE)[1:n_top]
  ]

  m$top5_BT <- m$node %in% top_nodes

  m
})
```

```{r}
monthly_plots_BTTOP5 <- lapply(monthly_top5_BT_multi, function(m) {

  df_long <- m |>
    pivot_longer(
      cols = c(outdeg_scaled, strength_scaled, betweenness_scaled),
      names_to = "metric",
      values_to = "value"
    )

  ggplot(
    df_long,
    aes(
      x = metric,
      y = value,
      group = node,
      colour = top5_BT
    )
  ) +
    geom_line(alpha = 0.6) +
    scale_colour_manual(
      values = c("FALSE" = "grey70", "TRUE" = "red")
    ) +
    theme_minimal() +
    ggtitle("Top 5% Farms Highlighted — Betweenness")
})
```
```{r}
# example
monthly_plots_BTTOP5[[6]]
```
```{r}
# AI use to make panel plots as per out-degree

# recreate betweenness plots with larger labels

monthly_plots_BTTOP5 <- lapply(
  seq_along(monthly_top5_BT_multi),
  function(i) {
    
    m <- monthly_top5_BT_multi[[i]]
    
    df_long <- m |>
      pivot_longer(
        cols = c(outdeg_scaled, strength_scaled, betweenness_scaled),
        names_to = "metric",
        values_to = "value"
      ) |>
      mutate(
        metric = recode(
          metric,
          outdeg_scaled = "Outdegree",
          strength_scaled = "Out-strength",
          betweenness_scaled = "Betweenness"
        )
      )
    
    ggplot(
      df_long,
      aes(
        x = metric,
        y = value,
        group = node,
        colour = top5_BT
      )
    ) +
      geom_line(alpha = 0.6, linewidth = 0.5) +
      
      scale_colour_manual(
        values = c(
          "FALSE" = "grey70",
          "TRUE" = "forestgreen"
        ),
        guide = "none"
      ) +
      
      scale_y_continuous(
        limits = c(0, 1),
        expand = c(0, 0.03)
      ) +
      
      labs(
        title = paste("Period", i),
        x = NULL,
        y = "Scaled value"
      ) +
      
      theme_minimal() +
      theme(
        plot.title = element_text(size = 11, hjust = 0.5),
        axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 10),
        axis.title.y = element_text(size = 12),
        panel.grid.minor = element_blank()
      )
  }
)
```

```{r}
# put periods 6-12 in panel

png(
  "parallel_betweenness_top5_periods6-12.png",
  width = 3600,
  height = 4200,
  res = 300
)

BT_top5_panel <- wrap_plots(
  monthly_plots_BTTOP5[6:12],
  ncol = 2
)

BT_top5_panel

dev.off()
```

Quantitative:

```{r}
# calculate the overlap for periods 6-12
monthly_overlap <- lapply(6:12, function(i) {
  
  m <- monthly_metrics_nodes[[i]]
  
  # Top 5% by outdegree
  n_top_OD <- ceiling(0.05 * nrow(m))
  top_OD <- m[order(-m$outdegree), "node"]
  top_OD <- top_OD[1:n_top_OD]
  
  # Top 5% by betweenness
  n_top_BT <- ceiling(0.05 * nrow(m))
  top_BT <- m[order(-m$betweenness), "node"]
  top_BT <- top_BT[1:n_top_BT]
  
  # Identify overlap
  both <- intersect(top_OD, top_BT)
  OD_only <- setdiff(top_OD, top_BT)
  BT_only <- setdiff(top_BT, top_OD)
  either <- union(top_OD, top_BT)
  
  data.frame(
    Period = paste0("Period_", i),
    OD_top5 = length(top_OD),
    BT_top5 = length(top_BT),
    Both = length(both),
    OD_only = length(OD_only),
    BT_only = length(BT_only),
    Either = length(either)
  )
})
```
```{r}
# make it a table
overlap_summary <- do.call(rbind, monthly_overlap)

overlap_summary
```
```{r}
# percentage overlap - what proportion of the farms identified by either  metric were identified by both?
overlap_summary$Both_percent_of_either <-
  100 * overlap_summary$Both / overlap_summary$Either

overlap_summary
```
```{r}
# outdegree unique farms
top5_outdegree_farms <- unique(unlist(
  lapply(monthly_metrics, function(m) {
    
    n_top <- ceiling(0.05 * nrow(m))
    
    top_DO <- m[order(-m$outdegree), ][1:n_top, ]
    
    top_DO$node
  })
))

length(top5_outdegree_farms)
```
The same number of farms was selected for the top 5% of each metric in each period. The table shows how many farms were identified by outdegree only, betweenness only, or both measures, with these groups making up the total farms identified by either metric. Od_only & BT_only are always the same because both metrics start with the same size top 5% group and share the same number of overlapping farms - e.g. 6 identified by both; OD-only = 13-6 = 7, and the same for BT-only


In Period 6, 6 of the 20 farms identified in the top 5% by either OD or BT were identified by both measures, while the remaining farms were identified by only one measure (OD-only or BT-only). This indicates that OD and BT captured overlapping but distinct dimensions of structural importance.

results:

  1. overlap table for periods 6-12: period; number of top 5% farms identified by both OD & BT; total identified by either measure; % overlap (both/eitherx 100)
  
  2. one representative period where the distinction between oD and BT is clear - show top 5% farms for one metric highlighted with the three standardised metrics (OD, BT, strength) - visually demonstrate that some structurally important farms have differnet profiles across the measures.

Asking whether structural importance is multidimensional, rather than assuming that a farm important according to one metric is neccesarily important according to the others

DO OUTDEGREE AND BETWEENNESS IDENTIFY THE SAME OR DIFFERENT STRUCTURALLY IMPORTANT FARMS - plots provide a visual check of this

so do the farms identified by outdegree also have high betweenness strength, or does each metric identify different structural positions?

essentially: which farms are structurally important, how persisent that importance is, and whether different centrality measures identify the same or different farms.

-->

RQ2 if they identify different types of farms, compare the dz sim effect of removing outdegree-important vs betweenness-important farms


