---
title: "RQ1"
author: "ESINCLAIR"
date: "2026-07-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#RQ1 
How does a synthetic temporal cattle-movement network generate structural heterogeneity, and what network features characterise highly connected farms (potential superspreaders) over time?

```{r}
# LIBRARIES
library(dplyr)
library(igraph)
```
Using pipeline created in previous network analysis '12 Month Networks.Rmd'

Data - 1 year synthetic cattle movement data, split into 12 equal dataframes using qcut in python. Entry point, packer removed.

# Copy of pipeline function
```{r}
node_colours <- c(
  "Cow-Calf" = "green",
  "Entry Point" = "lightblue",
  "Starter Feedlot" = "purple",
  "Finishing Feedlot" = "orange"
)

# margins
par(mar = c(0,0,0,0), oma = c(2,0,0,0)) # hash this is doing 12 plot layout

plot_movement_network <- function(df, month_label = "XX") {

  # convert IDs to character before building nodes and edges
  df$org_id  <- as.character(df$org_id)
  df$dest_id <- as.character(df$dest_id)

  # 1. Node table
  org_nodes  <- df[, c("org_id", "org_type")]
  colnames(org_nodes) <- c("name", "type")

  dest_nodes <- df[, c("dest_id", "dest_type")]
  colnames(dest_nodes) <- c("name", "type")

  nodes <- unique(rbind(org_nodes, dest_nodes))

  # 2. Edge list
  edges <- df[, c("org_id", "dest_id")]

  # 3. Graph
  g <- graph_from_data_frame(
    d = edges,
    vertices = nodes,
    directed = TRUE
  )

  # 4. Colours
  cols <- node_colours[V(g)$type]
  # margins
  #par(mar = c(0,0,0,0), oma = c(4,0,0,0)) # hash this is doing 12 plot layout

  # 5. Layers
  layer_order <- c("Cow-Calf",
                   "Entry Point",
                   "Starter Feedlot",
                   "Finishing Feedlot")

  V(g)$layer <- as.numeric(factor(V(g)$type, levels = layer_order))

  # Layout
  lay <- layout_with_sugiyama(g, layers = V(g)$layer)$layout

  # Plot
  plot(
    g,
    layout = lay,
    vertex.color = cols,
    vertex.label = V(g)$name,
    vertex.size = 25,
    vertex.label.cex = 0.9,
    edge.arrow.size = 0.5
  )

  # Title
  mtext(
    paste("Monthly Movement Network", month_label),
    side = 3, line = 0.5, cex = 1.5
  )

  # Legend
  legend(
    "topright",
    legend = names(node_colours),
    col = node_colours,
    pch = 19,
    pt.cex = 1.5,
    bty = "n"
  )
  # Return the graph object for metrics
  return(g)
}
```

# Import data #


