---
title: "RQ1"
author: "ESINCLAIR"
date: "2026-07-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#RQ1 
How does a synthetic temporal cattle-movement network generate structural heterogeneity, and what network features characterise highly connected farms (potential superspreaders) over time?

```{r}
# LIBRARIES
library(dplyr)
library(igraph)
```
Using pipeline created in previous network analysis '12 Month Networks.Rmd'

Data - 1 year synthetic cattle movement data, split into 12 equal dataframes using qcut in python. Entry point, packer removed. Aggregated so that all movement events in one day between 2 premises are 1 event with total_cattle

# Copy of GRAPH pipeline function (added edge weights - total cattle moved)
```{r}
node_colours <- c(
  "Cow-Calf" = "green",
  "Entry Point" = "lightblue",
  "Starter Feedlot" = "purple",
  "Finishing Feedlot" = "orange"
)

# margins
par(mar = c(0,0,0,0), oma = c(2,0,0,0)) # hash this is doing 12 plot layout

plot_movement_network <- function(df, month_label = "XX") {

  # convert IDs to character before building nodes and edges
  df$org_id  <- as.character(df$org_id)
  df$dest_id <- as.character(df$dest_id)

  # 1. Node table
  org_nodes  <- df[, c("org_id", "org_type")]
  colnames(org_nodes) <- c("name", "type")

  dest_nodes <- df[, c("dest_id", "dest_type")]
  colnames(dest_nodes) <- c("name", "type")

  nodes <- unique(rbind(org_nodes, dest_nodes))

  # 2. Edge list
  edges <- df[, c("org_id", "dest_id")]

  # 3. Graph
  g <- graph_from_data_frame(
    d = edges,
    vertices = nodes,
    directed = TRUE
  )

  # Add weights
  
  E(g)$weight <- df$total_cattle # 'E(g) is igraphs built-in edge sequence, meaning 'all edges in graph' - so this adds a new attribute called 'weight' to each edge
  
  # 4. Colours
  cols <- node_colours[V(g)$type]
  # margins
  #par(mar = c(0,0,0,0), oma = c(4,0,0,0)) # hash this is doing 12 plot layout

  # 5. Layers
  layer_order <- c("Cow-Calf",
                   "Entry Point",
                   "Starter Feedlot",
                   "Finishing Feedlot")

  V(g)$layer <- as.numeric(factor(V(g)$type, levels = layer_order))

  # Layout
  lay <- layout_with_sugiyama(g, layers = V(g)$layer)$layout

  # Plot
  plot(
    g,
    layout = lay,
    vertex.color = cols,
    vertex.label = V(g)$name,
    vertex.size = 25,
    vertex.label.cex = 0.9,
    edge.arrow.size = 0.5
  )

  # Title
  mtext(
    paste("Monthly Movement Network", month_label),
    side = 3, line = 0.5, cex = 1.5
  )

  # Legend
  legend(
    "topright",
    legend = names(node_colours),
    col = node_colours,
    pch = 19,
    pt.cex = 1.5,
    bty = "n"
  )
  # Return the graph object for metrics
  return(g)
}
```

# Import data #

```{r}
Q12_M1_RQ1 <- read.csv("Q12(D1_RQ1).csv", check.names = TRUE) # tells R to sanitise column names
#head(Q12_M1)
Q12_M2_RQ1 <- read.csv("Q12(D2_RQ1).csv", check.names = TRUE) 
#head(Q12_M2)
Q12_M3_RQ1 <- read.csv("Q12(D3_RQ1).csv", check.names = TRUE) 
#head(Q12_M3)
Q12_M4_RQ1 <- read.csv("Q12(D4_RQ1).csv", check.names = TRUE) 
#head(Q12_M4)
Q12_M5_RQ1 <- read.csv("Q12(D5_RQ1).csv", check.names = TRUE) 
#head(Q12_M5)
Q12_M6_RQ1 <- read.csv("Q12(D6_RQ1).csv", check.names = TRUE) 
#head(Q12_M6)
Q12_M7_RQ1 <- read.csv("Q12(D7_RQ1).csv", check.names = TRUE) 
#head(Q12_M7)
Q12_M8_RQ1 <- read.csv("Q12(D8_RQ1).csv", check.names = TRUE) 
#head(Q12_M18)
Q12_M9_RQ1 <- read.csv("Q12(D9_RQ1).csv", check.names = TRUE) 
#head(Q12_M9)
Q12_M10_RQ1 <- read.csv("Q12(D10_RQ1).csv", check.names = TRUE) 
#head(Q12_M10)
Q12_M11_RQ1 <- read.csv("Q12(D11_RQ1).csv", check.names = TRUE) 
#head(Q12_M11)
Q12_M12_RQ1 <- read.csv("Q12(D12_RQ1).csv", check.names = TRUE) 
#head(Q12_M12)
```
```{r}
head(Q12_M1_RQ1)
```
# Call function for all 12 months #

```{r}
#par(mfrow = c(6, 2), mar = c(0,0,0,0), oma = c(4,0,0,0)) # 3 rows, 4 columns, no margins

g_M1_RQ1 <- plot_movement_network(Q12_M1_RQ1, "Month 1")
g_M2_RQ1 <- plot_movement_network(Q12_M2_RQ1, "Month 2")
g_M3_RQ1 <- plot_movement_network(Q12_M3_RQ1, "Month 3")
g_M4_RQ1 <- plot_movement_network(Q12_M4_RQ1, "Month 4")
g_M5_RQ1 <- plot_movement_network(Q12_M5_RQ1, "Month 5")
g_M6_RQ1 <- plot_movement_network(Q12_M6_RQ1, "Month 6")
g_M7_RQ1 <- plot_movement_network(Q12_M7_RQ1, "Month 7")
g_M8_RQ1 <- plot_movement_network(Q12_M8_RQ1, "Month 8")
g_M9_RQ1 <- plot_movement_network(Q12_M9_RQ1, "Month 9")
g_M10_RQ1 <- plot_movement_network(Q12_M10_RQ1, "Month 10")
g_M11_RQ1 <- plot_movement_network(Q12_M11_RQ1, "Month 11")
g_M12_RQ1 <- plot_movement_network(Q12_M12_RQ1, "Month 12")

#par(mfrow = c(1, 1))   # reset afterwards
```

# compute metrics
```{r}
g <- g_M10_RQ1
d <- Q12_M10_RQ1

c1 <- vcount(g)
c2 <- ecount(g)

c3 <- degree(g, mode = "in") # remains unweighted. Degree = number of neighbours
c4 <- degree(g, mode = "out")

c5 <- nrow(d) # movements = number of rows

c6 <- betweenness(g, weights = E(g)$weight) # include weights so cattle volume taken into account - makes shortest path reflect movement volume - epidemiologically meaningful

c7 <- transitivity(g, type = "average") # clustering coefficient - expect to be zero as heirarchal and directed

c8 <- max(components(g, mode = "weak")$csize) # tells how many farms are reachable ignoring direction

c9 <- strength_in  <- strength(g, mode = "in",  weights = E(g)$weight)

c10 <-strength_out <- strength(g, mode = "out", weights = E(g)$weight)


summary_mets <- data.frame(
  node = c1,
  edge = c2,
  max_indegree = max(c3),
  min_indegree = min(c3),
  max_outdegree = max(c4),
  min_outdegree = min(c4),
  movements = c5,
  median_betweenness = median(c6),
  cluster_avg = c7,
  gwcc_size = c8,
  max_strength_in = max(c9),
  max_strength_out = max(c10)
)

summary_mets # gives network-level averages and extremes for the month - will need node metrics also
```

# Graph Metrics Table for Comparison #

| measurement         | g_M1 | g_M2 | g_M3 | g_M4 | g_M5 | g_M6 | g_M7 | g_M8 | g_M9 | g_M10 | g_M11 | g_M12 |
|---------------------|------|------|------|------|------|------|------|------|------|-------|-------|-------|
| node                |258   |  262 |241   |233   |243   |255   |229  |231    |215   |211   |231      |219       |
| edge                |  1132 |1035 |1035 | 1069  | 1019  |1054 |1080 |1048 | 1047  |1080   |1063     |1020       |
| max indegree        |  34   |  56 |40   |  38   | 33    |29   |22   |22   |33     |59     |54   | 56      |
| min indegree        |    0  |   0 |0    |0      | 0     |0    |0    |0    |0      |0      |0    | 0      |
| max outdegree       |   82   |  71|72   | 54    | 57    |100  |54   |56   |54     |42     |39   | 54      |
| min outdegree       |   0   |  0  |0    |0      | 0     |0    |0    |0    |0      |0      |0    | 0      |
| movements           |      |      | 1035 |1069  | 1019  |1054 |1080 |1048 |1047   |1080 |1063   | 1020      |
| median betweenness  |   0   |  0    |0   | 0    |0      |0    |0    |0    |0      |0    |0      |0       |
| cluster avg         |   0   |   0   | 0   | 0   |0      |0    |0    |0    |0      |0    |0      |0       |
| gwcc size           |  258    | 262 |241  |233  |243    |255  |229  |231  |215    |211  |231    |219       |
| max strength in     |   1999   |828 |652  |1264 |2899   |4448 |3859 |3122 |2191   |5403 |1851 |9077       |
| max strength out    |    7563  | 3313 |3223 |4887 |6505 |18453|10932|10287|7194| 8104 |2398   | 11228      |

# Node level metrics per month
Describe each farm individualy - gives one row per node, not one row per month as above

```{r}

g <- g_M1_RQ1
d <- Q12_M1_RQ1


node_metrics <- data.frame(
  node = V(g)$name,
  type = V(g)$type,
  indegree = degree(g, mode = "in"),
  outdegree = degree(g, mode = "out"),
  strength_in = strength(g, mode = "in", weights = E(g)$weight),
  strength_out = strength(g, mode = "out", weights = E(g)$weight),
  betweenness = betweenness(g, weights = E(g)$weight),
  eigen = eigen_centrality(g, weights = E(g)$weight)$vector,
  clustering = transitivity(g, type = "local")
)
```

```{r}
head(node_metrics)
```
```{r}
# histogram of strength out
hist(node_metrics$strength_out,
     main = "Distribution of Strength Out",
     xlab = "Total Cattle Sent Out",
     breaks = 30,
     col = "lightblue")

```
```{r}
# top farms, barplot of highest values
top <- node_metrics[order(-node_metrics$strength_out), ][1:10, ]

barplot(top$strength_out,
        names.arg = top$node,
        las = 2,
        col = "purple",
        main = "Top 10 Farms by Strength Out M1",
        ylab = "Total Cattle Sent Out")
```
```{r}
# top farms, barplot of highest values
top <- node_metrics[order(-node_metrics$outdegree), ][1:10, ]

barplot(top$outdegree,
        names.arg = top$node,
        las = 2,
        col = "purple",
        main = "Top 10 Farms by Outdegree M1",
        ylab = "Total Cattle Sent Out")
```
Eigenvector centrality (eigen) meaures how well connected a node is to other well-connected nodes - 'influence' sort of score rather than simplecount of conections.
A node will have a high eigen score if: it is connected to many nodes AND those nodes themselves have high connectivity

Degree = how many farms a node connects to
strength = how many cattle a node moves
Betweeness = how often the node sites between others
eigenvector = how influential a ndoes neighbours are - shows structurally influential farms, not just busy ones - i.e. low degree/strength/betweeness but high eigen-> it connects to a very important hub - can amplify disease even if its own movement volume is small

```{r}
# top farms, barplot of highest values
top <- node_metrics[order(-node_metrics$eigen), ][1:10, ]

barplot(top$eigen,
        names.arg = top$node,
        las = 2,
        col = "purple",
        main = "Top 10 Farms by Eigenvector M1",
        ylab = "Total Cattle Sent Out")
```
What other metrics to visualise - literature

So going to want:

?Superspreaders?

  strength out (weighted out-degree) - total cattle sent - potential infectious contacts
  Out-degree (unweighted) - number of distinct farms they send cattle to - spread reach
  Eigenvector centrality - measures influence through who a node is connected to - a farm selling cattle to important hubs is more dangerous

?Super recievers?

  strength in - total cattle recieved = direct measure of exposure
  in-degree - number of distinct farms they recieve cattle from - more sources = more exposure routes
  Eigen - high = structurally important recievers - they recieve from influencital hubs
  
THEN 
plot distributions, identify top x%, compare farm types & track farms across months - WEDNESDAY - LOOK AT PAPERS IN LIST THAT DO SIMILAR