---
title: "RQ2 RESULTS ANALYSIS ZIP"
author: "ESINCLAIR"
date: "2026-08-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# libraries
library(igraph)
library(ggplot2)
library(dplyr)
library(tidyr)
library(pbapply)
```

CONTAINS GWCC AND THRESHOLD ESTABLISHMENT + ANALYSIS

Import results of simulations for anlysis (RQ2 SOIR SIMS)


```{r}
# Baseline
scenario_1_results <- readRDS("scenario_1_results.rds") # baseline

# Interventions
scenario_2_results <- readRDS("scenario_2_results.rds") # Outdegree (high) >= 8 months
scenario_3_results <- readRDS("scenario_3_results.rds") # Outdegree (med) >= 6 months
scenario_4_results <- readRDS("scenario_4_results.rds") # Outdegree any month
scenario_5_results <- readRDS("scenario_5_results_corrected.rds") # Betweenness (high) >= 6 months
scenario_6_results <- readRDS("scenario_6_results_corrected.rds") # Betweenness (med) >= 4 months
scenario_7_results <- readRDS("scenario_7_results_corrected.rds") # Betweenness any month
```
```{r}
# Matched controls - Scenario 8: Control OD HIGH x10 (5 random farms removed)
scenario_8a_results <- readRDS("scenario_8P_control_1_results.rds")
scenario_8b_results <- readRDS("scenario_8P_control_2_results.rds")
scenario_8c_results <- readRDS("scenario_8P_control_3_results.rds")
scenario_8d_results <- readRDS("scenario_8P_control_4_results.rds")
scenario_8e_results <- readRDS("scenario_8P_control_5_results.rds")
scenario_8f_results <- readRDS("scenario_8P_control_6_results.rds")
scenario_8g_results <- readRDS("scenario_8P_control_7_results.rds")
scenario_8h_results <- readRDS("scenario_8P_control_8_results.rds")
scenario_8i_results <- readRDS("scenario_8P_control_9_results.rds")
scenario_8j_results <- readRDS("scenario_8P_control_10_results.rds")
  
```  
```{r}
# list of 2,000 simulation replicates & summary dataframe with each replicate
#class(scenario_8a_results)
#str(scenario_8a_results)
```
```{r}
# Matched controls - Scenario 9 Control OD Med x10, (random removal of 8 farms)
scenario_9a_results <- readRDS("scenario_9P_control_1_results.rds")
scenario_9b_results <- readRDS("scenario_9P_control_2_results.rds")
scenario_9c_results <- readRDS("scenario_9P_control_3_results.rds")
scenario_9d_results <- readRDS("scenario_9P_control_4_results.rds")
scenario_9e_results <- readRDS("scenario_9P_control_5_results.rds")
scenario_9f_results <- readRDS("scenario_9P_control_6_results.rds")
scenario_9g_results <- readRDS("scenario_9P_control_7_results.rds")
scenario_9h_results <- readRDS("scenario_9P_control_8_results.rds")
scenario_9i_results <- readRDS("scenario_9P_control_9_results.rds")
scenario_9j_results <- readRDS("scenario_9P_control_10_results.rds")
```
```{r}
# Matched controls - Scenario 10 Control OD all x10, (random removal of 43 farms)
scenario_10a_results <- readRDS("scenario_10P_control_1_results.rds")
scenario_10b_results <- readRDS("scenario_10P_control_2_results.rds")
scenario_10c_results <- readRDS("scenario_10P_control_3_results.rds")
scenario_10d_results <- readRDS("scenario_10P_control_4_results.rds")
scenario_10e_results <- readRDS("scenario_10P_control_5_results.rds")
scenario_10f_results <- readRDS("scenario_10P_control_6_results.rds")
scenario_10g_results <- readRDS("scenario_10P_control_7_results.rds")
scenario_10h_results <- readRDS("scenario_10P_control_8_results.rds")
scenario_10i_results <- readRDS("scenario_10P_control_9_results.rds")
scenario_10j_results <- readRDS("scenario_10P_control_10_results.rds")
```
```{r}
# Matched controls - Scenario 11 Control BT High x10, (random removal of 1 farms)
scenario_11a_results <- readRDS("scenario_11P_control_1_results.rds")
scenario_11b_results <- readRDS("scenario_11P_control_2_results.rds")
scenario_11c_results <- readRDS("scenario_11P_control_3_results.rds")
scenario_11d_results <- readRDS("scenario_11P_control_4_results.rds")
scenario_11e_results <- readRDS("scenario_11P_control_5_results.rds")
scenario_11f_results <- readRDS("scenario_11P_control_6_results.rds")
scenario_11g_results <- readRDS("scenario_11P_control_7_results.rds")
scenario_11h_results <- readRDS("scenario_11P_control_8_results.rds")
scenario_11i_results <- readRDS("scenario_11P_control_9_results.rds")
scenario_11j_results <- readRDS("scenario_11P_control_10_results.rds")
```
```{r}
# Matched controls - Scenario 12 Control BT Med x10, (random removal of 18 farms)
scenario_12a_results <- readRDS("scenario_12P_control_1_results.rds")
scenario_12b_results <- readRDS("scenario_12P_control_2_results.rds")
scenario_12c_results <- readRDS("scenario_12P_control_3_results.rds")
scenario_12d_results <- readRDS("scenario_12P_control_4_results.rds")
scenario_12e_results <- readRDS("scenario_12P_control_5_results.rds")
scenario_12f_results <- readRDS("scenario_12P_control_6_results.rds")
scenario_12g_results <- readRDS("scenario_12P_control_7_results.rds")
scenario_12h_results <- readRDS("scenario_12P_control_8_results.rds")
scenario_12i_results <- readRDS("scenario_12P_control_9_results.rds")
scenario_12j_results <- readRDS("scenario_12P_control_10_results.rds")
```
```{r}
# Matched controls - Scenario 13 Control BT All x10, (random removal of 35 farms)
scenario_13a_results <- readRDS("scenario_13P_control_1_results.rds")
scenario_13b_results <- readRDS("scenario_13P_control_2_results.rds")
scenario_13c_results <- readRDS("scenario_13P_control_3_results.rds")
scenario_13d_results <- readRDS("scenario_13P_control_4_results.rds")
scenario_13e_results <- readRDS("scenario_13P_control_5_results.rds")
scenario_13f_results <- readRDS("scenario_13P_control_6_results.rds")
scenario_13g_results <- readRDS("scenario_13P_control_7_results.rds")
scenario_13h_results <- readRDS("scenario_13P_control_8_results.rds")
scenario_13i_results <- readRDS("scenario_13P_control_9_results.rds")
scenario_13j_results <- readRDS("scenario_13P_control_10_results.rds")
```
  
# Pooled control & between-control variation

pooled control - 20,000 stochastic outcomes - gives the matched control estimate (median/mean/IQR)
between-control variation - 10 separate 2000-run simulations - demonstrates that the pooled result is reasonably consitent across the random farms selected

```{r}
# SCENARIO 8
# pool the 20,000 outcomes

control_8_results_pooled <- list(
  scenario_8a_results,
  scenario_8b_results,
  scenario_8c_results,
  scenario_8d_results,
  scenario_8e_results,
  scenario_8f_results,
  scenario_8g_results,
  scenario_8h_results,
  scenario_8i_results,
  scenario_8j_results
)

control_8_finalsizes <- unlist(
  lapply(control_8_results_pooled, function(x) {
    sapply(x, function(i) i$summary$final_infected)
  })
)

control_8_peakinfectious <- unlist(
  lapply(control_8_results_pooled, function(x) {
    sapply(x, function(i) i$summary$peak_infectious)
  })
)
```

```{r}
length(control_8_finalsizes)
```
```{r}
# pooled summary

control_8_summary_pooled <- data.frame(
  N = length(control_8_finalsizes),
  Mean = mean(control_8_finalsizes),
  Median = median(control_8_finalsizes),
  IQR = IQR(control_8_finalsizes),
  SD = sd(control_8_finalsizes),
  Min = min(control_8_finalsizes),
  Max = max(control_8_finalsizes)
)

control_8_summary_pooled
```
```{r}
# show isnt driven by one random control - calculate the summary repeatedly for each of the 10 controls

control_8_variance <- do.call(
  rbind,
  lapply(seq_along(control_8_results_pooled), function(i) {
    
    x <- sapply(
      control_8_results_pooled[[i]],
      function(j) j$summary$final_infected
    )
    
    data.frame(
      Control = i,
      Mean = mean(x),
      Median = median(x),
      IQR = IQR(x)
    )
  })
)

control_8_variance
```
```{r}
ggplot(control_8_variance, aes(x = Control, y = Mean)) +
  geom_point(size = 3) +
  geom_hline(
    yintercept = control_8_summary_pooled$Mean,
    linetype = "dashed"
  ) +
  theme_bw() +
  labs(
    title = "Scenario 8: variation in mean final outbreak size",
    x = "Random control simulation",
    y = "Mean final outbreak size"
  )
```
State - across the ten random control configurations, the median final outbreak size was 1 (IQR= 0) in all configurations, while mean oubtreak sizes were also similar, indicating limited variation between random control configurations - APPENDIX

```{r}
# SCENARIO 9
# pool the 20,000 outcomes

control_9_results_pooled <- list(
  scenario_9a_results,
  scenario_9b_results,
  scenario_9c_results,
  scenario_9d_results,
  scenario_9e_results,
  scenario_9f_results,
  scenario_9g_results,
  scenario_9h_results,
  scenario_9i_results,
  scenario_9j_results
)

control_9_finalsizes <- unlist(
  lapply(control_9_results_pooled, function(x) {
    sapply(x, function(i) i$summary$final_infected)
  })
)

control_9_peakinfectious <- unlist(
  lapply(control_9_results_pooled, function(x) {
    sapply(x, function(i) i$summary$peak_infectious)
  })
)
```

```{r}
length(control_9_finalsizes)
```
```{r}
# pooled summary

control_9_summary_pooled <- data.frame(
  N = length(control_9_finalsizes),
  Mean = mean(control_9_finalsizes),
  Median = median(control_9_finalsizes),
  IQR = IQR(control_9_finalsizes),
  SD = sd(control_9_finalsizes),
  Min = min(control_9_finalsizes),
  Max = max(control_9_finalsizes)
)

control_9_summary_pooled
```
```{r}
# show isnt driven by one random control - calculate the summary repeatedly for each of the 10 controls

control_9_variance <- do.call(
  rbind,
  lapply(seq_along(control_9_results_pooled), function(i) {
    
    x <- sapply(
      control_9_results_pooled[[i]],
      function(j) j$summary$final_infected
    )
    
    data.frame(
      Control = i,
      Mean = mean(x),
      Median = median(x),
      IQR = IQR(x)
    )
  })
)

control_9_variance
```
```{r}
ggplot(control_9_variance, aes(x = Control, y = Mean)) +
  geom_point(size = 3) +
  geom_hline(
    yintercept = control_9_summary_pooled$Mean,
    linetype = "dashed"
  ) +
  theme_bw() +
  labs(
    title = "Scenario 9: variation in mean final outbreak size",
    x = "Random control simulation",
    y = "Mean final outbreak size"
  )
```


```{r}
# SCENARIO 10
# pool the 20,000 outcomes

control_10_results_pooled <- list(
  scenario_10a_results,
  scenario_10b_results,
  scenario_10c_results,
  scenario_10d_results,
  scenario_10e_results,
  scenario_10f_results,
  scenario_10g_results,
  scenario_10h_results,
  scenario_10i_results,
  scenario_10j_results
)

control_10_finalsizes <- unlist(
  lapply(control_10_results_pooled, function(x) {
    sapply(x, function(i) i$summary$final_infected)
  })
)

control_10_peakinfectious <- unlist(
  lapply(control_10_results_pooled, function(x) {
    sapply(x, function(i) i$summary$peak_infectious)
  })
)
```

```{r}
length(control_10_finalsizes)
```
```{r}
# pooled summary

control_10_summary_pooled <- data.frame(
  N = length(control_10_finalsizes),
  Mean = mean(control_10_finalsizes),
  Median = median(control_10_finalsizes),
  IQR = IQR(control_10_finalsizes),
  SD = sd(control_10_finalsizes),
  Min = min(control_10_finalsizes),
  Max = max(control_10_finalsizes)
)

control_10_summary_pooled
```
```{r}
# show isnt driven by one random control - calculate the summary repeatedly for each of the 10 controls

control_10_variance <- do.call(
  rbind,
  lapply(seq_along(control_10_results_pooled), function(i) {
    
    x <- sapply(
      control_10_results_pooled[[i]],
      function(j) j$summary$final_infected
    )
    
    data.frame(
      Control = i,
      Mean = mean(x),
      Median = median(x),
      IQR = IQR(x)
    )
  })
)

control_10_variance
```
```{r}
ggplot(control_10_variance, aes(x = Control, y = Mean)) +
  geom_point(size = 3) +
  geom_hline(
    yintercept = control_10_summary_pooled$Mean,
    linetype = "dashed"
  ) +
  theme_bw() +
  labs(
    title = "Scenario 10: variation in mean final outbreak size",
    x = "Random control simulation",
    y = "Mean final outbreak size"
  )
```

```{r}
# SCENARIO 11
# pool the 20,000 outcomes

control_11_results_pooled <- list(
  scenario_11a_results,
  scenario_11b_results,
  scenario_11c_results,
  scenario_11d_results,
  scenario_11e_results,
  scenario_11f_results,
  scenario_11g_results,
  scenario_11h_results,
  scenario_11i_results,
  scenario_11j_results
)

control_11_finalsizes <- unlist(
  lapply(control_11_results_pooled, function(x) {
    sapply(x, function(i) i$summary$final_infected)
  })
)

control_11_peakinfectious <- unlist(
  lapply(control_11_results_pooled, function(x) {
    sapply(x, function(i) i$summary$peak_infectious)
  })
)
```

```{r}
length(control_11_finalsizes)
```
```{r}
# pooled summary

control_11_summary_pooled <- data.frame(
  N = length(control_11_finalsizes),
  Mean = mean(control_11_finalsizes),
  Median = median(control_11_finalsizes),
  IQR = IQR(control_11_finalsizes),
  SD = sd(control_11_finalsizes),
  Min = min(control_11_finalsizes),
  Max = max(control_11_finalsizes)
)

control_11_summary_pooled
```
```{r}
# show isnt driven by one random control - calculate the summary repeatedly for each of the 10 controls

control_11_variance <- do.call(
  rbind,
  lapply(seq_along(control_11_results_pooled), function(i) {
    
    x <- sapply(
      control_11_results_pooled[[i]],
      function(j) j$summary$final_infected
    )
    
    data.frame(
      Control = i,
      Mean = mean(x),
      Median = median(x),
      IQR = IQR(x)
    )
  })
)

control_11_variance
```
```{r}
ggplot(control_11_variance, aes(x = Control, y = Mean)) +
  geom_point(size = 3) +
  geom_hline(
    yintercept = control_11_summary_pooled$Mean,
    linetype = "dashed"
  ) +
  theme_bw() +
  labs(
    title = "Scenario 11: variation in mean final outbreak size",
    x = "Random control simulation",
    y = "Mean final outbreak size"
  )
```

```{r}
# SCENARIO 12
# pool the 20,000 outcomes

control_12_results_pooled <- list(
  scenario_12a_results,
  scenario_12b_results,
  scenario_12c_results,
  scenario_12d_results,
  scenario_12e_results,
  scenario_12f_results,
  scenario_12g_results,
  scenario_12h_results,
  scenario_12i_results,
  scenario_12j_results
)

control_12_finalsizes <- unlist(
  lapply(control_12_results_pooled, function(x) {
    sapply(x, function(i) i$summary$final_infected)
  })
)

control_12_peakinfectious <- unlist(
  lapply(control_12_results_pooled, function(x) {
    sapply(x, function(i) i$summary$peak_infectious)
  })
)
```

```{r}
length(control_12_finalsizes)
```
```{r}
# pooled summary

control_12_summary_pooled <- data.frame(
  N = length(control_12_finalsizes),
  Mean = mean(control_12_finalsizes),
  Median = median(control_12_finalsizes),
  IQR = IQR(control_12_finalsizes),
  SD = sd(control_12_finalsizes),
  Min = min(control_12_finalsizes),
  Max = max(control_12_finalsizes)
)

control_12_summary_pooled
```
```{r}
# show isnt driven by one random control - calculate the summary repeatedly for each of the 10 controls

control_12_variance <- do.call(
  rbind,
  lapply(seq_along(control_12_results_pooled), function(i) {
    
    x <- sapply(
      control_12_results_pooled[[i]],
      function(j) j$summary$final_infected
    )
    
    data.frame(
      Control = i,
      Mean = mean(x),
      Median = median(x),
      IQR = IQR(x)
    )
  })
)

control_12_variance
```
```{r}
ggplot(control_12_variance, aes(x = Control, y = Mean)) +
  geom_point(size = 3) +
  geom_hline(
    yintercept = control_12_summary_pooled$Mean,
    linetype = "dashed"
  ) +
  theme_bw() +
  labs(
    title = "Scenario 12: variation in mean final outbreak size",
    x = "Random control simulation",
    y = "Mean final outbreak size"
  )
```

```{r}
# SCENARIO 13
# pool the 20,000 outcomes

control_13_results_pooled <- list(
  scenario_13a_results,
  scenario_13b_results,
  scenario_13c_results,
  scenario_13d_results,
  scenario_13e_results,
  scenario_13f_results,
  scenario_13g_results,
  scenario_13h_results,
  scenario_13i_results,
  scenario_13j_results
)

control_13_finalsizes <- unlist(
  lapply(control_13_results_pooled, function(x) {
    sapply(x, function(i) i$summary$final_infected)
  })
)

control_13_peakinfectious <- unlist(
  lapply(control_13_results_pooled, function(x) {
    sapply(x, function(i) i$summary$peak_infectious)
  })
)
```

```{r}
length(control_13_finalsizes)
```
```{r}
# pooled summary

control_13_summary_pooled <- data.frame(
  N = length(control_13_finalsizes),
  Mean = mean(control_13_finalsizes),
  Median = median(control_13_finalsizes),
  IQR = IQR(control_13_finalsizes),
  SD = sd(control_13_finalsizes),
  Min = min(control_13_finalsizes),
  Max = max(control_13_finalsizes)
)

control_13_summary_pooled
```
```{r}
# show isnt driven by one random control - calculate the summary repeatedly for each of the 10 controls

control_13_variance <- do.call(
  rbind,
  lapply(seq_along(control_13_results_pooled), function(i) {
    
    x <- sapply(
      control_13_results_pooled[[i]],
      function(j) j$summary$final_infected
    )
    
    data.frame(
      Control = i,
      Mean = mean(x),
      Median = median(x),
      IQR = IQR(x)
    )
  })
)

control_13_variance
```
```{r}
ggplot(control_13_variance, aes(x = Control, y = Mean)) +
  geom_point(size = 3) +
  geom_hline(
    yintercept = control_13_summary_pooled$Mean,
    linetype = "dashed"
  ) +
  theme_bw() +
  labs(
    title = "Scenario 13: variation in mean final outbreak size",
    x = "Random control simulation",
    y = "Mean final outbreak size"
  )
```

```{r}
# combine the daily outputs from all 2000 sim replicates into one data frame

# daily result - epidemic curves for all 2000 stochastic runs
daily_results_SC1 <- bind_rows( 
  lapply(scenario_1_results, function(x) x$daily) 
)
# peak infectious and final outbreak size per run
summary_results_SC1 <- bind_rows(
  lapply(scenario_1_results, function(x) x$summary)
)
```
```{r}
# Put all scenario results into one list
all_scenarios <- list(
  SC1 = scenario_1_results,
  SC2 = scenario_2_results,
  SC3 = scenario_3_results,
  SC4 = scenario_4_results,
  SC5 = scenario_5_results,
  SC6 = scenario_6_results,
  SC7 = scenario_7_results
)

# Create empty lists to store outputs
daily_results <- list()
summary_results <- list()

# Loop through scenarios
for (scenario in names(all_scenarios)) {
  
  daily_results[[scenario]] <- bind_rows(
    lapply(all_scenarios[[scenario]], function(x) x$daily)
  )
  
  summary_results[[scenario]] <- bind_rows(
    lapply(all_scenarios[[scenario]], function(x) x$summary)
  )
}

# call individual with - summary_results$SC3 etc
```

# Read in data from RQ1 for number farms removed

```{r}
# OUTDEGREE
OD_all_candidates <- readRDS(
  "outdegree_all_candidates.rds"
)

OD_med_candidates <- readRDS(
  "outdegree_med_persistent.rds"
)

OD_high_candidates <- readRDS(
  "outdegree_high_persistent.rds"
)
```

```{r}
#OD_all_candidates
```
```{r}
# BETWEENNESS
BT_all_candidates <- readRDS(
  "BT_any_persistent_corrected.rds"
)

BT_high_candidates <- readRDS(
  "BT_high_persistent_corrected.rds"
)

BT_med_candidates <- readRDS(
  "BT_med_persistent_corrected.rds"
)
```
```{r}
#BT_med_candidates
```

# 1. Create scenario table - what was removed

```{r}
total_farms = 301
scenario_table <- data.frame(
  Scenario = c(
    "(1)",
    "(2)",
    "(3)",
    "(4)",
    "(5)",
    "(6)",
    "(7)",
    "(8)",
    "(9)",
    "(10)",
    "(11)",
    "(12)",
    "(13)"
  ),
  
  Type = c(
    "Baseline",
    "Targeted Intervention",
    "Targeted Intervention",
    "Targeted Intervention",
    "Targeted Intervention",
    "Targeted Intervention",
    "Targeted Intervention",
    "CONTROL POOLED OD HIGH",
    "CONTROL POOLED OD MED",
    "CONTROL POOLED OD ALL",
    "CONTROL POOLED BT HIGH",
    "CONTROL POOLED BT MED",
    "CONTROL POOLED BT ALL"
  ),
  
  Centrality = c(
    "None",
    "Out-degree",
    "Out-degree",
    "Out-degree",
    "Betweenness",
    "Betweenness",
    "Betweenness",
    "RandomOD",
    "RandomOD",
    "RandomOD",
    "RandomBT",
    "RandomBT",
    "RandomBT"
  ),
  
  Persistence = c(
    "None",
    "High >=8 months",
    "Medium >=6 months",
    ">=1 month",
    "High >=6 months",
    "Medium >=4 months",
    ">= 1 month",
    "Matched to OD HIGH",
    "Matched to OD MED",
    "Matched to OD ALL",
    "Matched to BT HIGH",
    "Matched to BT MED",
    "Matched to BT ALL"
  ),
  Number_of_farms_removed = c(
    0,
    length(OD_high_candidates),
    length(OD_med_candidates),
    length(OD_all_candidates),
    length(BT_high_candidates),
    length(BT_med_candidates),
    length(BT_all_candidates),
    5,
    8,
    43,
    1,
    8,
    35
  ),
  
  Percent_of_all_farms = c(
    0,
    length(OD_high_candidates)/total_farms*100,
    length(OD_med_candidates)/total_farms*100,
    length(OD_all_candidates)/total_farms*100,
    length(BT_high_candidates)/total_farms*100,
    length(BT_med_candidates)/total_farms*100,
    length(BT_all_candidates)/total_farms*100,
    5/total_farms*100,
    8/total_farms*100,
    43/total_farms*100,
    1/total_farms*100,
    8/total_farms*100,
    35/total_farms*100
  )
)

scenario_table
```
```{r}
total_farms
```

# 2. show change in largest connected component - GWCC - show removal of potential transmission pathways (undirected) - did removal change the movement network

load and construct the annual network
```{r}
# modify graph/network code from RQ1.Rmd
create_movement_network <- function(df) {

  # Convert IDs
  df$org_id  <- as.character(df$org_id)
  df$dest_id <- as.character(df$dest_id)

  # Node table
  org_nodes <- df[, c("org_id", "org_type")]
  colnames(org_nodes) <- c("name", "type")

  dest_nodes <- df[, c("dest_id", "dest_type")]
  colnames(dest_nodes) <- c("name", "type")

  nodes <- unique(rbind(org_nodes, dest_nodes))

  # Edge list
  edges <- df[, c("org_id", "dest_id")]

  # Create directed network
  g <- graph_from_data_frame(
    d = edges,
    vertices = nodes,
    directed = TRUE
  )

  # Add weights
  E(g)$weight <- df$total_cattle

  return(g)
}
```

```{r}
# import data
annual_movement <- read.csv("RQ1 NO EP.csv")
head(annual_movement)
```


```{r}
# create as graph network by calling function = construct annual movement network
annual_network = create_movement_network(annual_movement)
```
```{r}
# inspect the network
vcount(annual_network)

V(annual_network)$name
```

Calculate baseline GWCC
```{r}
# calculate basline GWCC
baseline_gwcc <- max(components(annual_network, mode = "weak")$csize) # csize = extract number of farms in each connected component
baseline_gwcc
```
Generate random removal controls

```{r}
# create random removal controls
# removal-size controls representing the typical number of farms removed across each persistence category

Control_high_random <- sample(
  V(annual_network)$name,
  size = 4,
  replace = FALSE
)

Control_medium_random <- sample(
  V(annual_network)$name,
  size = 9,
  replace = FALSE
)

Control_all_random <- sample(
  V(annual_network)$name,
  size = 40,
  replace = FALSE
)
# ADD HERE IF WANT TO PUT CONTROLS PER SCENARIO RATHER THAN AVERAGES
```
Define all removal scenarios

```{r}
# create loop to speed up
removal_scenarios <- list(
  S2 = OD_high_candidates,
  S3 = OD_med_candidates,
  S4 = OD_all_candidates,
  S5 = BT_high_candidates,
  S6 = BT_med_candidates,
  S7 = BT_all_candidates
  #S8 = CB_high_candidates,
  #S9 = CB_med_candidates,
  #S10 = CB_all_candidates,
  #S11 = Control_high_random,
  #S12 = Control_medium_random,
  #S13 = Control_all_random,
  #Annual_Bt_FIVE = annual_BT_top5,
  #Annual_OD_FIVE = annual_OD_top5,
  #Annual_OD_deg = annual_OD_top5_deg
  #SC14 = NEED TO ADD THESE
  #SC15 = 
  #SC16 = 
  #SC17 = 
  #SC18 = 
  #SC19 = 
)

```

Create empty results dataframe & apply each removal scenario & calculate network fragmentation

```{r}
gwcc_results <- data.frame(
  Scenario = character(),
  Farms_removed = numeric(),
  GWCC_size = numeric(),
  GWCC_percentage_remaining = numeric(),
  Number_components = numeric()
)

#total_farms_nw = vcount(annual_network)

for (s in names(removal_scenarios)) {
  
  removed_network <- delete_vertices(
    annual_network,
    removal_scenarios[[s]]
  )
  
  comp <- components(removed_network, mode = "weak")
  
  farms_remaining <- vcount(removed_network)
  
  gwcc_size <-max(comp$csize)
  
  gwcc_percent <- (gwcc_size / farms_remaining) * 100 # this is GWCC size (% of farms remaining), NOT percentage of original network
  
  gwcc_results <- rbind(
    gwcc_results,
    data.frame(
      Scenario = s,
      Farms_removed = length(removal_scenarios[[s]]),
      GWCC_size = gwcc_size,
      GWCC_percentage_remaining = gwcc_percent,
      Number_components = comp$no
    )
  )
}
```
Calculate baseline network components again & add the baseline network to the scenario results

```{r}
# calculate baseline network components
baseline_comp <- components(annual_network, mode = "weak")

#baseline_gwcc <- max(baseline_comp$csize)

#baseline_percent <- (baseline_gwcc / vcount(annual_network)) * 100

#baseline_components <- baseline_comp$no

# add baseline GWCC
gwcc_results <- rbind(
  data.frame(
    Scenario = "S1",
    Farms_removed = 0,
    GWCC_size = max(baseline_comp$csize),
    GWCC_percentage_remaining = (max(baseline_comp$csize) / vcount(annual_network))*100,
    Number_components = baseline_comp$no
  ),
  
  gwcc_results
)
```
View final conenctivity results
```{r}
gwcc_results
```
Clear the results object
```{r}
# REMOVE RESULTS
rm(gwcc_results)
```
Final network membership checks
```{r}
sum(annual_OD_top5 %in% V(annual_network)$name)
```
```{r}
length(annual_OD_top5)
length(intersect(annual_OD_top5, V(annual_network)$name))
```

## SOIR Epidemic outcome analysis ##

# Established epidemic definition: threshold sensitivity analysis

1. examine the baseline outbreak-size distribution

  supports why selected >=10 - baseline histogram

```{r}
# plot the distribution

ggplot(summary_results_SC1, aes(x = final_infected)) +
  geom_histogram(
    binwidth = 1,
    colour = "black",
    fill = "steelblue"
  ) +
  labs(
    x = "Final outbreak size (infected farms)",
    y = "Number of simulations",
    title = "Distribution of final outbreak sizes"
  ) +
  theme_classic()
```
Create the scenario outbreak size data

```{r}
summary_results_SC1 <- summary_results$SC1
summary_results_SC2 <- summary_results$SC2
summary_results_SC3 <- summary_results$SC3
summary_results_SC4 <- summary_results$SC4
summary_results_SC5 <- summary_results$SC5
summary_results_SC6 <- summary_results$SC6
summary_results_SC7 <- summary_results$SC7
#summary_results_SC8 <- summary_results$SC8
#summary_results_SC9 <- summary_results$SC9
#summary_results_SC10 <- summary_results$SC10
#summary_results_SC11 <- summary_results$SC11
#summary_results_SC12 <- summary_results$SC12
#summary_results_SC13 <- summary_results$SC13
summary_results_SC14 <- summary_results$SC14
summary_results_SC15 <- summary_results$SC15
summary_results_SC16 <- summary_results$SC16
summary_results_SC17 <- summary_results$SC17
summary_results_SC18 <- summary_results$SC18
summary_results_SC19 <- summary_results$SC19
```
```{r}
# create list of final outbreak sizes
outbreak_sizes <- list(
  Baseline = summary_results_SC1$final_infected,
  Outdegree_high = summary_results_SC2$final_infected,
  Out_degree_med = summary_results_SC3$final_infected,
  Outdegree_all = summary_results_SC4$final_infected,
  Betweenness_high = summary_results_SC5$final_infected,
  Betweenness_med =summary_results_SC6$final_infected,
  Betweenness_all = summary_results_SC7$final_infected
)

```
Threshold sensitivity analysis
  whether >=10 is robust
  
```{r}
thresholds2 <- c(1, 2, 3, 5, 7, 10, 15, 20)

threshold_results2 <- data.frame(
  scenario = character(),
  threshold = numeric(),
  proportion = numeric(),
  stringsAsFactors = FALSE
  )


for (s in names(outbreak_sizes)) {
  
  outbreaks <- outbreak_sizes[[s]]
  
  for (t in thresholds2) {
    threshold_results2 <- rbind(
      threshold_results2,
      data.frame(
        Scenario2 = s,
        Threshold2 = t,
        proportion2 = mean(outbreaks >= t) #* 100
      )
    )
  }
}
```
Sensitivity plot
```{r}
# plot sensitivity analysis

ggplot(threshold_results2,
       aes(x = Threshold2,
           y = proportion2,
           colour = Scenario2)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = "Threshold (infected farms)",
    y = "Probability of a major outbreak",
    colour = "Scenario"
  ) +
  theme_bw()
```


# RESULT 1

Likelihood of sustained transmission - replicats led to >=10 infected final infected farms

Define the scenarios and extract final outbreak size, this is the input to the analysis

```{r}
# list of scenarios to include - dropped combined - see prev
sustained_trans <- list(
  baseline = summary_results_SC1$final_infected, 
  OD_high = summary_results_SC2$final_infected,
  OD_med = summary_results_SC3$final_infected, 
  OD_all = summary_results_SC4$final_infected, 
  BT_high = summary_results_SC5$final_infected, 
  BT_med = summary_results_SC6$final_infected, 
  BT_all = summary_results_SC7$final_infected, 
  #CT_high = summary_results_SC11 
  #CT_med = summary_results_SC12 
  #CT_ALL = summary_results_SC13 
  CT_OD_HIGH = summary_results_SC14$final_infected,
  CT_OD_MED = summary_results_SC15$final_infected,
  CT_OD_ALL = summary_results_SC16$final_infected,
  CT_BT_HIGH = summary_results_SC17$final_infected,
  CT_BT_MED = summary_results_SC18$final_infected,
  CT_BT_ALL = summary_results_SC19$final_infected
)
```
Calculate the probability of sustained transmission and relative reduction

```{r}
# counts how many simulations exceeded the threshold & calculates number sustained outbreaks; proportion sustained outbreaks; relative reduction vs baseline
threshold_final <- 10

threshold_counts_final <- data.frame(
  Scenario_final = names(sustained_trans),
  Exceeding = sapply(sustained_trans, function(x) sum(x >= threshold_final)),
  Total = sapply(sustained_trans, length)
)

threshold_counts_final$Proportion_final <- threshold_counts_final$Exceeding / threshold_counts_final$Total

# extract baseline probability
baseline_prob <- threshold_counts_final$Proportion_final[threshold_counts_final$Scenario_final == "baseline"]

# calculate relatve reduction compared with baseline
threshold_counts_final$Relative_reduction <- (
  (baseline_prob - threshold_counts_final$Proportion_final) / baseline_prob
)

threshold_counts_final
```

Organise scenarios by intervention type and intensity

```{r}
# AI - how can i group this plot by intensity windows, maintaining controls
# group by intensity and intervention type

# grepl() - searches for a pattern in scenario name e.g. "High"
# factor(..., levels = ...) - tells ggplot the order to display categories instead of using alphabetical order

# create two empty columns to store grouping information for plotting
threshold_counts_final$Intensity <- NA
threshold_counts_final$Method <- NA

  # assign intervention intensity

# search scenario names (grepl) & assign each one to an intervention intensity label
threshold_counts_final$Intensity[grepl("HIGH", threshold_counts_final$Scenario_final, ignore.case = TRUE)] <- "High" # looks for high in every scenario name & assigns method
threshold_counts_final$Intensity[grepl("MED", threshold_counts_final$Scenario_final, ignore.case = TRUE)] <- "Medium"
threshold_counts_final$Intensity[grepl("ALL", threshold_counts_final$Scenario_final, ignore.case = TRUE)] <- "All"

# assign the baseline scenario its own intensity label
threshold_counts_final$Intensity[threshold_counts_final$Scenario_final == "baseline"] <- "Baseline"

  # assign intervention method

# Label taregted interventions (grepl) according to the network metric used
threshold_counts_final$Method[grepl("^OD", threshold_counts_final$Scenario_final)] <- "Out-degree"
threshold_counts_final$Method[grepl("^BT", threshold_counts_final$Scenario_final)] <- "Betweenness"

# Keep the random matched controls separate from the targeted interventions
threshold_counts_final$Method[grepl("CT_OD", threshold_counts_final$Scenario_final)] <- "Control (OD matched)"
threshold_counts_final$Method[grepl("CT_BT", threshold_counts_final$Scenario_final)] <- "Control (BT matched)"

# label the baseline scenario
threshold_counts_final$Method[threshold_counts_final$Scenario_final == "baseline"] <- "Baseline"


  # Set plotting order
# convert intensity to an ordered factor so the panels appear in order wanted rather than alphabetically
threshold_counts_final$Intensity <- factor(
  threshold_counts_final$Intensity,
  levels = c("High", "Medium", "All", "Baseline")
)

# convert method to an ordered factor to bars appear in order wanted rather than alphabetically 

threshold_counts_final$Method <- factor(
  threshold_counts_final$Method,
  levels = c(
    "Control (OD matched)",
    "Out-degree",
    "Control (BT matched)",
    "Betweenness",
    "Baseline"
  )
)
```

Create grouped probability plot

```{r}
# Duplicate baseline into each intensity panel

baseline_row <- threshold_counts_final[
  threshold_counts_final$Scenario_final == "baseline",
]

baseline_repeated <- baseline_row[rep(1, 3), ]

baseline_repeated$Intensity <- c("High", "Medium", "All")
baseline_repeated$Method <- "Baseline"

# Remove original baseline and add repeated baseline
plot_data <- rbind(
  threshold_counts_final[
    threshold_counts_final$Scenario_final != "baseline",
  ],
  baseline_repeated
)

# Set order of panels
plot_data$Intensity <- factor(
  plot_data$Intensity,
  levels = c("High", "Medium", "All")
)

# Set order of bars
plot_data$Method <- factor(
  plot_data$Method,
  levels = c(
    "Baseline",
    "Control (OD matched)",
    "Out-degree",
    "Control (BT matched)",
    "Betweenness"
  )
)
```

Plot figure

```{r, fig.width = 12, fig.height = 4}
# plot probability of sustained outbreak by intensity

# order of windows
threshold_counts_final$Intensity <- factor(
  threshold_counts_final$Intensity,
  levels = c("Baseline", "High", "Medium", "All")
)
ggplot(
  plot_data,
  aes(x = Method,
      y = Proportion_final,
      fill = Method)
) +
  geom_col(position = "dodge") +
  facet_wrap(~Intensity) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = "Intervention strategy",
    y = "Probability of sustained outbreak (≥10 infected farms)"
  ) +
  theme_bw()
```

Relative Reduction

```{r}
# Duplicate baseline into each intensity panel for relative reduction plot

baseline_reduction <- threshold_counts_final[
  threshold_counts_final$Scenario_final == "baseline",
]

baseline_reduction <- baseline_reduction[rep(1, 3), ]

baseline_reduction$Intensity <- c("High", "Medium", "All")
baseline_reduction$Method <- "Baseline"
baseline_reduction$Relative_reduction <- 0


# Remove original baseline and add repeated baseline
plot_reduction <- rbind(
  threshold_counts_final[
    threshold_counts_final$Scenario_final != "baseline",
  ],
  baseline_reduction
)


# Set panel order
plot_reduction$Intensity <- factor(
  plot_reduction$Intensity,
  levels = c("High", "Medium", "All")
)


# Set bar order
plot_reduction$Method <- factor(
  plot_reduction$Method,
  levels = c(
    "Baseline",
    "Control (OD matched)",
    "Out-degree",
    "Control (BT matched)",
    "Betweenness"
  )
)
```
```{r, fig.width=12, fig.height=4}
# plot relative reduction by intensity

# order of windows
#threshold_counts_final$Intensity <- factor(
  #threshold_counts_final$Intensity,
  #levels = c("High", "Medium", "All")
#)

ggplot(
  plot_reduction,
  aes(x = Method,
      y = Relative_reduction,
      fill = Method)
) +
  geom_col(position = "dodge") +
  facet_wrap(~Intensity) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = "Intervention strategy",
    y = "Relative outbreak reduction vs baseline"
  ) +
  theme_bw()
```

Paired fishers tests

# RESULT 2

if an epidemic becomes established (threshold >= 10), how large does it become/how many farms are infected?

Report - Median, IQR, % reduction relative to baseline

Filter established outbreaks:
```{r}
# filter simulations above threshold - keep sims where finout outbreak size >= 10

threshold_final <- 10

established_outbreaks <- lapply(
  sustained_trans,
  function(x) x[x >= threshold_final]
)
```
```{r}
# create a dataframe for plotting

plot_sizes <- stack(established_outbreaks)
colnames(plot_sizes) <- c("Final_size", "Scenario")

# final_size= final number of infected farms
# scenario = intervention scenario
```
```{r}
# add intervention method & intensity
plot_sizes <- merge(
  plot_sizes,
  threshold_counts_final[, c("Scenario_final", "Method", "Intensity")],
  by.x = "Scenario",
  by.y = "Scenario_final"
)
plot_sizes_stats <- plot_sizes # save a copy before duplication for stat analysis without baseline x3
```
Baseline duplication - needed for faceted figure of panels

```{r}
# duplicate baseline into every intensity panel

# extract baseline
baseline_sizes <- subset(plot_sizes, Scenario == "baseline")
```
```{r}
# duplicate baseline
baseline_sizes <- do.call(
  rbind,
  lapply(c("High", "Medium", "All"), function(i) {
    x <- baseline_sizes
    x$Intensity <- i
    x
  })
)

baseline_sizes$Method <- "Baseline"
```
```{r}
# remove original baseline & add duplicated baseline
plot_sizes <- rbind(
  subset(plot_sizes, Scenario != "baseline"),
  baseline_sizes
)
```
Set plotting order

```{r}
# set plotting order
plot_sizes$Intensity <- factor(
  plot_sizes$Intensity,
  levels = c("High", "Medium", "All")
)

plot_sizes$Method <- factor(
  plot_sizes$Method,
  levels = c(
    "Baseline",
    "Control (OD matched)",
    "Out-degree",
    "Control (BT matched)",
    "Betweenness"
  )
)
```
Boxplot

```{r}
# create boxplots

ggplot(
  plot_sizes,
  aes(
    x = Method,
    y = Final_size,
    fill = Method
  )
) +
  geom_boxplot() +
  facet_wrap(~Intensity) +
  labs(
    x = "Intervention strategy",
    y = "Final outbreak size (established outbreaks only)"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    )
  )
```
Descriptive summary - median/IQR/mean etc

```{r}
# summarise outbreak sizes

outbreak_summary <- data.frame(
  Scenario = names(established_outbreaks),
  N = sapply(established_outbreaks, length), # how many sims exceeded threshold
  Mean = sapply(established_outbreaks, mean), # average final outbreak size
  Median = sapply(established_outbreaks, median), # median final outbreak size
  SD = sapply(established_outbreaks, sd), # variability/standard deviation
  Min = sapply(established_outbreaks, min), # range
  Max = sapply(established_outbreaks, max) # range
)

outbreak_summary
```

STATS - poissons distribution

https://r-coder.com/poisson-distribution-r/

Visualise distributions - show that final outbreak sizes are not normally distributed - ?right skewed

```{r}
# histogram of outbreak sizes
ggplot(plot_sizes_stats,
       aes(x = Final_size)) +
  geom_histogram(
    bins = 30,
    fill = "steelblue",
    colour = "black"
  ) +
  facet_wrap(~Method) +
  labs(
    x = "Final outbreak size (established outbreaks only)",
    y = "Number of simulations"
  ) +
  theme_bw()
```


# RESULT 3
c) peak infectious farms - how intense is outbreak IF THERE IS AN OUTBREAK - explain as above - mean/median peak without <1 sims

When an epidemic occurs (threshold >=10 farms), how intense is it?

Only analyse successful epidemics

Calculate peak infectious farms for every replicate

```{r}
# calculate peak infectious farms for every simulation = max number of infectious farms reached in each simulation
peak_infectious <- lapply(
  daily_results,
  function(x) {
    x %>%
      group_by(replicate) %>%
      summarise(
        peak_infectious = max(infectious)
      )
  }
)
```
Restrict to established epidemics

```{r}
# keep sims where threshold transmission occured
# create dataframe with replicate/peak_inefctious/scenario

peak_infectious_established <- lapply(
  names(peak_infectious),
  function(scenario) {
    
    peaks <- peak_infectious[[scenario]]
    
    established <- summary_results[[scenario]] %>%
      filter(final_infected >= threshold_final) %>%
      select(replicate)
    
    peaks %>%
      filter(replicate %in% established$replicate) %>%
      mutate(Scenario = scenario)
  }
)

peak_infectious_established <- bind_rows(peak_infectious_established)
```

Create scenario look-up

```{r}
# create look up table to to map simulation names to scenario labels

scenario_lookup <- data.frame(
  Scenario = c(
    "SC1", "SC2", "SC3", "SC4", "SC5", "SC6", "SC7", "SC14", "SC15", "SC16", "SC17", "SC18", "SC19"
  ),
  Scenario_final = c(
    "baseline",  "OD_high", "OD_med", "OD_all", "BT_high", "BT_med", "BT_all", "CT_OD_HIGH", "CT_OD_MED", "CT_OD_ALL", "CT_BT_HIGH", "CT_BT_MED","CT_BT_ALL"
  )
)
```

Add method/intensity labels

```{r}
# add method & intensity labels
peak_infectious_established <- merge(
  peak_infectious_established,
  scenario_lookup,
  by = "Scenario"
)

peak_infectious_established <- merge(
  peak_infectious_established,
  threshold_counts_final[, c("Scenario_final", "Method", "Intensity")],
  by = "Scenario_final"
)

peak_infectious_stats <- peak_infectious_established # create copy for stats
```
Create the plotting copy

```{r}
#make separate copy for figures
peak_infectious_plot <- peak_infectious_established

nrow(peak_infectious_plot)
# lower because has dropped scenarios 8-13
```
Duplicate baseline for the three intensity panels

```{r}
# duplicate baseline so appears in all 3 panels

# extract basline peak infecitous values
baseline_peak <- subset(
  peak_infectious_plot,
  Method == "Baseline"
)

# Duplicate baseline into each intensity panel
baseline_peak <- baseline_peak[
  rep(1:nrow(baseline_peak), 3),
]

# assign intensity labels
baseline_peak$Intensity <- rep(
  c("High", "Medium", "All"),
  each = nrow(baseline_peak) / 3
)

# keep baseline label
baseline_peak$Method <- "Baseline"
```
```{r}
# remove original baseline and add duplicated baseline back
peak_infectious_plot <- rbind(
  subset(peak_infectious_plot, Method != "Baseline"),
  baseline_peak
)
```
Set plotting order

```{r}
# set plotting order
peak_infectious_plot$Intensity <- factor(
  peak_infectious_plot$Intensity,
  levels = c("High", "Medium", "All")
)

peak_infectious_plot$Method <- factor(
  peak_infectious_plot$Method,
  levels = c("Baseline", "Control (OD matched)",
             "Out-degree",
             "Control (BT matched)",
             "Betweenness"
             )
)
```
Final boxplot

```{r}
# plot
ggplot(
  peak_infectious_plot,
  aes(
    x = Method,
    y = peak_infectious,
    fill = Method
  )
) +
  geom_boxplot() +
  facet_wrap(~Intensity) +
  labs(
    x = "Intervention strategy",
    y = "Peak infectious farms (established outbreaks only)"
  )+
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```
Non parametric distribution/densities - stat test choices
Show densitiies non-parametric

```{r}
# density plots
ggplot(peak_infectious_plot,
       aes(x = peak_infectious,
           colour = Method,
           fill = Method)) +
  geom_density(alpha = 0.2) +
  facet_wrap(~Intensity) +
  labs(
    x = "Peak infectious size (established outbreaks only)",
    y = "Density"
  ) +
  theme_bw()
```

Stats - poissons distribution


