---
title: "RQ2 SOIR SIMS ZIP"
author: "ESINCLAIR"
date: "2026-08-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

WHEN SAVING TO SUBMIT - Make a copy specifically for knitting, where the expensive simulation chunks are set to eval=FALSE only in the copy. The code remains visible in the knitted document, but isn't executed.


RQ2 – epidemiological consequences: do structurally important farms actually matter for disease spread - How do structurally identified  superspreader candidate farms influence the scale and dynamics of simulated  bTB-like disease spread compared with random introductions and targeted removal scenarios. - i.e. do structurally important farms acutally matter for disease spread?

CONTAINS REPLICATE ANALYSIS & SIMS

# Scenarios - static network analysis - identified farms consistently high-risk - persistence graph across all 12 months - and run SOIR from day 1 - ?are farms that are consistently structurally important across the whole study period better targets for intervention?

Use data from RQ1 to define candidate farms - persistent high OD/BT farms. All simulations start same day, same movement period, same B, same SOIR parameters, same number of stochastic replicates


```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(pbapply)
```

```{r}
# from RQ2 SOIR SIMS
# only runs one replicate 
run_soir_weighted_RQ2 <- function(
    movements,
    seed_farms,
    beta = 0.005, 
    occult_period = 120,
    infectious_period = 180
){
  
  
  farms <- sort(unique(c(
    movements$source,
    movements$destination
  )))
  
  n_farms <- length(farms)

  farm_index <- seq_len(n_farms)
  names(farm_index) <- farms
  
 
  movements$source <- as.character(movements$source)
  movements$destination <- as.character(movements$destination)



  movements$source_id <- farm_index[movements$source]
  movements$dest_id   <- farm_index[movements$destination]
  
  
  
  state <- integer(n_farms) 

  days <- integer(n_farms) 



  state[farm_index[seed_farms]] <- 1 
  days[farm_index[seed_farms]] <- 0 



  max_day <- max(movements$day) 

  
  history <- data.frame( 
    day = integer(max_day), 
    S = integer(max_day),
    O = integer(max_day),
    I = integer(max_day),
    R = integer(max_day)
  )


  for(day in seq_len(max_day)){



    
    occult <- which(state == 1)

    if(length(occult) > 0){

      days[occult] <- days[occult] + 1 

      become_I <- occult[
        days[occult] >= occult_period
      ]

      state[become_I] <- 2
      days[become_I] <- 0

    }



    infectious <- which(state == 2)

    
    if(length(infectious) > 0){

      days[infectious] <- days[infectious] + 1

      become_R <- infectious[
        days[infectious] >= infectious_period
      ]

      state[become_R] <- 3
      days[become_R] <- 0

    }


    todays <- movements[movements$day == day, ]


    if(nrow(todays) > 0){

      for(i in seq_len(nrow(todays))){

        source <- todays$source_id[i]
        dest   <- todays$dest_id[i]

        
        if(state[source] != 2)
          next

        
        if(state[dest] != 0)
          next
       
        batch_size <- todays$total_cattle[i] 
        p_transmit <- 1 - (1 - beta)^batch_size 
        
        if(runif(1) < p_transmit){ 

          state[dest] <- 1
          days[dest] <- 0

        }

      }

    }

    
    history$day[day] <- day
    history$S[day] <- sum(state == 0)
    history$O[day] <- sum(state == 1)
    history$I[day] <- sum(state == 2)
    history$R[day] <- sum(state == 3)

  }

  list(

    history = history, 

    final_states = data.frame( 

      farm = farms,

      state = factor(
        state,
        levels = 0:3,
        labels = c("S","O","I","R")
      )

    )

  )

}
```

# Provide movement data
```{r}
# read in csv
movements_raw <- read.csv(
    "RQ1 NO EP.csv",
    stringsAsFactors = FALSE
)

#head(movements_raw)
```

```{r}
# create simulation movement  table

movements <- movements_raw %>%
    select(
        source = org_id,
        destination = dest_id,
        day,
        total_cattle
    )
#head(movements)
```

# Visual of probability x batch size
https://ggplot2.tidyverse.org/articles/ggplot2.html
```{r}

beta <- 0.005

df <- data.frame(
  batch_size = 1:800
)

df$p_transmit <- 1 - (1 - beta)^df$batch_size


ggplot(df, aes(x = batch_size, y = p_transmit)) +
  geom_line(
    colour = "firebrick",
    linewidth = 1.2
  ) +
  labs(
    x = "Cattle moved per batch",
    y = "Probability of transmission",
    title = paste0(
      "Effect of batch size on transmission probability (β = ",
      beta,
      ")"
    )
  ) +
  theme_classic()
```
```{r}
# TEST WORKS WITH OUTPUT
# N.B. without 'scenarios... replicates - this is only running ONCE
result_weighted_RQ2 <- run_soir_weighted_RQ2(
    movements,
    seed_farms = c("95")
)

head(result_weighted_RQ2$history)
```
```{r}
max(result_weighted_RQ2$history$I)
```
```{r}
# N.B.this is just one run
result_weighted_RQ2$history %>%
    tidyr::pivot_longer(
        cols = c(S,O,I,R),
        names_to = "state",
        values_to = "farms"
    ) %>%
    ggplot(
        aes(
            x = day,
            y = farms,
            colour = state
        )
    ) +
    geom_line(size = 1) +
    theme_minimal()
```

Create workflow to include replicate - modified from sensitivity analysis - COPY AND PASTES BELOW


MODIFIED VERSION FOR BASELINE MODEL
```{r}
# runs same model but with 10 independent stochastic simulation replicates
# run_soir_weighted_RQ2() = one SOIR epidemic simulation
# run_baseline() = one replicate, including randomly selecting seed farm
# baseline_scenario_grid = creates 10 replicate rows
# pblapply() calls run_baseline() once for each row


baseline_scenario_grid <- data.frame(
  replicate = 1:10
)

run_baseline <- function(scenario_row){
  
  
  seed_farm <- sample(unique(c(movements$source, movements$destination)),
                      1
                      )
  # Run SOIR model
  result <- run_soir_weighted_RQ2(
    movements = movements,
    seed_farms = seed_farm,
    beta = 0.005,
    occult_period = 120,
    infectious_period = 180
    )
  
  # Create daily output 
  daily_output <- data.frame(
    replicate = scenario_row$replicate,
    day = result$history$day,
    susceptible = result$history$S,
    infectious = result$history$I,
    occult = result$history$O,
    removed = result$history$R,
    total_infected = result$history$O +
      result$history$I + 
      result$history$R
  )
  
  # Create summary outputs - records key statistics for the simulation
  summary_output <- data.frame(
    
    replicate = scenario_row$replicate,
    seed_farm = seed_farm,
    final_infected = sum(result$final_states$state!= "S"),
    peak_infectious = max(result$history$I)
  )
  
  # Return both outputs
  list(
    daily = daily_output, # time series epidemic data
    summary = summary_output # overall simulation stats
  )
}
```
```{r}
# executes the 10 simulations 
# runs the run_baseline() function once for each row of baseline_scenario_grid and stores all the results
baseline_results <- pblapply(
  seq_len(nrow(baseline_scenario_grid)), # generates sequence 1,2,...10
  function(i){ # defines a temporary function that takes the row number 'i'
    run_baseline(baseline_scenario_grid[i, , drop = FALSE]) # extracts the 'ith row of the baseline_scenario_grid, drop=FALSE ensures it remains a data frame rather than being simplified to a vector - passes that row into run_baseline() as scenario_row
  }
)

```

```{r}
# organise/combine the outputs from the simulations that have already been run, into one dataframe


# daily result - epidemic curves for all 10 stochastic runs
daily_results <- bind_rows( 
  lapply(baseline_results, function(x) x$daily) 
)
# peak infectious and final outbreak size per run
summary_results <- bind_rows(
  lapply(baseline_results, function(x) x$summary)
)
```
```{r}
# daily_results
# summary_results
```

# Baseline with 50/100/300/500/1000/2000 runs - analysis to establish replicates

```{r}
replicate_tests <- c(10, 50, 100, 300, 500, 1000, 2000)

results <- list()

for (n in replicate_tests) {
  
  baseline_scenario_grid <- data.frame(
    replicate = 1:n
  )

# pblapply() calls run-baseline() function created earlier, repeatedly (contains the SOIR model)
  
  baseline_results <- pblapply(
    seq_len(nrow(baseline_scenario_grid)),
    function(i){
      run_baseline(baseline_scenario_grid[i, , drop = FALSE])
    }
  )

# just extract summary results because want to see mean(final_infected)  
    
  summary_results <- bind_rows(
    lapply(baseline_results, function(x) x$summary)
  )

# store mean final outbreak size for each tested number of replicates; converts 'n' to character to label mean_final_infected
  #results[[as.character(n)]] <- mean(summary_results$final_infected)
  results[[as.character(n)]] <- data.frame(
  replicates = n,
  mean_final = mean(summary_results$final_infected),
  median_final = median(summary_results$final_infected),
  prob_spread = mean(summary_results$final_infected > 1),
  p95_final = as.numeric(quantile(summary_results$final_infected, 0.95))
)
}
```
```{r}
# combine results
stability_results <- bind_rows(results)

stability_results
```


```{r}
results_df <- bind_rows(results)

results_df
```
```{r}
ggplot(results_df, aes(x = replicates, y = mean_final)) +
  geom_line() +
  geom_point() +
  labs(
    x = "Number of stochastic replicates",
    y = "Mean final infected farms",
    title = "Stabilisation of outbreak size estimate"
  ) +
  theme_classic()
```
```{r}
# Probability of spread
ggplot(results_df, aes(x = replicates, y = prob_spread)) +
  geom_line() +
  geom_point() +
  labs(
    x = "Number of stochastic replicates",
    y = "Probability of spread"
  ) +
  theme_classic()
```
```{r}
# 95th percentile
ggplot(results_df, aes(x = replicates, y = p95_final)) +
  geom_line() +
  geom_point() +
  labs(
    x = "Number of stochastic replicates",
    y = "95th percentile final infected farms"
  ) +
  theme_classic()
```

# SCENARIOS

Read in data from RQ1
```{r}
# OUTDEGREE
OD_all_candidates <- readRDS(
  "outdegree_all_candidates.rds"
)

OD_med_candidates <- readRDS(
  "outdegree_med_persistent.rds"
)

OD_high_candidates <- readRDS(
  "outdegree_high_persistent.rds"
)
```

```{r}
#head(OD_all_candidates)
```

```{r}
# BETWEENNESS
BT_all_candidates <- readRDS(
  "BT_any_persistent_corrected.rds"
) # 35 nodes

BT_high_candidates <- readRDS(
  "BT_high_persistent_corrected.rds"
) # 1 node

BT_med_candidates <- readRDS(
  "BT_med_persistent_corrected.rds"
) # 8 nodes
```
```{r}
BT_all_candidates
```
```{r}
# removed lengths - these are stored as character/lists not table so length appropriate
length(OD_all_candidates)
length(OD_high_candidates)
length(OD_med_candidates)
length(BT_all_candidates)
length(BT_high_candidates)
length(BT_med_candidates)
```
# Modify function for scenarios

Scenarios:

   (1) Baseline
  
  (2) Outdegree >= 8 months
  (3) Outdegree >= 6 months
  (4) Outdegree any month
  
  (5) Betweenness >= 6 months
  (6) Betweenness >= 4 months
  (7) Betweenness any month
  
  (8) Combined BT >= 6 months & OD >= 8 months
  (9) Combined BT>= 4 months & 0D >= 6 months
  (10) Combined any month
  
  CONTROLS - If I remove the same number of farms, is targeted removal (SC2-10) more effective than random removal?
  (11) random X farms (average of scenarios 2/5/8) = 9.33 (9)
  (12) random X farms (average of scenarios 3/6/9) = 4.33 (4)
  (13) random X farms (average of scenarios 4/7/10) = 40.33 (40)
  
Persistence thresholds were selected based on temporal persistence distributions of each centrality measure. Outdegree hubs emerged throughout the study period of network analysis, whereas betweeness farms did not emerge until approximately month 6 (when cow-calf started trading). Consequently, different persistence thresholds were used to identify sustained structural importance of each metric.

N.B. hierarchal network (usualy for livestock) - until cow-calf starts to trade, two levels and high outdegree farms present/informative of structural SS characteristics; betweenness develops with cow-calf trading. I.e. outdegree heterogeneity emerges early; betweenness heterogeneity emerges only once cow-calf layer developes

# Create new movement datasets with excluded farms

Change this to only scenarios used, add pooled controls, match to table in dissertation

```{r}
# SCENARIO 1 - baseline - no nodes removed
scenario_1 <- movements
```
```{r}
#dim(scenario_1)
#length(unique(c(scenario_1$source, scenario_1$destination))) # number of nodes
```
```{r}
# SCENARIO 2 - Outdegree >= 8 months
scenario_2 <- movements %>%
  filter(
    !(source %in% OD_high_candidates),
    !(destination %in% OD_high_candidates)
  )
```
```{r}
# debug
#sum(is.na(scenario_2$source))
#sum(is.na(scenario_2$destination))
```
```{r}
# check worked & df
#any(OD_high_candidates %in% scenario_2$source)
#any(OD_high_candidates %in% scenario_2$destination)
#class(scenario_2)
```
```{r}
# SCENARIO 3 - Outdegree >= 6 months
scenario_3 <- movements %>%
  filter(
    !(source %in% OD_med_candidates),
    !(destination %in% OD_med_candidates)
  )
```
```{r}
# check worked & df
#any(OD_med_candidates %in% scenario_3$source)
#any(OD_med_candidates %in% scenario_3$destination)
#class(scenario_3)
```
```{r}
# SCENARIO 4 - Outdegree any months
scenario_4 <- movements %>%
  filter(
    !(source %in% OD_all_candidates),
    !(destination %in% OD_all_candidates)
  )
```
```{r}
# check worked & df
#any(OD_all_candidates %in% scenario_4$source)
#any(OD_all_candidates %in% scenario_4$destination)
#class(scenario_4)
```
```{r}
# SCENARIO 5 - Betweeness >= 6 months
scenario_5 <- movements %>%
  filter(
    !(source %in% BT_high_candidates),
    !(destination %in% BT_high_candidates)
  )
```
```{r}
# check worked & df
#any(BT_high_candidates %in% scenario_5$source)
#any(BT_high_candidates %in% scenario_5$destination)
#class(scenario_5)
```
```{r}
# SCENARIO 6 - Betweeness >= 4 months
scenario_6 <- movements %>%
  filter(
    !(source %in% BT_med_candidates),
    !(destination %in% BT_med_candidates)
  )
```
```{r}
# check worked & df
#any(BT_med_candidates %in% scenario_6$source)
#any(BT_med_candidates %in% scenario_6$destination)
#class(scenario_6)
```
```{r}
# SCENARIO 7 - Betweeness any months
scenario_7 <- movements %>%
  filter(
    !(source %in% BT_all_candidates),
    !(destination %in% BT_all_candidates)
  )
```
```{r}
# check worked & df
#any(BT_all_candidates %in% scenario_7$source)
#any(BT_all_candidates %in% scenario_7$destination)
#class(scenario_7)
```
NEED TO REMOVE COMBINED AND REPEAT CONTROLS AS POOLED 

CREATING POOLED CONTROLS

Create network configurations Control_ODHIGH1/10 .... Control_ODHIGH10/10; set.seed needs to be different for every control - ensure different random farms are removed & is reproducible.
Run each control network configuration through the SOIR sim and get 10x results
Import the results to analysis notebook and ?pool them?

```{r}
all_farms <- unique(c(movements$source, movements$destination))
```
```{r}
# Control OD High 1/10 
# SCENARIO 14 - random removal of 5 farms = control HIGH

set.seed(14) # Set random seed to ensure the same random farms can be selected in future runs.

random_5 <- sample(all_farms, 5)

scenario_14 <- movements %>%
  filter(
    !(source %in% random_5),
    !(destination %in% random_5)
  )
```

```{r}
# prompt: make this code into a loop, with a different set.seed, to produce 10x random controls for scenario 14 (a->j)

# Control OD High 1/10 to 10/10
# SCENARIO 8 - random removal of 5 farms = control OD HIGH

scenario_8P_controls <- list()

for (i in 1:10) {
  
  set.seed(13 + i) # different seed for each random control
  
  random_5 <- sample(all_farms, 5)
  
  scenario_8P_controls[[i]] <- movements %>%
    filter(
      !(source %in% random_5),
      !(destination %in% random_5)
    )
}
```
```{r}
length(scenario_8P_controls)
```
```{r}
# individual movement datasets are:
scenario_8P_controls[[1]] # etc - create loop for SOIR? and save each control as separate .rds file
```

```{r}
# Control OD MED 
# SCENARIO 15 - random removal of 8 farms = control MED

set.seed(15) # Set random seed to ensure the same random farms can be selected in future runs.

random_8 <- sample(all_farms, 8)

scenario_15 <- movements %>%
  filter(
    !(source %in% random_8),
    !(destination %in% random_8)
  )
```
```{r}
# Control OD Med 1/10 to 10/10
# SCENARIO 9 - random removal of 8 farms = control OD MED

scenario_9P_controls <- list()

for (i in 1:10) {
  
  set.seed(23 + i) # different seed for each random control
  
  random_8 <- sample(all_farms, 8)
  
  scenario_9P_controls[[i]] <- movements %>%
    filter(
      !(source %in% random_8),
      !(destination %in% random_8)
    )
}
```
```{r}
length(scenario_9P_controls)
```
```{r}
# Control OD ALL 
# SCENARIO 16 - random removal of 43 farms = control all

set.seed(16) # Set random seed to ensure the same random farms can be selected in future runs.

random_43 <- sample(all_farms, 43)

scenario_16 <- movements %>%
  filter(
    !(source %in% random_43),
    !(destination %in% random_43)
  )
```


```{r}
# Control OD All 1/10 to 10/10
# SCENARIO 10 - random removal of 43 farms = control OD ALL

scenario_10P_controls <- list()

for (i in 1:10) {
  
  set.seed(33 + i) # different seed for each random control
  
  random_43 <- sample(all_farms, 43)
  
  scenario_10P_controls[[i]] <- movements %>%
    filter(
      !(source %in% random_43),
      !(destination %in% random_43)
    )
}
```
```{r}
length(scenario_10P_controls)
```

```{r}
# Control BT HIGH 
# SCENARIO 17 - random removal of 1 farms = control HIGH

set.seed(111) # Set random seed to ensure the same random farms can be selected in future runs.

all_farms <- unique(c(movements$source, movements$destination))

random_1 <- sample(all_farms, 1)

scenario_17 <- movements %>%
  filter(
    !(source %in% random_1),
    !(destination %in% random_1)
  )
```


```{r}
# Control BT HIGH 1/10 to 10/10
# SCENARIO 11 - random removal of 1 farms = control BT HIGH

scenario_11P_controls <- list()

for (i in 1:10) {
  
  set.seed(43 + i) # different seed for each random control
  
  random_1 <- sample(all_farms, 1)
  
  scenario_11P_controls[[i]] <- movements %>%
    filter(
      !(source %in% random_1),
      !(destination %in% random_1)
    )
}
```
```{r}
length(scenario_11P_controls)
```

```{r}
# Control BT MED 
# SCENARIO 18 - random removal of 14 farms = control MED

set.seed(118) # Set random seed to ensure the same random farms can be selected in future runs.

random_8 <- sample(all_farms, 18)

scenario_18 <- movements %>%
  filter(
    !(source %in% random_8),
    !(destination %in% random_8)
  )
```
```{r}
# check worked & df
any(random_8 %in% scenario_18$source)
any(random_8 %in% scenario_18$destination)
class(scenario_18)
```

```{r}
# Control BT MED 1/10 to 10/10
# SCENARIO 12 - random removal of 18 farms = control BT MED

scenario_12P_controls <- list()

for (i in 1:10) {
  
  set.seed(53 + i) # different seed for each random control
  
  random_18 <- sample(all_farms, 18)
  
  scenario_12P_controls[[i]] <- movements %>%
    filter(
      !(source %in% random_18),
      !(destination %in% random_18)
    )
}
```
```{r}
length(scenario_12P_controls)
```

```{r}
# Control BT ALL 
# SCENARIO 19 - random removal of 47 farms = control all

set.seed(119) # Set random seed to ensure the same random farms can be selected in future runs.

random_35 <- sample(all_farms, 35)

scenario_19 <- movements %>%
  filter(
    !(source %in% random_35),
    !(destination %in% random_35)
  )
```

```{r}
# Control BT ALL 1/10 to 10/10
# SCENARIO 13 - random removal of 35 farms = control BT ALL

scenario_13P_controls <- list()

for (i in 1:10) {
  
  set.seed(63 + i) # different seed for each random control
  
  random_35 <- sample(all_farms, 35)
  
  scenario_13P_controls[[i]] <- movements %>%
    filter(
      !(source %in% random_35),
      !(destination %in% random_35)
    )
}
```
```{r}
length(scenario_13P_controls)
```

CREATING POOLED CONTROLS

Create network configurations Control_ODHIGH1/10 .... Control_ODHIGH10/10; set.seed needs to be different for every control - ensure different random farms are removed & is reproducible.
Run each control network configuration through the SOIR sim and get 10x results
Import the results to analysis notebook and ?pool them?


```{r}
# check worked & df
any(random_35 %in% scenario_19$source)
any(random_35 %in% scenario_19$destination)
class(scenario_19)
```
```{r}
# look-up table documenting which scenario each number represents
scenario_table <- data.frame(
  scenario = paste0("scenario_", 1:13),
  description = c(
    "Baseline",
    "Outdegree >= 8 months",
    "Outdegree >= 6 months",
    "Outdegree any month",
    "Betweenness >= 6 months",
    "Betweenness >= 4 months",
    "Betweenness any month",
    "Control OD HIGH x10",
    "Control OD MED x10",
    "Cotnrol OD ALL x10",
    "Control BT HIGH X10",
    "Control BT MED X10",
    "Control BT All X10"
  )
)

scenario_table
```
# Modify function so can be used for all scenarios, with 2000 replicates
```{r}
# modify function so can be used for all 2000 scenarios - just need to change movement_data and scenario_name when call it for each scenario - next box
scenario_grid <- data.frame(
  replicate = 1:2000
)
# pass scenario name and movement network as inputs. Function needs 3 inputs:
# 1. one row from scenario_grid (scenario_row) - dont need to chanage this
# 2. a movement dataset (movement_data)
# 3. a label for the scenario (scenario_name)
run_scenario <- function(scenario_row, movement_data, scenario_name){
  
  # Randomly select seed farm from the available network
  seed_farm <- sample(
    unique(c(movement_data$source, movement_data$destination)),
    1
  )
  
  # Run SOIR model
  result <- run_soir_weighted_RQ2(
    movements = movement_data,
    seed_farms = seed_farm,
    beta = 0.005,
    occult_period = 120,
    infectious_period = 180
  )
  
  # Daily output
  daily_output <- data.frame(
    scenario = scenario_name,
    replicate = scenario_row$replicate,
    day = result$history$day,
    susceptible = result$history$S,
    infectious = result$history$I,
    occult = result$history$O,
    removed = result$history$R,
    total_infected = result$history$O +
      result$history$I +
      result$history$R
  )
  
  # Summary output
  summary_output <- data.frame(
    scenario = scenario_name,
    replicate = scenario_row$replicate,
    seed_farm = seed_farm,
    final_infected = sum(result$final_states$state != "S"),
    peak_infectious = max(result$history$I)
  )
  
  list(
    daily = daily_output,
    summary = summary_output
  )
}
```
# Scenario 1 - Baseline

```{r}
# calls the function 2000 times - one for each row of scenario_grid
scenario_1_results <- pblapply(
  seq_len(nrow(scenario_grid)),
  function(i){
    run_scenario(
      scenario_grid[i, , drop = FALSE], # becomes scenario_row inside funciton
      scenario_1,# movement dataset for this scenario_X 
      "Baseline" # description of scenario - see look up table
    )
  }
  
)

saveRDS(scenario_1_results, "scenario_1_results.rds")
```

```{r}
# combine the daily outputs from all 2000 sim replicates into one data frame

# daily result - epidemic curves for all 2000 stochastic runs
daily_results_SC1 <- bind_rows( 
  lapply(scenario_1_results, function(x) x$daily) 
)
# peak infectious and final outbreak size per run
summary_results_SC1 <- bind_rows(
  lapply(scenario_1_results, function(x) x$summary)
)
```
```{r}
head(daily_results_SC1)
head(summary_results_SC1)
```

# Scenario 2 - Outdegree (high) >= 8 months

YOU ARE HERE - ERROR WHENRUNNING SIMULATION 29/07/26

```{r}
# calls the function 2000 times - one for each row of scenario_grid
scenario_2_results <- pblapply(
  seq_len(nrow(scenario_grid)),
  function(i){
    run_scenario(
      scenario_grid[i, , drop = FALSE], # becomes scenario_row inside funciton
      scenario_2,# movement dataset for this scenario_X 
      "Outdegree >= 8 months" # description of scenario - see look up table
    )
  }
)

saveRDS(scenario_2_results, "scenario_2_results.rds")
```


```{r}
# combine the daily outputs from all 2000 sim replicates into one data frame

# daily result - epidemic curves for all 2000 stochastic runs
daily_results_SC2 <- bind_rows( 
  lapply(scenario_2_results, function(x) x$daily) 
)
# peak infectious and final outbreak size per run
summary_results_SC2 <- bind_rows(
  lapply(scenario_2_results, function(x) x$summary)
)
```
```{r}
head(daily_results_SC2)
head(summary_results_SC2)
```



# Scenario 3 - Outdegree (med) >= 6 months

```{r}
# calls the function 2000 times - one for each row of scenario_grid
scenario_3_results <- pblapply(
  seq_len(nrow(scenario_grid)),
  function(i){
    run_scenario(
      scenario_grid[i, , drop = FALSE], # becomes scenario_row inside funciton
      scenario_3,# movement dataset for this scenario_X 
      "Outdegree >= 6 months" # description of scenario - see look up table
    )
  }
)

saveRDS(scenario_3_results, "scenario_3_results.rds")
```


```{r}
# combine the daily outputs from all 2000 sim replicates into one data frame

# daily result - epidemic curves for all 2000 stochastic runs
daily_results_SC3 <- bind_rows( 
  lapply(scenario_3_results, function(x) x$daily) 
)
# peak infectious and final outbreak size per run
summary_results_SC3 <- bind_rows(
  lapply(scenario_3_results, function(x) x$summary)
)
```
```{r}
head(daily_results_SC3)
head(summary_results_SC3)
```

# Scenario 4 - Outdegree aLL month

```{r}
# calls the function 2000 times - one for each row of scenario_grid
scenario_4_results <- pblapply(
  seq_len(nrow(scenario_grid)),
  function(i){
    run_scenario(
      scenario_grid[i, , drop = FALSE], # becomes scenario_row inside funciton
      scenario_4,# movement dataset for this scenario_X 
      "Outdegree any month" # description of scenario - see look up table
    )
  }
)

saveRDS(scenario_4_results, "scenario_4_results.rds")
```


```{r}
# combine the daily outputs from all 2000 sim replicates into one data frame

# daily result - epidemic curves for all 2000 stochastic runs
daily_results_SC4 <- bind_rows( 
  lapply(scenario_4_results, function(x) x$daily) 
)
# peak infectious and final outbreak size per run
summary_results_SC4 <- bind_rows(
  lapply(scenario_4_results, function(x) x$summary)
)
```
```{r}
head(daily_results_SC4)
head(summary_results_SC4)
```

# Scenario 5 - Betweenness (high) >= 6 months
```{r}
# calls the function 2000 times - one for each row of scenario_grid
scenario_5_results <- pblapply(
  seq_len(nrow(scenario_grid)),
  function(i){
    run_scenario(
      scenario_grid[i, , drop = FALSE], # becomes scenario_row inside funciton
      scenario_5,# movement dataset for this scenario_X 
      "Betweenness >= 6 months" # description of scenario - see look up table
    )
  }
)
```
```{r}
saveRDS(scenario_5_results, "scenario_5_results_corrected.rds")
```


```{r}
# combine the daily outputs from all 2000 sim replicates into one data frame

# daily result - epidemic curves for all 2000 stochastic runs
daily_results_SC5 <- bind_rows( 
  lapply(scenario_5_results, function(x) x$daily) 
)
# peak infectious and final outbreak size per run
summary_results_SC5 <- bind_rows(
  lapply(scenario_5_results, function(x) x$summary)
)
```
```{r}
head(daily_results_SC5)
head(summary_results_SC5)
```

# Scenario 6 - Betweenness (med) >= 4 months

```{r}
# calls the function 2000 times - one for each row of scenario_grid
scenario_6_results <- pblapply(
  seq_len(nrow(scenario_grid)),
  function(i){
    run_scenario(
      scenario_grid[i, , drop = FALSE], # becomes scenario_row inside funciton
      scenario_6,# movement dataset for this scenario_X 
      "Betweenness >= 4 months" # description of scenario - see look up table
    )
  }
)

saveRDS(scenario_6_results, "scenario_6_results_corrected.rds")
```


```{r}
# combine the daily outputs from all 2000 sim replicates into one data frame

# daily result - epidemic curves for all 2000 stochastic runs
daily_results_SC6 <- bind_rows( 
  lapply(scenario_6_results, function(x) x$daily) 
)
# peak infectious and final outbreak size per run
summary_results_SC6 <- bind_rows(
  lapply(scenario_6_results, function(x) x$summary)
)
```
```{r}
head(daily_results_SC6)
head(summary_results_SC6)
```

# Scenario 7 - Betweenness aLL month

```{r}
# calls the function 2000 times - one for each row of scenario_grid
scenario_7_results <- pblapply(
  seq_len(nrow(scenario_grid)),
  function(i){
    run_scenario(
      scenario_grid[i, , drop = FALSE], # becomes scenario_row inside funciton
      scenario_7,# movement dataset for this scenario_X 
      "Betweenness any month" # description of scenario - see look up table
    )
  }
)
```
```{r}
saveRDS(scenario_7_results, "scenario_7_results_corrected.rds")
```


```{r}
# combine the daily outputs from all 2000 sim replicates into one data frame

# daily result - epidemic curves for all 2000 stochastic runs
daily_results_SC7 <- bind_rows( 
  lapply(scenario_7_results, function(x) x$daily) 
)
# peak infectious and final outbreak size per run
summary_results_SC7 <- bind_rows(
  lapply(scenario_7_results, function(x) x$summary)
)
```
```{r}
head(daily_results_SC7)
head(summary_results_SC7)
```

# Scenario 8	- Control OD HIGH x10
```{r}
# prompt: using this structure, show me code to loop through the 10 control scenarios and save each result as a separate .rds file

for (j in 1:10) {
  
  control_results <- pblapply(
    seq_len(nrow(scenario_grid)),
    function(i){
      run_scenario(
        scenario_grid[i, , drop = FALSE],
        scenario_8P_controls[[j]],
        paste0("Scenario 8P Control ", j)
      )
    }
  )
  
  saveRDS(
    control_results,
    paste0("scenario_8P_control_", j, "_results.rds")
  )
}
```



# Scenario 9 - Control OD MED x10
```{r}
for (j in 1:10) {
  
  control_results9 <- pblapply(
    seq_len(nrow(scenario_grid)),
    function(i){
      run_scenario(
        scenario_grid[i, , drop = FALSE],
        scenario_9P_controls[[j]],
        paste0("Scenario 9P Control ", j)
      )
    }
  )
  
  saveRDS(
    control_results9,
    paste0("scenario_9P_control_", j, "_results.rds")
  )
}
```


# Scenario 10 - Control OD ALL X10

```{r}
for (j in 1:10) {
  
  control_results10 <- pblapply(
    seq_len(nrow(scenario_grid)),
    function(i){
      run_scenario(
        scenario_grid[i, , drop = FALSE],
        scenario_10P_controls[[j]],
        paste0("Scenario 10P Control ", j)
      )
    }
  )
  
  saveRDS(
    control_results10,
    paste0("scenario_10P_control_", j, "_results.rds")
  )
}
```


# Scenario 11 - Control BT HIGH x10
```{r}
for (j in 1:10) {
  
  control_results11 <- pblapply(
    seq_len(nrow(scenario_grid)),
    function(i){
      run_scenario(
        scenario_grid[i, , drop = FALSE],
        scenario_11P_controls[[j]],
        paste0("Scenario 11P Control ", j)
      )
    }
  )
  
  saveRDS(
    control_results11,
    paste0("scenario_11P_control_", j, "_results.rds")
  )
}
```


# Scenario 12 - Control BT MED X10

```{r}
for (j in 1:10) {
  
  control_results12 <- pblapply(
    seq_len(nrow(scenario_grid)),
    function(i){
      run_scenario(
        scenario_grid[i, , drop = FALSE],
        scenario_12P_controls[[j]],
        paste0("Scenario 12P Control ", j)
      )
    }
  )
  
  saveRDS(
    control_results12,
    paste0("scenario_12P_control_", j, "_results.rds")
  )
}
```

# Scenario 13 - Control BT ALL X10

```{r}
for (j in 1:10) {
  
  control_results13 <- pblapply(
    seq_len(nrow(scenario_grid)),
    function(i){
      run_scenario(
        scenario_grid[i, , drop = FALSE],
        scenario_13P_controls[[j]],
        paste0("Scenario 13P Control ", j)
      )
    }
  )
  
  saveRDS(
    control_results13,
    paste0("scenario_13P_control_", j, "_results.rds")
  )
}
```

