---
title: "RQ2 SOIR SIMS"
author: "ESINCLAIR"
date: "2026-07-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
RQ2 – epidemiological consequences: do structurally important farms actually matter for disease spread - How do structurally identified  superspreader candidate farms influence the scale and dynamics of simulated  bTB-like disease spread compared with random introductions and targeted removal scenarios. - i.e. do structurally important farms acutally matter for disease spread?

# Scenarios - static network analysis - identified farms consistently high-risk - persistence graph accross all 12 months - and run SOIR from day 1 - ?are farms that are consistently structurally important across the whole study period better targets for intervention?

Use data drom RQ1 to define candidate farms - persistent high OD/BT farms. All simulations start same day, same movement period, same B, same SOIR parameters, same number of stochastic replicates

Baseline network - no removal

Remove top 5% out-degree farms

Remove top 5% Betweeness farms

Remove top 5% that appear in both BT & OD

Remove random X farms (control)

For each scenario:

  run 100-1000? stochastic replicates
  record outcomes - final infected farms, peak infectious farms, time to peak, outbreak duration, epidemic plot



```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(pbapply)
```

```{r}
# Weighted SOIR model: beta is a per-animal transmission probability, so larger cattle batches have a higher probability of transmitting infection.

run_soir_weighted_RQ2 <- function(
    movements,
    seed_farms, # can seed 1+ farms in one run
    beta = 0.005, # beta = 0.005, # per-animal transmission probability (varied in sensitivity analysis)
    occult_period = 120, # fixed from un-weighted sensitivity analysis
    infectious_period = 180 # fixed from un-weighted sensitivity analysis
){
  
# Build look-up table - pulls every unique farm ID from both the source and destination columns of movements  - combines, removes duplicates, sorts --> (farms) is a clean look-up table of all farms involved in the dataset
  
  farms <- sort(unique(c(
    movements$source,
    movements$destination
  )))
  
# Creates a numeric index for the farms, labels each with a number corresponding to the farm ID
  n_farms <- length(farms)

  farm_index <- seq_len(n_farms)
  names(farm_index) <- farms

# Convert movements to integer indices #

  movements$source_id <- farm_index[movements$source]
  movements$dest_id   <- farm_index[movements$destination]

# Set up vectors to track each farms disease status and how long it has been in its current state
  
  state <- integer(n_farms) # creates a vector of zeros - meaning all farms start susceptible

  days <- integer(n_farms) # creates a parallel vector to count how many days each farm has spent in its current state

# Seed infection - infect chosen seed farm by setting its state to 1 - farms to start as occult - representative of dz introduction
# Reset their disease timers to 0 days, indicating they have just entered the occult state.

  state[farm_index[seed_farms]] <- 1 # state[farm_index[seed_farms]] <- 1; change seeded farms from susceptible (0) to occult (1)

  days[farm_index[seed_farms]] <- 0 # for the same seeded farm(s), set no. days spent in current dz state to 0 - farm has just become occult at the start of the simulation

# Storage - prepares storage for the simulation results

  max_day <- max(movements$day) # finds last day in the movement data set

# History - creates a data frame with one row per day, and columns for the number of farms in each SOIR state, all columns are pre-allocated as integer vectors of length max_day for filling during the simulation
  
  history <- data.frame( 
    day = integer(max_day), 
    S = integer(max_day),
    O = integer(max_day),
    I = integer(max_day),
    R = integer(max_day)
  )


# Simulation - advances farms through the O -> I -> R stages as each day of the simulation passes
# The 'for' loop runs one iteration per day

  for(day in seq_len(max_day)){


# Progress occult farms:
# The day counters increase, and farms that have spent more than/equal to occult_period days in O become infectious (state == 2) and their day counters reset
    
    occult <- which(state == 1)

    if(length(occult) > 0){

      days[occult] <- days[occult] + 1 # increments until reaches occult_period -> becomes infectious

      become_I <- occult[
        days[occult] >= occult_period
      ]

      state[become_I] <- 2
      days[become_I] <- 0

    }

# Progress infectious farms

    infectious <- which(state == 2)

# Their day counters increase, any farms that have spent more than infectious_period i days in I become removed (state == 3) and their day counters reset    
    
    if(length(infectious) > 0){

      days[infectious] <- days[infectious] + 1

      become_R <- infectious[
        days[infectious] >= infectious_period
      ]

      state[become_R] <- 3
      days[become_R] <- 0

    }

# Today's movements - filters the 'movements' data frame to keep just the rows where 'day' matches the current loop day, giving the set of movements to process for that day.

    todays <- movements[movements$day == day, ]

    
# Process each movement event. Transmission is only considered from infectious source farms to susceptible destination farms.


    
# Processes each movement event and applies infection-spread rules.
# loop over all movements happening today, for each movement, get the source and destination farm indices; skip the movement if the source is not infectious (state != 2); skip the movement if the destination is not susceptible (state != 0). Only movements from the infectious -> susceptible farms proceed to the transmission step that follows.

    if(nrow(todays) > 0){

      for(i in seq_len(nrow(todays))){

        source <- todays$source_id[i]
        dest   <- todays$dest_id[i]

        # source not infectious
        if(state[source] != 2)
          next

        # destination already infected
        if(state[dest] != 0)
          next

        
# Transmission block:
# Calculate the movement-level transmission probability from the batch size.
# Each animal has an independent probability (beta) of initiating infection

# Stochastic transmission event:
# runif(1) generates a random number between 0 and 1.
# If the random value is less than p_transmit, the movement event results in transmission,
# and the destination farm changes from susceptible (S) to occult (O).
# This represents a Bernoulli trial where transmission either occurs (1) or does not occur (0).
# p_transmit is the batch-level probability of infection, derived from the per-animal
# transmission probability (beta) and number of cattle moved.
#https://www.rdocumentation.org/packages/stats/versions/3.5.2/topics/Uniform, Bernoulli

        # beta = individual animal-level probability
        # p_transmit = whole batch movement level probability
       
        batch_size <- todays$total_cattle[i] # no. cattle moved in this movement event
        p_transmit <- 1 - (1 - beta)^batch_size # Probability that at least one animal in the batch initiates infection.
        
        if(runif(1) < p_transmit){ # raw a random number between 0 and 1; if it is less than p_transmit, infection occurs.

          state[dest] <- 1
          days[dest] <- 0

        }

      }

    }

# ? eventually include the probability that a moved animal is actually infectious - not neccesserially important for research question

# Save epidemic - logs the SOIR count for the current simulation day - store the day number; count how many farms are S,O,I,R and write those values into the corresponding row of history
    
    history$day[day] <- day
    history$S[day] <- sum(state == 0)
    history$O[day] <- sum(state == 1)
    history$I[day] <- sum(state == 2)
    history$R[day] <- sum(state == 3)

  }

# Return - returns the simulation output as a structured list

  list(

    history = history, # stores the full day-by-day SOIR counts

    final_states = data.frame( # data frame listing each farm and its final SOIR state, converts numeric codes (0-3) into labelled factors S,O,I,R

      farm = farms,

      state = factor(
        state,
        levels = 0:3,
        labels = c("S","O","I","R")
      )

    )

  )

}
```

# Provide movement data
```{r}
# read in csv

movements_raw <- read.csv(
    "RQ1 NO EP.csv",
    stringsAsFactors = FALSE
)

#head(movements_raw)
```

```{r}
# create simulation movement  table

movements <- movements_raw %>%
    select(
        source = org_id,
        destination = dest_id,
        day,
        total_cattle
    )
#head(movements)
```

```{r}
str(movements)
```

# Visual of probability x batch size
https://ggplot2.tidyverse.org/articles/ggplot2.html
```{r}

beta <- 0.005

df <- data.frame(
  batch_size = 1:800
)

df$p_transmit <- 1 - (1 - beta)^df$batch_size


ggplot(df, aes(x = batch_size, y = p_transmit)) +
  geom_line(
    colour = "firebrick",
    linewidth = 1.2
  ) +
  labs(
    x = "Cattle moved per batch",
    y = "Probability of transmission",
    title = paste0(
      "Effect of batch size on transmission probability (β = ",
      beta,
      ")"
    )
  ) +
  theme_classic()
```
```{r}
# N.B. without 'scenarios... replicates - this is only running ONCE
result_weighted_RQ2 <- run_soir_weighted_RQ2(
    movements,
    seed_farms = c("95")
)

head(result_weighted_RQ2$history)
```
```{r}
max(result_weighted_RQ2$history$I)
```
```{r}
result_weighted_RQ2$history %>%
    tidyr::pivot_longer(
        cols = c(S,O,I,R),
        names_to = "state",
        values_to = "farms"
    ) %>%
    ggplot(
        aes(
            x = day,
            y = farms,
            colour = state
        )
    ) +
    geom_line(size = 1) +
    theme_minimal()
```
Epidemic curve- infected per day
```{r}

ggplot(result_weighted_RQ2$history,
       aes(x = day, y = I)) +
  geom_line(
    colour = "firebrick",
    linewidth = 1.2
  ) +
  labs(
    x = "Day",
    y = "Number of infectious farms",
    title = "Epidemic curve"
  ) +
  theme_classic()
```
Create workflow to include replicate - modified from sensitivity analysis - COPY AND PASTES BELOW


MODIFIED VERSION FOR BASELINE MODEL
```{r}
# baseline model parameters - as per  model/sensitivity analysis
#baseline_b <- 0.005
#basline_occult <- 120
#baseline_infect <- 180

# Create baseline scenarios (10 stochastic replicates - test run)

baseline_scenario_grid <- data.frame(
  replicate = 1:10
)

# Function to run one baseline simulation
run_baseline <- function(scenario_row){
  
  #print(class(x))
  #print(x)
  
  #Randomly select one seed farm
  seed_farm <- sample(unique(c(movements$source, movements$destination)),
                      1
                      )
  # Run SOIR model
  result <- run_soir_weighted_RQ2(
    movements = movements,
    seed_farms = seed_farm,
    beta = 0.005,
    occult_period = 120,
    infectious_period = 180
    )
  
  # Daily epidemic output
  daily_output <- data.frame(
    replicate = scenario_row$replicate,
    day = result$history$day,
    susceptible = result$history$S,
    infectious = result$history$I,
    occult = result$history$O,
    removed = result$history$R,
    total_infected = result$history$O +
      result$history$I + 
      result$history$R
  )
  
  # Summary output
  summary_output <- data.frame(
    #scenario = "baseline/scenario name", # change input depending on scenario - keep for stat tests later all in one dataframe
    replicate = scenario_row$replicate,
    seed_farm = seed_farm,
    final_infected = sum(result$final_states$state!= "S"),
    peak_infectious = max(result$history$I)
  )
  
  # Return both outputs
  list(
    daily = daily_output,
    summary = summary_output
  )
  # Return summary statistics
  #data.frame(
    #replicate = x$replicate,
    #seed_farm = seed_farm,
  #  final_infected = sum(result_baseline_scenarios$final_states$state != "S"),
   # peak_infectious = max(result_baseline_scenarios$history$I)
  #)
}
```

```{r}
baseline_results <- pblapply(
  seq_len(nrow(baseline_scenario_grid)),
  function(i){
    run_baseline(baseline_scenario_grid[i, , drop = FALSE])
  }
)
```


```{r}
# daily result - epidemic curves for all 10 stochastic runs
daily_results <- bind_rows(
  lapply(baseline_results, function(x) x$daily)
)
# peak infectious and final outbreak size per run
summary_results <- bind_rows(
  lapply(baseline_results, function(x) x$summary)
)
```
```{r}
daily_results
```
```{r}
summary_results
```

BotMEW paper

1. individual stochastic runs plot (example) - shows uncertainty and stochastic variation - demonstrates model variability. Each simulation starts with a randomly selected seed farm and includes random transmission events (runif(1) < p_transmit) - so even with identical parameters, outbreaks will differ.

https://pdodds.w3.uvm.edu/files/papers/others/2001/mackenzie2001.pdf

```{r}
# 1. plot individual stochastic runs (to show variability)
ggplot(
  daily_results,
  aes(x = day, y = infectious, group = replicate)
) +
  geom_line(alpha = 0.3) +
  labs(
    x = "Day",
    y = "Infectious farms",
    title = "Stochastic epidemic trajectories"
  ) +
  theme_classic()
```

2. Mean +/- SD (uncertainty) epidemic curve - summarises the expected epidemic trajectory across all stochastic replicates (shows expected outbreak dynamics) - calculate 'mean infectious farms on dat 't'' across all simulations. Shows average outbreak progression, when epidemic peaks, how much variability exists around the mean - represents thebehvaiour of the model rather than one random simulation - the mean epidemic curve was calculated across stochastic replicates, with variability represented by the standard deviation of infectious farm numbers at each time step.

https://www.sciencedirect.com/science/article/pii/S0167587797000020

```{r}
# 2. average number of infectious farms per day
mean_epidemic <- daily_results %>%
  group_by(day) %>%
  summarise(
    mean_infectious = mean(infectious),
    sd_infectious = sd(infectious)
  )

ggplot(mean_epidemic, aes(x = day, y = mean_infectious)) + geom_line(linewidth = 1.2, colour = "firebrick") + labs(
  x = "Day",
  y = "Mean epidemic curve across stochastic simulations"
) +
  theme_classic()
```

```{r}
# 2. add uncertainty
ggplot(mean_epidemic, aes(x = day, y = mean_infectious)) +
  geom_ribbon(
    aes(
      ymin = mean_infectious - sd_infectious,
      ymax = mean_infectious + sd_infectious
    ),
    alpha = 0.3,
    fill = "firebrick"
  ) +
  geom_line(
    linewidth = 1.2,
    colour = "firebrick"
  ) +
  labs(
    x = "Day",
    y = "Infectious farms",
    title = "Epidemic curve (mean ± SD)"
  ) +
  theme_classic()
```

3. Final outbreak size distribution - compares scenario impact

Uses output 'summary_results$final_infected - contains the number of farms in each replicate
Shows the probability distribution of outbreak sizes, whether outbreaks are usually small or occasionally large, whether changing model parameters changes epidemic risk

https://www.sciencedirect.com/science/article/pii/S0167587705000358

```{r}
ggplot(
  summary_results,
  aes(x = final_infected)
) +
  geom_density(
    fill = "steelblue",
    alpha = 0.4
  ) + 
  labs(x = "Final infected farms",
       y = "Density",
       title = "Distribution of final outbreak size"
       ) + 
  theme_classic()
```

4. 
To get line graph showing each compartment- need a mean of stochastic replications - to show mean trajectory across replicates

Need to create baseline_history containing day_by_day history from all 10 runs (with a replciate column) - then can create plot - currently only saved the summary statistics- need the history from each replicate to allow day-by-day epidemic plot

https://www.sciencedirect.com/science/article/pii/S1755436510000460

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

mean_SOIR <- daily_results %>%
  group_by(day) %>%
  summarise(
    mean_S = mean(susceptible),
    mean_O = mean(occult),
    mean_I = mean(infectious),
    mean_R = mean(removed)
  ) %>%
  pivot_longer(
    cols = c(mean_S, mean_O, mean_I, mean_R),
    names_to = "compartment",
    values_to = "mean_farms"
  )
```
```{r}
ggplot(
  mean_SOIR,
  aes(
    x = day,
    y = mean_farms,
    colour = compartment
  )
)+
  geom_line(
    linewidth = 1.2
  )+
  labs(
    x = "Day",
    y = "Mean number of farms",
    colour = "Disease state",
    title = "Mean SOIR epidemic dynamics across stochastic simulations"
  )+
  theme_classic()
```
5. Stat significance of different scenarios

For comparing scenarios (e.g. different seed farms, removing nodes, changing network structure), need 

(i) multiple stochastic replicates per scenario
(ii) summary outcome per replicate:
      final_infected (total outbreak size)
      peak_infectious
      ?epidemic duration
      ?cumulative infected over time
      
E.g. compare two scenarios by final outbreak size:
scenario A - baseline network
scenario B - network with x type of nodes removed

Would have - baseline_summary & removed_nodes_summary

Then - t-test? but if outbreak size is not normally distributed - likely (stochastic epidemics usually have - many small outbreaks/fewer larger outbreaks) - non-parametric test may be more appropriate - e.g. wilcox test

  wilcox.test(baseline_summary%final_infected, removed_nodes_summary$final_infected)

If compare several interventions - use ?anova? Kruskal-Wallis test?

COMPARE

  primary metric - final infected = overall epidemic impact
  secondary metric - peak infectious = maximum disease pressure & time to peak = epidemic speed





```{r}
# for month 7 - from sensitivity analysys
out_degree_top5 <- c("253", "129", "95", "255", "88", "119", "214", "213", "212", "202", "151", "299")
betweenness_top5 <- c("95", "296", "119", "41", "285", "300", "274", "178", "138", "235", "156")
random_farms <- c("2", "35", "180", "50", "260", "4", "230", "301", "92", "205", "167", "227")
```