---
title: "RQ2 RESULTS ANALYSIS ZIP"
author: "ESINCLAIR"
date: "2026-08-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# libraries
library(igraph)
library(ggplot2)
library(dplyr)
library(tidyr)
library(pbapply)
```

CONTAINS GWCC AND THRESHOLD ESTABLISHMENT + ANALYSIS

Import results of simulations for anlysis (RQ2 SOIR SIMS)


```{r}
scenario_1_results <- readRDS("scenario_1_results.rds") # baseline
scenario_2_results <- readRDS("scenario_2_results.rds") # Outdegree (high) >= 8 months
scenario_3_results <- readRDS("scenario_3_results.rds") # Outdegree (med) >= 6 months
scenario_4_results <- readRDS("scenario_4_results.rds") # Outdegree any month
scenario_5_results <- readRDS("scenario_5_results_corrected.rds") # Betweenness (high) >= 6 months
scenario_6_results <- readRDS("scenario_6_results_corrected.rds") # Betweenness (med) >= 4 months
scenario_7_results <- readRDS("scenario_7_results_corrected.rds") # Betweenness any month

scenario_14_results <- readRDS("scenario_14_results.rds") # OD HIGH CONTROL (5 removed)
scenario_15_results <- readRDS("scenario_15_results.rds") # OD MED CONTROL (8 removed)
scenario_16_results <- readRDS("scenario_16_results.rds") # OD ALL CONTROL (43 removed)
scenario_17_results <- readRDS("scenario_17_results_corrected.rds") # BT HIGH CONTROL (1 removed)
scenario_18_results <- readRDS("scenario_18_results_corrected.rds") # BT MED CONTROL (8 removed)
scenario_19_results <- readRDS("scenario_19_results_corrected.rds") # BT ALL CONTROL (35 removed)
```
```{r}
# combine the daily outputs from all 2000 sim replicates into one data frame

# daily result - epidemic curves for all 2000 stochastic runs
daily_results_SC1 <- bind_rows( 
  lapply(scenario_1_results, function(x) x$daily) 
)
# peak infectious and final outbreak size per run
summary_results_SC1 <- bind_rows(
  lapply(scenario_1_results, function(x) x$summary)
)
```
```{r}
# Put all scenario results into one list
all_scenarios <- list(
  SC1 = scenario_1_results,
  SC2 = scenario_2_results,
  SC3 = scenario_3_results,
  SC4 = scenario_4_results,
  SC5 = scenario_5_results,
  SC6 = scenario_6_results,
  SC7 = scenario_7_results,
  SC14 = scenario_14_results,
  SC15 = scenario_15_results,
  SC16 = scenario_16_results,
  SC17 = scenario_17_results,
  SC18 = scenario_18_results,
  SC19 = scenario_19_results
)

# Create empty lists to store outputs
daily_results <- list()
summary_results <- list()

# Loop through scenarios
for (scenario in names(all_scenarios)) {
  
  daily_results[[scenario]] <- bind_rows(
    lapply(all_scenarios[[scenario]], function(x) x$daily)
  )
  
  summary_results[[scenario]] <- bind_rows(
    lapply(all_scenarios[[scenario]], function(x) x$summary)
  )
}

# call individual with - summary_results$SC3 etc
```

# Read in data from RQ1 for number farms removed

```{r}
# OUTDEGREE
OD_all_candidates <- readRDS(
  "outdegree_all_candidates.rds"
)

OD_med_candidates <- readRDS(
  "outdegree_med_persistent.rds"
)

OD_high_candidates <- readRDS(
  "outdegree_high_persistent.rds"
)
```

```{r}
#OD_all_candidates
```
```{r}
# BETWEENNESS
BT_all_candidates <- readRDS(
  "BT_any_persistent_corrected.rds"
)

BT_high_candidates <- readRDS(
  "BT_high_persistent_corrected.rds"
)

BT_med_candidates <- readRDS(
  "BT_med_persistent_corrected.rds"
)
```
```{r}
#BT_med_candidates
```

# 1. Create scenario table - what was removed

```{r}
total_farms = 301
scenario_table <- data.frame(
  Scenario = c(
    "(1)",
    "(2)",
    "(3)",
    "(4)",
    "(5)",
    "(6)",
    "(7)",
    #"(8)",
    #"(9)",
    #"(10)",
    #"(11)",
    #"(12)",
    #"(13)",
    "(14)",
    "(15)",
    "(16)",
    "(17)",
    "(18)",
    "(19)"
  ),
  
  Type = c(
    "Baseline",
    "Targeted Intervention",
    "Targeted Intervention",
    "Targeted Intervention",
    "Targeted Intervention",
    "Targeted Intervention",
    "Targeted Intervention",
    #"Targeted Intervention",
    #"Targeted Intervention",
    #"Targeted Intervention",
    #"Random removal control",
    #"Random removal control",
    #"Random removal control",
    "CONTROL OD HIGH",
    "CONTROL OD MED",
    "CONTROL OD ALL",
    "CONTROL BT HIGH",
    "CONTROL BT MED",
    "CONTROL BT ALL"
  ),
  
  Centrality = c(
    "None",
    "Out-degree",
    "Out-degree",
    "Out-degree",
    "Betweenness",
    "Betweenness",
    "Betweenness",
    #"Combined",
    #"Combined",
   # "Combined",
   # "Random",
   # "Random",
   # "Random",
    "RandomOD",
    "RandomOD",
    "RandomOD",
    "RandomBT",
    "RandomBT",
    "RandomBT"
  ),
  
  Persistence = c(
    "None",
    "High >=8 months",
    "Medium >=6 months",
    ">=1 month",
    "High >=6 months",
    "Medium >=4 months",
    ">= 1 month",
   # "High BT >=6 AND OD >=8 months",
    #"Medium BT >=4 AND OD>=6 months",
    #">=1 month",
   # "Matched to 'high' candidates",
   # "Matched to 'medium' candidates",
   # "Matched to 'all' candidates",
    "Matched to OD HIGH",
    "Matched to OD MED",
    "Matched to OD ALL",
    "Matched to BT HIGH",
    "Matched to BT MED",
    "Matched to BT ALL"
  ),
  Number_of_farms_removed = c(
    0,
    length(OD_high_candidates),
    length(OD_med_candidates),
    length(OD_all_candidates),
    length(BT_high_candidates),
    length(BT_med_candidates),
    length(BT_all_candidates),
    #length(CB_high_candidates),
    #length(CB_med_candidates),
    #length(CB_all_candidates),
    #4,
    #9,
    #40,
    5,
    8,
    43,
    1,
    8,
    35
  ),
  
  Percent_of_all_farms = c(
    0,
    length(OD_high_candidates)/total_farms*100,
    length(OD_med_candidates)/total_farms*100,
    length(OD_all_candidates)/total_farms*100,
    length(BT_high_candidates)/total_farms*100,
    length(BT_med_candidates)/total_farms*100,
    length(BT_all_candidates)/total_farms*100,
    #length(CB_high_candidates)/total_farms*100,
    #length(CB_med_candidates)/total_farms*100,
    #length(CB_all_candidates)/total_farms*100,
    #4/total_farms*100,
    #9/total_farms*100,
    #40/total_farms*100,
    5/total_farms*100,
    8/total_farms*100,
    43/total_farms*100,
    1/total_farms*100,
    8/total_farms*100,
    35/total_farms*100
  )
)

scenario_table
```
```{r}
total_farms
```

# 2. show change in largest connected component - GWCC - show removal of potential transmission pathways (undirected) - did removal change the movement network

load and construct the annual network
```{r}
# modify graph/network code from RQ1.Rmd
create_movement_network <- function(df) {

  # Convert IDs
  df$org_id  <- as.character(df$org_id)
  df$dest_id <- as.character(df$dest_id)

  # Node table
  org_nodes <- df[, c("org_id", "org_type")]
  colnames(org_nodes) <- c("name", "type")

  dest_nodes <- df[, c("dest_id", "dest_type")]
  colnames(dest_nodes) <- c("name", "type")

  nodes <- unique(rbind(org_nodes, dest_nodes))

  # Edge list
  edges <- df[, c("org_id", "dest_id")]

  # Create directed network
  g <- graph_from_data_frame(
    d = edges,
    vertices = nodes,
    directed = TRUE
  )

  # Add weights
  E(g)$weight <- df$total_cattle

  return(g)
}
```

```{r}
# import data
annual_movement <- read.csv("RQ1 NO EP.csv")
head(annual_movement)
```


```{r}
# create as graph network by calling function = construct annual movement network
annual_network = create_movement_network(annual_movement)
```
```{r}
# inspect the network
vcount(annual_network)

V(annual_network)$name
```

Calculate baseline GWCC
```{r}
# calculate basline GWCC
baseline_gwcc <- max(components(annual_network, mode = "weak")$csize) # csize = extract number of farms in each connected component
baseline_gwcc
```
Generate random removal controls

```{r}
# create random removal controls
# removal-size controls representing the typical number of farms removed across each persistence category

Control_high_random <- sample(
  V(annual_network)$name,
  size = 4,
  replace = FALSE
)

Control_medium_random <- sample(
  V(annual_network)$name,
  size = 9,
  replace = FALSE
)

Control_all_random <- sample(
  V(annual_network)$name,
  size = 40,
  replace = FALSE
)
# ADD HERE IF WANT TO PUT CONTROLS PER SCENARIO RATHER THAN AVERAGES
```
Define all removal scenarios

```{r}
# create loop to speed up
removal_scenarios <- list(
  S2 = OD_high_candidates,
  S3 = OD_med_candidates,
  S4 = OD_all_candidates,
  S5 = BT_high_candidates,
  S6 = BT_med_candidates,
  S7 = BT_all_candidates
  #S8 = CB_high_candidates,
  #S9 = CB_med_candidates,
  #S10 = CB_all_candidates,
  #S11 = Control_high_random,
  #S12 = Control_medium_random,
  #S13 = Control_all_random,
  #Annual_Bt_FIVE = annual_BT_top5,
  #Annual_OD_FIVE = annual_OD_top5,
  #Annual_OD_deg = annual_OD_top5_deg
  #SC14 = NEED TO ADD THESE
  #SC15 = 
  #SC16 = 
  #SC17 = 
  #SC18 = 
  #SC19 = 
)

```

Create empty results dataframe & apply each removal scenario & calculate network fragmentation

```{r}
gwcc_results <- data.frame(
  Scenario = character(),
  Farms_removed = numeric(),
  GWCC_size = numeric(),
  GWCC_percentage_remaining = numeric(),
  Number_components = numeric()
)

#total_farms_nw = vcount(annual_network)

for (s in names(removal_scenarios)) {
  
  removed_network <- delete_vertices(
    annual_network,
    removal_scenarios[[s]]
  )
  
  comp <- components(removed_network, mode = "weak")
  
  farms_remaining <- vcount(removed_network)
  
  gwcc_size <-max(comp$csize)
  
  gwcc_percent <- (gwcc_size / farms_remaining) * 100 # this is GWCC size (% of farms remaining), NOT percentage of original network
  
  gwcc_results <- rbind(
    gwcc_results,
    data.frame(
      Scenario = s,
      Farms_removed = length(removal_scenarios[[s]]),
      GWCC_size = gwcc_size,
      GWCC_percentage_remaining = gwcc_percent,
      Number_components = comp$no
    )
  )
}
```
Calculate baseline network components again & add the baseline network to the scenario results

```{r}
# calculate baseline network components
baseline_comp <- components(annual_network, mode = "weak")

#baseline_gwcc <- max(baseline_comp$csize)

#baseline_percent <- (baseline_gwcc / vcount(annual_network)) * 100

#baseline_components <- baseline_comp$no

# add baseline GWCC
gwcc_results <- rbind(
  data.frame(
    Scenario = "S1",
    Farms_removed = 0,
    GWCC_size = max(baseline_comp$csize),
    GWCC_percentage_remaining = (max(baseline_comp$csize) / vcount(annual_network))*100,
    Number_components = baseline_comp$no
  ),
  
  gwcc_results
)
```
View final conenctivity results
```{r}
gwcc_results
```
Clear the results object
```{r}
# REMOVE RESULTS
rm(gwcc_results)
```
Final network membership checks
```{r}
sum(annual_OD_top5 %in% V(annual_network)$name)
```
```{r}
length(annual_OD_top5)
length(intersect(annual_OD_top5, V(annual_network)$name))
```

## SOIR Epidemic outcome analysis ##

# Established epidemic definition: threshold sensitivity analysis

1. examine the baseline outbreak-size distribution

  supports why selected >=10 - baseline histogram

```{r}
# plot the distribution

ggplot(summary_results_SC1, aes(x = final_infected)) +
  geom_histogram(
    binwidth = 1,
    colour = "black",
    fill = "steelblue"
  ) +
  labs(
    x = "Final outbreak size (infected farms)",
    y = "Number of simulations",
    title = "Distribution of final outbreak sizes"
  ) +
  theme_classic()
```
Create the scenario outbreak size data

```{r}
summary_results_SC1 <- summary_results$SC1
summary_results_SC2 <- summary_results$SC2
summary_results_SC3 <- summary_results$SC3
summary_results_SC4 <- summary_results$SC4
summary_results_SC5 <- summary_results$SC5
summary_results_SC6 <- summary_results$SC6
summary_results_SC7 <- summary_results$SC7
#summary_results_SC8 <- summary_results$SC8
#summary_results_SC9 <- summary_results$SC9
#summary_results_SC10 <- summary_results$SC10
#summary_results_SC11 <- summary_results$SC11
#summary_results_SC12 <- summary_results$SC12
#summary_results_SC13 <- summary_results$SC13
summary_results_SC14 <- summary_results$SC14
summary_results_SC15 <- summary_results$SC15
summary_results_SC16 <- summary_results$SC16
summary_results_SC17 <- summary_results$SC17
summary_results_SC18 <- summary_results$SC18
summary_results_SC19 <- summary_results$SC19
```
```{r}
# create list of final outbreak sizes
outbreak_sizes <- list(
  Baseline = summary_results_SC1$final_infected,
  Outdegree_high = summary_results_SC2$final_infected,
  Out_degree_med = summary_results_SC3$final_infected,
  Outdegree_all = summary_results_SC4$final_infected,
  Betweenness_high = summary_results_SC5$final_infected,
  Betweenness_med =summary_results_SC6$final_infected,
  Betweenness_all = summary_results_SC7$final_infected
)

```
Threshold sensitivity analysis
  whether >=10 is robust
  
```{r}
thresholds2 <- c(1, 2, 3, 5, 7, 10, 15, 20)

threshold_results2 <- data.frame(
  scenario = character(),
  threshold = numeric(),
  proportion = numeric(),
  stringsAsFactors = FALSE
  )


for (s in names(outbreak_sizes)) {
  
  outbreaks <- outbreak_sizes[[s]]
  
  for (t in thresholds2) {
    threshold_results2 <- rbind(
      threshold_results2,
      data.frame(
        Scenario2 = s,
        Threshold2 = t,
        proportion2 = mean(outbreaks >= t) #* 100
      )
    )
  }
}
```
Sensitivity plot
```{r}
# plot sensitivity analysis

ggplot(threshold_results2,
       aes(x = Threshold2,
           y = proportion2,
           colour = Scenario2)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = "Threshold (infected farms)",
    y = "Probability of a major outbreak",
    colour = "Scenario"
  ) +
  theme_bw()
```


# RESULT 1

Likelihood of sustained transmission - replicats led to >=10 infected final infected farms

Define the scenarios and extract final outbreak size, this is the input to the analysis

```{r}
# list of scenarios to include - dropped combined - see prev
sustained_trans <- list(
  baseline = summary_results_SC1$final_infected, 
  OD_high = summary_results_SC2$final_infected,
  OD_med = summary_results_SC3$final_infected, 
  OD_all = summary_results_SC4$final_infected, 
  BT_high = summary_results_SC5$final_infected, 
  BT_med = summary_results_SC6$final_infected, 
  BT_all = summary_results_SC7$final_infected, 
  #CT_high = summary_results_SC11 
  #CT_med = summary_results_SC12 
  #CT_ALL = summary_results_SC13 
  CT_OD_HIGH = summary_results_SC14$final_infected,
  CT_OD_MED = summary_results_SC15$final_infected,
  CT_OD_ALL = summary_results_SC16$final_infected,
  CT_BT_HIGH = summary_results_SC17$final_infected,
  CT_BT_MED = summary_results_SC18$final_infected,
  CT_BT_ALL = summary_results_SC19$final_infected
)
```
Calculate the probability of sustained transmission and relative reduction

```{r}
# counts how many simulations exceeded the threshold & calculates number sustained outbreaks; proportion sustained outbreaks; relative reduction vs baseline
threshold_final <- 10

threshold_counts_final <- data.frame(
  Scenario_final = names(sustained_trans),
  Exceeding = sapply(sustained_trans, function(x) sum(x >= threshold_final)),
  Total = sapply(sustained_trans, length)
)

threshold_counts_final$Proportion_final <- threshold_counts_final$Exceeding / threshold_counts_final$Total

# extract baseline probability
baseline_prob <- threshold_counts_final$Proportion_final[threshold_counts_final$Scenario_final == "baseline"]

# calculate relatve reduction compared with baseline
threshold_counts_final$Relative_reduction <- (
  (baseline_prob - threshold_counts_final$Proportion_final) / baseline_prob
)

threshold_counts_final
```

Organise scenarios by intervention type and intensity

```{r}
# AI - how can i group this plot by intensity windows, maintaining controls
# group by intensity and intervention type

# grepl() - searches for a pattern in scenario name e.g. "High"
# factor(..., levels = ...) - tells ggplot the order to display categories instead of using alphabetical order

# create two empty columns to store grouping information for plotting
threshold_counts_final$Intensity <- NA
threshold_counts_final$Method <- NA

  # assign intervention intensity

# search scenario names (grepl) & assign each one to an intervention intensity label
threshold_counts_final$Intensity[grepl("HIGH", threshold_counts_final$Scenario_final, ignore.case = TRUE)] <- "High" # looks for high in every scenario name & assigns method
threshold_counts_final$Intensity[grepl("MED", threshold_counts_final$Scenario_final, ignore.case = TRUE)] <- "Medium"
threshold_counts_final$Intensity[grepl("ALL", threshold_counts_final$Scenario_final, ignore.case = TRUE)] <- "All"

# assign the baseline scenario its own intensity label
threshold_counts_final$Intensity[threshold_counts_final$Scenario_final == "baseline"] <- "Baseline"

  # assign intervention method

# Label taregted interventions (grepl) according to the network metric used
threshold_counts_final$Method[grepl("^OD", threshold_counts_final$Scenario_final)] <- "Out-degree"
threshold_counts_final$Method[grepl("^BT", threshold_counts_final$Scenario_final)] <- "Betweenness"

# Keep the random matched controls separate from the targeted interventions
threshold_counts_final$Method[grepl("CT_OD", threshold_counts_final$Scenario_final)] <- "Control (OD matched)"
threshold_counts_final$Method[grepl("CT_BT", threshold_counts_final$Scenario_final)] <- "Control (BT matched)"

# label the baseline scenario
threshold_counts_final$Method[threshold_counts_final$Scenario_final == "baseline"] <- "Baseline"


  # Set plotting order
# convert intensity to an ordered factor so the panels appear in order wanted rather than alphabetically
threshold_counts_final$Intensity <- factor(
  threshold_counts_final$Intensity,
  levels = c("High", "Medium", "All", "Baseline")
)

# convert method to an ordered factor to bars appear in order wanted rather than alphabetically 

threshold_counts_final$Method <- factor(
  threshold_counts_final$Method,
  levels = c(
    "Control (OD matched)",
    "Out-degree",
    "Control (BT matched)",
    "Betweenness",
    "Baseline"
  )
)
```

Create grouped probability plot

```{r}
# Duplicate baseline into each intensity panel

baseline_row <- threshold_counts_final[
  threshold_counts_final$Scenario_final == "baseline",
]

baseline_repeated <- baseline_row[rep(1, 3), ]

baseline_repeated$Intensity <- c("High", "Medium", "All")
baseline_repeated$Method <- "Baseline"

# Remove original baseline and add repeated baseline
plot_data <- rbind(
  threshold_counts_final[
    threshold_counts_final$Scenario_final != "baseline",
  ],
  baseline_repeated
)

# Set order of panels
plot_data$Intensity <- factor(
  plot_data$Intensity,
  levels = c("High", "Medium", "All")
)

# Set order of bars
plot_data$Method <- factor(
  plot_data$Method,
  levels = c(
    "Baseline",
    "Control (OD matched)",
    "Out-degree",
    "Control (BT matched)",
    "Betweenness"
  )
)
```

Plot figure

```{r, fig.width = 12, fig.height = 4}
# plot probability of sustained outbreak by intensity

# order of windows
threshold_counts_final$Intensity <- factor(
  threshold_counts_final$Intensity,
  levels = c("Baseline", "High", "Medium", "All")
)
ggplot(
  plot_data,
  aes(x = Method,
      y = Proportion_final,
      fill = Method)
) +
  geom_col(position = "dodge") +
  facet_wrap(~Intensity) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = "Intervention strategy",
    y = "Probability of sustained outbreak (≥10 infected farms)"
  ) +
  theme_bw()
```

Relative Reduction

```{r}
# Duplicate baseline into each intensity panel for relative reduction plot

baseline_reduction <- threshold_counts_final[
  threshold_counts_final$Scenario_final == "baseline",
]

baseline_reduction <- baseline_reduction[rep(1, 3), ]

baseline_reduction$Intensity <- c("High", "Medium", "All")
baseline_reduction$Method <- "Baseline"
baseline_reduction$Relative_reduction <- 0


# Remove original baseline and add repeated baseline
plot_reduction <- rbind(
  threshold_counts_final[
    threshold_counts_final$Scenario_final != "baseline",
  ],
  baseline_reduction
)


# Set panel order
plot_reduction$Intensity <- factor(
  plot_reduction$Intensity,
  levels = c("High", "Medium", "All")
)


# Set bar order
plot_reduction$Method <- factor(
  plot_reduction$Method,
  levels = c(
    "Baseline",
    "Control (OD matched)",
    "Out-degree",
    "Control (BT matched)",
    "Betweenness"
  )
)
```
```{r, fig.width=12, fig.height=4}
# plot relative reduction by intensity

# order of windows
#threshold_counts_final$Intensity <- factor(
  #threshold_counts_final$Intensity,
  #levels = c("High", "Medium", "All")
#)

ggplot(
  plot_reduction,
  aes(x = Method,
      y = Relative_reduction,
      fill = Method)
) +
  geom_col(position = "dodge") +
  facet_wrap(~Intensity) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = "Intervention strategy",
    y = "Relative outbreak reduction vs baseline"
  ) +
  theme_bw()
```

Paired fishers tests

# RESULT 2

if an epidemic becomes established (threshold >= 10), how large does it become/how many farms are infected?

Report - Median, IQR, % reduction relative to baseline

Filter established outbreaks:
```{r}
# filter simulations above threshold - keep sims where finout outbreak size >= 10

threshold_final <- 10

established_outbreaks <- lapply(
  sustained_trans,
  function(x) x[x >= threshold_final]
)
```
```{r}
# create a dataframe for plotting

plot_sizes <- stack(established_outbreaks)
colnames(plot_sizes) <- c("Final_size", "Scenario")

# final_size= final number of infected farms
# scenario = intervention scenario
```
```{r}
# add intervention method & intensity
plot_sizes <- merge(
  plot_sizes,
  threshold_counts_final[, c("Scenario_final", "Method", "Intensity")],
  by.x = "Scenario",
  by.y = "Scenario_final"
)
plot_sizes_stats <- plot_sizes # save a copy before duplication for stat analysis without baseline x3
```
Baseline duplication - needed for faceted figure of panels

```{r}
# duplicate baseline into every intensity panel

# extract baseline
baseline_sizes <- subset(plot_sizes, Scenario == "baseline")
```
```{r}
# duplicate baseline
baseline_sizes <- do.call(
  rbind,
  lapply(c("High", "Medium", "All"), function(i) {
    x <- baseline_sizes
    x$Intensity <- i
    x
  })
)

baseline_sizes$Method <- "Baseline"
```
```{r}
# remove original baseline & add duplicated baseline
plot_sizes <- rbind(
  subset(plot_sizes, Scenario != "baseline"),
  baseline_sizes
)
```
Set plotting order

```{r}
# set plotting order
plot_sizes$Intensity <- factor(
  plot_sizes$Intensity,
  levels = c("High", "Medium", "All")
)

plot_sizes$Method <- factor(
  plot_sizes$Method,
  levels = c(
    "Baseline",
    "Control (OD matched)",
    "Out-degree",
    "Control (BT matched)",
    "Betweenness"
  )
)
```
Boxplot

```{r}
# create boxplots

ggplot(
  plot_sizes,
  aes(
    x = Method,
    y = Final_size,
    fill = Method
  )
) +
  geom_boxplot() +
  facet_wrap(~Intensity) +
  labs(
    x = "Intervention strategy",
    y = "Final outbreak size (established outbreaks only)"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    )
  )
```
Descriptive summary - median/IQR/mean etc

```{r}
# summarise outbreak sizes

outbreak_summary <- data.frame(
  Scenario = names(established_outbreaks),
  N = sapply(established_outbreaks, length), # how many sims exceeded threshold
  Mean = sapply(established_outbreaks, mean), # average final outbreak size
  Median = sapply(established_outbreaks, median), # median final outbreak size
  SD = sapply(established_outbreaks, sd), # variability/standard deviation
  Min = sapply(established_outbreaks, min), # range
  Max = sapply(established_outbreaks, max) # range
)

outbreak_summary
```

STATS - poissons distribution

Visualise distributions - show that final outbreak sizes are not normally distributed - ?right skewed

```{r}
# histogram of outbreak sizes
ggplot(plot_sizes_stats,
       aes(x = Final_size)) +
  geom_histogram(
    bins = 30,
    fill = "steelblue",
    colour = "black"
  ) +
  facet_wrap(~Method) +
  labs(
    x = "Final outbreak size (established outbreaks only)",
    y = "Number of simulations"
  ) +
  theme_bw()
```


# RESULT 3
c) peak infectious farms - how intense is outbreak IF THERE IS AN OUTBREAK - explain as above - mean/median peak without <1 sims

When an epidemic occurs (threshold >=10 farms), how intense is it?

Only analyse successful epidemics

Calculate peak infectious farms for every replicate

```{r}
# calculate peak infectious farms for every simulation = max number of infectious farms reached in each simulation
peak_infectious <- lapply(
  daily_results,
  function(x) {
    x %>%
      group_by(replicate) %>%
      summarise(
        peak_infectious = max(infectious)
      )
  }
)
```
Restrict to established epidemics

```{r}
# keep sims where threshold transmission occured
# create dataframe with replicate/peak_inefctious/scenario

peak_infectious_established <- lapply(
  names(peak_infectious),
  function(scenario) {
    
    peaks <- peak_infectious[[scenario]]
    
    established <- summary_results[[scenario]] %>%
      filter(final_infected >= threshold_final) %>%
      select(replicate)
    
    peaks %>%
      filter(replicate %in% established$replicate) %>%
      mutate(Scenario = scenario)
  }
)

peak_infectious_established <- bind_rows(peak_infectious_established)
```

Create scenario look-up

```{r}
# create look up table to to map simulation names to scenario labels

scenario_lookup <- data.frame(
  Scenario = c(
    "SC1", "SC2", "SC3", "SC4", "SC5", "SC6", "SC7", "SC14", "SC15", "SC16", "SC17", "SC18", "SC19"
  ),
  Scenario_final = c(
    "baseline",  "OD_high", "OD_med", "OD_all", "BT_high", "BT_med", "BT_all", "CT_OD_HIGH", "CT_OD_MED", "CT_OD_ALL", "CT_BT_HIGH", "CT_BT_MED","CT_BT_ALL"
  )
)
```

Add method/intensity labels

```{r}
# add method & intensity labels
peak_infectious_established <- merge(
  peak_infectious_established,
  scenario_lookup,
  by = "Scenario"
)

peak_infectious_established <- merge(
  peak_infectious_established,
  threshold_counts_final[, c("Scenario_final", "Method", "Intensity")],
  by = "Scenario_final"
)

peak_infectious_stats <- peak_infectious_established # create copy for stats
```
Create the plotting copy

```{r}
#make separate copy for figures
peak_infectious_plot <- peak_infectious_established

nrow(peak_infectious_plot)
# lower because has dropped scenarios 8-13
```
Duplicate baseline for the three intensity panels

```{r}
# duplicate baseline so appears in all 3 panels

# extract basline peak infecitous values
baseline_peak <- subset(
  peak_infectious_plot,
  Method == "Baseline"
)

# Duplicate baseline into each intensity panel
baseline_peak <- baseline_peak[
  rep(1:nrow(baseline_peak), 3),
]

# assign intensity labels
baseline_peak$Intensity <- rep(
  c("High", "Medium", "All"),
  each = nrow(baseline_peak) / 3
)

# keep baseline label
baseline_peak$Method <- "Baseline"
```
```{r}
# remove original baseline and add duplicated baseline back
peak_infectious_plot <- rbind(
  subset(peak_infectious_plot, Method != "Baseline"),
  baseline_peak
)
```
Set plotting order

```{r}
# set plotting order
peak_infectious_plot$Intensity <- factor(
  peak_infectious_plot$Intensity,
  levels = c("High", "Medium", "All")
)

peak_infectious_plot$Method <- factor(
  peak_infectious_plot$Method,
  levels = c("Baseline", "Control (OD matched)",
             "Out-degree",
             "Control (BT matched)",
             "Betweenness"
             )
)
```
Final boxplot

```{r}
# plot
ggplot(
  peak_infectious_plot,
  aes(
    x = Method,
    y = peak_infectious,
    fill = Method
  )
) +
  geom_boxplot() +
  facet_wrap(~Intensity) +
  labs(
    x = "Intervention strategy",
    y = "Peak infectious farms (established outbreaks only)"
  )+
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```
Non parametric distribution/densities - stat test choices
Show densitiies non-parametric

```{r}
# density plots
ggplot(peak_infectious_plot,
       aes(x = peak_infectious,
           colour = Method,
           fill = Method)) +
  geom_density(alpha = 0.2) +
  facet_wrap(~Intensity) +
  labs(
    x = "Peak infectious size (established outbreaks only)",
    y = "Density"
  ) +
  theme_bw()
```

Stats - poissons distribution


