---
title: "RQ2 SOIR SIMS ZIP"
author: "ESINCLAIR"
date: "2026-08-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# import libraries
library(ggplot2)
library(dplyr)
library(tidyr)
library(pbapply)
```
(original notebook = TEST SIM 2 (initial gen model))

PROMPT (ChatGPT)
show me basic complete R code for an SEIR model that tracks btb-like disease spread across nodes (farms). each movement event has a total_cattle batch size. Each farm can send or recieve several batches of cattle per day. I want to be able to seed an infection in a node an view the potential disease spread. I want to be able to remove certain  nodes  identified as superspreaders and compare simulation outcomes. Possibly in the future, I would like to compute probabiltiy of tranmission to be weighted by number of cattle

```{r}
# This is base model function from gpt prompt - a SEIR compartment stochastic model - farm level rather than animal level (issues with Movenet/DigiVet)

# Everything in this block is inside the run_seir function
# takes input:
  # movement data (movements) and a 
  # starting infected farm (seed_farm)
  # transmission probability (beta) 
  # latent_period - time between becoming infected and infectious - in compartment 'E' - incubating but CANNOT yet infect other farms through  movements - (?make latency move with animal - realistic infection developing further down contact chain?)
  # infectious period - time spent infectious and capable of transmitting infection before moving into R (removed) - model assumption based on time to detect infected farm. During these 80d - movements from the farm can transmit infection according to beta

run_seir <- function(
    movements,
    seed_farm,
    beta = 0.05, 
    latent_period = 60, 
    infectious_period = 180 
){


# Overall timeline if a susceptible farm becomes infected on day 100: 0-99d = susceptible, can't infect others; 100-159d = exposed, can't infect others, 160-339d = infectious, can infect others, 340d onwards = removed, cant infect others

# Build lookup table - pulls every unique farm ID from both the source and destination columns of movements  - combines, removes duplicates, sorts --> (farms) is a clean lookup table of all farms involved in the dataset
  
  farms <- sort(unique(c(
    movements$source,
    movements$destination
  )))
  
# creates a numeric index for the farms, labels each with a number corresponding to the farm ID
  n_farms <- length(farms)

  farm_index <- seq_len(n_farms)
  names(farm_index) <- farms

 
# Convert movements to integer indices #

  movements$source_id <- farm_index[movements$source]
  movements$dest_id   <- farm_index[movements$destination]

  
# Disease state definitions:
  
  # 0 = S = susceptible
  # 1 = E = exposed
  # 2 = I = infectious
  # 3 = R = removed

# set up vectors to track each farms disease status and how long it has been in its current state
  
  state <- integer(n_farms) # creates a vector of zeros - meaning all farms start susceptible

  days <- integer(n_farms) # creates a parallel vector to count how many days each farm has spent in its current state

 
# Seed infection - infect chosen seed farm by setting its state to 2(I)

  state[farm_index[seed_farm]] <- 2


# Storage - prepares storage for the simulation results

  max_day <- max(movements$day) # finds last day in the movement dataset

# History - creates a data frame with one row per day, and columns for the number of farms in each SEIR state, all columns are pre-allocated as integer vectors of length max_day for filling during the simulation
  
  history <- data.frame( 
    day = integer(max_day), 
    S = integer(max_day),
    E = integer(max_day),
    I = integer(max_day),
    R = integer(max_day)
  )
  


# Simulation - advances farms through the E -> I -> R stages as each day of the simulation passes

# the 'for' loop runs one iteration per day

  for(day in seq_len(max_day)){

 
# Progress exposed farms:
 
# the day counters increase, and farms that have spent more than latent_period days in E become infectious (state == 2) and their day counters reset
    
    exposed <- which(state == 1)

    if(length(exposed) > 0){

      days[exposed] <- days[exposed] + 1

      become_I <- exposed[
        days[exposed] >= latent_period
      ]

      state[become_I] <- 2
      days[become_I] <- 0

    }


# Progress infectious farms

    infectious <- which(state == 2)

# Their day counters increase, any farms that have spent more than infectious_period i days in I become removed (state == 3) and their day counters reset    
    
    if(length(infectious) > 0){

      days[infectious] <- days[infectious] + 1

      become_R <- infectious[
        days[infectious] >= infectious_period
      ]

      state[become_R] <- 3
      days[become_R] <- 0

    }


# Today's movements - filters the 'movements' data frame to keep just the rows where 'day' matches the current loop day, giving the set of movements to process for that day.

    todays <- movements[movements$day == day, ]


# Replay movement events - processes each movement and applies infection-spread rules.
# loop over all movements happening today, for each movement, get the source and destination farm indices; skip the movement if the source is not infectious (state != 2); skip the movement if the destination is not susceptible (state != 0). Only movements from the infectious -> susceptible farms proceed to the transmission step that follows
    

    if(nrow(todays) > 0){

      for(i in seq_len(nrow(todays))){

        source <- todays$source_id[i]
        dest   <- todays$dest_id[i]

        # source not infectious
        if(state[source] != 2)
          next

        # destination already infected - i.e. not susceptible
        if(state[dest] != 0)
          next


# Transmission - applies transmission step - an infectious source may infect a susceptible destination. With probability 'beta', the destination farm becomes exposed (state = 1) and is day counter resets to 0 
# https://www.rdocumentation.org/packages/stats/versions/3.5.2/topics/Uniform - runif() - generates random number between 0-1, if number is < beta - infection progresses = Bernoulli
        
        if(runif(1) < beta){

          state[dest] <- 1
          days[dest] <- 0

        }

      }

    }

    
# Save epidemic - logs the SEIR count for the current simulation day - store the day number; count how many farms are S,E,I,R and write those values into the corresponding row of history
    
    history$day[day] <- day
    history$S[day] <- sum(state == 0)
    history$E[day] <- sum(state == 1)
    history$I[day] <- sum(state == 2)
    history$R[day] <- sum(state == 3)

  }


# Return - returns the simulation output as a structured list
  
  list(

    history = history, # stores the full day-by-day SEIR counts

    final_states = data.frame( # data frame listing each farm and its final SEIR state, converts numeric codes (0-3) into labelled factors S,E,I,R

      farm = farms,

      state = factor(
        state,
        levels = 0:3,
        labels = c("S","E","I","R")
      )

    )

  )

}
```

# Provide movement data

```{r}
# read in csv
movements_raw <- read.csv(
    "RQ1 NO EP.csv",
    stringsAsFactors = FALSE
)
# head(movements_raw)
```

```{r}
# check the structure
str(movements_raw)
```
```{r}
# create simulation movement  table
movements <- movements_raw %>%
    select(
        source = org_id,
        destination = dest_id,
        day,
        total_cattle
    )
# head(movements)
```


```{r}
# convert  farm IDs to chr rather than integer to stop e.g. 1 being  treated differently from 001 in future datasets
movements <- movements %>%
    mutate(
        source = as.character(source),
        destination = as.character(destination)
    )
```

```{r}
# check structure
str(movements)
```
```{r}
# number of movement events
nrow(movements)
```
# Run simulation
```{r}
# N.B.this is only running one replicate of the simulation- being stochastic need to run 100-1000 replicates because each will be slightly different
result <- run_seir(
    movements,
    seed_farm = "285" # choose farm to seed infection - needs updated to allow several farms
)

head(result$history)
```
```{r}
# plot the curve - using ggplot

result$history %>%
    pivot_longer(
        cols = c(S,E,I,R),
        names_to = "state",
        values_to = "farms"
    ) %>%
    ggplot(
        aes(
            x = day,
            y = farms,
            colour = state
        )
    ) +
    geom_line(size = 1) +
    theme_minimal()
```
# Update to SOIR model

from 'SOIR MODEL'

State encoding            Movement transmission?
0 = S (Suscpetible farm)          No
1 = O (Occult infection)          No (evidence to show can shed during occult - left out for simplicity of model)
2 = I (Infectious/shedding farm)  Yes
3 = R (removed)                   No

Updated to SOIR because in bTB, infection does not immediately = infectiousness (https://pmc.ncbi.nlm.nih.gov/articles/PMC3475695/#s4)

```{r}
# model is unweighted by total_cattle - movement is either present or absent
run_soir <- function(
    movements,
    seed_farms, # can seed 1+ farms in one run - see update below
    beta = 0.05, # explore 0.001 - 0.05
    occult_period = 120, # explore 60-240
    infectious_period = 180 # explore 90 - 135
){
  
  
# Build lookup table - pulls every unique farm ID from both the source and destination columns of movements  - combines, removes duplicates, sorts --> (farms) is a clean lookup table of all farms involved in the dataset
  
  farms <- sort(unique(c(
    movements$source,
    movements$destination
  )))
  
# creates a numeric index for the farms, labels each with a number corresponding to the farm ID
  n_farms <- length(farms)

  farm_index <- seq_len(n_farms)
  names(farm_index) <- farms

 
# Convert movements to integer indices

  movements$source_id <- farm_index[movements$source]
  movements$dest_id   <- farm_index[movements$destination]

# Set up vectors to track each farms disease status and how long it has been in its current state
  
  state <- integer(n_farms) # creates a vector of zeros - meaning all farms start susceptible

  days <- integer(n_farms) # creates a parallel vector to count how many days each farm has spent in its current state

 
# Seed infection - infect chosen seed farm by setting its state to 1 - farms to start as occult - representative of dz introduction
# Reset their disease timers to 0 days, indicating they have just entered the infectious state
  
  state[farm_index[seed_farms]] <- 1  # find farm(s) listed in seed_farms, look up their position in the 'state' vector using 'farm_index'; change disease state S -> 0
  days[farm_index[seed_farms]] <- 0 # for the same seeded farm(s), set no. days spent in current dz state to 0 - farm has just become occult at the start of the simulation

# Storage - prepares storage for the simulation results

  max_day <- max(movements$day) # finds last day in the movement dataset

# history - creates a data frame with one row per day, and columns for the number of farms in each SOIR state, all columns are pre-allocated as integer vectors of length max_day for filling during the simulation
  
  history <- data.frame( 
    day = integer(max_day), 
    S = integer(max_day),
    O = integer(max_day),
    I = integer(max_day),
    R = integer(max_day)
  )


# Simulation - advances farms through the O -> I -> R stages as each day of the simulation passes
# the 'for' loop runs one iteration per day

  for(day in seq_len(max_day)){


# Progress exposed farms:
# The day counters increase, and farms that have spent more than occult_period days in O become infectious (state == 2) and their day counters reset
    
    occult <- which(state == 1)

    if(length(occult) > 0){

      days[occult] <- days[occult] + 1 # increments until reaches occult_period -> becomes infectious

      become_I <- occult[
        days[occult] >= occult_period
      ]

      state[become_I] <- 2
      days[become_I] <- 0

    }


# Progress infectious farms

    infectious <- which(state == 2)

# Their day counters increase, any farms that have spent more than infectious_period i days in I become removed (state == 3) and their day counters reset    
    
    if(length(infectious) > 0){

      days[infectious] <- days[infectious] + 1

      become_R <- infectious[
        days[infectious] >= infectious_period
      ]

      state[become_R] <- 3
      days[become_R] <- 0

    }

# Today's movements - filters the 'movements' data frame to keep just the rows where 'day' matches the current loop day, giving the set of movements to process for that day.

    todays <- movements[movements$day == day, ]

# Replay movement events - processes each movement and applies infection-spread rules.
# loop over all movements happening today, for each movement, get the source and destination farm indices; skip the movement if the source is not infectious (state != 2); skip the movement if the destination is not susceptible (state != 0). Only movements from the infectious -> susceptible farms proceed to the transmission step that follows.
    

    if(nrow(todays) > 0){

      for(i in seq_len(nrow(todays))){

        source <- todays$source_id[i]
        dest   <- todays$dest_id[i]

        # source not infectious
        if(state[source] != 2)
          next

        # destination already infected
        if(state[dest] != 0)
          next


# Transmission - applies transmission step - an infectious source may infect a susceptible destination. With probability 'beta', the destination farm becomes exposed (state = 1) and is day counter resets to 0 
# runif(1) generates a random number between 0 and 1, if that random number is less than beta, transmission occurs - changes destination farm's state from S -> O, resets the dz timer. So if beta = 0.05 = 5% chance that an infectious movement infects the destination farm. Bernoulli probability 0.05/0.95; https://www.rdocumentation.org/packages/stats/versions/3.5.2/topics/Uniform
       
        if(runif(1) < beta){

          state[dest] <- 1
          days[dest] <- 0

        }

      }

    }

# Save epidemic - logs the SOIR count for the current simulation day - store the day number; count how many farms are S,O,I,R and write those values into the corresponding row of history
    

    history$day[day] <- day
    history$S[day] <- sum(state == 0)
    history$O[day] <- sum(state == 1)
    history$I[day] <- sum(state == 2)
    history$R[day] <- sum(state == 3)

  }

# Return - returns the simulation output as a structured list

  list(

    history = history, # stores the full day-by-day SOIR counts

    final_states = data.frame( # data frame listing each farm and its final SOIR state, converts numeric codes (0-3) into labelled factors S,O,I,R

      farm = farms,

      state = factor(
        state,
        levels = 0:3,
        labels = c("S","O","I","R")
      )

    )

  )

}
```

Transmission events were simulated stochastically using Bernoulli trials. For each eligible movement, a uniform random number (runif(1)) was drawn and transmission occurred when the random draw was less than the calculated transmission probability (beta).

Provide movement data
```{r}
# read in csv
movements_raw <- read.csv(
    "RQ1 NO EP.csv",
    stringsAsFactors = FALSE
)
#head(movements_raw )
```
```{r}
# check the structure
str(movements_raw)
```
```{r}
# create simulation movement  table
movements <- movements_raw %>%
    select(
        source = org_id,
        destination = dest_id,
        day,
        total_cattle
    )
#head(movements)
```
```{r}
# convert  farm IDs to chr rather than integer to stop e.g. 1 being  treated differently from 001 in future datasets
movements <- movements %>%
    mutate(
        source = as.character(source),
        destination = as.character(destination)
    )
```
```{r}
str(movements)
```
```{r}
# number of movement events
nrow(movements)
```

```{r}
# N.B. this is only one replicate - is stochastic - need 100-1000 replicates
result <- run_soir(
    movements,
    seed_farms = c("2", "300", "193", "65")
)

#head(result$history)
```
```{r}
test <- run_soir(
  movements = movements,
  seed_farms = "95",
  beta = 0.05,
  occult_period = 120,
  infectious_period = 180
)
```
```{r}
# plot data - ggplot
test$history %>%
    tidyr::pivot_longer(
        cols = c(S,O,I,R),
        names_to = "state",
        values_to = "farms"
    ) %>%
    ggplot(
        aes(
            x = day,
            y = farms,
            colour = state
        )
    ) +
    geom_line(size = 1) +
    theme_minimal()
```
NEXT - SENSITIVITY ANALYSIS - MIGHT AS WELL DO AT SAME TIME AS WRITE UP LIKE FOR RQ1

