---
title: "RQ2 Results & Analysis"
author: "ESINCLAIR"
date: "2026-08-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
# libraries
library(igraph)
library(ggplot2)
library(dplyr)
library(tidyr)
library(pbapply)
```

Import results of simulations for anlysis (RQ2 SOIR SIMS)

```{r}
scenario_1_results <- readRDS("scenario_1_results.rds") # baseline
scenario_2_results <- readRDS("scenario_2_results.rds") # Outdegree (high) >= 8 months
scenario_3_results <- readRDS("scenario_3_results.rds") # Outdegree (med) >= 6 months
scenario_4_results <- readRDS("scenario_4_results.rds") # Outdegree any month
scenario_5_results <- readRDS("scenario_5_results_corrected.rds") # Betweenness (high) >= 6 months
scenario_6_results <- readRDS("scenario_6_results_corrected.rds") # Betweenness (med) >= 4 months
scenario_7_results <- readRDS("scenario_7_results_corrected.rds") # Betweenness any month
#scenario_8_results <- readRDS("scenario_8_results.rds") # Combined (high) BT >= 6 months & OD >= 8 months
#scenario_9_results <- readRDS("scenario_9_results.rds") # Combined (med) BT>= 4 months & 0D >= 6 months
#scenario_10_results <- readRDS("scenario_10_results.rds") # Combined any month
#scenario_11_results <- readRDS("scenario_11_results.rds") # control High - random 4 farms (average of scenarios 3/6/9) = 4.33 (4)
#scenario_12_results <- readRDS("scenario_12_results.rds") # control med - random 9 farms (average of scenarios 2/5/8) = 9.33 (9)
#scenario_13_results <- readRDS("scenario_13_results.rds") # random 40 farms (average of scenarios 4/7/10) = 40.33 (40)
scenario_14_results <- readRDS("scenario_14_results.rds") # OD HIGH CONTROL (5 removed)
scenario_15_results <- readRDS("scenario_15_results.rds") # OD MED CONTROL (8 removed)
scenario_16_results <- readRDS("scenario_16_results.rds") # OD ALL CONTROL (43 removed)
scenario_17_results <- readRDS("scenario_17_results_corrected.rds") # BT HIGH CONTROL (1 removed)
scenario_18_results <- readRDS("scenario_18_results_corrected.rds") # BT MED CONTROL (8 removed)
scenario_19_results <- readRDS("scenario_19_results_corrected.rds") # BT ALL CONTROL (35 removed)
```

```{r}
# combine the daily outputs from all 2000 sim replicates into one data frame

# daily result - epidemic curves for all 2000 stochastic runs
daily_results_SC1 <- bind_rows( 
  lapply(scenario_1_results, function(x) x$daily) 
)
# peak infectious and final outbreak size per run
summary_results_SC1 <- bind_rows(
  lapply(scenario_1_results, function(x) x$summary)
)
```

```{r}
# Put all scenario results into one list
all_scenarios <- list(
  SC1 = scenario_1_results,
  SC2 = scenario_2_results,
  SC3 = scenario_3_results,
  SC4 = scenario_4_results,
  SC5 = scenario_5_results,
  SC6 = scenario_6_results,
  SC7 = scenario_7_results,
  #SC8 = scenario_8_results,
  #SC9 = scenario_9_results,
  #SC10 = scenario_10_results,
  #SC11 = scenario_11_results,
  #SC12 = scenario_12_results,
  #SC13 = scenario_13_results,
  SC14 = scenario_14_results,
  SC15 = scenario_15_results,
  SC16 = scenario_16_results,
  SC17 = scenario_17_results,
  SC18 = scenario_18_results,
  SC19 = scenario_19_results
)

# Create empty lists to store outputs
daily_results <- list()
summary_results <- list()

# Loop through scenarios
for (scenario in names(all_scenarios)) {
  
  daily_results[[scenario]] <- bind_rows(
    lapply(all_scenarios[[scenario]], function(x) x$daily)
  )
  
  summary_results[[scenario]] <- bind_rows(
    lapply(all_scenarios[[scenario]], function(x) x$summary)
  )
}

# call individual with - summary_results$SC3 etc
```

Read in data from RQ1 for number farms removed
```{r}
# OUTDEGREE
OD_all_candidates <- readRDS(
  "outdegree_all_candidates.rds"
)

OD_med_candidates <- readRDS(
  "outdegree_med_persistent.rds"
)

OD_high_candidates <- readRDS(
  "outdegree_high_persistent.rds"
)
```

```{r}
OD_all_candidates
```

```{r}
# BETWEENNESS
BT_all_candidates <- readRDS(
  "BT_any_persistent_corrected.rds"
)

BT_high_candidates <- readRDS(
  "BT_high_persistent_corrected.rds"
)

BT_med_candidates <- readRDS(
  "BT_med_persistent_corrected.rds"
)
```
```{r}
BT_med_candidates
```

```{r}
# COMBINED BETWEENNESS & OUTDEGREE
#CB_all_candidates <- readRDS(
  "Combined_all_persistent.rds"
)

#CB_high_candidates <- readRDS(
  "Combined_high_persistent.rds"
)

#CB_med_candidates <- readRDS(
  "Combined_med_persistent.rds"
)
```
Core story:
Do temporally persistent structurally important farms represent effective targets for reducing bTB-like disease spread through cattle movements?
1.	Scenario description:
	Scenario table – see below – intervention type, persistence threshold, number of farms removed, % of farms removed
	
2.	Network validation: largest connected component
	Explains the mechanism – shows that farms removed were structurally important – which altered network connectivity & influenced transmission
	Single metric – size of largest connected component before/after removal – simple figure – baseline network LCC & intervention LCCs
3.	SOIR epidemic outcomes
a)	Probability of an epidemic establishment – proportion of simulations producing sustained transmission – define threshold
b)	Final outbreak size = how much disease occurs – median infected farms/IQR/% reduction relative to baseline – CARE here with median – report mean infected farms with uncertainty – SD AND distribution of final outbreak sizes to show stochastic nature of transmission – majority of infection seeds fade out and do not transmit. Could also add conditional outbreak size - 
c)	Peak infectious farms = how intense in the outbreak
d)	Show all results in one figure/table: probability of epidemic establishment; final outbreak size; peak infectious farms – interpret patterns

Drop combined scenarios? – mention in discussion/next steps – but is answering a different question - ? are combined centrality measures better intervention targets than single centrality measures – if add then make it – do farms occupying multiple important network positions provide even better intervention targets?
removing structurally important farms decreases the likelihood of epidemic establishment, limits outbreak size when transmission occurs, and decereases epidemic intensity

# 1. Create scenario table - what was removed

```{r}
total_farms = 301
scenario_table <- data.frame(
  Scenario = c(
    "(1)",
    "(2)",
    "(3)",
    "(4)",
    "(5)",
    "(6)",
    "(7)",
    #"(8)",
    #"(9)",
    #"(10)",
    #"(11)",
    #"(12)",
    #"(13)",
    "(14)",
    "(15)",
    "(16)",
    "(17)",
    "(18)",
    "(19)"
  ),
  
  Type = c(
    "Baseline",
    "Targeted Intervention",
    "Targeted Intervention",
    "Targeted Intervention",
    "Targeted Intervention",
    "Targeted Intervention",
    "Targeted Intervention",
    #"Targeted Intervention",
    #"Targeted Intervention",
    #"Targeted Intervention",
    #"Random removal control",
    #"Random removal control",
    #"Random removal control",
    "CONTROL OD HIGH",
    "CONTROL OD MED",
    "CONTROL OD ALL",
    "CONTROL BT HIGH",
    "CONTROL BT MED",
    "CONTROL BT ALL"
  ),
  
  Centrality = c(
    "None",
    "Out-degree",
    "Out-degree",
    "Out-degree",
    "Betweenness",
    "Betweenness",
    "Betweenness",
    #"Combined",
    #"Combined",
   # "Combined",
   # "Random",
   # "Random",
   # "Random",
    "RandomOD",
    "RandomOD",
    "RandomOD",
    "RandomBT",
    "RandomBT",
    "RandomBT"
  ),
  
  Persistence = c(
    "None",
    "High >=8 months",
    "Medium >=6 months",
    ">=1 month",
    "High >=6 months",
    "Medium >=4 months",
    ">= 1 month",
   # "High BT >=6 AND OD >=8 months",
    #"Medium BT >=4 AND OD>=6 months",
    #">=1 month",
   # "Matched to 'high' candidates",
   # "Matched to 'medium' candidates",
   # "Matched to 'all' candidates",
    "Matched to OD HIGH",
    "Matched to OD MED",
    "Matched to OD ALL",
    "Matched to BT HIGH",
    "Matched to BT MED",
    "Matched to BT ALL"
  ),
  Number_of_farms_removed = c(
    0,
    length(OD_high_candidates),
    length(OD_med_candidates),
    length(OD_all_candidates),
    length(BT_high_candidates),
    length(BT_med_candidates),
    length(BT_all_candidates),
    #length(CB_high_candidates),
    #length(CB_med_candidates),
    #length(CB_all_candidates),
    #4,
    #9,
    #40,
    5,
    8,
    43,
    1,
    8,
    35
  ),
  
  Percent_of_all_farms = c(
    0,
    length(OD_high_candidates)/total_farms*100,
    length(OD_med_candidates)/total_farms*100,
    length(OD_all_candidates)/total_farms*100,
    length(BT_high_candidates)/total_farms*100,
    length(BT_med_candidates)/total_farms*100,
    length(BT_all_candidates)/total_farms*100,
    #length(CB_high_candidates)/total_farms*100,
    #length(CB_med_candidates)/total_farms*100,
    #length(CB_all_candidates)/total_farms*100,
    #4/total_farms*100,
    #9/total_farms*100,
    #40/total_farms*100,
    5/total_farms*100,
    8/total_farms*100,
    43/total_farms*100,
    1/total_farms*100,
    8/total_farms*100,
    35/total_farms*100
  )
)

scenario_table
```
# 2. show change in largest connected component - GWCC - show removal of potential transmission pathways (undirected) - did removal change the movement network
Table and figure (x-axis: scenario; y_axis: GWCC size (% of farms)) - bars/points

  does betweeness removal fragment network more? does outdegree removal have less effect on connectivity; does combined have the largest effect?

Need to construct a whole year aggregated movement network, then apply each removal intervention to that network and calculate GWCC after removal

```{r}
# modify graph/network code from RQ1.Rmd
create_movement_network <- function(df) {

  # Convert IDs
  df$org_id  <- as.character(df$org_id)
  df$dest_id <- as.character(df$dest_id)

  # Node table
  org_nodes <- df[, c("org_id", "org_type")]
  colnames(org_nodes) <- c("name", "type")

  dest_nodes <- df[, c("dest_id", "dest_type")]
  colnames(dest_nodes) <- c("name", "type")

  nodes <- unique(rbind(org_nodes, dest_nodes))

  # Edge list
  edges <- df[, c("org_id", "dest_id")]

  # Create directed network
  g <- graph_from_data_frame(
    d = edges,
    vertices = nodes,
    directed = TRUE
  )

  # Add weights
  E(g)$weight <- df$total_cattle

  return(g)
}
```

```{r}
# import data
annual_movement <- read.csv("RQ1 NO EP.csv")
head(annual_movement)
```



```{r}
# create as graph network by calling function
annual_network = create_movement_network(annual_movement)
```
```{r}
vcount(annual_network)

V(annual_network)$name
```

```{r}
# calculate basline GWCC
baseline_gwcc <- max(components(annual_network, mode = "weak")$csize) # csize = extract number of farms in each connected component
baseline_gwcc
```
```{r}
# create random removal controls
# removal-size controls representing the typical number of farms removed across each persistence category

Control_high_random <- sample(
  V(annual_network)$name,
  size = 4,
  replace = FALSE
)

Control_medium_random <- sample(
  V(annual_network)$name,
  size = 9,
  replace = FALSE
)

Control_all_random <- sample(
  V(annual_network)$name,
  size = 40,
  replace = FALSE
)
# ADD HERE IF WANT TO PUT CONTROLS PER SCENARIO RATHER THAN AVERAGES
```
```{r}
# remove farms for each scenario
network_removed_ODhigh <- delete_vertices(
  annual_network,
  OD_high_candidates
)
```
```{r}
# calculate GWCC
GWCC_removed_ODhigh <- max(components(network_removed_ODhigh, mode = "weak")$csize)
GWCC_removed_ODhigh
```
```{r}
test <- components(network_removed_ODhigh, mode = "weak")$no
test
```
```{r}
# create loop to speed up
removal_scenarios <- list(
  S2 = OD_high_candidates,
  S3 = OD_med_candidates,
  S4 = OD_all_candidates,
  S5 = BT_high_candidates,
  S6 = BT_med_candidates,
  S7 = BT_all_candidates
  #S8 = CB_high_candidates,
  #S9 = CB_med_candidates,
  #S10 = CB_all_candidates,
  #S11 = Control_high_random,
  #S12 = Control_medium_random,
  #S13 = Control_all_random,
  #Annual_Bt_FIVE = annual_BT_top5,
  #Annual_OD_FIVE = annual_OD_top5,
  #Annual_OD_deg = annual_OD_top5_deg
  #SC14 = NEED TO ADD THESE
  #SC15 = 
  #SC16 = 
  #SC17 = 
  #SC18 = 
  #SC19 = 
)

```
```{r}
gwcc_results <- data.frame(
  Scenario = character(),
  Farms_removed = numeric(),
  GWCC_size = numeric(),
  GWCC_percentage_remaining = numeric(),
  Number_components = numeric()
)

#total_farms_nw = vcount(annual_network)

for (s in names(removal_scenarios)) {
  
  removed_network <- delete_vertices(
    annual_network,
    removal_scenarios[[s]]
  )
  
  comp <- components(removed_network, mode = "weak")
  
  farms_remaining <- vcount(removed_network)
  
  gwcc_size <-max(comp$csize)
  
  gwcc_percent <- (gwcc_size / farms_remaining) * 100
  
  gwcc_results <- rbind(
    gwcc_results,
    data.frame(
      Scenario = s,
      Farms_removed = length(removal_scenarios[[s]]),
      GWCC_size = gwcc_size,
      GWCC_percentage_remaining = gwcc_percent,
      Number_components = comp$no
    )
  )
}
```
```{r}
# calculate baseline network components
baseline_comp <- components(annual_network, mode = "weak")

#baseline_gwcc <- max(baseline_comp$csize)

#baseline_percent <- (baseline_gwcc / vcount(annual_network)) * 100

#baseline_components <- baseline_comp$no

# add baseline GWCC
gwcc_results <- rbind(
  data.frame(
    Scenario = "S1",
    Farms_removed = 0,
    GWCC_size = max(baseline_comp$csize),
    GWCC_percentage_remaining = (max(baseline_comp$csize) / vcount(annual_network))*100,
    Number_components = baseline_comp$no
  ),
  
  gwcc_results
)
```
```{r}
gwcc_results
```
```{r}
# REMOVE RESULTS
rm(gwcc_results)
```
```{r}
sum(annual_OD_top5 %in% V(annual_network)$name)
```
```{r}
length(annual_OD_top5)
length(intersect(annual_OD_top5, V(annual_network)$name))
```

Sc7 resulted in the greatest fragmentation of the annual movement network, reducing the GWCC to 193 farms and increasing the number of weakly connected compnents to 62. This suggests the farms targeted in this scenario occupied structurally important positions that maintained connectivity between groups

AND

Removal of farms that appeared in the top 5% of centrality measures for at least one month produced the greatest disruption to the network connectivity, with the strongest effect observed for betweeness-based removal, Persistent high-centrality farms identified over longer time period did not substantially fragment the annual network.

AND control result - supports interpretation that fragmentation is not simply because "many farms were removed"; it is because the removed farms occupied structurally important positions

AND - most interventions did not fragment the network. Fragmentation occurred only when all persistent betweenness/outdegree (and if applicable, combined) farms were removed. Random removal and annual top 5% removal did not produce comparable fragmentation.

So network fragmentation was limited under persistent-node removal scenarios, with the exception of removal of farms identified as high betweeness nodes at the lowest persistence threshold. Suggests that short-term emergence of structurally important bridging farms may have greater influence on annual network connectivity than farms consistently ranked as central over long periods

COMPARE TOP BETWEENNESS/OD WITH TEMPORAL AND ANNUAL STATIC NETWORKS
```{r}
# annual betweenness
annual_BT <- betweenness(
  annual_network,
  directed = TRUE,
  weights = E(annual_network)$weight
)

# top 5% annual farms
annual_BT_top5 <- names(
  sort(annual_BT, decreasing = TRUE)
)[1:ceiling(length(annual_BT)*0.05)]
```
```{r}
length(annual_BT_top5)
annual_BT_top5
```
```{r}
length(BT_all_candidates)
BT_all_candidates
```
```{r}
overlap_BT <- length(intersect(BT_all_candidates, annual_BT_top5)) / length(BT_all_candidates) * 100
overlap_BT
```
```{r}
annual_OD <- strength( # reflect cattle movement volume
  annual_network,
  mode = "out",
  weights = E(annual_network)$weight
)
```
```{r}
annual_OD_top5 <- names(
  sort(annual_OD, decreasing = TRUE)
)[1:ceiling(length(annual_OD) * 0.05)]
```
```{r}
length(annual_OD_top5)
annual_OD_top5
```
```{r}
annual_OD_deg <- degree( # if unweighted degree changes things
  annual_network,
  mode = "out"
)
```
```{r}
annual_OD_top5_deg <- names(
  sort(annual_OD_deg, decreasing = TRUE)
)[1:ceiling(length(annual_OD_deg)*0.05)]
```

All farms identified in the annual top 5% for betweenness centrality were also identified by the temporal persistence approach. Possible some of the observed reduction in network connectivity may reflect removal of farms that would also have been identified using a static annual network. The temporal approach identified an additional 31 farms exhibiting repeated monthly structural importance, although the independent contribution of these additional farms was not evaluated
```{r}
total_farms = 301
GWCC_table <- data.frame(
  Scenario = c(
    "(1)",
    "(2)",
    "(3)",
    "(4)",
    "(5)",
    "(6)",
    "(7)"
    #"(8)",
    #"(9)",
    #"(10)",
    #"(11)",
    #"(12)",
    #"(13)"
  ),
  
  Farms_removed = c(
    0,
    length(OD_high_candidates),
    length(OD_med_candidates),
    length(OD_all_candidates),
    length(BT_high_candidates),
    length(BT_med_candidates),
    length(BT_all_candidates)
    #length(CB_high_candidates),
    #length(CB_med_candidates),
    #length(CB_all_candidates),
    #4,
    #9,
    #40
  ),
  
  GWCC_size_farms = c(
    total_farms - 0,
    total_farms - length(OD_high_candidates),
    total_farms - length(OD_med_candidates),
    total_farms - length(OD_all_candidates),
    total_farms - length(BT_high_candidates),
    total_farms - length(BT_med_candidates),
    total_farms - length(BT_all_candidates)
    #total_farms - length(CB_high_candidates),
    #total_farms - length(CB_med_candidates),
    #total_farms - length(CB_all_candidates),
    #total_farms - 4,
    #total_farms - 9,
    #total_farms - 40
  ),
  
  GWCC_percentage_of_remaining_farms = c(# what are you calculating the percentage relative to?
    100,
    
    
  )
)

```
# 3. SOIR simulations - did the network changes alter disease spread

(a) probability of epidemic establishment - threshold = proportion of simulations --> sustained transmission
Does the intervention prevent epidemics from becoming established?


Report:
  number/proportion of simulations exceeding the threshold
  relatove reduction compared to baseline
  
Aim - capture the likelihood of sustained transmission

Define threshold - limited in literature - plot distribution of final outbreak sizes across all 2000 sims in baseline - see if natural cluster of larger outbreaks

```{r}
# plot distribution of final outbreak sizes across all 2000 sims in baseline

# name of column containing final outbreak size
names(summary_results_SC1)

```
```{r}
summary(summary_results_SC1$final_infected)
```
```{r}
# plot the distribution

ggplot(summary_results_SC1, aes(x = final_infected)) +
  geom_histogram(
    binwidth = 1,
    colour = "black",
    fill = "steelblue"
  ) +
  labs(
    x = "Final outbreak size (infected farms)",
    y = "Number of simulations",
    title = "Distribution of final outbreak sizes"
  ) +
  theme_classic()
```
```{r}
# zoom in on lower end
ggplot(summary_results_SC1, aes(x = final_infected)) +
  geom_histogram(
    binwidth = 1,
    colour = "black",
    fill = "steelblue"
  ) +
  coord_cartesian(xlim = c(1, 30)) +
  labs(
    x = "Final outbreak size (infected farms)",
    y = "Number of simulations"
  ) +
  theme_classic()
```
```{r}

outbreak_dist <- summary_results_SC1 %>%
  count(final_infected)
```
```{r}
ggplot(outbreak_dist, aes(x = final_infected, y = n)) +
  geom_area(fill = "steelblue", alpha = 0.6) +
  geom_line(colour = "steelblue4", linewidth = 1) +
  labs(
    x = "Final outbreak size (infected farms)",
    y = "Number of simulations"
  ) +
  theme_classic()
```
```{r}
# frequency table - shows exact numbers
table(summary_results_SC1$final_infected)
```
```{r}
as.data.frame(table(summary_results_SC1$final_infected))
```
```{r}
table(summary_results_SC1$final_infected)[1:20]
```
```{r}
summary_results_SC1 %>%
  count(final_infected)
```
No obvious break/peak to justify threshold - pick biologically representative threshold and then use sensitivity analysis across baseline/BT/OD to show preservation of ranking


```{r}
# threshold 1: final infected >5
mean(summary_results_SC1$final_infected >=5)

# 16% of infections results in epidemics involving 5+ farms
```
```{r}
# threshold 2: final infected >10
mean(summary_results_SC1$final_infected >= 10)

# 7% of infections results in epidemics involvinf 10+ farms - 140 outbreaks in 2000 sims
```
```{r}
# threshold 3: final infected >= 20
mean(summary_results_SC1$final_infected >=20)

# 3% of infections resulted in epidemics involving 20+ farms
```
Threshold >10 - reasonable middle threshold? capture substantial spread event while still leaving enough established outbreaks to compare final outbreak size and peak infectiousness? use sensitivity analysis ranking to check/support robustness

SENSITIVITY ANALYSIS OF THRESHOLD >= 10

Range of thresholds = 5, 10, 15, 20 infected farms
For each intervention scenario, calculate proportion of sims that exceed each threshold (out of the initial 2000 sims)
Check to see the ranking remains the same at every threshold - conclude that findings are robust to the choice of the threshold

```{r}
summary_results_SC1 <- summary_results$SC1
summary_results_SC2 <- summary_results$SC2
summary_results_SC3 <- summary_results$SC3
summary_results_SC4 <- summary_results$SC4
summary_results_SC5 <- summary_results$SC5
summary_results_SC6 <- summary_results$SC6
summary_results_SC7 <- summary_results$SC7
#summary_results_SC8 <- summary_results$SC8
#summary_results_SC9 <- summary_results$SC9
#summary_results_SC10 <- summary_results$SC10
#summary_results_SC11 <- summary_results$SC11
#summary_results_SC12 <- summary_results$SC12
#summary_results_SC13 <- summary_results$SC13
summary_results_SC14 <- summary_results$SC14
summary_results_SC15 <- summary_results$SC15
summary_results_SC16 <- summary_results$SC16
summary_results_SC17 <- summary_results$SC17
summary_results_SC18 <- summary_results$SC18
summary_results_SC19 <- summary_results$SC19
```

```{r}
# create list of final outbreak sizes
outbreak_sizes <- list(
  Baseline = summary_results_SC1$final_infected,
  Outdegree_high = summary_results_SC2$final_infected,
  Out_degree_med = summary_results_SC3$final_infected,
  Outdegree_all = summary_results_SC4$final_infected,
  Betweenness_high = summary_results_SC5$final_infected,
  Betweenness_med =summary_results_SC6$final_infected,
  Betweenness_all = summary_results_SC7$final_infected
)

```
```{r}
str(outbreak_sizes)
```
    YOU ARE HERE 06/08/26

```{r}
thresholds = c(5, 10, 15, 20)

threshold_results <- data.frame(
  scenario = character(),
  threshold = numeric(),
  proportion = numeric(),
  stringsAsFactors = FALSE
  )


for (s in names(outbreak_sizes)) {
  
  outbreaks <- outbreak_sizes[[s]]
  
  for (t in thresholds) {
    threshold_results <- rbind(
      threshold_results,
      data.frame(
        Scenario = s,
        Threshold = t,
        proportion = mean(outbreaks >= t) #* 100
      )
    )
  }
}
```
```{r}
threshold_results
```

```{r}
# plot sensitivity analysis

ggplot(threshold_results,
       aes(x = Threshold,
           y = proportion,
           colour = Scenario)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = "Threshold (infected farms)",
    y = "Probability of a major outbreak",
    colour = "Scenario"
  ) +
  theme_bw()
```


```{r}
thresholds2 <- c(1, 2, 3, 5, 7, 10, 15, 20)

threshold_results2 <- data.frame(
  scenario = character(),
  threshold = numeric(),
  proportion = numeric(),
  stringsAsFactors = FALSE
  )


for (s in names(outbreak_sizes)) {
  
  outbreaks <- outbreak_sizes[[s]]
  
  for (t in thresholds2) {
    threshold_results2 <- rbind(
      threshold_results2,
      data.frame(
        Scenario2 = s,
        Threshold2 = t,
        proportion2 = mean(outbreaks >= t) #* 100
      )
    )
  }
}
```

```{r}
threshold_results2
```

```{r}
# plot sensitivity analysis

ggplot(threshold_results2,
       aes(x = Threshold2,
           y = proportion2,
           colour = Scenario2)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = "Threshold (infected farms)",
    y = "Probability of a major outbreak",
    colour = "Scenario"
  ) +
  theme_bw()
```

Threshold of >= 10 infected farms operationally defines a major outbreak.
No natural separation between minor and major outbreaks is evident from the outbreak-size distribution (baseline), sensitivity analysis performed across thresholds of 1-20 infected farms was done to evaluate whether intervention rankings were dependent on this definition. CARE at one point betweenness and outdegree swap in the ranking, but the ranking of baseline vs intervention is consistent


Report:
  number/proportion of simulations exceeding the threshold per intervention
  relatove reduction compared to baseline
  
Aim - capture the likelihood of sustained transmission

```{r}
# list of scenarios to include - dropped combined - see prev
sustained_trans <- list(
  baseline = summary_results_SC1$final_infected, 
  OD_high = summary_results_SC2$final_infected,
  OD_med = summary_results_SC3$final_infected, 
  OD_all = summary_results_SC4$final_infected, 
  BT_high = summary_results_SC5$final_infected, 
  BT_med = summary_results_SC6$final_infected, 
  BT_all = summary_results_SC7$final_infected, 
  #CT_high = summary_results_SC11 
  #CT_med = summary_results_SC12 
  #CT_ALL = summary_results_SC13 
  CT_OD_HIGH = summary_results_SC14$final_infected,
  CT_OD_MED = summary_results_SC15$final_infected,
  CT_OD_ALL = summary_results_SC16$final_infected,
  CT_BT_HIGH = summary_results_SC17$final_infected,
  CT_BT_MED = summary_results_SC18$final_infected,
  CT_BT_ALL = summary_results_SC19$final_infected
)
```
```{r}
# counts how many simulations exceeded the threshold & calculates number sustained outbreaks; proportion sustained outbreaks; relative reduction vs baseline
threshold_final <- 10

threshold_counts_final <- data.frame(
  Scenario_final = names(sustained_trans),
  Exceeding = sapply(sustained_trans, function(x) sum(x >= threshold_final)),
  Total = sapply(sustained_trans, length)
)

threshold_counts_final$Proportion_final <- threshold_counts_final$Exceeding / threshold_counts_final$Total

# extract baseline probability
baseline_prob <- threshold_counts_final$Proportion_final[threshold_counts_final$Scenario_final == "baseline"]

# calculate relatve reduction compared with baseline
threshold_counts_final$Relative_reduction <- (
  (baseline_prob - threshold_counts_final$Proportion_final) / baseline_prob
)

threshold_counts_final
```

```{r}
# clear output
rm(threshold_counts_final)
```

Exceeding is the total number of sims that exceed the threshold (resulted in a sustained outbreak); total = total number of sims; proportion_final = proportion of sims that exceeded threshold (probability of a sustained outbreak)

Results show - OD removal slightly more effective than BT removal at high & med intervention levels, but both converge when all high-ranking farms are removed.

e.g. at the infected farm threshold >=10, 141/2000 baseliine simulations (7.1%) resulted in sustained outbreaks, compared with X/2000 (x%) following high out-degree removal - stochastic!

REPORT:
1. number/proportion of simulations exceeding the sustained outbreak threshold
2. relative reduction compared with baseline

Include control - helps to separate - effect of removing farms in general (control); effect of removing structurally important farms (OD/BT) - allows to state - targeted removal reduces outbreaks beyond what would be expected from removing a similar number of farms randomly

Figure 1 - probability of sustained outbreak (number of sims exceeding the sustained outbreak) - grouped by intervention type, with separate bars for high/med/all

  y-axis = prob of sustained outbreak (proportion_final)
  x-axis = intervention scenario
  bars grouped by - baseline/control/OD rmeoval/BT removal
  +/- separate by intervention intensity (high/med/all)

```{r, fig.width=12, fig.height=6}  
ggplot(threshold_counts_final,
       aes(x = Scenario_final,
           y = Proportion_final,
           fill = Scenario_final)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  labs(
     x = "Scenario",
     y = "Probability of sustained outbreak (>=10 threshold)"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
Figure 2 - relative reduction compared with baseline = effectiveness of each intervention relative to no intervention

  relative reduction = p(baseline) - p(scenario)/p(baseline)
  y-axis = relative reduction (%)
  x-axis = scenario
  same grouping as figure 1

```{r, fig.width=12, fig.height=6}
ggplot(threshold_counts_final,
       aes(x = Scenario_final,
           y = Relative_reduction,
           fill = Scenario_final)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  labs(
     x = "Scenario",
     y = "Relative outbreak reduction vs baseline"
  ) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
# AI - how can i group this plot by intensity windows, maintaining controls
# group by intensity and intervention type

# grepl() - searches for a pattern in scenario name e.g. "High"
# factor(..., levels = ...) - tells ggplot the order to display categories instead of using alphabetical order

# create two empty columns to store grouping information for plotting
threshold_counts_final$Intensity <- NA
threshold_counts_final$Method <- NA

  # assign intervention intensity

# search scenario names (grepl) & assign each one to an intervention intensity label
threshold_counts_final$Intensity[grepl("HIGH", threshold_counts_final$Scenario_final, ignore.case = TRUE)] <- "High" # looks for high in every scenario name & assigns method
threshold_counts_final$Intensity[grepl("MED", threshold_counts_final$Scenario_final, ignore.case = TRUE)] <- "Medium"
threshold_counts_final$Intensity[grepl("ALL", threshold_counts_final$Scenario_final, ignore.case = TRUE)] <- "All"

# assign the baseline scenario its own intensity label
threshold_counts_final$Intensity[threshold_counts_final$Scenario_final == "baseline"] <- "Baseline"

  # assign intervention method

# Label taregted interventions (grepl) according to the network metric used
threshold_counts_final$Method[grepl("^OD", threshold_counts_final$Scenario_final)] <- "Out-degree"
threshold_counts_final$Method[grepl("^BT", threshold_counts_final$Scenario_final)] <- "Betweenness"

# Keep the random matched controls separate from the targeted interventions
threshold_counts_final$Method[grepl("CT_OD", threshold_counts_final$Scenario_final)] <- "Control (OD matched)"
threshold_counts_final$Method[grepl("CT_BT", threshold_counts_final$Scenario_final)] <- "Control (BT matched)"

# label the baseline scenario
threshold_counts_final$Method[threshold_counts_final$Scenario_final == "baseline"] <- "Baseline"


  # Set plotting order
# convert intensity to an ordered factor so the panels appear in order wanted rather than alphabetically
threshold_counts_final$Intensity <- factor(
  threshold_counts_final$Intensity,
  levels = c("High", "Medium", "All", "Baseline")
)

# convert method to an ordered factor to bars appear in order wanted rather than alphabetically 

threshold_counts_final$Method <- factor(
  threshold_counts_final$Method,
  levels = c(
    "Control (OD matched)",
    "Out-degree",
    "Control (BT matched)",
    "Betweenness",
    "Baseline"
  )
)
```

```{r}
# Duplicate baseline into each intensity panel

baseline_row <- threshold_counts_final[
  threshold_counts_final$Scenario_final == "baseline",
]

baseline_repeated <- baseline_row[rep(1, 3), ]

baseline_repeated$Intensity <- c("High", "Medium", "All")
baseline_repeated$Method <- "Baseline"

# Remove original baseline and add repeated baseline
plot_data <- rbind(
  threshold_counts_final[
    threshold_counts_final$Scenario_final != "baseline",
  ],
  baseline_repeated
)

# Set order of panels
plot_data$Intensity <- factor(
  plot_data$Intensity,
  levels = c("High", "Medium", "All")
)

# Set order of bars
plot_data$Method <- factor(
  plot_data$Method,
  levels = c(
    "Baseline",
    "Control (OD matched)",
    "Out-degree",
    "Control (BT matched)",
    "Betweenness"
  )
)
```
```{r, fig.width = 12, fig.height = 4}
# plot probability of sustained outbreak by intensity

# order of windows
threshold_counts_final$Intensity <- factor(
  threshold_counts_final$Intensity,
  levels = c("Baseline", "High", "Medium", "All")
)
ggplot(
  plot_data,
  aes(x = Method,
      y = Proportion_final,
      fill = Method)
) +
  geom_col(position = "dodge") +
  facet_wrap(~Intensity) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = "Intervention strategy",
    y = "Probability of sustained outbreak (≥10 infected farms)"
  ) +
  theme_bw()
```


```{r}
# Duplicate baseline into each intensity panel for relative reduction plot

baseline_reduction <- threshold_counts_final[
  threshold_counts_final$Scenario_final == "baseline",
]

baseline_reduction <- baseline_reduction[rep(1, 3), ]

baseline_reduction$Intensity <- c("High", "Medium", "All")
baseline_reduction$Method <- "Baseline"
baseline_reduction$Relative_reduction <- 0


# Remove original baseline and add repeated baseline
plot_reduction <- rbind(
  threshold_counts_final[
    threshold_counts_final$Scenario_final != "baseline",
  ],
  baseline_reduction
)


# Set panel order
plot_reduction$Intensity <- factor(
  plot_reduction$Intensity,
  levels = c("High", "Medium", "All")
)


# Set bar order
plot_reduction$Method <- factor(
  plot_reduction$Method,
  levels = c(
    "Baseline",
    "Control (OD matched)",
    "Out-degree",
    "Control (BT matched)",
    "Betweenness"
  )
)
```
```{r, fig.width=12, fig.height=4}
# plot relative reduction by intensity

# order of windows
#threshold_counts_final$Intensity <- factor(
  #threshold_counts_final$Intensity,
  #levels = c("High", "Medium", "All")
#)

ggplot(
  plot_reduction,
  aes(x = Method,
      y = Relative_reduction,
      fill = Method)
) +
  geom_col(position = "dodge") +
  facet_wrap(~Intensity) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = "Intervention strategy",
    y = "Relative outbreak reduction vs baseline"
  ) +
  theme_bw()
```

NOTE baseline not visible bar because baseline relative outbreak reduction vs baseline = zero! - remove basleine bar?

# YOU ARE HERE 06/06/26

Note relatively small difference between BT control HIGH and BT HIGH
```{r}
table(plot_reduction$Intensity, plot_reduction$Method)
```
LOOK INTO lack of difference between BT control & intevention for 'HIGH'- true or coding anomaly?
!!!!For stats:!!!!
Factor 1: intervention type - baseline/OD/BT/CT - does intervention type matter? - yes/no
Factor 2: intervention intensity - high/med/all- does intensity matter?

OR simplify - keep all intensities (baseline/high/med/all) and compare OD & BT within each intensity

Stats looking to asnwer:

QUESTION: the probability of an infection of size >= 10 farms under different interventions/intensities vs baseline and matched controls

1. does probability of establishing an infection (yes/no) over threshold >=10 farms change with intervention type for intensities high/med

  outdegree targetiing:
  OD high vs baseline
  OD high vs matched control
  OD med vs baseline
  OD med vs matched control
  
  Betweenness targeting:
  BT high vs baseline
  BT high vs matched contorl
  BT med vs baseline
  BT med vs matched control

2. is the change in probability of outbreak change relative to baseline significat with intervention type for intensities high/med? - this does not need to be a statistical test - is an effect size- stat test will be on the underlying binary outcome - is the probability significantly different + relative reduction - how large is the effect.

STORY
1. probability bar plots
2. statistical tests - ?Fishers? pairwise targeted vs baseline/control; report p-values; report relative reduction (%) cf baseline = magnitude of effect

# YOU ARE HERE 8/08/26

Do an overall test then a pairwise test? to match other 2 questions - depends what want to answer - doing this version would allow to answer - is OD/BT more effective 

DATA - 

  plot and report probability e.g. 1600/2000 sims resulted in outbreak of >=10 farms = 0.8
  
  Fishers test - uses exact counts:
  Create 2x2 table - example: 
                Established   Not established
Baseline          1600            400
Out-degree         600           1400

# (b) final outbreak size --> how much dz occurs - use conditional on sustained transmission but show comparison with total/reference above question to show basis.

conditional on a sustained outbreak occuring, how severe is it?

if an epidemic becomes established (threshold >= 10), how large does it become/how many farms are infected?

the next result i want to generate is final outbreak size - if an epidemic becomes established (>=10), how large does it become/how many farms are infected? compare interventions to basline/control

Restrict analysis to successful epidemics (meet the establishment threshold >=10)

Report - Median, IQR, % reduction relative to baseline

Avoid the problem where median is always 1 because most introducitons die out

```{r}
# filter simulations above threshold - keep sims where finout outbreak size >= 10

threshold_final <- 10

established_outbreaks <- lapply(
  sustained_trans,
  function(x) x[x >= threshold_final]
)
```
```{r}
#length(established_outbreaks)
#established_outbreaks
```


```{r}
# create a dataframe for plotting

plot_sizes <- stack(established_outbreaks)
colnames(plot_sizes) <- c("Final_size", "Scenario")

# final_size= final number of infected farms
# scenario = intervention scenario
```
```{r}
# add intervention method & intensity
plot_sizes <- merge(
  plot_sizes,
  threshold_counts_final[, c("Scenario_final", "Method", "Intensity")],
  by.x = "Scenario",
  by.y = "Scenario_final"
)
plot_sizes_stats <- plot_sizes # save a copy before duplication for stat analysis without baseline x3
```
```{r}
# duplicate baseline into every intensity panel

# extract baseline
baseline_sizes <- subset(plot_sizes, Scenario == "baseline")
```
```{r}
# duplicate baseline
baseline_sizes <- do.call(
  rbind,
  lapply(c("High", "Medium", "All"), function(i) {
    x <- baseline_sizes
    x$Intensity <- i
    x
  })
)

baseline_sizes$Method <- "Baseline"
```
```{r}
table(baseline_sizes$Method, baseline_sizes$Intensity)
```
```{r}
# remove original baseline & add duplicated baseline
plot_sizes <- rbind(
  subset(plot_sizes, Scenario != "baseline"),
  baseline_sizes
)
```
```{r}
table(plot_sizes$Method, plot_sizes$Intensity)
```
```{r}
# set plotting order
plot_sizes$Intensity <- factor(
  plot_sizes$Intensity,
  levels = c("High", "Medium", "All")
)

plot_sizes$Method <- factor(
  plot_sizes$Method,
  levels = c(
    "Baseline",
    "Control (OD matched)",
    "Out-degree",
    "Control (BT matched)",
    "Betweenness"
  )
)
```
```{r}
# create boxplots

ggplot(
  plot_sizes,
  aes(
    x = Method,
    y = Final_size,
    fill = Method
  )
) +
  geom_boxplot() +
  facet_wrap(~Intensity) +
  labs(
    x = "Intervention strategy",
    y = "Final outbreak size (established outbreaks only)"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    )
  )
```
```{r}
# check the numebr of obs - note low number for OD/BT ALL - not suitable for stat testing, very small sample size?
table(
  plot_sizes$Method,
  plot_sizes$Intensity
)
```

Descriptive summary:
```{r}
# summarise outbreak sizes

outbreak_summary <- data.frame(
  Scenario = names(established_outbreaks),
  N = sapply(established_outbreaks, length), # how many sims exceeded threshold
  Mean = sapply(established_outbreaks, mean), # average final outbreak size
  Median = sapply(established_outbreaks, median), # median final outbreak size
  SD = sapply(established_outbreaks, sd), # variability/standard deviation
  Min = sapply(established_outbreaks, min), # range
  Max = sapply(established_outbreaks, max) # range
)

outbreak_summary
```
Statistical Analysis - need individual outbreak sizes - need distribution 
?given that an outbreak exceeded 10 farms, does the intervention reduce final outbreak size?

```{r}
# data source
plot_sizes_stats
```
```{r}
# outcome
#plot_sizes$Final_size
```

Use Kruskal-Wallis if skewed (overall comparison), then pairwise comparisons - Dunns? if significant - CARE not to include the duplicated baseline rows in statistical model - baseline should appear only once - use plot_sizes BEFORE baseline duplcation for stat analysis

Visualise distributions - show that final outbreak sizes are not normally distributed - ?right skewed

```{r}
# histogram of outbreak sizes
ggplot(plot_sizes_stats,
       aes(x = Final_size)) +
  geom_histogram(
    bins = 30,
    fill = "steelblue",
    colour = "black"
  ) +
  facet_wrap(~Method) +
  labs(
    x = "Final outbreak size (established outbreaks only)",
    y = "Number of simulations"
  ) +
  theme_bw()
```
```{r}
# density plots
ggplot(plot_sizes_stats,
       aes(x = Final_size,
           colour = Method,
           fill = Method)) +
  geom_density(alpha = 0.2) +
  facet_wrap(~Method) +
  labs(
    x = "Final outbreak size (established outbreaks only)",
    y = "Density"
  ) +
  theme_bw()
```
```{r}
# density plots
ggplot(peak_infectious_plot,
       aes(x = peak_infectious,
           colour = Method,
           fill = Method)) +
  geom_density(alpha = 0.2) +
  facet_wrap(~Intensity) +
  labs(
    x = "Peak outbreak size (established outbreaks only)",
    y = "Density"
  ) +
  theme_bw()
```
```{r}
ggplot(plot_sizes_stats,
       aes(x = Method,
           y = Final_size,
           fill = Method)) +
  geom_boxplot() +
  theme_bw()
```
Plots show final outbreak distributions were right skewed and contained extreme values - not normally distributed. Therefore, differences in outbreak size among intervention strategies will be assessed using a non-parametric test - e.g. Kruskal wallis +/- individual pairwise tests.

Null hypothesis - the distributions of final outbreak size are the same across all 5 methods

THIS IS A TEST RUN - HAS LUMPED ALL INTENSITIES INTO ONE
need to repeat analysis separately for each intensity level/include in statistical model

```{r}
# kruskal-wallis test (overall comparison)
kruskal.test(
  Final_size ~ Method,
  data = plot_sizes_stats
)
```
Reject null hypothesis - p <0.05 - shows that at least one group differs but not which interventions differ

DUNNS TEST - bonferroni
```{r}
#install.packages("FSA")
```
```{r}
library(FSA)

dunnTest(
  Final_size ~ Method,
  data = plot_sizes_stats,
 method = "bonferroni"
)
```
```{r}
#library(rstatix)

#plot_sizes_stats %>%
  #dunn_test(
    #Final_size ~ Method,
    #p.adjust.method = "BH"
 # )
```

Repeat by group by intensity
```{r}
High <- kruskal.test(
  Final_size ~ Method,
  data = subset(plot_sizes_stats, Intensity == "High")
)

Med <- kruskal.test(
  Final_size ~ Method,
  data = subset(plot_sizes_stats, Intensity == "Medium")
)

all <- kruskal.test(
  Final_size ~ Method,
  data = subset(plot_sizes_stats, Intensity == "All")
)
```
```{r}
print("High")
High
print("Medium")
Med
print("All")
all
```
Shows that within each intervention intensity, final outbreak size differs significantly among methods - pairwise comparison within each intensity

```{r}
# THIS DOESNT INCLUDE BASELINE
#high
dunnTest(
  Final_size ~ Method,
  data = subset(plot_sizes_stats, Intensity == "High"),
  method = "bonferroni"
)
```
```{r}
unique(plot_sizes_stats$Method)
```
YOU ARE HERE - UPDATE TO INCLUDE BASELINE IN STAT TEST

!!!!!STATS (b)!!!!

Stats looking to answer - how does targeted removal of structurally important farms (identified by OD & BT) affect the eventual scale of established outbreaks across different intervention intensities (high, med) - 'all' only visual as small sample size - SCALE

STORY IN REPORT
1. Density plot - shows distribution shape of peak infectious farms - demostrate skewed/non-normal --> use kruskal wallis
2. kruskal-wallis test - reports if overall significant difference between intervention methods - test statistic, df, p-value
3. dunns post hoc test (Bonferroni adjusted) - identifies which specific intervention pairs differ; P.adjusted
4. boxplot - 3 panels - shows direction & magnitude of change - describes - whether interventions reduced/increased peak infectious farms compared with baseline/contorls. discuss 'ALL' - insufficient sample size for stat test

```{r}
# 1. density plot
ggplot(plot_sizes,
       aes(x = Final_size,
           colour = Method,
           fill = Method)) +
  geom_density(alpha = 0.2) +
  facet_wrap(~Intensity) +
  labs(
    x = "Final outbreak size (established outbreaks only)",
    y = "Density"
  ) +
  theme_bw()
```
```{r}
names(plot_sizes_stats)
```
!!!HIGH!!!
```{r}
# 2. kruskal-Wallis test

# make baseline have intensity so will be picked up by filter
plot_sizes_stats_high <- plot_sizes_stats

plot_sizes_stats_high$Intensity[
  plot_sizes_stats_high$Method == "Baseline"
] <- "High"
```

```{r}
High_FS_KW <- kruskal.test(
  Final_size ~ Method,# numeric column inside dataframe being compared
  data = subset(
    plot_sizes_stats_high,
    Intensity == "High"
  )
)

High_FS_KW
```
p<0.05 - reject null- at least one intervention method differs in peak infectious farms under High intensity

Use post-hoc test (Dunn's test with Bonferroni correction) to identify specific differences.

```{r}
library(dunn.test)

High_FS_Dunn <- dunn.test(
  x = subset(
    plot_sizes_stats_high,
    Intensity == "High"
  )$Final_size,
  g = subset(
    plot_sizes_stats_high,
    Intensity == "High"
  )$Method,
  method = "bonferroni"
)

High_FS_Dunn
```
Interpret P.adjusted - because Dunn's test performs multiple pairwise comparisons - bonferroni correction
  
```{r}
  # AI - want to view p.adjusted as easy to read table
High_FS_Dunn_table <- data.frame(
  Comparison = High_FS_Dunn$comparisons,
  P_adjusted = round(High_FS_Dunn$P.adjusted, 5)
)

High_FS_Dunn_table
```

Dunn’s post-hoc test shows that final outbreak size differed significantly between several methods. Out-degree removal produced significantly smaller outbreaks than its matched control and baseline, while betweenness removal significantly reduced outbreak size compared with baseline but not its matched control. Both matched controls also differed significantly from baseline, suggesting that farm removal itself contributed to reductions in outbreak size.

!!!MEDIUM!!!


```{r}
# 2. kruskal-Wallis test

# make baseline have intensity so will be picked up by filter
plot_sizes_stats_med <- plot_sizes_stats

plot_sizes_stats_med$Intensity[
  plot_sizes_stats_med$Method == "Baseline"
] <- "Medium"
```

```{r}
Med_FS_KW <- kruskal.test(
  Final_size ~ Method,# numeric column inside dataframe being compared
  data = subset(
    plot_sizes_stats_med,
    Intensity == "Medium"
  )
)

Med_FS_KW
```
p<0.05 - reject null- at least one intervention method differs in peak infectious farms under High intensity

Use post-hoc test (Dunn's test with Bonferroni correction) to identify specific differences.

```{r}
#library(dunn.test)

Med_FS_Dunn <- dunn.test(
  x = subset(
    plot_sizes_stats_med,
    Intensity == "Medium"
  )$Final_size,
  g = subset(
    plot_sizes_stats_med,
    Intensity == "Medium"
  )$Method,
  method = "bonferroni"
)

Med_FS_Dunn
```
```{r}
  # AI - want to view p.adjusted as easy to read table
Med_FS_Dunn_table <- data.frame(
  Comparison = Med_FS_Dunn$comparisons,
  P_adjusted = round(Med_FS_Dunn$P.adjusted, 7)
)

Med_FS_Dunn_table
```
1. de

# DEBUGS
```{r}
table(plot_sizes$Method, plot_sizes$Intensity)
```
```{r}
unique(plot_sizes$Scenario)
```
```{r}
baseline_sizes <- subset(plot_sizes, Scenario == "baseline")

nrow(baseline_sizes)
```
```{r}
summary(baseline_sizes$Final_size)
```

(c) peak infectious farms - how intense is outbreak IF THERE IS AN OUTBREAK - explain as above - mean/median peak without <1 sims

When an epidemic occurs (threshold >=10 farms), how intense is it?

Only analyse successful epidemics

Report - median peak infectious farms; IQR; % redution relative to baseline

Aim - measure the epidemic burden on the system; independent of whether an epidemic occured in the first place or not
Conditional on an epidemic becoming established (>=10 infected farms), what is the maximum number on infectious farms at any one time (peak infectious) - need sim output that stores infectious farms over time = epidemic curve

```{r}
names(daily_results$SC1)
# confirm contains the epidemic curves, relevant column = 'infectious'
```
```{r}
head(daily_results$SC1)
```
```{r}
# calculate peak infectious farms for every simulation = max number of infectious farms reached in each simulation
peak_infectious <- lapply(
  daily_results,
  function(x) {
    x %>%
      group_by(replicate) %>%
      summarise(
        peak_infectious = max(infectious)
      )
  }
)
```
```{r}
names(peak_infectious)
```

```{r}
# keep sims where threshold transmission occured
# create dataframe with replicate/peak_inefctious/scenario

peak_infectious_established <- lapply(
  names(peak_infectious),
  function(scenario) {
    
    peaks <- peak_infectious[[scenario]]
    
    established <- summary_results[[scenario]] %>%
      filter(final_infected >= threshold_final) %>%
      select(replicate)
    
    peaks %>%
      filter(replicate %in% established$replicate) %>%
      mutate(Scenario = scenario)
  }
)

peak_infectious_established <- bind_rows(peak_infectious_established)
```
```{r}
head(peak_infectious_established)
```
```{r}
nrow(peak_infectious_established)
```
```{r}
unique(peak_infectious_established$Scenario)
```
```{r}
unique(threshold_counts_final$Scenario_final)
```
```{r}
# create look up table to to map simulation names to scenario labels

scenario_lookup <- data.frame(
  Scenario = c(
    "SC1", "SC2", "SC3", "SC4", "SC5", "SC6", "SC7", "SC14", "SC15", "SC16", "SC17", "SC18", "SC19"
  ),
  Scenario_final = c(
    "baseline",  "OD_high", "OD_med", "OD_all", "BT_high", "BT_med", "BT_all", "CT_OD_HIGH", "CT_OD_MED", "CT_OD_ALL", "CT_BT_HIGH", "CT_BT_MED","CT_BT_ALL"
  )
)
```


```{r}
# add method & intensity labels
peak_infectious_established <- merge(
  peak_infectious_established,
  scenario_lookup,
  by = "Scenario"
)

peak_infectious_established <- merge(
  peak_infectious_established,
  threshold_counts_final[, c("Scenario_final", "Method", "Intensity")],
  by = "Scenario_final"
)

peak_infectious_stats <- peak_infectious_established
```
```{r}
#make separate copy for figures
peak_infectious_plot <- peak_infectious_established

nrow(peak_infectious_plot)
# lower because has dropped scenarios 8-13
```
```{r}
head(peak_infectious_plot)
```
```{r}
unique(peak_infectious_established$Scenario)
```
```{r}
head(peak_infectious_established)
```
```{r}
unique(peak_infectious_established$Scenario_final)
```
```{r}
table(
  peak_infectious_established$Method,
  peak_infectious_established$Intensity
)
```
```{r}
# duplicate baseline so appears in all 3 panels

# extract basline peak infecitous values
baseline_peak <- subset(
  peak_infectious_plot,
  Method == "Baseline"
)

# Duplicate baseline into each intensity panel
baseline_peak <- baseline_peak[
  rep(1:nrow(baseline_peak), 3),
]

# assign intensity labels
baseline_peak$Intensity <- rep(
  c("High", "Medium", "All"),
  each = nrow(baseline_peak) / 3
)

# keep baseline label
baseline_peak$Method <- "Baseline"
```
```{r}
# remove original baseline and add duplicated baseline back
peak_infectious_plot <- rbind(
  subset(peak_infectious_plot, Method != "Baseline"),
  baseline_peak
)
```
```{r}
# set plotting order
peak_infectious_plot$Intensity <- factor(
  peak_infectious_plot$Intensity,
  levels = c("High", "Medium", "All")
)

peak_infectious_plot$Method <- factor(
  peak_infectious_plot$Method,
  levels = c("Baseline", "Control (OD matched)",
             "Out-degree",
             "Control (BT matched)",
             "Betweenness"
             )
)
```

```{r}
unique(peak_infectious_plot$Method)
```
```{r}
table(peak_infectious_plot$Method, useNA = "always")
```
```{r}
# plot
ggplot(
  peak_infectious_plot,
  aes(
    x = Method,
    y = peak_infectious,
    fill = Method
  )
) +
  geom_boxplot() +
  facet_wrap(~Intensity) +
  labs(
    x = "Intervention strategy",
    y = "Peak infectious farms (established outbreaks only)"
  )+
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

box = interquartile range (25th - 75th percentile)
line inside box = median
whiskers = extend to most extreme values within 1.5x IQR of the box
dots = values beyond whiskers = outliers

Baseline boxplot - long upper whisker and black dots -> most established outbreaks have moderate peaks, but some produce much larger peaks (black dots are outliers)

E.G.
outdegree - smaller box with long upper whisker and no upper dots -> peak infectious farms are more tightly clustered, there are no sims producing unusually large epidemic peaks under that intervention (within the established outbreaks that occurred)

no whiskers - very little variation (whisker length essensitlaly zero); the upper/lower whisker overlaps with the box - happens when the largest non-outlier value is close to the upper quartile - whisker is hidden by edge of box/ small sample size - only a few established intervention/intensity combos

```{r}
# check number of observations - note low number at ALL for BT/OD - no whiskers because small sample size - include this alongside the figure to give context. CARE with stats - very small sample - just report 'ALL' intensity descriptively
table(
  peak_infectious_plot$Method,
  peak_infectious_plot$Intensity
)
```
```{r}
subset(
  peak_infectious_plot,
  Method == "Out-degree" & Intensity == "All"
)
```
```{r}
head(peak_infectious_stats)
```

STATS - peak infectious

Answers: at high and medium intensities, did targeted removal useing OD & BT result in a significant reduciton in peak infectious farms compared with baseline and matched random controls?

Addresses scale/severity/dynamics/intensity component of RQ2 - a reduction in peak infectious farms indicates that targeted removal reduces the epidemic burden...

Look at farms in 'high' and compare to those in top5% for annual network - same farms? was temporal useful. Essentially without checking if the random removal plays a part will never know because was significant cf baseline

Show densitiies non-parametric

```{r}
# density plots
ggplot(peak_infectious_plot,
       aes(x = peak_infectious,
           colour = Method,
           fill = Method)) +
  geom_density(alpha = 0.2) +
  facet_wrap(~Intensity) +
  labs(
    x = "Peak infectious size (established outbreaks only)",
    y = "Density"
  ) +
  theme_bw()
```
Skewed distribution and heavy right tail all intensities
Kruskall-Wallis test for intensities high & med for all methods (All - to few samples)
```{r}
table(
  peak_infectious_stats$Method,
  peak_infectious_stats$Intensity
)
```
```{r}
unique(peak_infectious_stats$Method)
```
!!!!!HIGH!!!
```{r}
# make baseline have intensity so will be picked up by filter
peak_infectious_stats_high <- peak_infectious_stats

peak_infectious_stats_high$Intensity[
  peak_infectious_stats_high$Method == "Baseline"
] <- "High"
```

```{r}
High_PI_KW <- kruskal.test(
  peak_infectious ~ Method,# numeric column inside dataframe being compared
  data = subset(
    peak_infectious_stats_high,
    Intensity == "High"
  )
)

High_PI_KW
```
p<0.05 - reject null- at least one intervention method differs in peak infectious farms under High intensity

Use post-hoc test (Dunn's test with Bonferroni correction) to identify specific differences.

```{r}
#library(dunn.test)

High_PI_Dunn <- dunn.test(
  x = subset(
    peak_infectious_stats_high,
    Intensity == "High"
  )$peak_infectious,
  g = subset(
    peak_infectious_stats_high,
    Intensity == "High"
  )$Method,
  method = "bonferroni"
)

High_PI_Dunn
```
Interpret P.adjusted - because Dunn's test performs multiple pairwise comparisons - bonferroni correction
  
```{r}
  # AI - want to view p.adjusted as easy to read table
High_PI_Dunn_table <- data.frame(
  Comparison = High_PI_Dunn$comparisons,
  P_adjusted = round(High_PI_Dunn$P.adjusted, 5)
)

High_PI_Dunn_table
```
Significant results, using P.adjusted & significance threshold of p<0.05:
  Control(OD matched) - out degree = 0.0094
  Out-degree - baseline = 0.000031
  Betweenness - baseline = 0.0320
  
!!!MEDIUM!!!
```{r}
# make baseline have intensity so will be picked up by filter
peak_infectious_stats_med <- peak_infectious_stats

peak_infectious_stats_med$Intensity[
  peak_infectious_stats_med$Method == "Baseline"
] <- "Medium"
```

```{r}
MED_PI_KW <- kruskal.test(
  peak_infectious ~ Method,# numeric column inside dataframe being compared
  data = subset(
    peak_infectious_stats_med,
    Intensity == "Medium"
  )
)

MED_PI_KW
```
```{r}
#library(dunn.test)

MED_PI_Dunn <- dunn.test(
  x = subset(
    peak_infectious_stats_med,
    Intensity == "Medium"
  )$peak_infectious,
  g = subset(
    peak_infectious_stats_med,
    Intensity == "Medium"
  )$Method,
  method = "bonferroni"
)

MED_PI_Dunn
```
```{r}
  # AI - want to view p.adjusted as easy to read table
Med_PI_Dunn_table <- data.frame(
  Comparison = MED_PI_Dunn$comparisons,
  P_adjusted = MED_PI_Dunn$P.adjusted
)

Med_PI_Dunn_table
```

PLAIN ENGLISH INTERPRETATION
across all interventions and inensities - only OD-HIGH provided significant improvement in peak infectious compared to both baseline and controls. 

Betweenness targeting significantly reduced peak infectious farms compared with baseline, but was not significantly different from its matched control. The fact that the matched control also did not differ significantly from baseline suggests that the apparent reduction may be related to targeted removal, but it does not prove it, because the direct comparison between betweenness and its matched contol was not significant!


Random removal controls were generated using a fixed random seed to ensure reproducibility. However, because sensitivity anayses across multiple random removal replicates were not performed, uncertainty associated with the choice of random control farms cannot be fully assessed.
```{r}
# confirms control matches are correct- this is all from 'high' intensity
table(
  subset(
    peak_infectious_stats_high,
    Intensity == "High"
  )$Scenario,
  subset(
    peak_infectious_stats_high,
    Intensity == "High"
  )$Method
)
```
  
Statistical significance is assessed at alpha = 0.05. For post-hoc pairwise ocmparisons following significant Kruskal-Wallis tests, Bonferroni-adjusted p-values were used to account for multiple comparisons. Report both the adjusted p-values and the direction of change (e.g. lower peak infectious farms under intervention).

Essentially - in intervention HIGH

STORY IN REPORT
1. Density plot - shows distribution shape of peak infectious farms - demostrate skewed/non-normal --> use kruskal wallis
2. kruskal-wallis test - reports if overall significant difference between intervention methods - test statistic, df, p-value
3. dunns post hoc test (Bonferroni adjusted) - identifies which specific intervention pairs differ; P.adjusted
4. boxplot - 3 panels - shows direction & magnitude of change - describes - whether interventions reduced/increased peak infectious farms compared with baseline/contorls. discuss 'ALL' - insufficient sample size for stat test

Kruskal-Wallis - https://numiqo.com/video/l86wEhUzkY4 

  hypothesis test used to check whether there is a difference between several independent groups
  
  Not analysis of variance because data is not normally distributed and the assumptions for the analysis of variance are not met - use Kruskal-Wallis test
  
  is the non-parametric counterpart to single factor analysis of variance
  
  Difference between the two tests: 
    Analysis of variance test - is there a difference between the two means?
    Kruskal-Wallis test - is there a difference in the rank totals - across all groups - assigns rank in ascending order, computes rank total per group, compares rank totals across all groups - data does not have to satisfy any distribution of form and also dont need it to be normally distributed
  
  ASSUMPTIONS  
  when to use - when have nominal or ordinal variable wth more than 2 expressions/values and a metric or ordinal variable
  
  only several independent random samples with at least ordinally scaled characteristics must be available; variables do not have to satisfy a distribution curve.
  
  Null hypothesis = the indepdent samples all have hte same central tendency and therefore come from the same population - i.e. there is no difference in the rank sum
  
  Alternative hypothesis - at least one of the indepednent samples does not have the same central tendency as the other samples, and therefore comes from a different population - i.e. at least one group differs in rank sum

# ANALYSIS
Look at anova that is non-parametric - compare BT/OD/baseline and will highlight is there is a significant difference between them - then there is another test that can help find the difference if there is one. NOT using high/med/all - incoporate this?

Suggestion from google:
  Report median & OQR for each scenario
  Use Kruskal-Wallis test to assess whether there is an     overall difference among scenarios
 If significant - perform pairwise Wilcoxon test (or Dunns test) comparing each intervention to the baseline, with multiple-testing correction
 
FIRST - check for distribution - normal/not? CARE to remember and explain if have removed <1 sims and what the distribution is