---
title: "RQ2 SOIR SIMS ZIP"
author: "ESINCLAIR"
date: "2026-08-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# import libraries
library(ggplot2)
library(dplyr)
library(tidyr)
library(pbapply)
```
(original notebook = TEST SIM 2 (initial gen model))

PROMPT summary (ChatGPT)
show me basic complete R code for an stochastic compartmental SEIR model that tracks bTB-like disease spread across nodes (farms). Each farm can send or receive several batches of cattle per day. I want to be able to seed an infection in a node and view the potential disease spread. I also want to be able to remove certain nodes identified as superspreaders and compare simulation outcomes. I want the model to work on a daily time scale, movements between farms will be the opportunities for disease transmission. The transmission should be stochastic, with a specified probability of transmission (beta) between an infected and susceptible farm. The disease should progress through the SEIR compartments over defined exposed and infectious periods. Possibly in the future I would like to compute the probability of transmission to be weighted by the number of cattle.

```{r}
# This is base model function from gpt prompt - a SEIR compartment stochastic model - farm level rather than animal level (issues with Movenet/DigiVet)

# Everything in this block is inside the run_seir function
# takes input:
  # movement data (movements) and a 
  # starting infected farm (seed_farm)
  # transmission probability (beta) 
  # latent_period - time between becoming infected and infectious - in compartment 'E' - incubating but CANNOT yet infect other farms through  movements - (?make latency move with animal - realistic infection developing further down contact chain?)
  # infectious period - time spent infectious and capable of transmitting infection before moving into R (removed) - model assumption based on time to detect infected farm. During these 80d - movements from the farm can transmit infection according to beta

run_seir <- function(
    movements,
    seed_farm,
    beta = 0.05, 
    latent_period = 60, 
    infectious_period = 180 
){


# Overall timeline if a susceptible farm becomes infected on day 100: 0-99d = susceptible, can't infect others; 100-159d = exposed, can't infect others, 160-339d = infectious, can infect others, 340d onwards = removed, cant infect others

# Build lookup table - pulls every unique farm ID from both the source and destination columns of movements  - combines, removes duplicates, sorts --> (farms) is a clean lookup table of all farms involved in the dataset
  
  farms <- sort(unique(c(
    movements$source,
    movements$destination
  )))
  
# creates a numeric index for the farms, labels each with a number corresponding to the farm ID
  n_farms <- length(farms)

  farm_index <- seq_len(n_farms)
  names(farm_index) <- farms

 
# Convert movements to integer indices #

  movements$source_id <- farm_index[movements$source]
  movements$dest_id   <- farm_index[movements$destination]

  
# Disease state definitions:
  
  # 0 = S = susceptible
  # 1 = E = exposed
  # 2 = I = infectious
  # 3 = R = removed

# set up vectors to track each farms disease status and how long it has been in its current state
  
  state <- integer(n_farms) # creates a vector of zeros - meaning all farms start susceptible

  days <- integer(n_farms) # creates a parallel vector to count how many days each farm has spent in its current state

 
# Seed infection - infect chosen seed farm by setting its state to 2(I)

  state[farm_index[seed_farm]] <- 2


# Storage - prepares storage for the simulation results

  max_day <- max(movements$day) # finds last day in the movement dataset

# History - creates a data frame with one row per day, and columns for the number of farms in each SEIR state, all columns are pre-allocated as integer vectors of length max_day for filling during the simulation
  
  history <- data.frame( 
    day = integer(max_day), 
    S = integer(max_day),
    E = integer(max_day),
    I = integer(max_day),
    R = integer(max_day)
  )
  


# Simulation - advances farms through the E -> I -> R stages as each day of the simulation passes

# the 'for' loop runs one iteration per day

  for(day in seq_len(max_day)){

 
# Progress exposed farms:
 
# the day counters increase, and farms that have spent more than latent_period days in E become infectious (state == 2) and their day counters reset
    
    exposed <- which(state == 1)

    if(length(exposed) > 0){

      days[exposed] <- days[exposed] + 1

      become_I <- exposed[
        days[exposed] >= latent_period
      ]

      state[become_I] <- 2
      days[become_I] <- 0

    }


# Progress infectious farms

    infectious <- which(state == 2)

# Their day counters increase, any farms that have spent more than infectious_period i days in I become removed (state == 3) and their day counters reset    
    
    if(length(infectious) > 0){

      days[infectious] <- days[infectious] + 1

      become_R <- infectious[
        days[infectious] >= infectious_period
      ]

      state[become_R] <- 3
      days[become_R] <- 0

    }


# Today's movements - filters the 'movements' data frame to keep just the rows where 'day' matches the current loop day, giving the set of movements to process for that day.

    todays <- movements[movements$day == day, ]


# Replay movement events - processes each movement and applies infection-spread rules.
# loop over all movements happening today, for each movement, get the source and destination farm indices; skip the movement if the source is not infectious (state != 2); skip the movement if the destination is not susceptible (state != 0). Only movements from the infectious -> susceptible farms proceed to the transmission step that follows
    

    if(nrow(todays) > 0){

      for(i in seq_len(nrow(todays))){

        source <- todays$source_id[i]
        dest   <- todays$dest_id[i]

        # source not infectious
        if(state[source] != 2)
          next

        # destination already infected - i.e. not susceptible
        if(state[dest] != 0)
          next


# Transmission - applies transmission step - an infectious source may infect a susceptible destination. With probability 'beta', the destination farm becomes exposed (state = 1) and is day counter resets to 0 
# https://www.rdocumentation.org/packages/stats/versions/3.5.2/topics/Uniform - runif() - generates random number between 0-1, if number is < beta - infection progresses = Bernoulli
        
        if(runif(1) < beta){

          state[dest] <- 1
          days[dest] <- 0

        }

      }

    }

    
# Save epidemic - logs the SEIR count for the current simulation day - store the day number; count how many farms are S,E,I,R and write those values into the corresponding row of history
    
    history$day[day] <- day
    history$S[day] <- sum(state == 0)
    history$E[day] <- sum(state == 1)
    history$I[day] <- sum(state == 2)
    history$R[day] <- sum(state == 3)

  }


# Return - returns the simulation output as a structured list
  
  list(

    history = history, # stores the full day-by-day SEIR counts

    final_states = data.frame( # data frame listing each farm and its final SEIR state, converts numeric codes (0-3) into labelled factors S,E,I,R

      farm = farms,

      state = factor(
        state,
        levels = 0:3,
        labels = c("S","E","I","R")
      )

    )

  )

}
```

# Provide movement data

```{r}
# read in csv
movements_raw <- read.csv(
    "RQ1 NO EP.csv",
    stringsAsFactors = FALSE
)
# head(movements_raw)
```

```{r}
# check the structure
str(movements_raw)
```
```{r}
# create simulation movement  table
movements <- movements_raw %>%
    select(
        source = org_id,
        destination = dest_id,
        day,
        total_cattle
    )
# head(movements)
```


```{r}
# convert  farm IDs to chr rather than integer to stop e.g. 1 being  treated differently from 001 in future datasets
movements <- movements %>%
    mutate(
        source = as.character(source),
        destination = as.character(destination)
    )
```

```{r}
# check structure
str(movements)
```
```{r}
# number of movement events
nrow(movements)
```
# Run simulation
```{r}
# N.B.this is only running one replicate of the simulation- being stochastic need to run 100-1000 replicates because each will be slightly different
result <- run_seir(
    movements,
    seed_farm = "285" # choose farm to seed infection - needs updated to allow several farms
)

head(result$history)
```
```{r}
# plot the curve - using ggplot

result$history %>%
    pivot_longer(
        cols = c(S,E,I,R),
        names_to = "state",
        values_to = "farms"
    ) %>%
    ggplot(
        aes(
            x = day,
            y = farms,
            colour = state
        )
    ) +
    geom_line(size = 1) +
    theme_minimal()
```
# Update to SOIR model

from 'SOIR MODEL'

State encoding            Movement transmission?
0 = S (Suscpetible farm)          No
1 = O (Latent infection)          No (period of ?seroconversion, farm has become infected but is not yet infectious)
2 = I (Infectious/shedding farm)  Yes
3 = R (removed)                   No

Updated to SOIR because in bTB, infection does not immediately = infectiousness (https://pmc.ncbi.nlm.nih.gov/articles/PMC3475695/#s4)

```{r}
# model is unweighted by total_cattle - movement is either present or absent
run_soir <- function(
    movements,
    seed_farms, # can seed 1+ farms in one run - see update below
    beta = 0.05, # explore 0.001 - 0.05
    occult_period = 120, # explore 60-240
    infectious_period = 180 # explore 90 - 135
){
  
  
# Build lookup table - pulls every unique farm ID from both the source and destination columns of movements  - combines, removes duplicates, sorts --> (farms) is a clean lookup table of all farms involved in the dataset
  
  farms <- sort(unique(c(
    movements$source,
    movements$destination
  )))
  
# creates a numeric index for the farms, labels each with a number corresponding to the farm ID
  n_farms <- length(farms)

  farm_index <- seq_len(n_farms)
  names(farm_index) <- farms

 
# Convert movements to integer indices

  movements$source_id <- farm_index[movements$source]
  movements$dest_id   <- farm_index[movements$destination]

# Set up vectors to track each farms disease status and how long it has been in its current state
  
  state <- integer(n_farms) # creates a vector of zeros - meaning all farms start susceptible

  days <- integer(n_farms) # creates a parallel vector to count how many days each farm has spent in its current state

 
# Seed infection - infect chosen seed farm by setting its state to 1 - farms to start as occult - representative of dz introduction
# Reset their disease timers to 0 days, indicating they have just entered the infectious state
  
  state[farm_index[seed_farms]] <- 1  # find farm(s) listed in seed_farms, look up their position in the 'state' vector using 'farm_index'; change disease state S -> 0
  days[farm_index[seed_farms]] <- 0 # for the same seeded farm(s), set no. days spent in current dz state to 0 - farm has just become occult at the start of the simulation

# Storage - prepares storage for the simulation results

  max_day <- max(movements$day) # finds last day in the movement dataset

# history - creates a data frame with one row per day, and columns for the number of farms in each SOIR state, all columns are pre-allocated as integer vectors of length max_day for filling during the simulation
  
  history <- data.frame( 
    day = integer(max_day), 
    S = integer(max_day),
    O = integer(max_day),
    I = integer(max_day),
    R = integer(max_day)
  )


# Simulation - advances farms through the O -> I -> R stages as each day of the simulation passes
# the 'for' loop runs one iteration per day

  for(day in seq_len(max_day)){


# Progress exposed farms:
# The day counters increase, and farms that have spent more than occult_period days in O become infectious (state == 2) and their day counters reset
    
    occult <- which(state == 1)

    if(length(occult) > 0){

      days[occult] <- days[occult] + 1 # increments until reaches occult_period -> becomes infectious

      become_I <- occult[
        days[occult] >= occult_period
      ]

      state[become_I] <- 2
      days[become_I] <- 0

    }


# Progress infectious farms

    infectious <- which(state == 2)

# Their day counters increase, any farms that have spent more than infectious_period i days in I become removed (state == 3) and their day counters reset    
    
    if(length(infectious) > 0){

      days[infectious] <- days[infectious] + 1

      become_R <- infectious[
        days[infectious] >= infectious_period
      ]

      state[become_R] <- 3
      days[become_R] <- 0

    }

# Today's movements - filters the 'movements' data frame to keep just the rows where 'day' matches the current loop day, giving the set of movements to process for that day.

    todays <- movements[movements$day == day, ]

# Replay movement events - processes each movement and applies infection-spread rules.
# loop over all movements happening today, for each movement, get the source and destination farm indices; skip the movement if the source is not infectious (state != 2); skip the movement if the destination is not susceptible (state != 0). Only movements from the infectious -> susceptible farms proceed to the transmission step that follows.
    

    if(nrow(todays) > 0){

      for(i in seq_len(nrow(todays))){

        source <- todays$source_id[i]
        dest   <- todays$dest_id[i]

        # source not infectious
        if(state[source] != 2)
          next

        # destination already infected
        if(state[dest] != 0)
          next


# Transmission - applies transmission step - an infectious source may infect a susceptible destination. With probability 'beta', the destination farm becomes exposed (state = 1) and is day counter resets to 0 
# runif(1) generates a random number between 0 and 1, if that random number is less than beta, transmission occurs - changes destination farm's state from S -> O, resets the dz timer. So if beta = 0.05 = 5% chance that an infectious movement infects the destination farm. Bernoulli probability 0.05/0.95; https://www.rdocumentation.org/packages/stats/versions/3.5.2/topics/Uniform
       
        if(runif(1) < beta){

          state[dest] <- 1
          days[dest] <- 0

        }

      }

    }

# Save epidemic - logs the SOIR count for the current simulation day - store the day number; count how many farms are S,O,I,R and write those values into the corresponding row of history
    

    history$day[day] <- day
    history$S[day] <- sum(state == 0)
    history$O[day] <- sum(state == 1)
    history$I[day] <- sum(state == 2)
    history$R[day] <- sum(state == 3)

  }

# Return - returns the simulation output as a structured list

  list(

    history = history, # stores the full day-by-day SOIR counts

    final_states = data.frame( # data frame listing each farm and its final SOIR state, converts numeric codes (0-3) into labelled factors S,O,I,R

      farm = farms,

      state = factor(
        state,
        levels = 0:3,
        labels = c("S","O","I","R")
      )

    )

  )

}
```

Transmission events were simulated stochastically using Bernoulli trials. For each eligible movement, a uniform random number (runif(1)) was drawn and transmission occurred when the random draw was less than the calculated transmission probability (beta).

Provide movement data
```{r}
# read in csv
movements_raw <- read.csv(
    "RQ1 NO EP.csv",
    stringsAsFactors = FALSE
)
#head(movements_raw )
```
```{r}
# check the structure
str(movements_raw)
```
```{r}
# create simulation movement  table
movements <- movements_raw %>%
    select(
        source = org_id,
        destination = dest_id,
        day,
        total_cattle
    )
#head(movements)
```
```{r}
# convert  farm IDs to chr rather than integer to stop e.g. 1 being  treated differently from 001 in future datasets
movements <- movements %>%
    mutate(
        source = as.character(source),
        destination = as.character(destination)
    )
```
```{r}
str(movements)
```
```{r}
# number of movement events
nrow(movements)
```

```{r}
# N.B. this is only one replicate - is stochastic - need 100-1000 replicates
result <- run_soir(
    movements,
    seed_farms = c("2", "300", "193", "65")
)

#head(result$history)
```
```{r}
test <- run_soir(
  movements = movements,
  seed_farms = "95",
  beta = 0.05,
  occult_period = 120,
  infectious_period = 180
)
```
```{r}
# plot data - ggplot
test$history %>%
    tidyr::pivot_longer(
        cols = c(S,O,I,R),
        names_to = "state",
        values_to = "farms"
    ) %>%
    ggplot(
        aes(
            x = day,
            y = farms,
            colour = state
        )
    ) +
    geom_line(size = 1) +
    theme_minimal()
```
NEXT - PARAMETER SENSITIVITY ANALYSIS - MIGHT AS WELL DO AT SAME TIME AS WRITE UP LIKE FOR RQ1



Want to show: Do the farms identified as structural super spreaders consistently generate larger or faster outbreaks across a plausible range of bTB transmission parameters?
If the relative importance of the super spreader farms remains similar across parameter combinations, conclude that the findings are robust to uncertainty in the disease parameters.

Choose to seed infection in known super spreader nodes & random nodes: top 5% of farms by super spreader metric (RQ1 temporal network analysis). For each replicate, randomly choose one farm from this group as the seed. For the comparison, randomly choose one farm from the remainder of the network.

Because split into 12 months for top 5% ID - choose 1 representative month with typical movement pattern (month 7 - all layers and mid year movements); run sensitivity scenarios - alter B/occult period/infectious period; for each parameter combination run 100 sims in top 5% seeder farms in OD and BT (choose one at random) and run 100 sims in random seed farm (not top 5%); compare mean outbreak outcomes across all three - expect to see ranking.

```{r}
# for PERIOD 7
out_degree_top5 <- c("253", "129", "95", "255", "88", "119", "214", "213", "212", "202", "151", "299")
betweenness_top5 <- c("95", "296", "119", "41", "285", "300", "274", "178", "138", "235", "156")
random_farms <- c("2", "35", "180", "50", "260", "4", "230", "301", "92", "205", "167", "227")
```


```{r}
# Define sensitivity ranges
beta_values <- c(0.001, 0.005, 0.01)
occult_values <- c(60, 120, 240)
infectious_values <- c(90, 180, 365) # N.B. robust against 365 day (longer that data set)

# Define seed groups for one month
seed_groups <- list(
  out_degree = out_degree_top5,
  betweenness = betweenness_top5,
  random = random_farms
)

# Create simulation scenarios
# https://www.rdocumentation.org/packages/base/versions/3.3.0/topics/expand.grid
scenarios <- expand.grid( # creates 100 rows for each parameter combination, lapply runs every row
  beta = beta_values,
  occult_period = occult_values,
  infectious_period = infectious_values,
  seed_type = names(seed_groups),
  replicate = 1:100 # creates 100 identical copies of every unique parameter combination - 
)


# Function to run one simulation

# create a function called run_sensitivity, 'x' is one row from the 'scenarios' table (b, occult_p, infectious_p, seed_t, replicate no.)
run_sensitivity <- function(x){

  # select seed farm - looks at the seed type for this scenario, e.g. out_degree - finds corresponding list of farms, randomly selects one from the list to be the initial infected farm
  seed_farm <- sample(
    seed_groups[[as.character(x$seed_type)]],
    1
  )

  # run SOIR model - runs one complete SOIR sim
  result <- run_soir(
    movements = movements,
    seed_farms = seed_farm,# randomly selected seed farm
    beta = x$beta, # parameter values from scenario 'x'
    occult_period = x$occult_period,
    infectious_period = x$infectious_period
  )


  # Return only summary statistics
  data.frame(
    beta = x$beta,
    occult_period = x$occult_period,
    infectious_period = x$infectious_period,
    seed_type = x$seed_type,
    replicate = x$replicate,

# Final epidemic size - counts how many farms not susceptible at the end of the simulation = total number of farms infected during the outbreak
    final_infected = sum(
      result$final_states$state != "S"
    ),
# peak epidemic size - finds largest number of infectious farms on any single day
    peak_infectious = max(
      result$history$I
    )
  )
}
```

```{r}
# Run all simulations - loops through every simulation scenario, runs the run_sensitivity() function for each one, stores the results as a list
# add a progress bar - replace above block

results <- pblapply(
  seq_len(nrow(scenarios)), # creates list from 1 to no. simulation scenarios
  function(i){ # takes each scenario in turn
    run_sensitivity(scenarios[i, ])  # runs 1 simulation using that scenario's parameter values (lapply repeats this for every scenario andstores each result in a list called 'results - so even if the parameter values are identical in all 100 replicates, each run is different because is a stochastic model)
  }
)
```
```{r}
# Combine results - stacks all the one-row dataframes to create one combined data frame
# bind_rows = dplyr
results_RB <- bind_rows(results) 
#results
```
Mean outbreak size by seed type:
  Gives the overall effect across all tested beta/occult/infectious values & stochastic replicates
```{r}
OB_by_seedtype <- results_RB %>%
  group_by(seed_type) %>%
  summarise(
    mean_outbreak_size = mean(final_infected),
    sd_outbreak_size = sd(final_infected),
    mean_peak_infectious = mean(peak_infectious)
  )
OB_by_seedtype
```
Interpret - Across X (check how many run!) simulations, dz introductions into structurally important farms produced larger outbreaks than random introductions, irrespective of parameter changes. 

Robustness across transmission parameter space:
  Shows the proportion of parameter combinations where      structural farms outperform random farms

```{r}
# reshape so each parameter becomes a row and seed types become columns - robustness check

ranking_table3 <- results_RB %>%
  group_by(beta, occult_period, infectious_period, seed_type) %>%
  summarise(
    mean_infected = mean(final_infected),
    .groups = "drop"
  ) %>%
  pivot_wider(
    names_from = seed_type,
    values_from = mean_infected
  )
```

```{r}
ranking_table3 %>%
  summarise(
    out_degree_better = mean(out_degree > random),
    betweenness_better = mean(betweenness > random)
  )
```
The advantage of structurally important farms was maintained across the tested disease parameter space. Outdegree and betweenness farms generated larger outbreaks than random introductions in >90% of parameter combinations

Show effect of changing disease parameters - table grouped by beta because likely to be the biggest driver
```{r}
results_RB %>%
  group_by(beta, seed_type) %>%
  summarise(
    mean_outbreak_size = mean(final_infected),
    .groups = "drop"
  )
```
Across all tested combinations of transmission probability, occult period, and infectious period, structurally important farms consistenly generated larger outbreaks than random introductions. Although absolute outbreaksizes increased with higher transmission probabilitues and longer infectious periods, the relative advantage of structurally important farms (ranking) remained consistent, supporting the robustness of the selected parameter values for subsequent epidemic sims - the relationship of structrally important farms is robust to uncertainty in disease parameters.


Sensitivity analysis conclusion:
  The ranking of high-risk farms was robust to uncertainty in disease parameters: both structural metrics consistently outperformed random farm selection.
  
```{r}
# box plot across all 8100 simulations to visually show the separation between structural and random introductions
ggplot(results_RB, aes(
  x = seed_type,
  y = final_infected,
  fill = seed_type
)) +
  geom_boxplot() +
  labs(
    x = "Seed farm type",
    y = "Final number of infected farms",
    title = "Outbreak size following introduction into structurally important farms"
  ) +
  theme_classic() +
  theme(
    legend.position = "none"
  )
```

# Change to transmission weight per animal and repeat sensitivity analysis on beta
https://link.springer.com/article/10.1007/s41109-021-00354-x?utm_source=chatgpt.com

```{r}
# Weighted SOIR model: beta is a per-animal transmission probability, so larger cattle batches have a higher probability of transmitting infection.

run_soir_weighted <- function(
    movements,
    seed_farms, # can seed 1+ farms in one run
    beta = 0.005, # per-animal transmission probability (varied in sensitivity analysis)
    occult_period = 120, # fixed from un-weighted sensitivity analysis
    infectious_period = 180 # fixed from un-weighted sensitivity analysis
){
  
# Build look-up table - pulls every unique farm ID from both the source and destination columns of movements  - combines, removes duplicates, sorts --> (farms) is a clean look-up table of all farms involved in the dataset
  
  farms <- sort(unique(c(
    movements$source,
    movements$destination
  )))
  
# Creates a numeric index for the farms, labels each with a number corresponding to the farm ID
  n_farms <- length(farms)

  farm_index <- seq_len(n_farms)
  names(farm_index) <- farms

# Convert movements to integer indices #

  movements$source_id <- farm_index[movements$source]
  movements$dest_id   <- farm_index[movements$destination]

# Set up vectors to track each farms disease status and how long it has been in its current state
  
  state <- integer(n_farms) # creates a vector of zeros - meaning all farms start susceptible

  days <- integer(n_farms) # creates a parallel vector to count how many days each farm has spent in its current state

# Seed infection - infect chosen seed farm by setting its state to 1 - farms to start as occult - representative of dz introduction
# Reset their disease timers to 0 days, indicating they have just entered the occult state.

  state[farm_index[seed_farms]] <- 1 # state[farm_index[seed_farms]] <- 1; change seeded farms from susceptible (0) to occult (1)

  days[farm_index[seed_farms]] <- 0 # for the same seeded farm(s), set no. days spent in current dz state to 0 - farm has just become occult at the start of the simulation

# Storage - prepares storage for the simulation results

  max_day <- max(movements$day) # finds last day in the movement data set

# History - creates a data frame with one row per day, and columns for the number of farms in each SOIR state, all columns are pre-allocated as integer vectors of length max_day for filling during the simulation
  
  history <- data.frame( 
    day = integer(max_day), 
    S = integer(max_day),
    O = integer(max_day),
    I = integer(max_day),
    R = integer(max_day)
  )


# Simulation - advances farms through the O -> I -> R stages as each day of the simulation passes
# The 'for' loop runs one iteration per day

  for(day in seq_len(max_day)){


# Progress occult farms:
# The day counters increase, and farms that have spent more than/equal to occult_period days in O become infectious (state == 2) and their day counters reset
    
    occult <- which(state == 1)

    if(length(occult) > 0){

      days[occult] <- days[occult] + 1 # increments until reaches occult_period -> becomes infectious

      become_I <- occult[
        days[occult] >= occult_period
      ]

      state[become_I] <- 2
      days[become_I] <- 0

    }

# Progress infectious farms

    infectious <- which(state == 2)

# Their day counters increase, any farms that have spent more than infectious_period i days in I become removed (state == 3) and their day counters reset    
    
    if(length(infectious) > 0){

      days[infectious] <- days[infectious] + 1

      become_R <- infectious[
        days[infectious] >= infectious_period
      ]

      state[become_R] <- 3
      days[become_R] <- 0

    }

# Today's movements - filters the 'movements' data frame to keep just the rows where 'day' matches the current loop day, giving the set of movements to process for that day.

    todays <- movements[movements$day == day, ]

    
# Process each movement event. Transmission is only considered from infectious source farms to susceptible destination farms.


    
# Processes each movement event and applies infection-spread rules.
# loop over all movements happening today, for each movement, get the source and destination farm indices; skip the movement if the source is not infectious (state != 2); skip the movement if the destination is not susceptible (state != 0). Only movements from the infectious -> susceptible farms proceed to the transmission step that follows.

    if(nrow(todays) > 0){

      for(i in seq_len(nrow(todays))){

        source <- todays$source_id[i]
        dest   <- todays$dest_id[i]

        # source not infectious
        if(state[source] != 2)
          next

        # destination already infected
        if(state[dest] != 0)
          next

        
# Transmission block:
# Calculate the movement-level transmission probability from the batch size.
# Each animal has an independent probability (beta) of initiating infection

# Stochastic transmission event:
# runif(1) generates a random number between 0 and 1.
# If the random value is less than p_transmit, the movement event results in transmission,
# and the destination farm changes from susceptible (S) to occult (O).
# This represents a Bernoulli trial where transmission either occurs (1) or does not occur (0).
# p_transmit is the batch-level probability of infection, derived from the per-animal
# transmission probability (beta) and number of cattle moved.
#https://www.rdocumentation.org/packages/stats/versions/3.5.2/topics/Uniform, Bernoulli

        # beta = individual animal-level probability
        # p_transmit = whole batch movement level probability
       
        batch_size <- todays$total_cattle[i] # no. cattle moved in this movement event
        p_transmit <- 1 - (1 - beta)^batch_size # Probability that at least one animal in the batch initiates  PLOT PROBABILITY AGINST BATCH SIZE FOR WRITE UP
        if(runif(1) < p_transmit){ # raw a random number between 0 and 1; if it is less than p_transmit, infection occurs.

          state[dest] <- 1
          days[dest] <- 0

        }

      }

    }

# ? eventually include the probability that a moved animal is actually infectious - not neccesserially important for research question

# Save epidemic - logs the SOIR count for the current simulation day - store the day number; count how many farms are S,O,I,R and write those values into the corresponding row of history
    
    history$day[day] <- day
    history$S[day] <- sum(state == 0)
    history$O[day] <- sum(state == 1)
    history$I[day] <- sum(state == 2)
    history$R[day] <- sum(state == 3)

  }

# Return - returns the simulation output as a structured list

  list(

    history = history, # stores the full day-by-day SOIR counts

    final_states = data.frame( # data frame listing each farm and its final SOIR state, converts numeric codes (0-3) into labelled factors S,O,I,R

      farm = farms,

      state = factor(
        state,
        levels = 0:3,
        labels = c("S","O","I","R")
      )

    )

  )

}
```
# Provide movement data

```{r}
# read in csv
movements_raw <- read.csv(
    "RQ1 NO EP.csv",
    stringsAsFactors = FALSE
)

#head(movements_raw)
```
```{r}
# create simulation movement  table

movements <- movements_raw %>%
    select(
        source = org_id,
        destination = dest_id,
        day,
        total_cattle
    )
#head(movements)
```

```{r}
str(movements)
```
```{r}
# N.B. this is only running one replicate
result_weighted <- run_soir_weighted(
    movements,
    seed_farms = c("95")
)

head(result_weighted$history)
```
```{r}
# maximum infected farms
max(result_weighted$history$I)
```
```{r}
# plot result (uses tidyr - pivot_longer, ggplot), N.B. is only one simulation
result_weighted$history %>%
    pivot_longer(
        cols = c(S,O,I,R),
        names_to = "state",
        values_to = "farms"
    ) %>%
    ggplot(
        aes(
            x = day,
            y = farms,
            colour = state
        )
    ) +
    geom_line(size = 1) +
    theme_minimal()
```
```{r}
# final states
table(result_weighted$final_states$state)
```

Previous sensitivity analysis demonstrated that model outcomes were robust to variation in occult and infectious periods. Therefore, these parameters were fixed at intermediate values, and subsequent sensitivity analysis focused on uncertainty in the newly introduced batch-weighted transmission parameter. Avoid unnecessarily expanding parameter space

```{r}
# for period 7 - representative month - identical to last sensitivity analysis
out_degree_top5 <- c("253", "129", "95", "255", "88", "119", "214", "213", "212", "202", "151", "299")
betweenness_top5 <- c("95", "296", "119", "41", "285", "300", "274", "178", "138", "235", "156")
random_farms <- c("2", "35", "180", "50", "260", "4", "230", "301", "92", "205", "167", "227")
```

```{r}
# Define sensitivity ranges
beta_values <- c(0.001, 0.005, 0.01, 0.02, 0.05) # sensitivity range for effective per-animal movement transmission probability; selected to represent low transmission intensity expected for bTB
occult_values <- c(120) # same as non-weighted values
infectious_values <- c(180) # same as non-weighted values

# Define seed groups for one month
seed_groups <- list(
  out_degree = out_degree_top5,
  betweenness = betweenness_top5,
  random = random_farms
)

# Create simulation scenarios - run 100 iterations - stochastic process
scenarios_weighted <- expand.grid(
  beta = beta_values,
  occult_period = occult_values,
  infectious_period = infectious_values,
  seed_type = names(seed_groups),
  replicate = 1:100
)


# Function to run one simulation
run_sensitivity <- function(x){

  # select seed farm
  seed_farm <- sample(
    seed_groups[[as.character(x$seed_type)]],
    1
  )

  # run SOIR model
  result <- run_soir_weighted(
    movements = movements,
    seed_farms = seed_farm,
    beta = x$beta,
    occult_period = x$occult_period,
    infectious_period = x$infectious_period
  )


  # Return only summary statistics
  data.frame(
    beta = x$beta,
    occult_period = x$occult_period,
    infectious_period = x$infectious_period,
    seed_type = x$seed_type,
    replicate = x$replicate,

    final_infected = sum(
      result$final_states$state != "S"
    ),

    peak_infectious = max(
      result$history$I
    )
  )
}
```

```{r}
# lapply() runs run_sensitivity once for each row in scenarios weigthed (expand.grid already made 100 copies of each); do.call(rbind...) combines all individual simulation results into single results table. pblapply - adds progress bar
results2 <- do.call(
  rbind,
  pblapply(
    seq_len(nrow(scenarios_weighted)),
    function(i){
      run_sensitivity(scenarios_weighted[i, ])
    }
  )
)
```
```{r}
# Combine results - stacks all the one-row dataframes to create one combined data frame
# bind_rows = dplyr
results2 <- bind_rows(results2) 
#results
```

Mean outbreak size by seed type:
  Gives the overall effect across all tested beta/occult/infectious values & stochastic replicates
```{r}
OB_by_seedtype <- results2 %>%
  group_by(seed_type) %>%
  summarise(
    mean_outbreak_size = mean(final_infected),
    sd_outbreak_size = sd(final_infected),
    mean_peak_infectious = mean(peak_infectious)
  )
OB_by_seedtype
```
Interpret - Across X (check how many run!) simulations, dz introductions into structurally important farms produced larger outbreaks than random introductions, irrespective of parameter changes. 

Robustness across transmission parameter space:
  Shows the proportion of parameter combinations where structural farms outperform random farms

```{r}
# reshape so each parameter becomes a row and seed types become columns - robustness check

ranking_table3 <- results2 %>%
  group_by(beta, occult_period, infectious_period, seed_type) %>%
  summarise(
    mean_infected = mean(final_infected),
    .groups = "drop"
  ) %>%
  pivot_wider(
    names_from = seed_type,
    values_from = mean_infected
  )
```

```{r}
ranking_table3 %>%
  summarise(
    out_degree_better = mean(out_degree > random),
    betweenness_better = mean(betweenness > random)
  )
```
In 100% of tested β scenarios, the mean outbreak size from out-degree seeding was greater than random seeding.
In 100% of tested β scenarios, the mean outbreak size from betweenness seeding was greater than random seeding.



The advantage of structurally important farms was maintained across the tested disease parameter space. Outdegree and betweenness farms generated larger outbreaks than random introductions in >99% of parameter combinations

Show effect of changing disease parameters - table grouped by beta because likely to be the biggest driver
```{r}
results2 %>%
  group_by(beta, seed_type) %>%
  summarise(
    mean_outbreak_size = mean(final_infected),
    .groups = "drop"
  )
```

Across all tested combinations of transmission probability, (using already assessed values for occult period, and infectious period), structurally important farms consistently generated larger outbreaks than random introductions. Although absolute outbreak sizes increased with higher transmission probabilities and longer infectious periods, the relative advantage of structurally important farms (ranking) remained consistent, supporting the robustness of the selected parameter values for subsequent epidemic sims - the relationship of structurally important farms is robust to uncertainty in disease parameters.


Sensitivity analysis conclusion:
  The ranking of high-risk farms was robust to uncertainty in disease parameters: both structural metrics consistently    outperformed random farm selection.
  
```{r}
# box plot across all 8100 simulations to visually show the separation between structural and random introductions
ggplot(results2, aes(
  x = seed_type,
  y = final_infected,
  fill = seed_type
)) +
  geom_boxplot() +
  labs(
    x = "Seed farm type",
    y = "Final number of infected farms",
    title = "Outbreak size following introduction into structurally important farms"
  ) +
  theme_classic() +
  theme(
    legend.position = "none"
  )
```

# Choosing parameter:

Increasing β increases outbreak size
This is expected.
Higher transmission probability produces larger epidemics.
All β values maintain the expected ordering
Targeted high-risk farms produce larger outbreaks than random farms.
The model responds sensibly to changes in transmission intensity.
Very high β values create very large outbreaks
β = 0.05 gives mean outbreaks of ~73 farms for out-degree seeding.
Given bTB is relatively slow-spreading, this may represent a relatively aggressive transmission scenario.

Choose beta = 0.005

A sensitivity analysis was conducted across β values from 0.001 to 0.05. Increasing β resulted in progressively larger outbreaks, while the qualitative pattern of greater spread following seeding of structurally important farms remained consistent. A β value of 0.005 was selected for the main analysis as an intermediate transmission scenario, representing a low transmission probability consistent with the relatively slow spread of bTB

Final model = run_soir_weighted

```{r}
# Weighted SOIR model: beta is a per-animal transmission probability, so larger cattle batches have a higher probability of transmitting infection - see graph showing relationship.

run_soir_weighted_RQ2 <- function(
    movements,
    seed_farms, 
    beta = 0.005, 
    occult_period = 120,
    infectious_period = 180 
){
  
  farms <- sort(unique(c(
    movements$source,
    movements$destination
  )))

  n_farms <- length(farms)

  farm_index <- seq_len(n_farms)
  names(farm_index) <- farms
  
  movements$source <- as.character(movements$source)
  movements$destination <- as.character(movements$destination)



  movements$source_id <- farm_index[movements$source]
  movements$dest_id   <- farm_index[movements$destination]
  
  
  state <- integer(n_farms) 

  days <- integer(n_farms) 

  state[farm_index[seed_farms]] <- 1

  days[farm_index[seed_farms]] <- 0 

  max_day <- max(movements$day) 
  
  history <- data.frame( 
    day = integer(max_day), 
    S = integer(max_day),
    O = integer(max_day),
    I = integer(max_day),
    R = integer(max_day)
  )

  for(day in seq_len(max_day)){
    
    occult <- which(state == 1)

    if(length(occult) > 0){

      days[occult] <- days[occult] + 1 

      become_I <- occult[
        days[occult] >= occult_period
      ]

      state[become_I] <- 2
      days[become_I] <- 0

    }


    infectious <- which(state == 2)

    
    if(length(infectious) > 0){

      days[infectious] <- days[infectious] + 1

      become_R <- infectious[
        days[infectious] >= infectious_period
      ]

      state[become_R] <- 3
      days[become_R] <- 0

    }


    todays <- movements[movements$day == day, ]

    if(nrow(todays) > 0){

      for(i in seq_len(nrow(todays))){

        source <- todays$source_id[i]
        dest   <- todays$dest_id[i]

        if(state[source] != 2)
          next

        if(state[dest] != 0)
          next
       
        batch_size <- todays$total_cattle[i] 
        p_transmit <- 1 - (1 - beta)^batch_size 
        
        if(runif(1) < p_transmit){ 

          state[dest] <- 1
          days[dest] <- 0

        }

      }

    }

    
    history$day[day] <- day
    history$S[day] <- sum(state == 0)
    history$O[day] <- sum(state == 1)
    history$I[day] <- sum(state == 2)
    history$R[day] <- sum(state == 3)

  }

  list(

    history = history, 

    final_states = data.frame( 

      farm = farms,

      state = factor(
        state,
        levels = 0:3,
        labels = c("S","O","I","R")
      )

    )

  )

}
```

PLOT OF probability of transmisison by batch size?

# Visual of probability x batch size
https://ggplot2.tidyverse.org/articles/ggplot2.html
```{r}

beta <- 0.005

df <- data.frame(
  batch_size = 1:800
)

df$p_transmit <- 1 - (1 - beta)^df$batch_size


ggplot(df, aes(x = batch_size, y = p_transmit)) +
  geom_line(
    colour = "firebrick",
    linewidth = 1.2
  ) +
  labs(
    x = "Cattle moved per batch",
    y = "Probability of transmission",
    title = paste0(
      "Effect of batch size on transmission probability (β = ",
      beta,
      ")"
    )
  ) +
  theme_classic()
```

Replicates sensitivity analysis

