---
title: "RQ2 RESULTS ANALYSIS ZIP"
author: "ESINCLAIR"
date: "2026-08-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# libraries
library(igraph)
library(ggplot2)
library(dplyr)
library(tidyr)
library(pbapply)
```

CONTAINS GWCC AND THRESHOLD ESTABLISHMENT + ANALYSIS

Import results of simulations for anlysis (RQ2 SOIR SIMS)


```{r}
# Baseline
scenario_1_results <- readRDS("scenario_1_results.rds") # baseline

# Interventions
scenario_2_results <- readRDS("scenario_2_results.rds") # Outdegree (high) >= 8 months
scenario_3_results <- readRDS("scenario_3_results.rds") # Outdegree (med) >= 6 months
scenario_4_results <- readRDS("scenario_4_results.rds") # Outdegree any month
scenario_5_results <- readRDS("scenario_5_results_corrected.rds") # Betweenness (high) >= 6 months
scenario_6_results <- readRDS("scenario_6_results_corrected.rds") # Betweenness (med) >= 4 months
scenario_7_results <- readRDS("scenario_7_results_corrected.rds") # Betweenness any month
```

```{r}
# CHECK: baseline imported results
sum(sapply(
  scenario_1_results,
  function(x) x$summary$final_infected #>= 10
))
```

```{r}
# Matched controls - Scenario 8: Control OD HIGH x10 (5 random farms removed)
scenario_8a_results <- readRDS("scenario_8P_control_1_results.rds")
scenario_8b_results <- readRDS("scenario_8P_control_2_results.rds")
scenario_8c_results <- readRDS("scenario_8P_control_3_results.rds")
scenario_8d_results <- readRDS("scenario_8P_control_4_results.rds")
scenario_8e_results <- readRDS("scenario_8P_control_5_results.rds")
scenario_8f_results <- readRDS("scenario_8P_control_6_results.rds")
scenario_8g_results <- readRDS("scenario_8P_control_7_results.rds")
scenario_8h_results <- readRDS("scenario_8P_control_8_results.rds")
scenario_8i_results <- readRDS("scenario_8P_control_9_results.rds")
scenario_8j_results <- readRDS("scenario_8P_control_10_results.rds")
  
```  

```{r}
# CHECK: SC8 imported results
sum(sapply(
  list(
    scenario_8a_results,
    scenario_8b_results,
    scenario_8c_results,
    scenario_8d_results,
    scenario_8e_results,
    scenario_8f_results,
    scenario_8g_results,
    scenario_8h_results,
    scenario_8i_results,
    scenario_8j_results
  ),
  function(x) {
    sum(sapply(x, function(i) i$summary$final_infected >= 10))
  }
))
```

```{r}
# list of 2,000 simulation replicates & summary dataframe with each replicate
#class(scenario_8a_results)
#str(scenario_8a_results)
```
```{r}
# Matched controls - Scenario 9 Control OD Med x10, (random removal of 8 farms)
scenario_9a_results <- readRDS("scenario_9P_control_1_results.rds")
scenario_9b_results <- readRDS("scenario_9P_control_2_results.rds")
scenario_9c_results <- readRDS("scenario_9P_control_3_results.rds")
scenario_9d_results <- readRDS("scenario_9P_control_4_results.rds")
scenario_9e_results <- readRDS("scenario_9P_control_5_results.rds")
scenario_9f_results <- readRDS("scenario_9P_control_6_results.rds")
scenario_9g_results <- readRDS("scenario_9P_control_7_results.rds")
scenario_9h_results <- readRDS("scenario_9P_control_8_results.rds")
scenario_9i_results <- readRDS("scenario_9P_control_9_results.rds")
scenario_9j_results <- readRDS("scenario_9P_control_10_results.rds")
```
```{r}
# Matched controls - Scenario 10 Control OD all x10, (random removal of 43 farms)
scenario_10a_results <- readRDS("scenario_10P_control_1_results.rds")
scenario_10b_results <- readRDS("scenario_10P_control_2_results.rds")
scenario_10c_results <- readRDS("scenario_10P_control_3_results.rds")
scenario_10d_results <- readRDS("scenario_10P_control_4_results.rds")
scenario_10e_results <- readRDS("scenario_10P_control_5_results.rds")
scenario_10f_results <- readRDS("scenario_10P_control_6_results.rds")
scenario_10g_results <- readRDS("scenario_10P_control_7_results.rds")
scenario_10h_results <- readRDS("scenario_10P_control_8_results.rds")
scenario_10i_results <- readRDS("scenario_10P_control_9_results.rds")
scenario_10j_results <- readRDS("scenario_10P_control_10_results.rds")
```
```{r}
# Matched controls - Scenario 11 Control BT High x10, (random removal of 1 farms)
scenario_11a_results <- readRDS("scenario_11P_control_1_results.rds")
scenario_11b_results <- readRDS("scenario_11P_control_2_results.rds")
scenario_11c_results <- readRDS("scenario_11P_control_3_results.rds")
scenario_11d_results <- readRDS("scenario_11P_control_4_results.rds")
scenario_11e_results <- readRDS("scenario_11P_control_5_results.rds")
scenario_11f_results <- readRDS("scenario_11P_control_6_results.rds")
scenario_11g_results <- readRDS("scenario_11P_control_7_results.rds")
scenario_11h_results <- readRDS("scenario_11P_control_8_results.rds")
scenario_11i_results <- readRDS("scenario_11P_control_9_results.rds")
scenario_11j_results <- readRDS("scenario_11P_control_10_results.rds")
```
```{r}
# Matched controls - Scenario 12 Control BT Med x10, (random removal of 18 farms)
scenario_12a_results <- readRDS("scenario_12P_control_1_results.rds")
scenario_12b_results <- readRDS("scenario_12P_control_2_results.rds")
scenario_12c_results <- readRDS("scenario_12P_control_3_results.rds")
scenario_12d_results <- readRDS("scenario_12P_control_4_results.rds")
scenario_12e_results <- readRDS("scenario_12P_control_5_results.rds")
scenario_12f_results <- readRDS("scenario_12P_control_6_results.rds")
scenario_12g_results <- readRDS("scenario_12P_control_7_results.rds")
scenario_12h_results <- readRDS("scenario_12P_control_8_results.rds")
scenario_12i_results <- readRDS("scenario_12P_control_9_results.rds")
scenario_12j_results <- readRDS("scenario_12P_control_10_results.rds")
```
```{r}
# Matched controls - Scenario 13 Control BT All x10, (random removal of 35 farms)
scenario_13a_results <- readRDS("scenario_13P_control_1_results.rds")
scenario_13b_results <- readRDS("scenario_13P_control_2_results.rds")
scenario_13c_results <- readRDS("scenario_13P_control_3_results.rds")
scenario_13d_results <- readRDS("scenario_13P_control_4_results.rds")
scenario_13e_results <- readRDS("scenario_13P_control_5_results.rds")
scenario_13f_results <- readRDS("scenario_13P_control_6_results.rds")
scenario_13g_results <- readRDS("scenario_13P_control_7_results.rds")
scenario_13h_results <- readRDS("scenario_13P_control_8_results.rds")
scenario_13i_results <- readRDS("scenario_13P_control_9_results.rds")
scenario_13j_results <- readRDS("scenario_13P_control_10_results.rds")
```

```{r}
# Put all scenario results into one list
all_scenarios <- list(
  SC1 = scenario_1_results,
  SC2 = scenario_2_results,
  SC3 = scenario_3_results,
  SC4 = scenario_4_results,
  SC5 = scenario_5_results,
  SC6 = scenario_6_results,
  SC7 = scenario_7_results
)

# Create empty lists to store outputs
daily_results <- list()
summary_results <- list()

# Loop through scenarios
for (scenario in names(all_scenarios)) {
  
  daily_results[[scenario]] <- bind_rows(
    lapply(all_scenarios[[scenario]], function(x) x$daily)
  )
  
  summary_results[[scenario]] <- bind_rows(
    lapply(all_scenarios[[scenario]], function(x) x$summary)
  )
}

# call individual with - summary_results$SC3 etc
```
  
# Pooled control & between-control variation

pooled control - 20,000 stochastic outcomes - gives the matched control estimate (median/mean/IQR)
between-control variation - 10 separate 2000-run simulations - demonstrates that the pooled result is reasonably consitent across the random farms selected

Pooled results - used to estimate the overall distribution of outbreak outcomes under matched random removal

Between control variation - used to demonstrate that this pooled distribution isn't being triven disproportionately by one particular random-control network

```{r}
# SCENARIO 8
# pool the 20,000 outcomes

control_8_results_pooled <- list(
  scenario_8a_results,
  scenario_8b_results,
  scenario_8c_results,
  scenario_8d_results,
  scenario_8e_results,
  scenario_8f_results,
  scenario_8g_results,
  scenario_8h_results,
  scenario_8i_results,
  scenario_8j_results
)

control_8_finalsizes <- unlist(
  lapply(control_8_results_pooled, function(x) {
    sapply(x, function(i) i$summary$final_infected)
  })
)

control_8_peakinfectious <- unlist(
  lapply(control_8_results_pooled, function(x) {
    sapply(x, function(i) i$summary$peak_infectious)
  })
)
```
```{r}
# create summary results for analysis - contains the 20000 individual pooled simulation summaries - adds it to summary_results created above
# AI code - fixed nesting problem to keep column names
summary_results$SC8 <- bind_rows(
  lapply(control_8_results_pooled, function(x) {
    bind_rows(lapply(x, function(i) i$summary))
  })
)
```

```{r}
length(control_8_finalsizes)
```
```{r}
# pooled summary

control_8_summary_pooled <- data.frame(
  N = length(control_8_finalsizes),
  Mean = mean(control_8_finalsizes),
  Median = median(control_8_finalsizes),
  IQR = IQR(control_8_finalsizes),
  SD = sd(control_8_finalsizes),
  Min = min(control_8_finalsizes),
  Max = max(control_8_finalsizes)
)

control_8_summary_pooled
```
```{r}
# show isnt driven by one random control - calculate the summary repeatedly for each of the 10 controls

control_8_variance <- do.call(
  rbind,
  lapply(seq_along(control_8_results_pooled), function(i) {
    
    x <- sapply(
      control_8_results_pooled[[i]],
      function(j) j$summary$final_infected
    )
    
    data.frame(
      Control = i,
      N = length(x),
      Mean = mean(x),
      Median = median(x),
      IQR = IQR(x),
      SD = sd(x)
    )
  })
)

control_8_variance
```
```{r}
# calculate variability between the 10 controls - how consistent are the matched controls?

sd(control_8_variance$Mean)
```
```{r}
# range of control means

range(control_8_variance$Mean)
```
Across the 10 matched random controls, mean final outbreak size ranged from 2.47 to 2.96, with a SD of 0.16

```{r}
ggplot(control_8_variance, aes(x = Control, y = Mean)) +
  geom_point(size = 3) +
  geom_hline(
    yintercept = control_8_summary_pooled$Mean,
    linetype = "dashed"
  ) +
  theme_bw() +
  labs(
    title = "Scenario 8: variation in mean final outbreak size",
    x = "Random control simulation",
    y = "Mean final outbreak size"
  )



```
```{r}
# add boxplot to show distribution (AI)
boxplot_c8 <- ggplot(control_8_variance, aes(x = 1, y = Mean)) +
  geom_boxplot(width = 0.25, alpha = 0.2) +
  geom_jitter(width = 0.05, size = 3) +
  geom_hline(
    yintercept = control_8_summary_pooled$Mean,
    linetype = "dashed"
  ) +
  theme_bw() +
  labs(
    title = "Scenario 8: consistency across matched random controls",
    x = NULL,
    y = "Mean final outbreak size"
  ) +
  scale_x_continuous(
    breaks = 1,
    labels = "10 matched random controls"
  )

boxplot_c8

ggsave(
  filename = "boxplot_cx_var.png",
  plot = boxplot_c8,
  width = 6,
  height = 4,
  units = "in",
  dpi = 300
)
```
State - across the ten random control configurations, the median final outbreak size was 1 (IQR= 0) in all configurations, while mean oubtreak sizes were also similar, indicating limited variation between random control configurations - APPENDIX

```{r}
# SCENARIO 9
# pool the 20,000 outcomes

control_9_results_pooled <- list(
  scenario_9a_results,
  scenario_9b_results,
  scenario_9c_results,
  scenario_9d_results,
  scenario_9e_results,
  scenario_9f_results,
  scenario_9g_results,
  scenario_9h_results,
  scenario_9i_results,
  scenario_9j_results
)

control_9_finalsizes <- unlist(
  lapply(control_9_results_pooled, function(x) {
    sapply(x, function(i) i$summary$final_infected)
  })
)

control_9_peakinfectious <- unlist(
  lapply(control_9_results_pooled, function(x) {
    sapply(x, function(i) i$summary$peak_infectious)
  })
)
```
```{r}
# create summary results for analysis - contains the 20000 individual pooled simulation summaries
#summary_results$SC9 <- bind_rows(
 # lapply(control_9_results_pooled, function(x) x$summary)
#)
```
```{r}
# create summary results for analysis - contains the 20000 individual pooled simulation summaries
# AI code - fixed nesting problem to keep column names
summary_results$SC9 <- bind_rows(
  lapply(control_9_results_pooled, function(x) {
    bind_rows(lapply(x, function(i) i$summary))
  })
)
```

```{r}
length(control_9_finalsizes)
```
```{r}
# pooled summary

control_9_summary_pooled <- data.frame(
  N = length(control_9_finalsizes),
  Mean = mean(control_9_finalsizes),
  Median = median(control_9_finalsizes),
  IQR = IQR(control_9_finalsizes),
  SD = sd(control_9_finalsizes),
  Min = min(control_9_finalsizes),
  Max = max(control_9_finalsizes)
)

control_9_summary_pooled
```
```{r}
# show isnt driven by one random control - calculate the summary repeatedly for each of the 10 controls

control_9_variance <- do.call(
  rbind,
  lapply(seq_along(control_9_results_pooled), function(i) {
    
    x <- sapply(
      control_9_results_pooled[[i]],
      function(j) j$summary$final_infected
    )
    
    data.frame(
      Control = i,
      N = length(x),
      Mean = mean(x),
      Median = median(x),
      IQR = IQR(x),
      SD = sd(x)
    )
  })
)

control_9_variance
```
```{r}
# calculate variability between the 10 controls - how consistent are the matched controls?

sd(control_9_variance$Mean)
```
```{r}
# range of control means

range(control_9_variance$Mean)
```
```{r}
ggplot(control_9_variance, aes(x = Control, y = Mean)) +
  geom_point(size = 3) +
  geom_hline(
    yintercept = control_9_summary_pooled$Mean,
    linetype = "dashed"
  ) +
  theme_bw() +
  labs(
    title = "Scenario 9: variation in mean final outbreak size",
    x = "Random control simulation",
    y = "Mean final outbreak size"
  )
```
```{r}
# add boxplot to show distribution (AI)
boxplot_c9 <- ggplot(control_9_variance, aes(x = 1, y = Mean)) +
  geom_boxplot(width = 0.25, alpha = 0.2) +
  geom_jitter(width = 0.05, size = 3) +
  geom_hline(
    yintercept = control_9_summary_pooled$Mean,
    linetype = "dashed"
  ) +
  theme_bw() +
  labs(
    title = "Scenario 9: consistency across random controls",
    x = NULL,
    y = "Mean final outbreak size"
  ) +
  scale_x_continuous(
    breaks = 1,
    labels = "10 random controls"
  )

boxplot_c9

ggsave(
  filename = "boxplot_c9_var.png",
  plot = boxplot_c9,
  width = 6,
  height = 4,
  units = "in",
  dpi = 300
)
```


```{r}
# SCENARIO 10
# pool the 20,000 outcomes

control_10_results_pooled <- list(
  scenario_10a_results,
  scenario_10b_results,
  scenario_10c_results,
  scenario_10d_results,
  scenario_10e_results,
  scenario_10f_results,
  scenario_10g_results,
  scenario_10h_results,
  scenario_10i_results,
  scenario_10j_results
)

control_10_finalsizes <- unlist(
  lapply(control_10_results_pooled, function(x) {
    sapply(x, function(i) i$summary$final_infected)
  })
)

control_10_peakinfectious <- unlist(
  lapply(control_10_results_pooled, function(x) {
    sapply(x, function(i) i$summary$peak_infectious)
  })
)
```
```{r}
# create summary results for analysis - contains the 20000 individual pooled simulation summaries
#summary_results$SC10 <- bind_rows(
  #lapply(control_10_results_pooled, function(x) x$summary)
#)
```
```{r}
# create summary results for analysis - contains the 20000 individual pooled simulation summaries
# AI code - fixed nesting problem to keep column names
summary_results$SC10 <- bind_rows(
  lapply(control_10_results_pooled, function(x) {
    bind_rows(lapply(x, function(i) i$summary))
  })
)
```

```{r}
length(control_10_finalsizes)
```
```{r}
# pooled summary

control_10_summary_pooled <- data.frame(
  N = length(control_10_finalsizes),
  Mean = mean(control_10_finalsizes),
  Median = median(control_10_finalsizes),
  IQR = IQR(control_10_finalsizes),
  SD = sd(control_10_finalsizes),
  Min = min(control_10_finalsizes),
  Max = max(control_10_finalsizes)
)

control_10_summary_pooled
```
```{r}
# show isnt driven by one random control - calculate the summary repeatedly for each of the 10 controls

control_10_variance <- do.call(
  rbind,
  lapply(seq_along(control_10_results_pooled), function(i) {
    
    x <- sapply(
      control_10_results_pooled[[i]],
      function(j) j$summary$final_infected
    )
    
    data.frame(
      Control = i,
      N = length(x),
      Mean = mean(x),
      Median = median(x),
      IQR = IQR(x),
      SD = sd(x)
    )
  })
)

control_10_variance
```
```{r}
# calculate variability between the 10 controls - how consistent are the matched controls?

sd(control_10_variance$Mean)
```
```{r}
# range of control means

range(control_10_variance$Mean)
```
```{r}
# SEE CODE ABOVE TO MAKE BOX PLOT TO BETTER SHOW DIST
ggplot(control_10_variance, aes(x = Control, y = Mean)) +
  geom_point(size = 3) +
  geom_hline(
    yintercept = control_10_summary_pooled$Mean,
    linetype = "dashed"
  ) +
  theme_bw() +
  labs(
    title = "Scenario 10: variation in mean final outbreak size",
    x = "Random control simulation",
    y = "Mean final outbreak size"
  )
```
```{r}
# add boxplot to show distribution (AI)
boxplot_c10 <- ggplot(control_10_variance, aes(x = 1, y = Mean)) +
  geom_boxplot(width = 0.25, alpha = 0.2) +
  geom_jitter(width = 0.05, size = 3) +
  geom_hline(
    yintercept = control_10_summary_pooled$Mean,
    linetype = "dashed"
  ) +
  theme_bw() +
  labs(
    title = "Scenario 10: consistency across random controls",
    x = NULL,
    y = "Mean final outbreak size"
  ) +
  scale_x_continuous(
    breaks = 1,
    labels = "10 random controls"
  )

boxplot_c10

ggsave(
  filename = "boxplot_c10_var.png",
  plot = boxplot_c10,
  width = 6,
  height = 4,
  units = "in",
  dpi = 300
)
```

```{r}
# SCENARIO 11
# pool the 20,000 outcomes

control_11_results_pooled <- list(
  scenario_11a_results,
  scenario_11b_results,
  scenario_11c_results,
  scenario_11d_results,
  scenario_11e_results,
  scenario_11f_results,
  scenario_11g_results,
  scenario_11h_results,
  scenario_11i_results,
  scenario_11j_results
)

control_11_finalsizes <- unlist(
  lapply(control_11_results_pooled, function(x) {
    sapply(x, function(i) i$summary$final_infected)
  })
)

control_11_peakinfectious <- unlist(
  lapply(control_11_results_pooled, function(x) {
    sapply(x, function(i) i$summary$peak_infectious)
  })
)
```
```{r}
# create summary results for analysis - contains the 20000 individual pooled simulation summaries
#summary_results$SC11 <- bind_rows(
  #lapply(control_11_results_pooled, function(x) x$summary)
#)
```
```{r}
# create summary results for analysis - contains the 20000 individual pooled simulation summaries
# AI code - fixed nesting problem to keep column names
summary_results$SC11 <- bind_rows(
  lapply(control_11_results_pooled, function(x) {
    bind_rows(lapply(x, function(i) i$summary))
  })
)
```

```{r}
length(control_11_finalsizes)
```
```{r}
# pooled summary

control_11_summary_pooled <- data.frame(
  N = length(control_11_finalsizes),
  Mean = mean(control_11_finalsizes),
  Median = median(control_11_finalsizes),
  IQR = IQR(control_11_finalsizes),
  SD = sd(control_11_finalsizes),
  Min = min(control_11_finalsizes),
  Max = max(control_11_finalsizes)
)

control_11_summary_pooled
```
```{r}
# show isnt driven by one random control - calculate the summary repeatedly for each of the 10 controls

control_11_variance <- do.call(
  rbind,
  lapply(seq_along(control_11_results_pooled), function(i) {
    
    x <- sapply(
      control_11_results_pooled[[i]],
      function(j) j$summary$final_infected
    )
    
    data.frame(
      Control = i,
      N = length(x),
      Mean = mean(x),
      Median = median(x),
      IQR = IQR(x),
      SD = sd(x)
    )
  })
)

control_11_variance
```
```{r}
# calculate variability between the 10 controls - how consistent are the matched controls?

sd(control_11_variance$Mean)
```
```{r}
# range of control means

range(control_11_variance$Mean)
```

```{r}
ggplot(control_11_variance, aes(x = Control, y = Mean)) +
  geom_point(size = 3) +
  geom_hline(
    yintercept = control_11_summary_pooled$Mean,
    linetype = "dashed"
  ) +
  theme_bw() +
  labs(
    title = "Scenario 11: variation in mean final outbreak size",
    x = "Random control simulation",
    y = "Mean final outbreak size"
  )
```
```{r}
# add boxplot to show distribution (AI)
boxplot_c11 <- ggplot(control_11_variance, aes(x = 1, y = Mean)) +
  geom_boxplot(width = 0.25, alpha = 0.2) +
  geom_jitter(width = 0.05, size = 3) +
  geom_hline(
    yintercept = control_11_summary_pooled$Mean,
    linetype = "dashed"
  ) +
  theme_bw() +
  labs(
    title = "Scenario 11: consistency across random controls",
    x = NULL,
    y = "Mean final outbreak size"
  ) +
  scale_x_continuous(
    breaks = 1,
    labels = "10 random controls"
  )

boxplot_c11

ggsave(
  filename = "boxplot_c11_var.png",
  plot = boxplot_c11,
  width = 6,
  height = 4,
  units = "in",
  dpi = 300
)
```


```{r}
# SCENARIO 12
# pool the 20,000 outcomes

control_12_results_pooled <- list(
  scenario_12a_results,
  scenario_12b_results,
  scenario_12c_results,
  scenario_12d_results,
  scenario_12e_results,
  scenario_12f_results,
  scenario_12g_results,
  scenario_12h_results,
  scenario_12i_results,
  scenario_12j_results
)

control_12_finalsizes <- unlist(
  lapply(control_12_results_pooled, function(x) {
    sapply(x, function(i) i$summary$final_infected)
  })
)

control_12_peakinfectious <- unlist(
  lapply(control_12_results_pooled, function(x) {
    sapply(x, function(i) i$summary$peak_infectious)
  })
)
```
```{r}
# create summary results for analysis - contains the 20000 individual pooled simulation summaries
#summary_results$SC12 <- bind_rows(
  #lapply(control_12_results_pooled, function(x) x$summary)
#)
```
```{r}
# create summary results for analysis - contains the 20000 individual pooled simulation summaries
# AI code - fixed nesting problem to keep column names
summary_results$SC12 <- bind_rows(
  lapply(control_12_results_pooled, function(x) {
    bind_rows(lapply(x, function(i) i$summary))
  })
)
```

```{r}
length(control_12_finalsizes)
```
```{r}
# pooled summary

control_12_summary_pooled <- data.frame(
  N = length(control_12_finalsizes),
  Mean = mean(control_12_finalsizes),
  Median = median(control_12_finalsizes),
  IQR = IQR(control_12_finalsizes),
  SD = sd(control_12_finalsizes),
  Min = min(control_12_finalsizes),
  Max = max(control_12_finalsizes)
)

control_12_summary_pooled
```
```{r}
# show isnt driven by one random control - calculate the summary repeatedly for each of the 10 controls

control_12_variance <- do.call(
  rbind,
  lapply(seq_along(control_12_results_pooled), function(i) {
    
    x <- sapply(
      control_12_results_pooled[[i]],
      function(j) j$summary$final_infected
    )
    
    data.frame(
      Control = i,
      N = length(x),
      Mean = mean(x),
      Median = median(x),
      IQR = IQR(x),
      SD = sd(x)
    )
  })
)

control_12_variance
```
```{r}
# calculate variability between the 10 controls - how consistent are the matched controls?

sd(control_12_variance$Mean)
```
```{r}
# range of control means

range(control_12_variance$Mean)
```
```{r}
ggplot(control_12_variance, aes(x = Control, y = Mean)) +
  geom_point(size = 3) +
  geom_hline(
    yintercept = control_12_summary_pooled$Mean,
    linetype = "dashed"
  ) +
  theme_bw() +
  labs(
    title = "Scenario 12: variation in mean final outbreak size",
    x = "Random control simulation",
    y = "Mean final outbreak size"
  )
```
```{r}
# add boxplot to show distribution (AI)
boxplot_c12 <- ggplot(control_12_variance, aes(x = 1, y = Mean)) +
  geom_boxplot(width = 0.25, alpha = 0.2) +
  geom_jitter(width = 0.05, size = 3) +
  geom_hline(
    yintercept = control_12_summary_pooled$Mean,
    linetype = "dashed"
  ) +
  theme_bw() +
  labs(
    title = "Scenario 12: consistency across random controls",
    x = NULL,
    y = "Mean final outbreak size"
  ) +
  scale_x_continuous(
    breaks = 1,
    labels = "10 random controls"
  )

boxplot_c12

ggsave(
  filename = "boxplot_c12_var.png",
  plot = boxplot_c12,
  width = 6,
  height = 4,
  units = "in",
  dpi = 300
)
```

```{r}
# SCENARIO 13
# pool the 20,000 outcomes

control_13_results_pooled <- list(
  scenario_13a_results,
  scenario_13b_results,
  scenario_13c_results,
  scenario_13d_results,
  scenario_13e_results,
  scenario_13f_results,
  scenario_13g_results,
  scenario_13h_results,
  scenario_13i_results,
  scenario_13j_results
)

control_13_finalsizes <- unlist(
  lapply(control_13_results_pooled, function(x) {
    sapply(x, function(i) i$summary$final_infected)
  })
)

control_13_peakinfectious <- unlist(
  lapply(control_13_results_pooled, function(x) {
    sapply(x, function(i) i$summary$peak_infectious)
  })
)
```
```{r}
# create summary results for analysis - contains the 20000 individual pooled simulation summaries
#summary_results$SC13 <- bind_rows(
  #lapply(control_13_results_pooled, function(x) x$summary)
#)
```
```{r}
# create summary results for analysis - contains the 20000 individual pooled simulation summaries
# AI code - fixed nesting problem to keep column names
summary_results$SC13 <- bind_rows(
  lapply(control_13_results_pooled, function(x) {
    bind_rows(lapply(x, function(i) i$summary))
  })
)
```

```{r}
length(control_13_finalsizes)
```
```{r}
# pooled summary

control_13_summary_pooled <- data.frame(
  N = length(control_13_finalsizes),
  Mean = mean(control_13_finalsizes),
  Median = median(control_13_finalsizes),
  IQR = IQR(control_13_finalsizes),
  SD = sd(control_13_finalsizes),
  Min = min(control_13_finalsizes),
  Max = max(control_13_finalsizes)
)

control_13_summary_pooled
```
```{r}
# show isnt driven by one random control - calculate the summary repeatedly for each of the 10 controls

control_13_variance <- do.call(
  rbind,
  lapply(seq_along(control_13_results_pooled), function(i) {
    
    x <- sapply(
      control_13_results_pooled[[i]],
      function(j) j$summary$final_infected
    )
    
    data.frame(
      Control = i,
      N = length(x),
      Mean = mean(x),
      Median = median(x),
      IQR = IQR(x),
      SD = sd(x)
    )
  })
)

control_13_variance
```
```{r}
# calculate variability between the 10 controls - how consistent are the matched controls?

sd(control_13_variance$Mean)
```
```{r}
# range of control means

range(control_13_variance$Mean)
```
```{r}
ggplot(control_13_variance, aes(x = Control, y = Mean)) +
  geom_point(size = 3) +
  geom_hline(
    yintercept = control_13_summary_pooled$Mean,
    linetype = "dashed"
  ) +
  theme_bw() +
  labs(
    title = "Scenario 13: variation in mean final outbreak size",
    x = "Random control simulation",
    y = "Mean final outbreak size"
  )
```
```{r}
names(summary_results)
```
```{r}
sapply(summary_results, nrow)
```
```{r}
# add boxplot to show distribution (AI)
boxplot_c13 <- ggplot(control_13_variance, aes(x = 1, y = Mean)) +
  geom_boxplot(width = 0.25, alpha = 0.2) +
  geom_jitter(width = 0.05, size = 3) +
  geom_hline(
    yintercept = control_13_summary_pooled$Mean,
    linetype = "dashed"
  ) +
  theme_bw() +
  labs(
    title = "Scenario 13: consistency across random controls",
    x = NULL,
    y = "Mean final outbreak size"
  ) +
  scale_x_continuous(
    breaks = 1,
    labels = "10 random controls"
  )

boxplot_c13

ggsave(
  filename = "boxplot_c13_var.png",
  plot = boxplot_c13,
  width = 6,
  height = 4,
  units = "in",
  dpi = 300
)
```

# Read in data from RQ1 for number farms removed

```{r}
# OUTDEGREE
OD_all_candidates <- readRDS(
  "outdegree_all_candidates.rds"
)

OD_med_candidates <- readRDS(
  "outdegree_med_persistent.rds"
)

OD_high_candidates <- readRDS(
  "outdegree_high_persistent.rds"
)
```

```{r}
#OD_all_candidates
```
```{r}
# BETWEENNESS
BT_all_candidates <- readRDS(
  "BT_any_persistent_corrected.rds"
)

BT_high_candidates <- readRDS(
  "BT_high_persistent_corrected.rds"
)

BT_med_candidates <- readRDS(
  "BT_med_persistent_corrected.rds"
)
```
```{r}
#BT_med_candidates
```

# 1. Create scenario table - what was removed

```{r}
total_farms = 301
scenario_table <- data.frame(
  Scenario = c(
    "(1)",
    "(2)",
    "(3)",
    "(4)",
    "(5)",
    "(6)",
    "(7)",
    "(8)",
    "(9)",
    "(10)",
    "(11)",
    "(12)",
    "(13)"
  ),
  
  Type = c(
    "Baseline",
    "Targeted Intervention",
    "Targeted Intervention",
    "Targeted Intervention",
    "Targeted Intervention",
    "Targeted Intervention",
    "Targeted Intervention",
    "CONTROL POOLED OD HIGH",
    "CONTROL POOLED OD MED",
    "CONTROL POOLED OD ALL",
    "CONTROL POOLED BT HIGH",
    "CONTROL POOLED BT MED",
    "CONTROL POOLED BT ALL"
  ),
  
  Centrality = c(
    "None",
    "Out-degree",
    "Out-degree",
    "Out-degree",
    "Betweenness",
    "Betweenness",
    "Betweenness",
    "RandomOD",
    "RandomOD",
    "RandomOD",
    "RandomBT",
    "RandomBT",
    "RandomBT"
  ),
  
  Persistence = c(
    "None",
    "High >=8 months",
    "Medium >=6 months",
    ">=1 month",
    "High >=6 months",
    "Medium >=4 months",
    ">= 1 month",
    "Matched to OD HIGH",
    "Matched to OD MED",
    "Matched to OD ALL",
    "Matched to BT HIGH",
    "Matched to BT MED",
    "Matched to BT ALL"
  ),
  Number_of_farms_removed = c(
    0,
    length(OD_high_candidates),
    length(OD_med_candidates),
    length(OD_all_candidates),
    length(BT_high_candidates),
    length(BT_med_candidates),
    length(BT_all_candidates),
    5,
    8,
    43,
    1,
    8,
    35
  ),
  
  Percent_of_all_farms = c(
    0,
    length(OD_high_candidates)/total_farms*100,
    length(OD_med_candidates)/total_farms*100,
    length(OD_all_candidates)/total_farms*100,
    length(BT_high_candidates)/total_farms*100,
    length(BT_med_candidates)/total_farms*100,
    length(BT_all_candidates)/total_farms*100,
    5/total_farms*100,
    8/total_farms*100,
    43/total_farms*100,
    1/total_farms*100,
    8/total_farms*100,
    35/total_farms*100
  )
)

scenario_table
```
```{r}
total_farms
```

## SOIR Epidemic outcome analysis ##

# Established epidemic definition: threshold sensitivity analysis

1. examine the baseline outbreak-size distribution

  supports why selected >=10 - baseline histogram

```{r}
# plot the distribution

p_hist <- ggplot(summary_results_SC1, aes(x = final_infected)) +
  geom_histogram(
    binwidth = 1,
    colour = "black",
    fill = "forestgreen"
  ) +
  labs(
    x = "Final outbreak size (infected farms)",
    y = "Number of simulations",
    title = "Distribution of final outbreak sizes in baseline simulations"
  ) +
  theme_classic()

ggsave(
  "baseline_outbreak_dists.png", # can change to pdf
  plot = p_hist,
  width = 8,
  height = 5,
  units = "in",
  dpi = 300
)

p_hist
```
Histogram does not show obvious break to select as threshold. 

Create the scenario outbreak size data
```{r}
# extract outbreak-size results for each scenario
summary_results_SC1 <- summary_results$SC1
summary_results_SC2 <- summary_results$SC2
summary_results_SC3 <- summary_results$SC3
summary_results_SC4 <- summary_results$SC4
summary_results_SC5 <- summary_results$SC5
summary_results_SC6 <- summary_results$SC6
summary_results_SC7 <- summary_results$SC7
# pooled controls
summary_results_SC8 <- summary_results$SC8
summary_results_SC9 <- summary_results$SC9
summary_results_SC10 <- summary_results$SC10
summary_results_SC11 <- summary_results$SC11
summary_results_SC12 <- summary_results$SC12
summary_results_SC13 <- summary_results$SC13

```
```{r}
# create list of final outbreak sizes
outbreak_sizes <- list(
  Baseline = summary_results$SC1$final_infected,
  
  Outdegree_high = summary_results$SC2$final_infected,
  Out_degree_high_random = summary_results$SC8$final_infected,
  
  Out_degree_med = summary_results$SC3$final_infected,
  Out_degree_med_random = summary_results$SC9$final_infected,
  
  Outdegree_all = summary_results$SC4$final_infected,
  Outdegree_all_random = summary_results$SC10$final_infected,
  
  Betweenness_high = summary_results$SC5$final_infected,
  Betweenness_high_random = summary_results$SC11$final_infected,
  
  Betweenness_med =summary_results$SC6$final_infected,
  Betweenness_med_random = summary_results$SC12$final_infected,
  
  Betweenness_all = summary_results$SC7$final_infected,
  Betweenness_all_random = summary_results$SC13$final_infected
)

```
```{r}
names(summary_results$SC8)
names(summary_results$SC9)
names(summary_results$SC10)
names(summary_results$SC11)
names(summary_results$SC12)
names(summary_results$SC13)
```
Threshold sensitivity analysis
  whether >=10 is robust
  
```{r}
# calculates, for each scenario/control and threshold, the proportion of simulations where final infected farms >= thresholds2
thresholds2 <- c(1, 2, 3, 5, 7, 10, 15, 20)

threshold_results2 <- data.frame(
  scenario = character(),
  threshold = numeric(),
  proportion = numeric(),
  stringsAsFactors = FALSE
  )


for (s in names(outbreak_sizes)) {
  
  outbreaks <- outbreak_sizes[[s]]
  
  for (t in thresholds2) {
    threshold_results2 <- rbind(
      threshold_results2,
      data.frame(
        Scenario2 = s,
        Threshold2 = t,
        proportion2 = mean(outbreaks >= t) #* 100
      )
    )
  }
}
```
```{r}
threshold_results2
```

Sensitivity plot
```{r, width = 12, height = 5}
# plot sensitivity analysis - lines show the % of simulations that have exceeded the threshold of x final farms infected

p_thresh_sens <- ggplot(threshold_results2,
       aes(x = Threshold2,
           y = proportion2,
           colour = Scenario2)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    title = "Established epidemic threshold sensitivity analysis",
    x = "Outbreak size threshold (no. infected farms)",
    y = "Probability of exceeding threshold",
    colour = "Scenario"
  ) +
  theme_bw()

ggsave(
  "threshold_establish.png", # can change to pdf
  plot = p_thresh_sens,
  width = 8,
  height = 5,
  units = "in",
  dpi = 300
)

p_thresh_sens 
```


Use for result 1?
```{r}
# plot intervention/control pairs

threshold_10 <- threshold_results2 %>%
  filter(Threshold2 == 10)

ggplot(threshold_10,
       aes(x = Scenario2, y = proportion2, fill = Scenario2)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  labs(
    x = NULL,
    y = "Probability of outbreak ≥10 infected farms"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )
```

```{r}
#AI - how to make grouped bar plot to show pairs
```

# RESULT 1

Probability of sustained outbreak - replicates led to >=10 infected final infected farms
Q = probability of a sustained outbreak over conditional threshold >= 10 final infected farms

Define the scenarios and extract final outbreak size, this is the input to the analysis

Need to consider - keep pooled controls = 20,000 vs 2,000 - should be ok if looking at overall probability of a >=10 outbreak under matched random removal BUT could keep distribution of 10 control probabilties and compare to the intervention probability (2,000 to 2,000).

PLAN - stick with first option but report variance

```{r}
# list of scenarios to include - dropped combined - see prev

sustained_trans <- list(
  Baseline = summary_results$SC1$final_infected,
  
  Outdegree_high = summary_results$SC2$final_infected, # contains final_infected result of 2000 simulations
  Out_degree_high_random = summary_results$SC8$final_infected, # contains final_infected result of 20,000 simulations
  
  Out_degree_med = summary_results$SC3$final_infected,
  Out_degree_med_random = summary_results$SC9$final_infected,
  
  Outdegree_all = summary_results$SC4$final_infected,
  Outdegree_all_random = summary_results$SC10$final_infected,
  
  Betweenness_high = summary_results$SC5$final_infected,
  Betweenness_high_random = summary_results$SC11$final_infected,
  
  Betweenness_med =summary_results$SC6$final_infected,
  Betweenness_med_random = summary_results$SC12$final_infected,
  
  Betweenness_all = summary_results$SC7$final_infected,
  Betweenness_all_random = summary_results$SC13$final_infected
)

```
Calculate the probability of sustained transmission and relative reduction

```{r}
# counts how many simulations exceeded the threshold & calculates number sustained outbreaks; proportion sustained outbreaks; relative reduction vs baseline
threshold_final <- 10

# calculate proportions
threshold_counts_final <- data.frame(
  Scenario_final = names(sustained_trans),
  Exceeding = sapply(sustained_trans, function(x) sum(x >= threshold_final)),
  Total = sapply(sustained_trans, length)
)

threshold_counts_final$Proportion_final <- threshold_counts_final$Exceeding / threshold_counts_final$Total

# extract baseline probability
baseline_prob <- threshold_counts_final$Proportion_final[
  threshold_counts_final$Scenario_final == "Baseline"
]

# Relative reduction random matchec control vs baseline
threshold_counts_final$Relative_reduction_vs_baseline <- NA_real_

random_controls <- c(
  "Out_degree_high_random",
  "Out_degree_med_random",
  "Outdegree_all_random",
  "Betweenness_high_random",
  "Betweenness_med_random",
  "Betweenness_all_random"
)

for (control in random_controls) {
  
  control_prob <- threshold_counts_final$Proportion_final[
    threshold_counts_final$Scenario_final == control
  ]
  
  threshold_counts_final$Relative_reduction_vs_baseline[
    threshold_counts_final$Scenario_final == control
  ] <- (baseline_prob - control_prob) / baseline_prob
}

#baseline_prob <- threshold_counts_final$Proportion_final[threshold_counts_final$Scenario_final == "Baseline"]

#threshold_counts_final$Relative_reduction <- (
 # (baseline_prob - threshold_counts_final$Proportion_final) / #baseline_prob
#)

# Relative reduction intervention vs matched control (AI added code)

threshold_counts_final$Relative_reduction_vs_random <- NA_real_

pairs <- list(
  c("Outdegree_high", "Out_degree_high_random"),
  c("Out_degree_med", "Out_degree_med_random"),
  c("Outdegree_all", "Outdegree_all_random"),
  c("Betweenness_high", "Betweenness_high_random"),
  c("Betweenness_med", "Betweenness_med_random"),
  c("Betweenness_all", "Betweenness_all_random")
)

for (pair in pairs) {
  
  intervention <- pair[1]
  control <- pair[2]
  
  intervention_prob <- threshold_counts_final$Proportion_final[
    threshold_counts_final$Scenario_final == intervention
  ]
  
  control_prob <- threshold_counts_final$Proportion_final[
    threshold_counts_final$Scenario_final == control
  ]
  
  threshold_counts_final$Relative_reduction_vs_random[
    threshold_counts_final$Scenario_final == intervention
  ] <- (control_prob - intervention_prob) / control_prob
}

colnames(threshold_counts_final) <- c(
  "Scenario",
  "Number of simulations >=10 infected farms",
  "Total simulations",
  "Probability of >=10 infected farms",
  "Relative reduction matched Cx vs baseline",
  "Relative reduction Int vs matched random control"
)

threshold_counts_final
```
```{r}
# save table
write.csv(
  threshold_counts_final,
  "established_outbreak_results.csv",
  row.names = FALSE
)
```

Present results:
(a) paired plot (bar?) of probabitliy of >=10 infected farms, intervention beside matched conttol

(b) main table - scenario, number sims >=10; probability >=10, reduction: random contrl vs baseline; reduction int vs matched control.

? add 95% confidence interval


Organise scenarios by intervention type and intensity
```{r}
# AI - how can i group this plot by intensity windows, maintaining controls

# Create empty columns
threshold_counts_final$Intensity <- NA_character_
threshold_counts_final$Method <- NA_character_

# Assign intensity

threshold_counts_final$Intensity[
  grepl("high", threshold_counts_final$Scenario, ignore.case = TRUE)
] <- "High"

threshold_counts_final$Intensity[
  grepl("med", threshold_counts_final$Scenario, ignore.case = TRUE)
] <- "Medium"

threshold_counts_final$Intensity[
  grepl("all", threshold_counts_final$Scenario, ignore.case = TRUE)
] <- "All"

threshold_counts_final$Intensity[
  threshold_counts_final$Scenario == "Baseline"
] <- "Baseline"

# Assign method

# Random matched controls
threshold_counts_final$Method[
  grepl("out.*random", threshold_counts_final$Scenario, ignore.case = TRUE)
] <- "Random control (OD matched)"

threshold_counts_final$Method[
  grepl("betweenness.*random", threshold_counts_final$Scenario, ignore.case = TRUE)
] <- "Random control (BT matched)"


# Targeted interventions
threshold_counts_final$Method[
  grepl("^outdegree|^out_degree",
        threshold_counts_final$Scenario,
        ignore.case = TRUE) &
  !grepl("random",
         threshold_counts_final$Scenario,
         ignore.case = TRUE)
] <- "Out-degree"

threshold_counts_final$Method[
  grepl("^betweenness",
        threshold_counts_final$Scenario,
        ignore.case = TRUE) &
  !grepl("random",
         threshold_counts_final$Scenario,
         ignore.case = TRUE)
] <- "Betweenness"


# Baseline
threshold_counts_final$Method[
  threshold_counts_final$Scenario == "Baseline"
] <- "Baseline"


# Set plotting order

threshold_counts_final$Intensity <- factor(
  threshold_counts_final$Intensity,
  levels = c("High", "Medium", "All", "Baseline")
)

threshold_counts_final$Method <- factor(
  threshold_counts_final$Method,
  levels = c(
    "Baseline",
    "Random control (OD matched)",
    "Out-degree",
    "Random control (BT matched)",
    "Betweenness"
  )
)


# Check
threshold_counts_final[, c("Scenario", "Intensity", "Method")]
```
```{r, fig.width = 14, fig.height = 6}
# Main probability plot - grouped bar plot - int & matched ct beside each other with each intensity
# AI - how to add baseline to each plot

# Duplicate baseline into each intensity panel

baseline_row <- threshold_counts_final[
  threshold_counts_final$Scenario == "Baseline", 
]

baseline_repeated <- baseline_row[rep(1, 3), ]

baseline_repeated$Intensity <- factor(
  c("High", "Medium", "All"),
  levels = c("High", "Medium", "All")
)

baseline_repeated$Method <- "Baseline"


# Remove original baseline from plotting data
# and add baseline to each intensity panel

plot_data <- rbind(
  threshold_counts_final[
    threshold_counts_final$Scenario != "Baseline",
  ],
  baseline_repeated
)


# Set plotting order
plot_data$Intensity <- factor(
  plot_data$Intensity,
  levels = c("High", "Medium", "All")
)

plot_data$Method <- factor(
  plot_data$Method,
  levels = c(
    "Baseline",
    "Random control (OD matched)",
    "Out-degree",
    "Random control (BT matched)",
    "Betweenness"
  )
)


# Plot

Prob_infected_plot <- ggplot(
  plot_data,
  aes(
    x = Method,
    y = `Probability of >=10 infected farms`,
    fill = Method
  )
) +
  geom_col(width = 0.7) +
  facet_wrap(~Intensity,
             labeller = as_labeller(c(
               High = "Highly persistent",
               Medium = "Persistent",
               All= "Transient"
             ))
             ) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    title = "Probability of established outbreak by intervention strategy",
    x = "Intervention strategy",
    y = "Probability of ≥10 infected farms",
    fill = NULL
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 50, hjust = 1),
    plot.margin = margin(10, 10, 25, 30)
  )
Prob_infected_plot
```
```{r}
# main results table
threshold_counts_final
```
```{r}
ggsave(
  "Prob_infected_plot.png", # can change to pdf
  plot = Prob_infected_plot,
  width = 8,
  height = 5,
  units = "in",
  dpi = 300
)
```


https://search.r-project.org/R/refmans/stats/html/fisher.test.html

Fishers exact test comparing outbreak proportions (not paired because using a pooled sim?)

Pooled random control vs baseline
Pooledrandom control vs targeted intervention

create 2x2 table for each comparison - epi(>=10), no epi (<10)

Create variables for fishers test tables:

```{r}
# ct vs int
control_exceeding <- threshold_counts_final$`Number of simulations >=10 infected farms`[
  threshold_counts_final$Scenario == "Out_degree_high_random"
]

control_total <- threshold_counts_final$`Total simulations`[
  threshold_counts_final$Scenario == "Out_degree_high_random"
]

intervention_exceeding <- threshold_counts_final$`Number of simulations >=10 infected farms`[
  threshold_counts_final$Scenario == "Outdegree_high"
]

intervention_total <- threshold_counts_final$`Total simulations`[
  threshold_counts_final$Scenario == "Outdegree_high"
]

# ORGANISED IN DIRECTION OF TEST
test_matrix_ODhigh <- matrix(
  c(
    control_exceeding,
    control_total - control_exceeding,
    intervention_exceeding,
    intervention_total - intervention_exceeding
  ),
  nrow = 2
)

fisher.test(test_matrix_ODhigh)

# stat sig at 5% level
```
```{r}
# AI code - how to automate to complete all 12 comparisons & present in one table

run_fisher <- function(group1, group2) {
  
  x1 <- threshold_counts_final[
    threshold_counts_final$Scenario == group1, ]
  
  x2 <- threshold_counts_final[
    threshold_counts_final$Scenario == group2, ]
  
  test_matrix <- matrix(
    c(
      x1$`Number of simulations >=10 infected farms`,
      x1$`Total simulations` - x1$`Number of simulations >=10 infected farms`,
      x2$`Number of simulations >=10 infected farms`,
      x2$`Total simulations` - x2$`Number of simulations >=10 infected farms`
    ),
    nrow = 2,
    byrow = TRUE
  )
  
  test <- fisher.test(test_matrix)
  
  c(
    Odds_ratio = unname(test$estimate),
    CI_lower = test$conf.int[1],
    CI_upper = test$conf.int[2],
    P_value = test$p.value
  )
}
```
```{r}
# Intervention vs matched random control
intervention_control <- do.call(rbind, lapply(
  pairs, function(x) run_fisher(x[1], x[2])
))

# Random control vs baseline
control_baseline <- do.call(rbind, lapply(
  random_controls, function(x) run_fisher(x, "Baseline")
))

intervention_control
control_baseline
```
```{r}
# AI - as exportable table
all_fisher_tests <- rbind(
  do.call(rbind, lapply(pairs, function(x) {
    result <- run_fisher(x[1], x[2])
    data.frame(
      Comparison = paste(x[1], "vs", x[2]),
      Odds_ratio = result["Odds_ratio"],
      CI_lower = result["CI_lower"],
      CI_upper = result["CI_upper"],
      P_value = result["P_value"]
    )
  })),
  
  do.call(rbind, lapply(random_controls, function(x) {
    result <- run_fisher(x, "Baseline")
    data.frame(
      Comparison = paste(x, "vs Baseline"),
      Odds_ratio = result["Odds_ratio"],
      CI_lower = result["CI_lower"],
      CI_upper = result["CI_upper"],
      P_value = result["P_value"]
    )
  }))
)

# Make the table easier to read
all_fisher_tests$Odds_ratio <- round(as.numeric(all_fisher_tests$Odds_ratio), 2)
all_fisher_tests$CI_lower <- round(as.numeric(all_fisher_tests$CI_lower), 2)
all_fisher_tests$CI_upper <- round(as.numeric(all_fisher_tests$CI_upper), 2)
all_fisher_tests$P_value <- signif(as.numeric(all_fisher_tests$P_value), 3)

all_fisher_tests
```
```{r}
# export
write.csv(
  all_fisher_tests,
  "fisher_exact_results.csv",
  row.names = FALSE
)
```
? consider multiple comparison correction if comparing p-values?

1. effect of removing nodes at all = random ct vs baseline
2. effect of which nodes are removed = int vs ct

so e.g. OD high vs OD high random (p<0.05) AND OD high random vs baseline p>0.05
Evidence suggests that the selection of high outdegree nodes contributed to the reduction beyond the effect of removing nodes at random etc

# RESULT 2

if an epidemic becomes established (threshold >= 10), how large does it become/how many farms are infected?

Report - Median, IQR, % reduction relative to baseline

Filter established outbreaks:
```{r}
# filter simulations above threshold - keep sims where finout outbreak size >= 10

threshold_final <- 10

established_outbreaks <- lapply(
  sustained_trans,# already created for Q1
  function(x) x[x >= threshold_final]
)
```
```{r}
# create a dataframe for plotting

plot_sizes <- stack(established_outbreaks)
colnames(plot_sizes) <- c("Final_size", "Scenario")

# final_size= final number of infected farms
# scenario = intervention scenario
```
```{r}
# add intervention method and intensity from Q1

plot_sizes <- merge(
  plot_sizes,
  threshold_counts_final[, c("Scenario", "Method", "Intensity")],
  by = "Scenario"
)

# Save before duplicating baseline
plot_sizes_stats <- plot_sizes
```

```{r}
# names(threshold_counts_final)
```
Baseline duplication - needed for 3 figure of panels

```{r}
# duplicate baseline into every intensity panel

# extract baseline
baseline_sizes <- subset(plot_sizes, Scenario == "Baseline")
```
```{r}
# duplicate baseline into high/med/all panels
baseline_sizes <- do.call(
  rbind,
  lapply(c("High", "Medium", "All"), function(i) {
    x <- baseline_sizes
    x$Intensity <- i
    x
  })
)

baseline_sizes$Method <- "Baseline"
```
```{r}
# remove original baseline & add duplicated baseline
plot_sizes <- rbind(
  subset(plot_sizes, Scenario != "Baseline"),
  baseline_sizes
)
```
Set plotting order

```{r}
# set plotting order
plot_sizes$Intensity <- factor(
  plot_sizes$Intensity,
  levels = c("High", "Medium", "All")
)

plot_sizes$Method <- factor(
  plot_sizes$Method,
  levels = c(
    "Baseline",
    "Random control (OD matched)",
    "Out-degree",
    "Random control (BT matched)",
    "Betweenness"
  )
)
```
Boxplot

```{r, width = 14, height = 5}
# create boxplots

result2_plot <- ggplot(
  plot_sizes,
  aes(
    x = Method,
    y = Final_size,
    fill = Method
  )
) +
  geom_boxplot() +
  facet_wrap(~Intensity,
             labeller = as_labeller(c(
               High = "Highly persistent",
               Medium = "Persistent",
               All = "Transient"
             ))
             ) +
  labs(
    title = "Final outbreak size by intevention strategy",
    x = "Intervention strategy",
    y = "Final outbreak size (established outbreaks)"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(
      angle = 50,
      hjust = 1
    ),
    plot.margin = margin(10,10,25,30)
  )

# display plot
result2_plot

ggsave(
  filename = "Result2_boxplots.png",
  plot = result2_plot,
  width = 14,
  height = 5,
  units = "in",
  dpi = 300
)

```
Descriptive summary - median/IQR/mean etc

```{r}
# summarise outbreak sizes 

outbreak_summary <- data.frame(
  Scenario = names(established_outbreaks),
  N_established = sapply(established_outbreaks, length), # how many sims exceeded threshold
  Mean = sapply(established_outbreaks, mean), # average final outbreak size
  Variance = sapply(established_outbreaks, var),
  Median = sapply(established_outbreaks, median), # median final outbreak size
  #SD = sapply(established_outbreaks, sd), # variability/standard deviation
  Q1 = sapply(established_outbreaks, quantile, probs = 0.25),
  Q3 = sapply(established_outbreaks, quantile, probs = 0.75),
  P95 = sapply(established_outbreaks,quantile,probs = 0.95,names = FALSE
  ),
  Min = sapply(established_outbreaks, min), # range
  Max = sapply(established_outbreaks, max) # range
)

outbreak_summary
```
```{r}
write.csv(
  outbreak_summary,
  "result_two_outbreak_summary.csv",
  row.names = FALSE
)
```
ADD 95% QUANTILE - plot as hist and red line at 0.95 - report median and 95th percentile.

among established outbreaks in OD_high, the median final outbreak size was 12 (IQR 11 - 15), compared to CT_OH_HIGH, with median final outbreak size of 14 (IQR 11 - 22)

Check distributions - see if still strongly right skewed with filter - relevance of median/SD

looking at Q3 - baseline to ODrandom reduces almost 50% - good! etc
ecdf(data) - if plot = empirical distirbution function. cumulative plot - can look at plot and se percentage at e.g 30. Put a grid on and plot comparison data in RED - look at maximum difference between the two lines. ANY DISTIRBUTION. 
USE ks.test (x, j) - gives p-value PAIRWISE 
plot(ecdf(kk),col='red', add TRUE)

STATS  smirnoff test - ks.test(), same logic as fishers

```{r}
head(plot_sizes_stats)
```
```{r}
unique(plot_sizes_stats$Scenario)
```


mention that student t-test is not applicable because is not Guassian. 

INCLUDE THE NULL HYPOTHESIS IN WRITE UP
H0= they are THE same distribution
H1 - they are different

?ks.test - to get refernces - birnbaum - first one OR USE book underneath. two sample smirnoff test https://www.geeksforgeeks.org/machine-learning/kolmogorov-smirnov-test-ks-test/

DOESNT BASE ON MEAN

STATS - use plot_sizes_stats
want to compare:
  matched control to baseline
  matched control to intervention
STATS - poissons distribution

```{r}
plot_sizes_stats
```
```{r}
jj <- plot_sizes_stats[plot_sizes_stats$Method == "Betweenness", ] # [rows, columns] - so giv me rows where intensity = baseline, and all columns
#jj
dput(jj$Final_size)
#table(plot_sizes_stats$Intensity)
```
```{r}
plot(jj$Final_size)
# HIST()
# PLOT(SORT()) - CHANGE CLASSES/BINS NCLASS = 
```
```{r}
hist(jj$Final_size)
```
```{r}
plot(sort(jj$Final_size))
```
```{r}
kk <- plot_sizes_stats[plot_sizes_stats$Intensity == "Outdegree_high", ]
kk
```

KS test - compares distirbution of final outbreak sizes between two groups. Given that an outbreak became established (>=10 farms), does the intervenetion/control change hte distribution of outbreak size?

using established outbreaks only:

```{r}
# check established_outbreaks

established_outbreaks <- lapply(
  sustained_trans,
  function(x) x[x >= threshold_final]
)
#established_outbreaks
```
```{r}
# create the KS function - adapted from run_fisher() function

# tells the max difference between two cumulative distributions of final outbreak size

run_ks <- function(group1, group2) {
  
  x1 <- established_outbreaks[[group1]]
  x2 <- established_outbreaks[[group2]]
  
  test <- ks.test(
    x1,
    x2,
    alternative = "two.sided",
    exact = FALSE # don't calculate exact p-value, there are replicates in distribution numbers see note at end
  )
  
  c(
    KS_statistic = unname(test$statistic),
    P_value = test$p.value
  )
}
```
```{r}
# intervention vs matched random control - use existing 'pairs' from fishers tests
ks_intervention_control <- do.call(
  rbind,
  lapply(pairs, function(x) {
    run_ks(x[1], x[2])
  })
)

ks_intervention_control
```
```{r}
# random control vs baseline
ks_control_baseline <- do.call(
  rbind,
  lapply(random_controls, function(x) {
    run_ks(x, "Baseline")
  })
)

ks_control_baseline
```
```{r}
# put everything into one tabl
all_ks_tests <- rbind(
  
  # Intervention vs matched random control
  do.call(rbind, lapply(pairs, function(x) {
    
    result <- run_ks(x[1], x[2])
    
    data.frame(
      Comparison = paste(x[1], "vs", x[2]),
      KS_statistic = result["KS_statistic"],
      P_value = result["P_value"]
    )
  })),
  
  # Random control vs baseline
  do.call(rbind, lapply(random_controls, function(x) {
    
    result <- run_ks(x, "Baseline")
    
    data.frame(
      Comparison = paste(x, "vs Baseline"),
      KS_statistic = result["KS_statistic"],
      P_value = result["P_value"]
    )
  }))
)

# Round for presentation
all_ks_tests$KS_statistic <- round(
  as.numeric(all_ks_tests$KS_statistic), 3
)

all_ks_tests$P_value <- signif(
  as.numeric(all_ks_tests$P_value), 3
)

all_ks_tests
```
```{r}
write.csv(
  all_ks_tests,
  "all_ks_tests1.csv",
  row.names = FALSE
)
```
this error: p-value will be approximate.... not true error but cautious of presenting KS p-values as if they are exact. occurs because final outbreak size is a discrete count of farms, so some values can occur many times - the repeated values are ties, and standard KS test's exact p-values calculation assumes continous distribution without ties

Differences in the distributions of final outbreak size among established outbreaks were asesesd using two-sided two-sample Kolmogorov-Smirnoff tests. as final outbreak size was a discrete count outcome with tied values, asymptomatic p-values were used. 

```{r}
scenario_labels <- c( # SEE TABLE IN METHODOLGOY FOR PAIRS
  "A",
  "B",
  "C",
  "D",
  "E",
  "F"
)
```
```{r}
# Create data for intervention vs matched random-control pairs
ecdf_data <- do.call(
  rbind,
  lapply(seq_along(pairs), function(i) {
    
    intervention <- pairs[[i]][1]
    control <- pairs[[i]][2]
    
    data.frame(
      Final_size = c(
        established_outbreaks[[intervention]],
        established_outbreaks[[control]]
      ),
      Scenario = c(
        rep(
          "Targeted intervention",
          length(established_outbreaks[[intervention]])
        ),
        rep(
          "Corresponding random control",
          length(established_outbreaks[[control]])
        )
      ),
      Comparison = scenario_labels[i]
    )
  })
)

# ECDF plot
ecdf_intervention_control <- ggplot(
  ecdf_data,
  aes(
    x = Final_size,
    colour = Scenario
  )
) +
  stat_ecdf(linewidth = 1) +
  facet_wrap(
    ~Comparison,
    ncol = 3
  ) +
  coord_cartesian(xlim = c(10, NA)) +
  labs(
    title = "Final outbreak size by intervention strategy",
    x = "Final outbreak size (established outbreaks)",
    y = "Cumulative proportion (established outbreaks)",
    colour = NULL
  ) +
  theme_bw() +
  theme(
    strip.text = element_text(size = 11)
  )

ecdf_intervention_control
```
```{r}
ggsave(
  filename = "ECDF_intervention_vs_control.png",
  plot = ecdf_intervention_control,
  width = 8,
  height = 4,
  units = "in",
  dpi = 300
)
```
e.g. at 20 infected farms:
Intervention = 1 => 100% of established intervention simulations had a final outbreak size of 20 farms or fewer.
Control = 0.6 => 60% of established control simulations had a final outbreak size of 20 farms or fewer

i.e. what proportion of simulated outbreaks are at or below this outbreak size 
# RESULT 3 - YOU ARE HERE ISSUES WITH CODE IDing PEAK INFECTIOUS
c) peak infectious farms - how intense is outbreak IF THERE IS AN OUTBREAK - explain as above - mean/median peak without <1 sims

When an epidemic occurs (threshold >=10 farms), how intense is it?

Only analyse successful epidemics

Calculate peak infectious farms for every replicate

```{r}
threshold_peak_analysis <- 10
```

```{r}
# create scenario mapping
scenario_lookup <- c(
  Baseline = "SC1",
  
  Outdegree_high = "SC2",
  Out_degree_high_random = "SC8",
  
  Out_degree_med = "SC3",
  Out_degree_med_random = "SC9",
  
  Outdegree_all = "SC4",
  Outdegree_all_random = "SC10",
  
  Betweenness_high = "SC5",
  Betweenness_high_random = "SC11",
  
  Betweenness_med = "SC6",
  Betweenness_med_random = "SC12",
  
  Betweenness_all = "SC7",
  Betweenness_all_random = "SC13"
)
```
Restrict to established epidemics:
For each of hte 13 scenarios it:
1. finds the appropriate 'summaryy$results$SC#'
2. removes simulaitons where final_infected <10
3. keeps the replicate and its corresponding peak_infectious
4. labels it with the scenario name
5. combines everything into one dataframe

```{r}
# keep sims where threshold transmission occured
# create dataframe with replicate/peak_inefctious/scenario

peak_infectious_established <- lapply(
  names(scenario_lookup),
  function(scenario) {
    
    # contains both final_infected and peak_infectious for same replicate - canfilter on final_infected and retain the corresponding peak_infectious
    
    peaks <- summary_results[[scenario_lookup[[scenario]]]]
    
    peaks %>%
      filter(final_infected >= threshold_peak_analysis) %>%
      select(replicate, peak_infectious) %>%
      mutate(Scenario = scenario)
  }
)

peak_infectious_established <- bind_rows(peak_infectious_established)
```
```{r}
names(scenario_lookup)
scenario_lookup
```
```{r}
lapply(
  names(scenario_lookup),
  function(scenario) {
    list(
      scenario = scenario,
      lookup = scenario_lookup[[scenario]],
      class = class(summary_results[[scenario_lookup[[scenario]]]]),
      is_null = is.null(summary_results[[scenario_lookup[[scenario]]]])
    )
  }
)
```
```{r}
head(peak_infectious_established)
```
```{r}
table(peak_infectious_established$Scenario)
```

Add method and intensity labels to the peak_infectious data by matching with scenario
```{r}
peak_infectious_established <- merge(
  peak_infectious_established,
  threshold_counts_final[, c("Scenario", "Method", "Intensity")], # from threshold_counts_final only take these 3 columns
  by = "Scenario" # & match the two dataframes wherever their scenario values are the same, and attach method & intensity
)
```
```{r}
tail(peak_infectious_established)
```


Create copies

```{r}
# make separate copy of original dataset for stats 
peak_infectious_stats <- peak_infectious_established
#make separate copy for figures
peak_infectious_plot <- peak_infectious_established
```
DIAGNOSTICS
```{r}
#head(summary_results$SC1)
```
```{r}
#sum(summary_results$SC1$final_infected >= 10)
```
```{r}
#table(summary_results$SC1$final_infected >= 10)
```
```{r}
#sapply(
#  summary_results[c("SC1","SC2","SC3","SC4","SC5","SC6","SC7",
#                    "SC8","SC9","SC10","SC11","SC12","SC13")],
#  function(x) sum(x$final_infected >= 10)
#)
```
```{r}
#names(daily_results)
#names(summary_results)
```
Duplicate baseline for the three intensity panels

```{r}
# duplicate baseline so appears in all 3 panels

# extract basline peak infecitous values
baseline_peak <- subset(
  peak_infectious_plot,
  Method == "Baseline"
)

# Duplicate baseline into each intensity panel (3 copies)
baseline_peak <- baseline_peak[
  rep(1:nrow(baseline_peak), 3),
]

# assign intensity labels to copies
baseline_peak$Intensity <- rep(
  c("High", "Medium", "All"),
  each = nrow(baseline_peak) / 3
)

# keep baseline label
baseline_peak$Method <- "Baseline"
```
```{r}
# remove original baseline and replace with 3 duplicated baseline versions
peak_infectious_plot <- rbind(
  subset(peak_infectious_plot, Method != "Baseline"),
  baseline_peak
)
```
Set plotting order

```{r}
# set plotting intensity order
peak_infectious_plot$Intensity <- factor(
  peak_infectious_plot$Intensity,
  levels = c("High", "Medium", "All")
)

# sets method order
peak_infectious_plot$Method <- factor(
  peak_infectious_plot$Method,
  levels = c("Baseline", 
             "Random control (OD matched)",
             "Out-degree",
             "Random control (BT matched)",
             "Betweenness"
             )
)
```
Final boxplot

```{r}
# plot
boxplot_PI <- ggplot(
  peak_infectious_plot,
  aes(
    x = Method, # one boxplot per intervention/control method
    y = peak_infectious, # distirbution of peak infectious farms
    fill = Method # each method gets own colour
  )
) +
  geom_boxplot() +
  facet_wrap(~Intensity,
             labeller = as_labeller(c(
               High = "Highly persistent",
               Medium = "Persistent",
               All = "Transient"
             ))
             ) + #creates 3 panels by intensity
  labs(
    x = "Intervention strategy",
    y = "Peak number of infectious farms"
  )+
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

boxplot_PI

ggsave(
  filename = "peak_infectious_boxplot.png",
  plot = boxplot_PI,
  width = 8,
  height = 4,
  units = "in",
  dpi = 300
)
```

```{r}
unique(peak_infectious_established$Method)
```
```{r}
unique(peak_infectious_established$Intensity)
```
```{r}
table(
  peak_infectious_established$Method,
  useNA = "ifany"
)
```
```{r}
peak_infectious_summary <- peak_infectious_stats %>%
  group_by(Scenario) %>%
  summarise(
    N_established = n(),
    Mean = mean(peak_infectious),
    Variance = var(peak_infectious),
    Median = median(peak_infectious),
    Q1 = quantile(peak_infectious, 0.25),
    Q3 = quantile(peak_infectious, 0.75),
    Min = min(peak_infectious),
    Max = max(peak_infectious),
    .groups = "drop"
  )

peak_infectious_summary

# N.B. the differning number of N_established in controls will not bias the mean etc (the random control is likely a more precise estimate)
```

Non parametric distribution/densities - stat test choices
Show densitiies non-parametric

```{r}
# density plots
ggplot(peak_infectious_plot,
       aes(x = peak_infectious,
           colour = Method,
           fill = Method)) +
  geom_density(alpha = 0.2) +
  facet_wrap(~Intensity) +
  labs(
    x = "Peak infectious size (established outbreaks only)",
    y = "Density"
  ) +
  theme_bw()
```

 KS TEST STATS
 1. summary statistics
 
```{r}
 peak_infectious_summary <- peak_infectious_stats %>%
  group_by(Scenario) %>%
  summarise(
    N_established = n(),
    Median = median(peak_infectious),
    P95 = quantile(peak_infectious, 0.95),
    .groups = "drop"
  )

peak_infectious_summary
```
```{r}
write.csv(
  peak_infectious_summary,
  "peak_infectious_summary.csv",
  row.names = FALSE
)
```
 
 2. ECDF plots: targeted intervetion vs random control 
 
```{r}
scenario_labels <- c(
  "A",
  "B",
  "C",
  "D",
  "E",
  "F"
)

ecdf_peak_data <- do.call(
  rbind,
  lapply(seq_along(pairs), function(i) {
    
    intervention <- pairs[[i]][1]
    control <- pairs[[i]][2]
    
    data.frame(
      Peak_infectious = c(
        peak_infectious_established$peak_infectious[
          peak_infectious_established$Scenario == intervention
        ],
        peak_infectious_established$peak_infectious[
          peak_infectious_established$Scenario == control
        ]
      ),
      Scenario = c(
        rep(
          "Targeted intervention",
          sum(peak_infectious_established$Scenario == intervention)
        ),
        rep(
          "Matched random control",
          sum(peak_infectious_established$Scenario == control)
        )
      ),
      Comparison = scenario_labels[i]
    )
  })
)
```
```{r}
ecdf_peak <- ggplot(
  ecdf_peak_data,
  aes(
    x = Peak_infectious,
    colour = Scenario
  )
) +
  stat_ecdf(linewidth = 1) +
  facet_wrap(~Comparison, ncol = 3) +
  labs(
    title = "Peak number of infectious farms by intervention strategy",
    x = "Peak number of infectious farms (established outbreaks)",
    y = "Cumulative proportion (established outbreaks)",
    colour = NULL
  ) +
  theme_bw()

ecdf_peak
```
```{r}
ggsave(
  "ECDF_peak_infectious_intervention_vs_control.png",
  ecdf_peak,
  width = 8,
  height = 4,
  units = "in",
  dpi = 300
)
```
 
 3. KS tests: same pairwise ocmparisons
```{r}
run_ks_peak <- function(group1, group2) {
  
  x1 <- peak_infectious_established$peak_infectious[
    peak_infectious_established$Scenario == group1
  ]
  
  x2 <- peak_infectious_established$peak_infectious[
    peak_infectious_established$Scenario == group2
  ]
  
  test <- ks.test(
    x1,
    x2,
    alternative = "two.sided",
    exact = FALSE
  )
  
  c(
    KS_statistic = unname(test$statistic),
    P_value = test$p.value
  )
}
```
```{r}
# 6 intervention vs matched- control comparisons
ks_peak_intervention_control <- do.call(
  rbind,
  lapply(pairs, function(x) {
    run_ks_peak(x[1], x[2])
  })
)

ks_peak_intervention_control
```
```{r}
# 6 random control v baseline
ks_peak_control_baseline <- do.call(
  rbind,
  lapply(random_controls, function(x) {
    run_ks_peak(x, "Baseline")
  })
)

ks_peak_control_baseline
```
```{r}
# combine
all_ks_peak_tests <- rbind(
  
  do.call(rbind, lapply(pairs, function(x) {
    
    result <- run_ks_peak(x[1], x[2])
    
    data.frame(
      Comparison = paste(x[1], "vs", x[2]),
      KS_statistic = result["KS_statistic"],
      P_value = result["P_value"]
    )
  })),
  
  do.call(rbind, lapply(random_controls, function(x) {
    
    result <- run_ks_peak(x, "Baseline")
    
    data.frame(
      Comparison = paste(x, "vs Baseline"),
      KS_statistic = result["KS_statistic"],
      P_value = result["P_value"]
    )
  }))
)

all_ks_peak_tests$KS_statistic <- round(
  as.numeric(all_ks_peak_tests$KS_statistic), 3
)

all_ks_peak_tests$P_value <- signif(
  as.numeric(all_ks_peak_tests$P_value), 3
)

all_ks_peak_tests
```
```{r}
write.csv(
  all_ks_peak_tests,
  "result3_all_ks_peak_KS_tests.csv",
  row.names = FALSE
)
```
