---
title: "RQ2 SOIR SIMS"
author: "ESINCLAIR"
date: "2026-07-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
RQ2 – epidemiological consequences: do structurally important farms actually matter for disease spread - How do structurally identified  superspreader candidate farms influence the scale and dynamics of simulated  bTB-like disease spread compared with random introductions and targeted removal scenarios. - i.e. do structurally important farms acutally matter for disease spread?

# Scenarios - static network analysis - identified farms consistently high-risk - persistence graph across all 12 months - and run SOIR from day 1 - ?are farms that are consistently structurally important across the whole study period better targets for intervention?

Use data from RQ1 to define candidate farms - persistent high OD/BT farms. All simulations start same day, same movement period, same B, same SOIR parameters, same number of stochastic replicates

Baseline network - no removal

Remove top 5% out-degree farms

Remove top 5% Betweeness farms

Remove top 5% that appear in both BT & OD

Remove random X farms (control)

Sensitivity analysis for intro at 1 farm vs intro at 5-10 farms - infection load?

For each scenario:

  run 100-1000? stochastic replicates
  record outcomes - final infected farms, peak infectious farms, time to peak, outbreak duration, epidemic plot



```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(pbapply)
```

```{r}
# Weighted SOIR model: beta is a per-animal transmission probability, so larger cattle batches have a higher probability of transmitting infection - see graph showing relationship.

run_soir_weighted_RQ2 <- function(
    movements,
    seed_farms, # can seed 1+ farms in one run
    beta = 0.005, # beta = 0.005, # per-animal transmission probability (varied in sensitivity analysis)
    occult_period = 120, # fixed from un-weighted sensitivity analysis
    infectious_period = 180 # fixed from un-weighted sensitivity analysis
){
  
# Build look-up table - pulls every unique farm ID from both the source and destination columns of movements  - combines, removes duplicates, sorts --> (farms) is a clean look-up table of all farms involved in the dataset
  
  farms <- sort(unique(c(
    movements$source,
    movements$destination
  )))
  
# Creates a numeric index for the farms, labels each with a number corresponding to the farm ID
  n_farms <- length(farms)

  farm_index <- seq_len(n_farms)
  names(farm_index) <- farms
  
  # ensures ID types match from different 'movements' datasets
  movements$source <- as.character(movements$source)
  movements$destination <- as.character(movements$destination)

# Convert movements to integer indices #

  movements$source_id <- farm_index[movements$source]
  movements$dest_id   <- farm_index[movements$destination]
  
  # debug line
  #print(sum(is.na(movements$source_id)))
  #print(sum(is.na(movements$dest_id)))
  
  # debug line
  #print(head(names(farm_index)))
  #print(head(movements$source))
  #print(head(movements$destination))

# Set up vectors to track each farms disease status and how long it has been in its current state
  
  state <- integer(n_farms) # creates a vector of zeros - meaning all farms start susceptible

  days <- integer(n_farms) # creates a parallel vector to count how many days each farm has spent in its current state

# Seed infection - infect chosen seed farm by setting its state to 1 - farms to start as occult - representative of dz introduction
# Reset their disease timers to 0 days, indicating they have just entered the occult state.

  state[farm_index[seed_farms]] <- 1 # state[farm_index[seed_farms]] <- 1; change seeded farms from susceptible (0) to occult (1)

  days[farm_index[seed_farms]] <- 0 # for the same seeded farm(s), set no. days spent in current dz state to 0 - farm has just become occult at the start of the simulation

# Storage - prepares storage for the simulation results

  max_day <- max(movements$day) # finds last day in the movement data set

# History - creates a data frame with one row per day, and columns for the number of farms in each SOIR state, all columns are pre-allocated as integer vectors of length max_day for filling during the simulation
  
  history <- data.frame( 
    day = integer(max_day), 
    S = integer(max_day),
    O = integer(max_day),
    I = integer(max_day),
    R = integer(max_day)
  )


# Simulation - advances farms through the O -> I -> R stages as each day of the simulation passes
# The 'for' loop runs one iteration per day

  for(day in seq_len(max_day)){


# Progress occult farms:
# The day counters increase, and farms that have spent more than/equal to occult_period days in O become infectious (state == 2) and their day counters reset
    
    occult <- which(state == 1)

    if(length(occult) > 0){

      days[occult] <- days[occult] + 1 # increments until reaches occult_period -> becomes infectious

      become_I <- occult[
        days[occult] >= occult_period
      ]

      state[become_I] <- 2
      days[become_I] <- 0

    }

# Progress infectious farms

    infectious <- which(state == 2)

# Their day counters increase, any farms that have spent more than infectious_period i days in I become removed (state == 3) and their day counters reset    
    
    if(length(infectious) > 0){

      days[infectious] <- days[infectious] + 1

      become_R <- infectious[
        days[infectious] >= infectious_period
      ]

      state[become_R] <- 3
      days[become_R] <- 0

    }

# Today's movements - filters the 'movements' data frame to keep just the rows where 'day' matches the current loop day, giving the set of movements to process for that day.

    todays <- movements[movements$day == day, ]

    
# Process each movement event. Transmission is only considered from infectious source farms to susceptible destination farms.


    
# Processes each movement event and applies infection-spread rules.
# loop over all movements happening today, for each movement, get the source and destination farm indices; skip the movement if the source is not infectious (state != 2); skip the movement if the destination is not susceptible (state != 0). Only movements from the infectious -> susceptible farms proceed to the transmission step that follows.

    if(nrow(todays) > 0){

      for(i in seq_len(nrow(todays))){

        source <- todays$source_id[i]
        dest   <- todays$dest_id[i]

        # source not infectious
        if(state[source] != 2)
          next

        # destination already infected
        if(state[dest] != 0)
          next

        
# Transmission block:
# Calculate the movement-level transmission probability from the batch size.
# Each animal has an independent probability (beta) of initiating infection

# Stochastic transmission event:
# runif(1) generates a random number between 0 and 1.
# If the random value is less than p_transmit, the movement event results in transmission,
# and the destination farm changes from susceptible (S) to occult (O).
# This represents a Bernoulli trial where transmission either occurs (1) or does not occur (0).
# p_transmit is the batch-level probability of infection, derived from the per-animal
# transmission probability (beta) and number of cattle moved.
#https://www.rdocumentation.org/packages/stats/versions/3.5.2/topics/Uniform, Bernoulli

        # beta = individual animal-level probability
        # p_transmit = whole batch movement level probability
       
        batch_size <- todays$total_cattle[i] # no. cattle moved in this movement event
        p_transmit <- 1 - (1 - beta)^batch_size # Probability that at least one animal in the batch initiates infection.
        
        if(runif(1) < p_transmit){ # raw a random number between 0 and 1; if it is less than p_transmit, infection occurs.

          state[dest] <- 1
          days[dest] <- 0

        }

      }

    }

# ? eventually include the probability that a moved animal is actually infectious - not neccesserially important for research question

# Save epidemic - logs the SOIR count for the current simulation day - store the day number; count how many farms are S,O,I,R and write those values into the corresponding row of history
    
    history$day[day] <- day
    history$S[day] <- sum(state == 0)
    history$O[day] <- sum(state == 1)
    history$I[day] <- sum(state == 2)
    history$R[day] <- sum(state == 3)

  }

# Return - returns the simulation output as a structured list

  list(

    history = history, # stores the full day-by-day SOIR counts

    final_states = data.frame( # data frame listing each farm and its final SOIR state, converts numeric codes (0-3) into labelled factors S,O,I,R

      farm = farms,

      state = factor(
        state,
        levels = 0:3,
        labels = c("S","O","I","R")
      )

    )

  )

}
```

# Provide movement data
```{r}
# read in csv

movements_raw <- read.csv(
    "RQ1 NO EP.csv",
    stringsAsFactors = FALSE
)

#head(movements_raw)
```

```{r}
# create simulation movement  table

movements <- movements_raw %>%
    select(
        source = org_id,
        destination = dest_id,
        day,
        total_cattle
    )
#head(movements)
```

```{r}
str(movements)
```

# Visual of probability x batch size
https://ggplot2.tidyverse.org/articles/ggplot2.html
```{r}

beta <- 0.005

df <- data.frame(
  batch_size = 1:800
)

df$p_transmit <- 1 - (1 - beta)^df$batch_size


ggplot(df, aes(x = batch_size, y = p_transmit)) +
  geom_line(
    colour = "firebrick",
    linewidth = 1.2
  ) +
  labs(
    x = "Cattle moved per batch",
    y = "Probability of transmission",
    title = paste0(
      "Effect of batch size on transmission probability (β = ",
      beta,
      ")"
    )
  ) +
  theme_classic()
```
```{r}
# N.B. without 'scenarios... replicates - this is only running ONCE
result_weighted_RQ2 <- run_soir_weighted_RQ2(
    movements,
    seed_farms = c("95")
)

head(result_weighted_RQ2$history)
```
```{r}
max(result_weighted_RQ2$history$I)
```
```{r}
# N.B.this is just one run
result_weighted_RQ2$history %>%
    tidyr::pivot_longer(
        cols = c(S,O,I,R),
        names_to = "state",
        values_to = "farms"
    ) %>%
    ggplot(
        aes(
            x = day,
            y = farms,
            colour = state
        )
    ) +
    geom_line(size = 1) +
    theme_minimal()
```
Epidemic curve- infected per day
```{r}

ggplot(result_weighted_RQ2$history,
       aes(x = day, y = I)) +
  geom_line(
    colour = "firebrick",
    linewidth = 1.2
  ) +
  labs(
    x = "Day",
    y = "Number of infectious farms",
    title = "Epidemic curve"
  ) +
  theme_classic()
```
Create workflow to include replicate - modified from sensitivity analysis - COPY AND PASTES BELOW


MODIFIED VERSION FOR BASELINE MODEL
```{r}
# baseline model parameters - as per  model/sensitivity analysis
#baseline_b <- 0.005
#basline_occult <- 120
#baseline_infect <- 180

# Stochastic - randomly selected seed farm & randomness inside SOIR model = 

# Creates dataframe with 10 simulation replicates (10 stochastic replicates - test run)

baseline_scenario_grid <- data.frame(
  replicate = 1:10
)

# Defines function that runs ONE stochastic epidemic simulation for one row (one replicate) and returns daily & summary outputs. The 10 replicates come from repeatedly calling the function using the rows of baseline_scenario_grid.
run_baseline <- function(scenario_row){
  
  # scenario_row exists solely as the functions input parameter - the function is intended to receive one row from the scenario grid. Scenario_row & x refert to the same object
  
  #print(class(x))
  #print(x)
  
  #Randomly select one seed farm from all farms in movement data
  seed_farm <- sample(unique(c(movements$source, movements$destination)),
                      1
                      )
  # Run SOIR model
  result <- run_soir_weighted_RQ2(
    movements = movements,
    seed_farms = seed_farm,
    beta = 0.005,
    occult_period = 120,
    infectious_period = 180
    )
  
  # Create daily output - stores epidemic count for each day. result$history$S etc all generated by the run_soir_weighted_RQ2() function
  daily_output <- data.frame(
    replicate = scenario_row$replicate,
    day = result$history$day,
    susceptible = result$history$S,
    infectious = result$history$I,
    occult = result$history$O,
    removed = result$history$R,
    total_infected = result$history$O +
      result$history$I + 
      result$history$R
  )
  
  # Create summary outputs - records key statistics for the simulation
  summary_output <- data.frame(
    #scenario = "baseline/scenario name", # change input depending on scenario - keep for stat tests later all in one dataframe
    replicate = scenario_row$replicate,
    seed_farm = seed_farm,
    final_infected = sum(result$final_states$state!= "S"),
    peak_infectious = max(result$history$I)
  )
  
  # Return both outputs
  list(
    daily = daily_output, # time series epidemic data
    summary = summary_output # overall simulation stats
  )
}
```

```{r}
# runs the run_baseline() function once for each row of baseline_scenario_grid and stores all the results
baseline_results <- pblapply(
  seq_len(nrow(baseline_scenario_grid)), # generates sequence 1,2,...10
  function(i){ # defines a temporary function that takes the row number 'i'
    run_baseline(baseline_scenario_grid[i, , drop = FALSE]) # extracts the 'ith row of the baseline_scenario_grid, drop=FALSE ensures it remains a data frame rather than being simplified to a vector - passes that row into run_baseline() as scenario_row
  }
)

# iterations are:
# run_baseline(baseline_scenario_grid[1, ]), repeated to [10, ]. 
# each call runs one stochastic epidemic simulation, the returned list (daily & summary) is stored in baseline results.
```


```{r}
# combine the daily outputs from all 10 sim replicates into one data frame

# daily result - epidemic curves for all 10 stochastic runs
daily_results <- bind_rows( # bind_rows (dpylr) stacks multiple data frames vertically into single data frame
  lapply(baseline_results, function(x) x$daily) # loops through each element of baseline_results, extracts only the 'daily' data frame from each replicate
)
# peak infectious and final outbreak size per run
summary_results <- bind_rows(
  lapply(baseline_results, function(x) x$summary)
)
```
```{r}
daily_results
```
```{r}
summary_results
```

BotMEW paper

1. individual stochastic runs plot (example) - shows uncertainty and stochastic variation - demonstrates model variability. Each simulation starts with a randomly selected seed farm and includes random transmission events (runif(1) < p_transmit) - so even with identical parameters, outbreaks will differ.

https://pdodds.w3.uvm.edu/files/papers/others/2001/mackenzie2001.pdf

```{r}
# 1. plot individual stochastic runs (to show variability)
ggplot(
  daily_results,
  aes(x = day, y = infectious, group = replicate)
) +
  geom_line(alpha = 0.3) +
  labs(
    x = "Day",
    y = "Infectious farms",
    title = "Stochastic epidemic trajectories"
  ) +
  theme_classic()
```

2. Mean +/- SD (uncertainty) epidemic curve - summarises the expected epidemic trajectory across all stochastic replicates (shows expected outbreak dynamics) - calculate 'mean infectious farms on dat 't'' across all simulations. Shows average outbreak progression, when epidemic peaks, how much variability exists around the mean - represents thebehvaiour of the model rather than one random simulation - the mean epidemic curve was calculated across stochastic replicates, with variability represented by the standard deviation of infectious farm numbers at each time step.

https://www.sciencedirect.com/science/article/pii/S0167587797000020

```{r}
# 2. average number of infectious farms per day
mean_epidemic <- daily_results %>%
  group_by(day) %>%
  summarise(
    mean_infectious = mean(infectious),
    sd_infectious = sd(infectious)
  )

ggplot(mean_epidemic, aes(x = day, y = mean_infectious)) + geom_line(linewidth = 1.2, colour = "firebrick") + labs(
  x = "Day",
  y = "Mean epidemic curve across stochastic simulations"
) +
  theme_classic()
```

```{r}
# 2. add uncertainty
ggplot(mean_epidemic, aes(x = day, y = mean_infectious)) +
  geom_ribbon(
    aes(
      ymin = mean_infectious - sd_infectious,
      ymax = mean_infectious + sd_infectious
    ),
    alpha = 0.3,
    fill = "firebrick"
  ) +
  geom_line(
    linewidth = 1.2,
    colour = "firebrick"
  ) +
  labs(
    x = "Day",
    y = "Infectious farms",
    title = "Epidemic curve (mean ± SD)"
  ) +
  theme_classic()
```

3. Final outbreak size distribution - compares scenario impact

Uses output 'summary_results$final_infected - contains the number of farms in each replicate
Shows the probability distribution of outbreak sizes, whether outbreaks are usually small or occasionally large, whether changing model parameters changes epidemic risk

https://www.sciencedirect.com/science/article/pii/S0167587705000358

```{r}
ggplot(
  summary_results,
  aes(x = final_infected)
) +
  geom_density(
    fill = "steelblue",
    alpha = 0.4
  ) + 
  labs(x = "Final infected farms",
       y = "Density",
       title = "Distribution of final outbreak size"
       ) + 
  theme_classic()
```

4. 
To get line graph showing each compartment- need a mean of stochastic replications - to show mean trajectory across replicates

Need to create baseline_history containing day_by_day history from all 10 runs (with a replciate column) - then can create plot - currently only saved the summary statistics- need the history from each replicate to allow day-by-day epidemic plot

https://www.sciencedirect.com/science/article/pii/S1755436510000460

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

mean_SOIR <- daily_results %>%
  group_by(day) %>%
  summarise(
    mean_S = mean(susceptible),
    mean_O = mean(occult),
    mean_I = mean(infectious),
    mean_R = mean(removed)
  ) %>%
  pivot_longer(
    cols = c(mean_S, mean_O, mean_I, mean_R),
    names_to = "compartment",
    values_to = "mean_farms"
  )
```
```{r}
ggplot(
  mean_SOIR,
  aes(
    x = day,
    y = mean_farms,
    colour = compartment
  )
)+
  geom_line(
    linewidth = 1.2
  )+
  labs(
    x = "Day",
    y = "Mean number of farms",
    colour = "Disease state",
    title = "Mean SOIR epidemic dynamics across stochastic simulations"
  )+
  theme_classic()
```
5. Stat significance of different scenarios

For comparing scenarios (e.g. different seed farms, removing nodes, changing network structure), need 

(i) multiple stochastic replicates per scenario
(ii) summary outcome per replicate:
      final_infected (total outbreak size)
      peak_infectious
      ?epidemic duration
      ?cumulative infected over time
      
E.g. compare two scenarios by final outbreak size:
scenario A - baseline network
scenario B - network with x type of nodes removed

Would have - baseline_summary & removed_nodes_summary

Then - t-test? but if outbreak size is not normally distributed - likely (stochastic epidemics usually have - many small outbreaks/fewer larger outbreaks) - non-parametric test may be more appropriate - e.g. wilcox test

  wilcox.test(baseline_summary%final_infected, removed_nodes_summary$final_infected)

If compare several interventions - use ?anova? Kruskal-Wallis test?

COMPARE

  primary metric - final infected = overall epidemic impact
  secondary metric - peak infectious = maximum disease pressure & time to peak = epidemic speed

# Baseline with 50/100/300/500/1000/2000 runs - analysis to establish replicates

compare summary stats - 
Mean final outbreak size/mean epidemic duration - check whether stabalise - do estimates change  very little when add more replicates - plot cumalitive mean against number of runs

```{r}
replicate_tests <- c(10, 50, 100, 300, 500, 1000, 2000)

results <- list()

for (n in replicate_tests) {
  
  baseline_scenario_grid <- data.frame(
    replicate = 1:n
  )

# pblapply() calls run-baseline() function created earlier, repeatedly (contains the SOIR model)
  
  baseline_results <- pblapply(
    seq_len(nrow(baseline_scenario_grid)),
    function(i){
      run_baseline(baseline_scenario_grid[i, , drop = FALSE])
    }
  )

# just extract summary results because want to see mean(final_infected)  
    
  summary_results <- bind_rows(
    lapply(baseline_results, function(x) x$summary)
  )

# store mean final outbreak size for each tested number of replicates; converts 'n' to character to label mean_final_infected
  #results[[as.character(n)]] <- mean(summary_results$final_infected)
  results[[as.character(n)]] <- data.frame(
  replicates = n,
  mean_final = mean(summary_results$final_infected),
  median_final = median(summary_results$final_infected),
  prob_spread = mean(summary_results$final_infected > 1),
  p95_final = as.numeric(quantile(summary_results$final_infected, 0.95))
)
}
```
```{r}
# combine results
stability_results <- bind_rows(results)

stability_results
```
Estimates are stable around 1000-2000 replicates
mean final infected/probabiilty of spread/95th percentile all little/no change between 1000 & 2000 - suggests main summary stats have converged.

Median stays at 1 accross all replicates - indicates most common outcome is no onward spread beyond the initial farm, while occasional larger outbreaks increase the mean.

The number of stochastic simulations were determiined by assessing convergence of outbreak summary statistics. Estimates of mean outbreak size, probability of spread, and the 95th percentile outbreak size (value below which 95% of simulation outcomes fall - e.g. 95% of simulated outbreaks infected 11 farms of fewer, only the largest 5% of outbreaks exceeded 11 farms) stabalised between 1000 & 2000 replicates, 2000 replicates is seleceted for subsequent scenrario analysis.


```{r}
results_df <- bind_rows(results)

results_df
```

```{r}
ggplot(results_df, aes(x = replicates, y = mean_final)) +
  geom_line() +
  geom_point() +
  labs(
    x = "Number of stochastic replicates",
    y = "Mean final infected farms",
    title = "Stabilisation of outbreak size estimate"
  ) +
  theme_classic()
```
```{r}
# Probability of spread
ggplot(results_df, aes(x = replicates, y = prob_spread)) +
  geom_line() +
  geom_point() +
  labs(
    x = "Number of stochastic replicates",
    y = "Probability of spread"
  ) +
  theme_classic()
```

```{r}
# 95th percentile
ggplot(results_df, aes(x = replicates, y = p95_final)) +
  geom_line() +
  geom_point() +
  labs(
    x = "Number of stochastic replicates",
    y = "95th percentile final infected farms"
  ) +
  theme_classic()
```

# SCENARIOS

Now have choosen 2000 replicates 
For each scenario - final_infected (total outbreak size); peak_infectious (max number infectious at one time); probability of spread (proportion where final_infected > 1); p95_final (larger outbreak behaviour)

1. Run baseline scenario() 2000 times & store daily epidemic outputs (daily_results) & summary statistics (summary_results)
2. Intervention/scenario analyses - apply changes - create modified movement datasets - remove top 5% OD & BT (identified in network analyses); run modified model for 2000 replicates. compare against the baseline using the same summary metrics

?If farms that repeatedly emerge as structurally important are removed, does epidemic size and spread probability decrease compared with random introductions in the unmodified network??

Need to justify what threshold for monthly persistence will use - do across range of persistence- show how much intervention is required to reduce epidemic risk. With 'all identified candidates even if only appearing in 1-2 months' - essentially answering that a farm that only appears in 1-2 months is not representing a true SS in the network - interpret as - what is the maximum effect of targeting all farms that were ever structurally imporant.

When fill below table, will answer RQ2 hypothesis test - does structural node removal affect epidemic outcome? 
AND
Does the definition of superspreader matter - using the persistence thresholds test how sensitive the result is to the definition of superspreader



Read in data from RQ1
```{r}
# OUTDEGREE
OD_all_candidates <- readRDS(
  "outdegree_all_candidates.rds"
)

OD_med_candidates <- readRDS(
  "outdegree_med_persistent.rds"
)

OD_high_candidates <- readRDS(
  "outdegree_high_persistent.rds"
)
```

```{r}
head(OD_all_candidates)
```

```{r}
# BETWEENNESS
BT_all_candidates <- readRDS(
  "BT_any_persistent.rds"
)

BT_high_candidates <- readRDS(
  "BT_high_persistent.rds"
)

BT_med_candidates <- readRDS(
  "BT_med_persistent.rds"
)
```

```{r}
# COMBINED BETWEENNESS & OUTDEGREE
CB_all_candidates <- readRDS(
  "Combined_all_persistent.rds"
)

CB_high_candidates <- readRDS(
  "Combined_high_persistent.rds"
)

CB_med_candidates <- readRDS(
  "Combined_med_persistent.rds"
)
```

```{r}
head(CB_high_candidates)
class(CB_high_candidates)
```

```{r}
# removed lengths - these are stored as character/lists not table so length appropriate
length(OD_all_candidates)
length(OD_high_candidates)
length(OD_med_candidates)
length(BT_all_candidates)
length(BT_high_candidates)
length(BT_med_candidates)
length(CB_all_candidates)
length(CB_med_candidates)
length(CB_high_candidates)
```
Create Scenario Table
```{r}
scenario_table <- data.frame(
  Scenario = c(
    "(1) Baseline",
    "(2) Outdegree_>=4",
    "(3) Outdegree_>=8",
    "(4) Outdegree_all",
    "(5) Betweenness_>=4",
    "(6) Betweenness_>=6",
    "(7) Betweenness_all",
    "(8) Combined_>=6",
    "(9) Combined_>=8",
    "(10) Combined_all"
  ),
  Target_farms_removed = c(
    "None",
    "Outdegree medium candidates",
    "Outdegree high candidates",
    "Outdegree any candidates",
    "Betweenness medium candidates",
    "Betweenness high candidates",
    "Betweenness all candidates",
    "Overlap medium candidates",
    "Overlap high candidates",
    "Overlap any candidates"
  ),
  Persistence_threshold = c(
    "None",
    ">=6 months",
    ">=8 months",
    "Any month",
    ">=4 months",
    ">=6 months",
    "Any month",
    "BT >=4 & OD >=6 months",
    "BT >= 6 & OD>= 8 months",
    "Any month"
  ),
  Number_of_farms_removed = c(
    0,
    length(OD_med_candidates),
    length(OD_high_candidates),
    length(OD_all_candidates),
    length(BT_med_candidates),
    length(BT_high_candidates),
    length(BT_all_candidates),
    length(CB_med_candidates),
    length(CB_high_candidates),
    length(CB_all_candidates)
  )
  %_of_all_farms
)

scenario_table
```

# Modify function for scenarios

Scenarios:

   (1) Baseline
  
  (2) Outdegree >= 8 months
  (3) Outdegree >= 6 months
  (4) Outdegree any month
  
  (5) Betweenness >= 6 months
  (6) Betweenness >= 4 months
  (7) Betweenness any month
  
  (8) Combined BT >= 6 months & OD >= 8 months
  (9) Combined BT>= 4 months & 0D >= 6 months
  (10) Combined any month
  
  CONTROLS - If I remove the same number of farms, is targeted removal (SC2-10) more effective than random removal?
  (11) random X farms (average of scenarios 2/5/8) = 9.33 (9)
  (12) random X farms (average of scenarios 3/6/9) = 4.33 (4)
  (13) random X farms (average of scenarios 4/7/10) = 40.33 (40)

Persistence thresholds were selected based on temporal persistence distributions of each centrality measure. Outdegree hubs emerged throughout the study period of network analysis, whereas betweeness farms did not emerge until approximately month 6 (when cow-calf started trading). Consequently, different persistence thresholds were used to identify sustained structural importance of each metric.

N.B. hierarchal network (usualy for livestock) - until cow-calf starts to trade, two levels and high outdegree farms present/informative of structural SS characteristics; betweenness develops with cow-calf trading. I.e. outdegree heterogeneity emerges early; betweenness heterogeneity emerges only once cow-calf layer developes

Create new movement datasets with excluded farms

```{r}
# baseline - no nodes removed
scenario_1 <- movements
```
```{r}
dim(scenario_1)
length(unique(c(scenario_1$source, scenario_1$destination))) # number of nodes
```
```{r}
# Outdegree >= 8 months
scenario_2 <- movements %>%
  filter(
    !(source %in% OD_high_candidates),
    !(destination %in% OD_high_candidates)
  )
```
```{r}
# debug
sum(is.na(scenario_2$source))
sum(is.na(scenario_2$destination))
```
```{r}
# check worked & df
any(OD_high_candidates %in% scenario_2$source)
any(OD_high_candidates %in% scenario_2$destination)
class(scenario_2)
```
```{r}
# Outdegree >= 6 months
scenario_3 <- movements %>%
  filter(
    !(source %in% OD_med_candidates),
    !(destination %in% OD_med_candidates)
  )
```
```{r}
# check worked & df
any(OD_med_candidates %in% scenario_3$source)
any(OD_med_candidates %in% scenario_3$destination)
class(scenario_3)
```
```{r}
# Outdegree any months
scenario_4 <- movements %>%
  filter(
    !(source %in% OD_all_candidates),
    !(destination %in% OD_all_candidates)
  )
```
```{r}
# check worked & df
any(OD_all_candidates %in% scenario_4$source)
any(OD_all_candidates %in% scenario_4$destination)
class(scenario_4)
```
```{r}
# Betweeness >= 6 months
scenario_5 <- movements %>%
  filter(
    !(source %in% BT_high_candidates),
    !(destination %in% BT_high_candidates)
  )
```
```{r}
# check worked & df
any(BT_high_candidates %in% scenario_5$source)
any(BT_high_candidates %in% scenario_5$destination)
class(scenario_5)
```
```{r}
# Betweeness >= 4 months
scenario_6 <- movements %>%
  filter(
    !(source %in% BT_med_candidates),
    !(destination %in% BT_med_candidates)
  )
```
```{r}
# check worked & df
any(BT_med_candidates %in% scenario_6$source)
any(BT_med_candidates %in% scenario_6$destination)
class(scenario_6)
```
```{r}
# Betweeness any months
scenario_7 <- movements %>%
  filter(
    !(source %in% BT_all_candidates),
    !(destination %in% BT_all_candidates)
  )
```
```{r}
# check worked & df
any(BT_all_candidates %in% scenario_7$source)
any(BT_all_candidates %in% scenario_7$destination)
class(scenario_7)
```
```{r}
# Combined BT >= 6 months & OD >= 8 months
scenario_8 <- movements %>%
  filter(
    !(source %in% CB_high_candidates),
    !(destination %in% CB_high_candidates)
  )
```
```{r}
# check worked & df
any(CB_high_candidates %in% scenario_8$source)
any(CB_high_candidates %in% scenario_8$destination)
class(scenario_8)
length(scenario_8)
```
```{r}
# Combined BT>= 4 months & 0D >= 6 months
scenario_9 <- movements %>%
  filter(
    !(source %in% CB_med_candidates),
    !(destination %in% CB_med_candidates)
  )
```
```{r}
# check worked & df
any(CB_med_candidates %in% scenario_9$source)
any(CB_med_candidates %in% scenario_9$destination)
class(scenario_9)
dim(scenario_9)
```
```{r}
# Combined any month
scenario_10 <- movements %>%
  filter(
    !(source %in% CB_all_candidates),
    !(destination %in% CB_all_candidates)
  )
```
```{r}
# check worked & df
any(CB_all_candidates %in% scenario_10$source)
any(CB_all_candidates %in% scenario_10$destination)
class(scenario_10)
```

```{r}
# control 1 (high)
# Scenario 11 - random removal of 4 farms

set.seed(11) # Set random seed to ensure the same random farms can be selected in future runs.

all_farms <- unique(c(movements$source, movements$destination))

random_4 <- sample(all_farms, 4) 

scenario_11 <- movements %>%
  filter(
    !(source %in% random_4),
    !(destination %in% random_4)
  )
```
```{r}
# check worked & df
any(random_4 %in% scenario_11$source)
any(random_4 %in% scenario_11$destination)
class(scenario_11)
dim(scenario_11) # - number of rows (movements)
length(unique(c(scenario_11$source, scenario_11$destination))) # number of nodes
```

```{r}
# control 2 (med)
# Scenario 12 - random removal of 9 farms

set.seed(12) # Set random seed to ensure the same random farms can be selected in future runs.

random_9 <- sample(all_farms, 9)

scenario_12 <- movements %>%
  filter(
    !(source %in% random_9),
    !(destination %in% random_9)
  )
```
```{r}
# check worked & df
any(random_9 %in% scenario_12$source)
any(random_9 %in% scenario_12$destination)
class(scenario_12)

dim(scenario_12) # - number of rows (movements)
length(unique(c(scenario_12$source, scenario_12$destination))) # number of nodes
```

```{r}
# Control 3 (all)
# Scenario 13 - random removal of 40 farms = control all

set.seed(13) # Set random seed to ensure the same random farms can be selected in future runs.

random_40 <- sample(all_farms, 40)

scenario_13 <- movements %>%
  filter(
    !(source %in% random_40),
    !(destination %in% random_40)
  )
```
```{r}
# check worked & df
any(random_40 %in% scenario_13$source)
any(random_40 %in% scenario_13$destination)
class(scenario_13)
```
```{r}
dim(scenario_13) # - number of rows (movements)
length(unique(c(scenario_13$source, scenario_13$destination))) # number of nodes
```

```{r}
# look-up table documenting which scenario each number represents
scenario_table <- data.frame(
  scenario = paste0("scenario_", 1:13),
  description = c(
    "Baseline",
    "Outdegree >= 8 months",
    "Outdegree >= 6 months",
    "Outdegree any month",
    "Betweenness >= 6 months",
    "Betweenness >= 4 months",
    "Betweenness any month",
    "Combined BT >= 6 months & OD >= 8 months",
    "Combined BT >= 4 months & OD >= 6 months",
    "Combined any month",
    "Control High",
    "Control Med",
    "Control All"
  )
)

scenario_table
```

```{r}
# modify function so can be used for all 2000 scenarios - just need to change movement_data and scenario_name when call it for each scenario - next box
scenario_grid <- data.frame(
  replicate = 1:2000
)
# pass scenario name and movement network as inputs. Function needs 3 inputs:
# 1. one row from scenario_grid (scenario_row) - dont need to chanage this
# 2. a movement dataset (movement_data)
# 3. a label for the scenario (scenario_name)
run_scenario <- function(scenario_row, movement_data, scenario_name){
  
  # Randomly select seed farm from the available network
  seed_farm <- sample(
    unique(c(movement_data$source, movement_data$destination)),
    1
  )
  
  # Run SOIR model
  result <- run_soir_weighted_RQ2(
    movements = movement_data,
    seed_farms = seed_farm,
    beta = 0.005,
    occult_period = 120,
    infectious_period = 180
  )
  
  # Daily output
  daily_output <- data.frame(
    scenario = scenario_name,
    replicate = scenario_row$replicate,
    day = result$history$day,
    susceptible = result$history$S,
    infectious = result$history$I,
    occult = result$history$O,
    removed = result$history$R,
    total_infected = result$history$O +
      result$history$I +
      result$history$R
  )
  
  # Summary output
  summary_output <- data.frame(
    scenario = scenario_name,
    replicate = scenario_row$replicate,
    seed_farm = seed_farm,
    final_infected = sum(result$final_states$state != "S"),
    peak_infectious = max(result$history$I)
  )
  
  list(
    daily = daily_output,
    summary = summary_output
  )
}
```
Run the function for each scenario - pblapply gives progress bar

  (1) Baseline
  
  (2) Outdegree >= 8 months
  (3) Outdegree >= 6 months
  (4) Outdegree any month
  
  (5) Betweenness >= 6 months
  (6) Betweenness >= 4 months
  (7) Betweenness any month
  
  (8) Combined BT >= 6 months & OD >= 8 months
  (9) Combined BT>= 4 months & 0D >= 6 months
  (10) Combined any month
  
  CONTROLS - If I remove the same number of farms, is targeted removal (SC2-10) more effective than random removal?
  (11) random X farms (average of scenarios 2/5/8) = 9.33 (9)
  (12) random X farms (average of scenarios 3/6/9) = 4.33 (4)
  (13) random X farms (average of scenarios 4/7/10) = 40.33 (40)
  
For each simulation replicate, an index farm was selected uniformly at random from the farms present in the scenario-specific movement network. Following the removal of structurally important farms in intervention scenarios, the pool of eligible seed farms therefore reflected the remaining network. This approach was chosen to represent random disease introduction while avoiding systematic bias towards particular network positions.

There are two different types of randomness (stochasticity) in the model:

  The disease seed farm - decides where the outbreak starts - randomly selected in each simulation replicate (code in SOIR)
  
  The random number generator stochasticity - the R random number generator controls the random processes in the model, including selection of seed farm (sample()) and transmission events (runif(1)). therefore, differneces between scenarios are influenced by both the change in network structure (farm removal) and random variation in disease introduction and transmission events. Because SC2 etc contains different set of farms after node remval, using identical random sequences would not result in identical outbreak pathways between scenarions because the pool of available farms has changed - so setting the seed would not create a perfectly paired comparison.

Two sources of stochasticity were incorporated in the SOIR model. First, disease introduction was stochastic, with the initial infected farm selected randomly from the farms present in each scenario-specific network. Second, transmission events were stochastic, with infection determined through random Bernoulli trials. Therefore, differences between scenarios reflect both changes to network structure following farm removal and inherent variation due to random disease introduction and transmission. A large number of replicate simulations (n = 2,000 per scenario) were used to minimise the influence of stochastic variation.

# Scenario 1 - Baseline

```{r}
# calls the function 2000 times - one for each row of scenario_grid
scenario_1_results <- pblapply(
  seq_len(nrow(scenario_grid)),
  function(i){
    run_scenario(
      scenario_grid[i, , drop = FALSE], # becomes scenario_row inside funciton
      scenario_1,# movement dataset for this scenario_X 
      "Baseline" # description of scenario - see look up table
    )
  }
  
)

saveRDS(scenario_1_results, "scenario_1_results.rds")
```

```{r}
# combine the daily outputs from all 2000 sim replicates into one data frame

# daily result - epidemic curves for all 2000 stochastic runs
daily_results_SC1 <- bind_rows( 
  lapply(scenario_1_results, function(x) x$daily) 
)
# peak infectious and final outbreak size per run
summary_results_SC1 <- bind_rows(
  lapply(scenario_1_results, function(x) x$summary)
)
```
```{r}
head(daily_results_SC1)
head(summary_results_SC1)
```

# Scenario 2 - Outdegree (high) >= 8 months

YOU ARE HERE - ERROR WHENRUNNING SIMULATION 29/07/26

```{r}
# calls the function 2000 times - one for each row of scenario_grid
scenario_2_results <- pblapply(
  seq_len(nrow(scenario_grid)),
  function(i){
    run_scenario(
      scenario_grid[i, , drop = FALSE], # becomes scenario_row inside funciton
      scenario_2,# movement dataset for this scenario_X 
      "Outdegree >= 8 months" # description of scenario - see look up table
    )
  }
)

saveRDS(scenario_2_results, "scenario_2_results.rds")
```


```{r}
# combine the daily outputs from all 2000 sim replicates into one data frame

# daily result - epidemic curves for all 2000 stochastic runs
daily_results_SC2 <- bind_rows( 
  lapply(scenario_2_results, function(x) x$daily) 
)
# peak infectious and final outbreak size per run
summary_results_SC2 <- bind_rows(
  lapply(scenario_2_results, function(x) x$summary)
)
```
```{r}
head(daily_results_SC2)
head(summary_results_SC2)
```



# Scenario 3 - Outdegree (med) >= 6 months

```{r}
# calls the function 2000 times - one for each row of scenario_grid
scenario_3_results <- pblapply(
  seq_len(nrow(scenario_grid)),
  function(i){
    run_scenario(
      scenario_grid[i, , drop = FALSE], # becomes scenario_row inside funciton
      scenario_3,# movement dataset for this scenario_X 
      "Outdegree >= 6 months" # description of scenario - see look up table
    )
  }
)

saveRDS(scenario_3_results, "scenario_3_results.rds")
```


```{r}
# combine the daily outputs from all 2000 sim replicates into one data frame

# daily result - epidemic curves for all 2000 stochastic runs
daily_results_SC3 <- bind_rows( 
  lapply(scenario_3_results, function(x) x$daily) 
)
# peak infectious and final outbreak size per run
summary_results_SC3 <- bind_rows(
  lapply(scenario_3_results, function(x) x$summary)
)
```
```{r}
head(daily_results_SC3)
head(summary_results_SC3)
```

# Scenario 4 - Outdegree any month

```{r}
# calls the function 2000 times - one for each row of scenario_grid
scenario_4_results <- pblapply(
  seq_len(nrow(scenario_grid)),
  function(i){
    run_scenario(
      scenario_grid[i, , drop = FALSE], # becomes scenario_row inside funciton
      scenario_4,# movement dataset for this scenario_X 
      "Outdegree any month" # description of scenario - see look up table
    )
  }
)

saveRDS(scenario_4_results, "scenario_4_results.rds")
```


```{r}
# combine the daily outputs from all 2000 sim replicates into one data frame

# daily result - epidemic curves for all 2000 stochastic runs
daily_results_SC4 <- bind_rows( 
  lapply(scenario_4_results, function(x) x$daily) 
)
# peak infectious and final outbreak size per run
summary_results_SC4 <- bind_rows(
  lapply(scenario_4_results, function(x) x$summary)
)
```
```{r}
head(daily_results_SC4)
head(summary_results_SC4)
```

# Scenario 5 - Betweenness (high) >= 6 months
```{r}
# calls the function 2000 times - one for each row of scenario_grid
scenario_5_results <- pblapply(
  seq_len(nrow(scenario_grid)),
  function(i){
    run_scenario(
      scenario_grid[i, , drop = FALSE], # becomes scenario_row inside funciton
      scenario_5,# movement dataset for this scenario_X 
      "Betweenness >= 6 months" # description of scenario - see look up table
    )
  }
)

saveRDS(scenario_5_results, "scenario_5_results.rds")
```


```{r}
# combine the daily outputs from all 2000 sim replicates into one data frame

# daily result - epidemic curves for all 2000 stochastic runs
daily_results_SC5 <- bind_rows( 
  lapply(scenario_5_results, function(x) x$daily) 
)
# peak infectious and final outbreak size per run
summary_results_SC5 <- bind_rows(
  lapply(scenario_5_results, function(x) x$summary)
)
```
```{r}
head(daily_results_SC5)
head(summary_results_SC5)
```

# Scenario 6 - Betweenness (med) >= 4 months

```{r}
# calls the function 2000 times - one for each row of scenario_grid
scenario_6_results <- pblapply(
  seq_len(nrow(scenario_grid)),
  function(i){
    run_scenario(
      scenario_grid[i, , drop = FALSE], # becomes scenario_row inside funciton
      scenario_6,# movement dataset for this scenario_X 
      "Betweenness >= 4 months" # description of scenario - see look up table
    )
  }
)

saveRDS(scenario_6_results, "scenario_6_results.rds")
```


```{r}
# combine the daily outputs from all 2000 sim replicates into one data frame

# daily result - epidemic curves for all 2000 stochastic runs
daily_results_SC6 <- bind_rows( 
  lapply(scenario_6_results, function(x) x$daily) 
)
# peak infectious and final outbreak size per run
summary_results_SC6 <- bind_rows(
  lapply(scenario_6_results, function(x) x$summary)
)
```
```{r}
head(daily_results_SC6)
head(summary_results_SC6)
```

# Scenario 7 - Betweenness any month

```{r}
# calls the function 2000 times - one for each row of scenario_grid
scenario_7_results <- pblapply(
  seq_len(nrow(scenario_grid)),
  function(i){
    run_scenario(
      scenario_grid[i, , drop = FALSE], # becomes scenario_row inside funciton
      scenario_7,# movement dataset for this scenario_X 
      "Betweenness any month" # description of scenario - see look up table
    )
  }
)

saveRDS(scenario_7_results, "scenario_7_results.rds")
```


```{r}
# combine the daily outputs from all 2000 sim replicates into one data frame

# daily result - epidemic curves for all 2000 stochastic runs
daily_results_SC7 <- bind_rows( 
  lapply(scenario_7_results, function(x) x$daily) 
)
# peak infectious and final outbreak size per run
summary_results_SC7 <- bind_rows(
  lapply(scenario_7_results, function(x) x$summary)
)
```
```{r}
head(daily_results_SC7)
head(summary_results_SC7)
```

# Scenario 8 - Combined (high) BT >= 6 months & OD >= 8 months
```{r}
# calls the function 2000 times - one for each row of scenario_grid
scenario_8_results <- pblapply(
  seq_len(nrow(scenario_grid)),
  function(i){
    run_scenario(
      scenario_grid[i, , drop = FALSE], # becomes scenario_row inside funciton
      scenario_8,# movement dataset for this scenario_X 
      "Combined BT >= 6 months & OD >= 8 months" # description of scenario - see look up table
    )
  }
)

saveRDS(scenario_8_results, "scenario_8_results.rds")
```


```{r}
# combine the daily outputs from all 2000 sim replicates into one data frame

# daily result - epidemic curves for all 2000 stochastic runs
daily_results_SC8 <- bind_rows( 
  lapply(scenario_8_results, function(x) x$daily) 
)
# peak infectious and final outbreak size per run
summary_results_SC8 <- bind_rows(
  lapply(scenario_8_results, function(x) x$summary)
)
```
```{r}
head(daily_results_SC8)
head(summary_results_SC8)
```

# Scenario 9 - Combined (med) BT>= 4 months & 0D >= 6 months

```{r}
# calls the function 2000 times - one for each row of scenario_grid
scenario_9_results <- pblapply(
  seq_len(nrow(scenario_grid)),
  function(i){
    run_scenario(
      scenario_grid[i, , drop = FALSE], # becomes scenario_row inside funciton
      scenario_9,# movement dataset for this scenario_X 
      "Combined BT>= 4 months & 0D >= 6 months" # description of scenario - see look up table
    )
  }
)

saveRDS(scenario_9_results, "scenario_9_results.rds")
```


```{r}
# combine the daily outputs from all 2000 sim replicates into one data frame

# daily result - epidemic curves for all 2000 stochastic runs
daily_results_SC9 <- bind_rows( 
  lapply(scenario_9_results, function(x) x$daily) 
)
# peak infectious and final outbreak size per run
summary_results_SC9 <- bind_rows(
  lapply(scenario_9_results, function(x) x$summary)
)
```
```{r}
head(daily_results_SC9)
head(summary_results_SC9)
```

# Scenario 10 - Combined any month

```{r}
# calls the function 2000 times - one for each row of scenario_grid
scenario_10_results <- pblapply(
  seq_len(nrow(scenario_grid)),
  function(i){
    run_scenario(
      scenario_grid[i, , drop = FALSE], # becomes scenario_row inside funciton
      scenario_10,# movement dataset for this scenario_X 
      "Combined any month" # description of scenario - see look up table
    )
  }
)

saveRDS(scenario_10_results, "scenario_10_results.rds")
```


```{r}
# combine the daily outputs from all 2000 sim replicates into one data frame

# daily result - epidemic curves for all 2000 stochastic runs
daily_results_SC10 <- bind_rows( 
  lapply(scenario_10_results, function(x) x$daily) 
)
# peak infectious and final outbreak size per run
summary_results_SC10 <- bind_rows(
  lapply(scenario_10_results, function(x) x$summary)
)
```
```{r}
head(daily_results_SC10)
head(summary_results_SC10)
```

CONTROLS - If I remove the same number of farms, is targeted removal (SC2-10) more effective than random removal?
# Scenario 11 - control High - random 4 farms (average of scenarios 3/6/9) = 4.33 (4) 
```{r}
# calls the function 2000 times - one for each row of scenario_grid
scenario_11_results <- pblapply(
  seq_len(nrow(scenario_grid)),
  function(i){
    run_scenario(
      scenario_grid[i, , drop = FALSE], # becomes scenario_row inside funciton
      scenario_11,# movement dataset for this scenario_X 
      "Control High" # description of scenario - see look up table
    )
  }
)

saveRDS(scenario_11_results, "scenario_11_results.rds")
```
```{r}
# combine the daily outputs from all 2000 sim replicates into one data frame

# daily result - epidemic curves for all 2000 stochastic runs
daily_results_SC11 <- bind_rows( 
  lapply(scenario_11_results, function(x) x$daily) 
)
# peak infectious and final outbreak size per run
summary_results_SC11 <- bind_rows(
  lapply(scenario_11_results, function(x) x$summary)
)
```
```{r}
head(daily_results_SC11)
head(summary_results_SC11)
```
# Scenario 12 - control med - random 9 farms (average of scenarios 2/5/8) = 9.33 (9)

```{r}
# calls the function 2000 times - one for each row of scenario_grid
scenario_12_results <- pblapply(
  seq_len(nrow(scenario_grid)),
  function(i){
    run_scenario(
      scenario_grid[i, , drop = FALSE], # becomes scenario_row inside funciton
      scenario_12,# movement dataset for this scenario_X 
      "Control Med" # description of scenario - see look up table
    )
  }
)

saveRDS(scenario_12_results, "scenario_12_results.rds")
```
```{r}
# combine the daily outputs from all 2000 sim replicates into one data frame

# daily result - epidemic curves for all 2000 stochastic runs
daily_results_SC12 <- bind_rows( 
  lapply(scenario_12_results, function(x) x$daily) 
)
# peak infectious and final outbreak size per run
summary_results_SC12 <- bind_rows(
  lapply(scenario_12_results, function(x) x$summary)
)
```
```{r}
head(daily_results_SC12)
head(summary_results_SC12)
```
# Scenario 13 - random 40 farms (average of scenarios 4/7/10) = 40.33 (40)

```{r}
# calls the function 2000 times - one for each row of scenario_grid
scenario_13_results <- pblapply(
  seq_len(nrow(scenario_grid)),
  function(i){
    run_scenario(
      scenario_grid[i, , drop = FALSE], # becomes scenario_row inside funciton
      scenario_13,# movement dataset for this scenario_X 
      "Control All" # description of scenario - see look up table
    )
  }
)

saveRDS(scenario_13_results, "scenario_13_results.rds")
```
```{r}
# combine the daily outputs from all 2000 sim replicates into one data frame

# daily result - epidemic curves for all 2000 stochastic runs
daily_results_SC13 <- bind_rows( 
  lapply(scenario_13_results, function(x) x$daily) 
)
# peak infectious and final outbreak size per run
summary_results_SC13 <- bind_rows(
  lapply(scenario_13_results, function(x) x$summary)
)
```
```{r}
head(daily_results_SC13)
head(summary_results_SC13)
```

# SCENARIO SIM CHECKS


```{r}
# check 2000 sims ran
nrow(summary_results_SC2)
```

```{r}
# check how many simulations ended with each number of infected farms
table(summary_results_SC2$final_infected)

# e.g. 23 sims infected 0 farms (beyond the initial seed); 1,620 sims infected 1 farm etc

# if = 2000 then all sims produced a summary result
sum(table(summary_results_SC2$final_infected))
```

```{r}
summary(summary_results_SC2$final_infected)
```

```{r}
quantile(
  summary_results_SC2$final_infected,
  probs = c(0, 0.25, 0.5, 0.75, 0.95, 1)
)
```
# SCENARIO COMPARISONS

```{r}
# mean final outbreak size
# relative reduction = ((2.8756 - 2.0625)/2.8765)x100
print("Baseline")
mean(summary_results_SC1$final_infected)
print("Outdegree: High, Med, All")
mean(summary_results_SC2$final_infected)
mean(summary_results_SC3$final_infected)
mean(summary_results_SC4$final_infected)
print("Betweenness: High, Med, All")
mean(summary_results_SC5$final_infected)
mean(summary_results_SC6$final_infected)
mean(summary_results_SC7$final_infected)
print("Combined: High, Med, All")
mean(summary_results_SC8$final_infected)
mean(summary_results_SC9$final_infected)
mean(summary_results_SC10$final_infected)
print("Control: High (-4), Med (-9), All (-40)")
mean(summary_results_SC11$final_infected)
mean(summary_results_SC12$final_infected)
mean(summary_results_SC13$final_infected)
```

```{r}
# proportion of sims with onward transmission - gives probability that a randomly introduced infection spreads beyond the initial farm
print("Baseline")
mean(summary_results_SC1$final_infected >1)
print("Outdegree: High, Med, All")
mean(summary_results_SC2$final_infected >1)
mean(summary_results_SC3$final_infected >1)
mean(summary_results_SC4$final_infected >1)
print("Betweenness: High, Med, All")
mean(summary_results_SC5$final_infected >1)
mean(summary_results_SC6$final_infected >1)
mean(summary_results_SC7$final_infected >1)
print("Combined: High, Med, All")
mean(summary_results_SC8$final_infected >1)
mean(summary_results_SC9$final_infected >1)
mean(summary_results_SC10$final_infected >1)
print("Control: High(-4), Med(-9), All(-40")
mean(summary_results_SC11$final_infected >1)
mean(summary_results_SC12$final_infected >1)
mean(summary_results_SC13$final_infected >1)
```
```{r}
# probability of a large outbreak - with defined threshold
print("Baseline")
mean(summary_results_SC1$final_infected >6)
print("Outdegree: High, Med, All")
mean(summary_results_SC2$final_infected >6)
mean(summary_results_SC3$final_infected >6)
mean(summary_results_SC4$final_infected >6)
print("Betweenness: High, Med, All")
mean(summary_results_SC5$final_infected >6)
mean(summary_results_SC6$final_infected >6)
mean(summary_results_SC7$final_infected >6)
print("Combined: High, Med, All")
mean(summary_results_SC8$final_infected >6)
mean(summary_results_SC9$final_infected >6)
mean(summary_results_SC10$final_infected >6)
print("Control: High (-4), Med (-9), All (-40)")
mean(summary_results_SC11$final_infected >6)
mean(summary_results_SC12$final_infected >6)
mean(summary_results_SC13$final_infected >6)
```


```{r}
# upper tail outbreak size - does removing important farms reduce rare but large superspreading events
print("Baseline")
quantile(
  summary_results_SC1$final_infected,
  probs = c(0.9, 0.95, 0.99)
)
print("Outdegree: High, Med, All")
quantile(
  summary_results_SC2$final_infected,
  probs = c(0.9, 0.95, 0.99)
)

quantile(
  summary_results_SC3$final_infected,
  probs = c(0.9, 0.95, 0.99)
)

quantile(
  summary_results_SC4$final_infected,
  probs = c(0.9, 0.95, 0.99)
)
print("Betweenness: High, Med, All")
quantile(
  summary_results_SC5$final_infected,
  probs = c(0.9, 0.95, 0.99)
)

quantile(
  summary_results_SC6$final_infected,
  probs = c(0.9, 0.95, 0.99)
)

quantile(
  summary_results_SC7$final_infected,
  probs = c(0.9, 0.95, 0.99)
)
print("Combined: High, Med, All")

quantile(
  summary_results_SC8$final_infected,
  probs = c(0.9, 0.95, 0.99)
)

quantile(
  summary_results_SC9$final_infected,
  probs = c(0.9, 0.95, 0.99)
)

quantile(
  summary_results_SC10$final_infected,
  probs = c(0.9, 0.95, 0.99)
)
print("Control: High, Med, All")
quantile(
  summary_results_SC11$final_infected,
  probs = c(0.9, 0.95, 0.99)
)

print("Control: High (-4), Med (-9), All (-40)")
quantile(
  summary_results_SC12$final_infected,
  probs = c(0.9, 0.95, 0.99)
)

quantile(
  summary_results_SC13$final_infected,
  probs = c(0.9, 0.95, 0.99)
)
```


```{r}
# peak infectious farms
print("Baseline")
summary(summary_results_SC1$peak_infectious)
print("Outdegree: High, Med, All")
summary(summary_results_SC2$peak_infectious)
summary(summary_results_SC3$peak_infectious)
summary(summary_results_SC4$peak_infectious)
print("Betweenness: High, Med, All")
summary(summary_results_SC5$peak_infectious)
summary(summary_results_SC6$peak_infectious)
summary(summary_results_SC7$peak_infectious)
print("Combined: High, Med, All")
summary(summary_results_SC8$peak_infectious)
summary(summary_results_SC9$peak_infectious)
summary(summary_results_SC10$peak_infectious)
print("Control: High (-4), Med (-9), All (-40)")
summary(summary_results_SC11$peak_infectious)
summary(summary_results_SC12$peak_infectious)
summary(summary_results_SC13$peak_infectious)
```

The targeted removals consistently reduce outbreak size compared with baseline. In contrast, the random removals have much smaller effects.

Controls support interpretation that simply removing farms from the network has little effect on epidemic size. In contrast, targeted removal of structurally important farms substantially reduces epidemic size, indicating that the network position of the removed farms, rather than the number of farms removed, drives the reduction in dz spread.

Use confidence intervals/non-parametric test (e.g. Wilcox rank-sum test - heavy right tail) to demonstrate differences between some are meaningful vs not

# DEBUGS

```{r}
test_SC2 <- scenario_2 %>%
  filter(
    is.na(source) | is.na(destination)
  )

nrow(test_SC2)
```
```{r}
farms_SC2 <- unique(c(scenario_2$source, scenario_2$destination))

length(farms_SC2)
sum(is.na(farms_SC2))
```
```{r}
sum(scenario_2$source %in% OD_high_candidates)
sum(scenario_2$destination %in% OD_high_candidates)
```
```{r}
class(scenario_2$source)
class(scenario_2$destination)
```
```{r}
# combine the daily outputs from all 2000 sim replicates into one data frame

# daily result - epidemic curves for all 2000 stochastic runs
daily_results_SC2 <- bind_rows( 
  lapply(scenario_2_results, function(x) x$daily) 
)
# peak infectious and final outbreak size per run
summary_results_SC1 <- bind_rows(
  lapply(scenario_1_results, function(x) x$summary)
)
```
```{r}
head(daily_results_SC1)
head(summary_results_SC1)
```
```{r}
nrow(movements)
nrow(scenario_11)
```
```{r}
max(movements$day)

max(scenario_11$day)
```
```{r}
length(unique(c(movements$source, movements$destination)))

length(unique(c(scenario_11$source, scenario_11$destination)))
```

## NEXT STEPS 31/07/26

1. compare epidemic curves (use daily_results_SCx) - plot mean number infectious farms through time (?+confidence intervals) for each scenario --> how big outbreak becomes & how quickly spread when peak. Care with mean & median as heavy right tail here. Also consider mean of 2000 sims - outbreaks will behave differently

2. compare the distrbution of outbreak sizes - looking at mean, median, interquartile range, boxplot, histogram - ?does targeted removal reduce the frequency of the larger outbreaks - relate to superspreading

3. Peak infectious farms - compare across scenarios - lower peak = fewer farms simultaneously infectious = surveillance control

4. probability of a large outbreak - define threshold for mean outbreak size = "large outbreak", compare the probability of a "large outbreak" across scenarios

  create comparison table to show importance of structurally important farms driving superspreading - calculate proportion of simulations exceeding different outbreak-size threshold (e.g. >5/10/20 farms) for each scenario - ?does targeted removal reduce the likelihood of the largest outbreaks?

5. show control comparison - compare to baseline/random removal/targeted removal

6. calculate relative reduction from baseline, percentage reductions rather than just mean

7. silent spreader index - ?structural network metric - capture farms with high bridging influence (high betweenness) but low outward movement volume (low outdegree). Farms that don't look risky from movement counts alone, but are structurally positioned to spread disease quietly. SSI = betweeness/(outdegree +1) - +1 prevents division by zero & emphasise low outdegree farms

7. statistical analysis - see above note - want effect size & p-values. Look at paper discussing reporting of epidemic outcomes 


# PLOTS
```{r}
# 1. plot individual stochastic epidemic trajectories (to show variability)
ggplot(
  daily_results_SC1,
  aes(x = day, y = infectious, group = replicate)
) +
  geom_line(alpha = 0.3) +
  labs(
    x = "Day",
    y = "Infectious farms",
    title = "SC1 stochastic epidemic trajectories"
  ) +
  theme_classic()
```
```{r}
# proportion of introductions/sims that produced an epidemic with more than 5 simultanrously infected farms
# e.g. 0.0645 = 6.45% of the stochastic introductions produced a simulation where more than 5 farms were simultaneously infected at some point
mean(tapply(
  daily_results_SC1$infectious,
  daily_results_SC1$replicate,
  max
) > 5)
```





```{r}
# 2(a). mean/median epidemic curve with 5th-95th percentile uncertainty ribbon
# Calculate mean and uncertainty bounds across stochastic simulations
mean_epidemic_SC1 <- daily_results_SC1 %>%
  group_by(day) %>%
  summarise(
    mean_infectious = mean(infectious),
    median_infectious = median(infectious),
    p5_infectious = quantile(infectious, 0.05),
    p95_infectious = quantile(infectious, 0.95)
  )

# Plot mean epidemic curve with 5-95% stochastic range
ggplot(mean_epidemic_SC1, aes(x = day, y = mean_infectious)) +
  geom_ribbon(
    aes(
      ymin = p5_infectious,
      ymax = p95_infectious
    ),
    alpha = 0.3,
    fill = "firebrick"
  ) +
  geom_line(
    linewidth = 1.2,
    colour = "firebrick"
  ) +
    geom_line(
    aes(y = median_infectious),
    linewidth = 1.2,
    colour = "forestgreen",
    linetype = "dashed"
  ) +
  labs(
    x = "Day",
    y = "Mean infectious farms",
    title = "SC1 epidemic curve (mean with 5th–95th percentile range)"
  ) +
  theme_classic()
```
red line = avg no. infectious farms per day accross 2000 sims
green line - typical epidemic trajectory (50% of sims above and below)
shaded area = range containing the 90% of stochastic outcomes
bottom of shaded area = 5th percentile - 
top of the shaded area = 95th percentile = the number of infectious farms that 95% of simulations are below that day
e.g. - on day 300 - upper red area = 4 --> 95% of the 2000 stochastic simulations had 4 or fewer farms in the infectious compartment on that day



  1. plot median alongside mean - know data has high right skewed distribution
  
```{r}
# 2(b). similar to 2 - but shows total_infected rather than infectious - more useful - capture whether removing structurally important farms reduces final outbreak magnitude &e epidemic growth? add lower percentile ribbon?
daily_summary_SC1 <- daily_results_SC1 %>%
  group_by(day) %>%
  summarise(
    mean_infected = mean(total_infected),
    median_infected = median(total_infected),
    p95_infected = quantile(total_infected, 0.95) # greyed out area
  )
```
```{r}
ggplot(daily_summary_SC1, aes(day)) +
  geom_line(aes(y = mean_infected), colour = "blue") +
  geom_line(aes(y = median_infected), colour = "black") +
  geom_ribbon(
    aes(ymin = median_infected,
        ymax = p95_infected),
    alpha = 0.2
  )
```
Black median line staying at ~1 --> most sims little or no onward spread beyond initial infected farm
Blue mean line rising to ~3 --> smaller number of larger outbreaks pulling average upwards
Grey ribbon expanding upwards --> a minority of simulations produce sybstantially larger outbreaks - check 95th percentile classification

```{r}
# 3. final outbreak size distribution
ggplot(
  summary_results_SC1,
  aes(x = final_infected) # uses the final number of infected farms at the end of each sim
) +
  geom_density(
    fill = "steelblue",
    alpha = 0.4
  ) + 
  labs(x = "Final infected farms", # per sim
       y = "Density", # how common each outbreak size is
       title = "SC_1 Distribution of final outbreak size"
       ) + 
  theme_classic()
```
A large peak near 1 - most sims ended with only the initial farm infected = no onward spread
A long tail to right - small number of sims produced much larger outbreaks
Overlay this to compare scenarios! shows whether removing candidate farms changes the entire outbreak size distribution, not just the mean


```{r}
# 4. scenario comparison - compare SC1-10
# median final infected
# mean final infected
# probability of spread
# 95th percentile outbreak size
```


```{r}
# for month 7 - from sensitivity analysys
out_degree_top5 <- c("253", "129", "95", "255", "88", "119", "214", "213", "212", "202", "151", "299")
betweenness_top5 <- c("95", "296", "119", "41", "285", "300", "274", "178", "138", "235", "156")
random_farms <- c("2", "35", "180", "50", "260", "4", "230", "301", "92", "205", "167", "227")
```