---
title: "RQ2"
author: "ESINCLAIR"
date: "2026-07-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# LIBRARIES
library(dplyr)
library(igraph)
library(movenet)
library(SimInf)
```

Network weighted SEIR (edge weighted) - uses BFS and roll dice on probability of infection (based on weigthts - n_cattle, n_movements)
Compartment-based SEIR (SimInf's built-in SEIR - compartment based) - doesn't perform BFS, does not roll a probability per edge, does not use movement intensity, does not use n_cattle or n_movements, does not weight edges, does not compute infection probability per movement.SimInf knows who moves to whom, when they move, how many individuals move, global beta, epsilon, gamma

So either need to:
1. Farm level comparment-based model using SimInf inbuilt SEIR (where disease spreads, when disease spreads, how big the outbreak becomes) & separate risk layer analysis (which farms have high movement intesity, which farms are superspreaders, which edges have high transmission potential). Then compare structural risk vs dynamic infetion = epidemiological insight. SEIR compartment model aligns with this paper https://www.mdpi.com/2076-2607/9/2/227 - also parameters - bTB reflective - susceptible, occult (infected but not infectious), infected, removed; https://www.sciencedirect.com/science/article/pii/S0034528814001404#s0015

2. build a custom SimInf model - lots of code

Phase 1: build SimInf SEIR on synthetic cattle movement network and confirm model runs and outputs are sensible; tune parameters for generic SEIR - using plausible values beta/epsilon/gamma to get a functioning euidemic crve; make it bTB reaslitic - adjust parameters; possibly refine compartment meanings to S-occult-I-removed - check results

Phase 2: structural risk based analysis - build movement based risk layer using data - identify OD superspreaders, BT "silent" spreaders, and their persistence over time - will look for references for this concept - need movement networks contain superspreaders; movement intensity predicts epidemic risk; network centrality identifies high risk farms; risk based surveillance uses movement derived risk,removing high risk nodes reduces epidemic size, livestock movement risk analysis is standard pactice - ?but not in combination with SimInf

Phase 3 - superspreader removal scenrios - remove high outdegree superspreader farms; remove high betweeness superspreader farms; re-run siminf and compare epidemic dynamics (final size, peak, timing, duration) against the baseline

Import data
```{r}
data_year <- read.csv("RQ1 NO EP.csv", check.names = TRUE) # tells R to sanitise column names
```
```{r}
head(data_year)
```

```{r}

movements_per_node <- data_year %>%
  group_by(org_id) %>% # groups all rows (movements) by the origin farm
  summarise(n_moves = n()) # counts how many rows each farm has -> number of movements per farm

max_moves <- max(movements_per_node$n_moves) # extracts the single largest count accros all farms
max_moves
```

# NODE LIST


```{r}
# 1. Unique farm IDs
all_ids <- sort(unique(c(data_year$org_id, data_year$dest_id)))

# 2. Type lookup
type_lookup <- rbind(
  data.frame(id = data_year$org_id,  type = data_year$org_type),
  data.frame(id = data_year$dest_id, type = data_year$dest_type)
) |> dplyr::distinct(id, .keep_all = TRUE)

# 3. Build nodes table, #left_join never drops IDs (occasioallly merge will)
nodes <- left_join(
  data.frame(
  id = all_ids),
  type_lookup, by = "id"
)

# 4. Final node list using transmute (matches movenet test_sim example)
nodes <- nodes |>
  transmute(
    id   = id,
    type = type,
    size = max_moves # # 'size' = S = movement/infection capacity per farm (global value = max total movements)
  )
# “Because farms frequently make multiple same-day movements, we set a global node size 𝑆=10 to represent movement/infection capacity rather than literal individuals. This allows each farm to generate multiple infection opportunities per day, consistent with the empirical movement network.”

```
```{r}
# double check
length(all_ids)
nrow(nodes)
```

In SimInf:
  The node table defines how many infdividuals exist in each node
  The u0 table initial state is built from that node table
  The size column is interpreted as the number of individuals in the node
  Therefore size =1 -> S=1 at model start
```{r}
tail(nodes)
```

# EDGES/MOVEMENTS table
needs to match SimInf events - doesnt care what column names arebut needs origin farm ID, dest farm ID, time of movement (date/day - needs to be an int timestep), number of individuals moved (total_cattle) - see TEST SIM doc

```{r}
events <- data_year %>%
  dplyr::transmute(
    event = "extTrans",
    time  = day,                     # already integer days
    node  = match(org_id,  nodes$id),
    dest  = match(dest_id, nodes$id),
    n     = 0, # n = 0 in SimInf means ' dont move a fixed number of individuals' - so switches to using the proportion field instead, so all S,E and I individuals move to the destination farm
    proportion = 0.5,# means move 100% of SEI individuals - every movement carries infectiousn if the farm is infectious - this means that each movement empties the origin farm (all S,E,I leave) so node e.g. 157 stops beng infectious as soon as it 'moves'
    select = 2, # means use row 2 of the select matrix
    shift = 0
  )

# proportion = fraction of the farm’s infectious pressure that moves during a movement event. e.g. on average 10% ofa farms animals move during a movement event - even though model is farm-level, this gives a realistic scaling
```
```{r}
head(events)
```

event = "extrans"= movement between nodes

time = day = day of movement

node= origin

dest = destination

n = number of cattle moved

proportion - controls whether the movement removes a fixed number (n) or a proportion of individuals - proportion = 0 = use n (batch ize) - correct for farm level model

selected - tells siminf which compartments to sample from when removing individuals - it indexes the select matrix that is set later - because select = 2 it means - remove individuals from SE and I, do not remove from R - it matches the example siminf - animals can move whether susceptible, exposed, or infectious

shift - controls whether individuals change compartment during the event: shift = 0 - no compartment change; shift >0 - use the shift matrix to move individuals into a different compartment
```{r}
time_unit <- 1 #every how much time / how many days to report results for - because in events table - already in weeks - then time step of '1' = 1 week
beta = 1.0 #transmission rate
#Guinat et al 2016: 0.3 between-pen, 0.6 within-pen;
#used in Halasa 2016a as low & high tm levels
epsilon = 0.2 #incubation rate (1/incubation period)
#Zhang et al 2021: 0.1 (0.05-0.25);
#based on Barongo et al 2016, Nielsen et al 2017
gamma = 0.1 #mortality rate
#Zhang et al 2021: 0.25 (0.07-0.5)
#based on Guinat et al 2016
```

# Build SEIR model

1. build the initial conditions (u0)
2. Build hte model timespan (tspan)
3. construct the SEI model
4. set the select_matrix
5. run the model
6 summarise the results

N.B. SimInf's SEIR is not farm level but individual level - need to make each farm 1 individual
In write up need to explain why chose this more simplistic model

```{r}
# build initial conditions - have one individual per farm (size = 1), so each node starts with: S=, E=0, I=0
# then seed infection in one random node
n <- length(nodes$id)

u0 <- data.frame(
  S = nodes$size,
  E = rep(0, n),
  I = rep(0, n),
  R = rep(0, n)
)
```

```{r}
# Seed infection in one random node
infected_node <- sample(n, 1)
u0[infected_node, "S"] <- u0[infected_node, "S"] - 1
u0[infected_node, "I"] <- 1
```
```{r}
# Build the model timespan

# events$time column already in integer days
tspan <- seq(
  from = min(events$time),
  to   = max(events$time),
  by   = time_unit   # time_unit = 1 day
)
```
```{r}
#Build the SEIR model - SimInf built in SEIR model expects 4 compartments - S, E, I, R - R is anything that has left the #infectious pool - for this scenario will treat is as removed/culled.

#gamma (y) - rate at which infected are culled/removed
#select = 0 - so R never moves - matches "removed"


model <- SEIR(
  u0 = u0, # initial conditions
  tspan = tspan, # timespan
  events = events, #moves between nodes
  beta = beta, #transmission rate
  epsilon = epsilon, #incubation rate (1/incubation period)
  gamma = gamma
  ) #mortality rate
```
```{r}
# Set the select matrix

# Tells SimInf which compartments animals can be removed from during movement - column 2 - remove from SEI, do not remove from R

select_matrix(model) <- matrix(
  c(1, 0, 0, 0,
    1, 1, 1, 0),
  nrow = 4
)
```

Run the model

```{r}
result <- run(model)
```

Extract the trajectory - gives the time series of SEIR for every node
```{r}
traj <- trajectory(result)

# contains time, node, SEIR counts per node at each time step
```
```{r}
head(traj)
```

Plot the epidemic/disease spread curve (total infectious over time)

```{r}
library(dplyr)
library(ggplot2)

epi_curve <- traj %>%
  group_by(time) %>%
  summarise(total_I = sum(I))

ggplot(epi_curve, aes(time, total_I)) +
  geom_line(colour = "red", size = 1) +
  labs(
    title = "Total Infectious Over Time",
    x = "Time",
    y = "Number Infectious"
  )

# shows outbreak start, peak infection, decline, final fade out
```
Plot epidemic size (cumulative number infected) - need to compute as SimInf doesnt track directly
```{r}
epi_farms <- traj %>%
  group_by(time) %>%
  summarise(
    infected_farms = sum(I > 0)  # count farms with at least one infectious individual
  )

ggplot(epi_farms, aes(time, infected_farms)) +
  geom_line(colour = "red", size = 1) +
  labs(
    title = "Number of Infectious Farms Over Time",
    x = "Time",
    y = "Infectious farms"
  )

# shows how many farms infected at each time point
```
Visualise spread across nodes (heatmap)
```{r}
# shows which farms were infected and when
ggplot(traj, aes(time, factor(node), fill = I)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "red") +
  labs(
    title = "Infection Heatmap by Node",
    x = "Time",
    y = "Farm ID"
  )
```
Identify peak infection time and size
```{r}
# gives peak number infectous, exact time occurs
peak <- epi_curve %>% filter(total_I == max(total_I))
peak
```
Identify which farms were infected
```{r}
# useful to compare baseline simulation vs simulation with SS nodes removed
infected_nodes <- traj %>%
  group_by(node) %>%
  summarise(ever_infected = any(I > 0 | E > 0 | R > 0))

infected_nodes
```
# Trouble shooting codes
```{r}
traj <- trajectory(result)

traj %>%
  group_by(time) %>%
  summarise(total_I = sum(I)) %>%
  head()
```

```{r}
# check initial state of node
model@u0[157, ]
```
```{r}
# check for NA values
summary(events$node)
summary(events$dest)
```

```{r}
head(nodes$id)
tail(nodes$id)
```
```{r}
events %>% filter(node == 157) %>% head()
```
```{r}
sum(is.na(events$node))
sum(is.na(events$dest))
```
```{r}
which(nodes$id == 157)
```
```{r}
events %>% filter(node == 157) %>% arrange(time)
```
