---
title: "SOIR"
author: "ESINCLAIR"
date: "2026-07-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

https://pmc.ncbi.nlm.nih.gov/articles/PMC3475695/ - SORI model
https://pmc.ncbi.nlm.nih.gov/articles/instance/2600903/bin/rspb20071601s01.pdf 
https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1004038#pcbi-1004038-g001
https://pmc.ncbi.nlm.nih.gov/articles/PMC6122770/#Sec5
https://efsa.onlinelibrary.wiley.com/doi/full/10.2903/j.efsa.2023.8028
https://efsa.onlinelibrary.wiley.com/doi/10.2903/j.efsa.2017.4959 - bTB informaiton

https://www.rdocumentation.org/packages/stats/versions/3.5.2/topics/Uniform - R Documentation - runif()

file:///C:/Users/errai/Downloads/978-3-540-78911-6_3.pdf - chapter on maths behind models
Built from work in TEST SIM 2 Rmd

Choosing parameters (involves sensitivity analysis to give plausible values)

Beta:

  https://www.nature.com/articles/nature13529#MOESM74 & supplementary information https://static-content.springer.com/esm/art%3A10.1038%2Fnature13529/MediaObjects/41586_2014_BFnature13529_MOESM74_ESM.pdf
  
  Paper shows beta is low - however my model does not have environment etc - test range of beta
  β = 0.001
  β = 0.005
  β = 0.01
  β = 0.02
  


Occult period	          Brooks-Pollock/Conlan (SORI framework and discussion of uncertainty)

Movement transmission	  Green et al.; BoTMEW (movement-based transmission)

Transmission probability (β)	see ref in beta This paper, emphasising that direct transmission is low and uncertain

Infectious period	  EFSA review and experimental infection literature, because this paper does not estimate a fixed infectious duration


Parameter	Literature evidence	Suggested baseline	Sensitivity range	Rationale
Occult period (S → O → I)	Brooks-Pollock/Conlan: progression highly uncertain; many cattle may never become infectious. Skin-test conversion typically occurs within 1–9 weeks after infection, but progression to shedding may take months or longer.	120 days	60–240 days	Represents delayed onset of infectiousness while still allowing transmission within a 1-year simulation.
Infectious period (I → R)	EFSA review and experimental studies indicate infected cattle can shed for extended periods until detected/removed. Movement restrictions eventually remove infectious herds.	180 days	90–365 days	Six months is a reasonable compromise between rapid detection and prolonged infectiousness.
```{r}
# load libraries
library(ggplot2)
library(dplyr)
library(tidyr)
```
State encoding            Movement transmission?
0 = S (Suscpetible farm)          No
1 = O (Occult infection)          No (evidence to show can shed during occult - left out for simplicity of model)
2 = I (Infectious/shedding farm)  Yes
3 = R (removed)                   No

Updated to SOIR because in bTB, infection does not immediately = infectiousness (https://pmc.ncbi.nlm.nih.gov/articles/PMC3475695/#s4)

```{r}
# model is unweighted by total_cattle - movement is either present or absent
run_soir <- function(
    movements,
    seed_farms, # can seed 1+ farms in one run - see update below
    beta = 0.05, # explore 0.001 - 0.05
    occult_period = 120, # explore 60-240
    infectious_period = 180 # explore 90 - 135
){
  
  
# Build lookup table - pulls every unique farm ID from both the source and destination columns of movements  - combines, removes duplicates, sorts --> (farms) is a clean lookup table of all farms involved in the dataset
  
  farms <- sort(unique(c(
    movements$source,
    movements$destination
  )))
  
# creates a numeric index for the farms, labels each with a number corresponding to the farm ID
  n_farms <- length(farms)

  farm_index <- seq_len(n_farms)
  names(farm_index) <- farms

 
# Convert movements to integer indices

  movements$source_id <- farm_index[movements$source]
  movements$dest_id   <- farm_index[movements$destination]

# Set up vectors to track each farms disease status and how long it has been in its current state
  
  state <- integer(n_farms) # creates a vector of zeros - meaning all farms start susceptible

  days <- integer(n_farms) # creates a parallel vector to count how many days each farm has spent in its current state

 
# Seed infection - infect chosen seed farm by setting its state to 1 - farms to start as occult - representative of dz introduction
# Reset their disease timers to 0 days, indicating they have just entered the infectious state
  
  state[farm_index[seed_farms]] <- 1  # find farm(s) listed in seed_farms, look up their position in the 'state' vector using 'farm_index'; change disease state S -> 0
  days[farm_index[seed_farms]] <- 0 # for the same seeded farm(s), set no. days spent in current dz state to 0 - farm has just become occult at the start of the simulation

# Storage - prepares storage for the simulation results

  max_day <- max(movements$day) # finds last day in the movement dataset

# history - creates a data frame with one row per day, and columns for the number of farms in each SOIR state, all columns are pre-allocated as integer vectors of length max_day for filling during the simulation
  
  history <- data.frame( 
    day = integer(max_day), 
    S = integer(max_day),
    O = integer(max_day),
    I = integer(max_day),
    R = integer(max_day)
  )


# Simulation - advances farms through the O -> I -> R stages as each day of the simulation passes
# the 'for' loop runs one iteration per day

  for(day in seq_len(max_day)){


# Progress exposed farms:
# The day counters increase, and farms that have spent more than occult_period days in O become infectious (state == 2) and their day counters reset
    
    occult <- which(state == 1)

    if(length(occult) > 0){

      days[occult] <- days[occult] + 1 # increments until reaches occult_period -> becomes infectious

      become_I <- occult[
        days[occult] >= occult_period
      ]

      state[become_I] <- 2
      days[become_I] <- 0

    }


# Progress infectious farms

    infectious <- which(state == 2)

# Their day counters increase, any farms that have spent more than infectious_period i days in I become removed (state == 3) and their day counters reset    
    
    if(length(infectious) > 0){

      days[infectious] <- days[infectious] + 1

      become_R <- infectious[
        days[infectious] >= infectious_period
      ]

      state[become_R] <- 3
      days[become_R] <- 0

    }

# Today's movements - filters the 'movements' data frame to keep just the rows where 'day' matches the current loop day, giving the set of movements to process for that day.

    todays <- movements[movements$day == day, ]

# Replay movement events - processes each movement and applies infection-spread rules.
# loop over all movements happening today, for each movement, get the source and destination farm indices; skip the movement if the source is not infectious (state != 2); skip the movement if the destination is not susceptible (state != 0). Only movements from the infectious -> susceptible farms proceed to the transmission step that follows.
    

    if(nrow(todays) > 0){

      for(i in seq_len(nrow(todays))){

        source <- todays$source_id[i]
        dest   <- todays$dest_id[i]

        # source not infectious
        if(state[source] != 2)
          next

        # destination already infected
        if(state[dest] != 0)
          next


# Transmission - applies transmission step - an infectious source may infect a susceptible destination. With probability 'beta', the destination farm becomes exposed (state = 1) and is day counter resets to 0 
# runif(1) generates a random number between 0 and 1, if that random number is less than beta, transmission occurs - changes destination farm's state from S -> O, resets the dz timer. So if beta = 0.05 = 5% chance that an infectious movement infects the destination farm. Bernoulli probability 0.05/0.95; https://www.rdocumentation.org/packages/stats/versions/3.5.2/topics/Uniform
       
        if(runif(1) < beta){

          state[dest] <- 1
          days[dest] <- 0

        }

      }

    }

# Save epidemic - logs the SOIR count for the current simulation day - store the day number; count how many farms are S,O,I,R and write those values into the corresponding row of history
    

    history$day[day] <- day
    history$S[day] <- sum(state == 0)
    history$O[day] <- sum(state == 1)
    history$I[day] <- sum(state == 2)
    history$R[day] <- sum(state == 3)

  }

# Return - returns the simulation output as a structured list

  list(

    history = history, # stores the full day-by-day SOIR counts

    final_states = data.frame( # data frame listing each farm and its final SOIR state, converts numeric codes (0-3) into labelled factors S,O,I,R

      farm = farms,

      state = factor(
        state,
        levels = 0:3,
        labels = c("S","O","I","R")
      )

    )

  )

}
```


Transmission events were simulated stochastically using Bernoulli trials. For each eligible movement, a uniform random number (runif(1)) was drawn and transmission occurred when the random draw was less than the calculated transmission probability (beta).

?Consider effective reproductive number - but not typical SIR model with R(0)?

# Provide movement data
```{r}
# read in csv
movements_raw <- read.csv(
    "RQ1 NO EP.csv",
    stringsAsFactors = FALSE
)
#head(movements_raw )
```
```{r}
# check the structure
str(movements_raw)
```
```{r}
# create simulation movement  table
movements <- movements_raw %>%
    select(
        source = org_id,
        destination = dest_id,
        day,
        total_cattle
    )
#head(movements)
```

```{r}
# convert  farm IDs to chr rather than integer to stop e.g. 1 being  treated differently from 001 in future datasets
movements <- movements %>%
    mutate(
        source = as.character(source),
        destination = as.character(destination)
    )
```
```{r}
str(movements)
```
```{r}
# check time period
range(movements$day)
```
```{r}
# number of movement events
nrow(movements)
```
```{r}
# check movement intensity
movements %>%
    count(day) %>%
    summary()

```
```{r}
# N.B. this is only one replicate - is stochastic - need 100-1000 replicates
result <- run_soir(
    movements,
    seed_farms = c("2", "300", "193", "65")
)

#head(result$history)
```
```{r}
test <- run_soir(
  movements = movements,
  seed_farms = "95",
  beta = 0.05,
  occult_period = 120,
  infectious_period = 180
)
```
```{r}
# maximum farms infected
max(test$history$I)
```
```{r}
# plot data - ggplot
test$history %>%
    tidyr::pivot_longer(
        cols = c(S,O,I,R),
        names_to = "state",
        values_to = "farms"
    ) %>%
    ggplot(
        aes(
            x = day,
            y = farms,
            colour = state
        )
    ) +
    geom_line(size = 1) +
    theme_minimal()
```
```{r}
# Final counts in each compartment
table(test$final_states$state)
```

# Sensitivity analysis

https://aab.copernicus.org/articles/53/529/2010/aab-53-529-2010.pdf
https://www.sciencedirect.com/science/article/pii/S1755436510000460
https://arxiv.org/pdf/2607.01476

Because stochastic epidemic models incorporate uncertainty in both disease processes and transmission dynamics, sensitivity analyses are required to determine whether model outputs are robust to variation in uncertain parameters. Similar approaches have been applied in livestock disease simulations, where key epidemiological parameters were varied to evaluate their influence on simulated outbreaks.

Do the farms identified as structural super spreaders in the temporal network analysis (RQ1) consistently generate larger or faster outbreaks across a plausible range of bTB transmission parameters?

The absolute outbreak size changes because beta has changed, but the relative ranking remains the same: super spreader farms still produce larger outbreaks than random farms.

Choose baseline parameters from the literature
  Select biologically plausible values for beta, occult_period and infectious_period based on published bTB studies.
Run a sensitivity analysis
  Test a range of plausible values for each parameter.
  Run 100 stochastic simulations for each parameter combination.
  Calculate summary statistics (mean outbreak size, peak infectious farms, epidemic duration).
  Assess robustness of the results - mean/distribution
  Compare epidemic outcomes across parameter combinations.
  Determine whether farms identified as structural super spreaders consistently produce larger outbreaks than randomly    selected farms acrross range of parameters

If the relative importance of the super spreader farms remains similar across parameter combinations, conclude that the findings are robust to uncertainty in the disease parameters.

Choose to seed infection in known super spreader nodes & random nodes: top 5% of farms by super spreader metric (RQ1 temporal network analysis). For each replicate, randomly choose one farm from this group as the seed. For the comparison, randomly choose one farm from the remainder of the network.

Because split into 12 months for top 5% ID - choose 1 representative month with typical movement pattern (month 7 - all layers and mid year movements); run sensitivity scenarios - alter B/occult period/infectious period; for each parameter combination run 100 sims in top 5% seeder farms in OD and BT (choose one at random) and run 100 sims in random seed farm (not top 5%); compare mean outbreak outcomes across all three - expect to see ranking

```{r}
# for month 7
out_degree_top5 <- c("253", "129", "95", "255", "88", "119", "214", "213", "212", "202", "151", "299")
betweenness_top5 <- c("95", "296", "119", "41", "285", "300", "274", "178", "138", "235", "156")
random_farms <- c("2", "35", "180", "50", "260", "4", "230", "301", "92", "205", "167", "227")
```

```{r}
# Define sensitivity ranges
beta_values <- c(0.001, 0.005, 0.01)
occult_values <- c(60, 120, 240)
infectious_values <- c(90, 180, 365) # N.B. robust against 365 day (longer that data set)

# Define seed groups for one month
seed_groups <- list(
  out_degree = out_degree_top5,
  betweenness = betweenness_top5,
  random = random_farms
)

# Create simulation scenarios
# https://www.rdocumentation.org/packages/base/versions/3.3.0/topics/expand.grid
scenarios <- expand.grid( # creates 100 rows for each parameter combination, lapply runs every row
  beta = beta_values,
  occult_period = occult_values,
  infectious_period = infectious_values,
  seed_type = names(seed_groups),
  replicate = 1:100 # creates 100 identical copies of every unique parameter combination - 
)


# Function to run one simulation

# create a function called run_sensitivity, 'x' is one row from the 'scenarios' table (b, occult_p, infectious_p, seed_t, replicate no.)
run_sensitivity <- function(x){

  # select seed farm - looks at the seed type for this scenario, e.g. out_degree - finds corresponding list of farms, randomly selects one from the list to be the initial infected farm
  seed_farm <- sample(
    seed_groups[[as.character(x$seed_type)]],
    1
  )

  # run SOIR model - runs one complete SOIR sim
  result <- run_soir(
    movements = movements,
    seed_farms = seed_farm,# randomly selected seed farm
    beta = x$beta, # parameter values from scenario 'x'
    occult_period = x$occult_period,
    infectious_period = x$infectious_period
  )


  # Return only summary statistics
  data.frame(
    beta = x$beta,
    occult_period = x$occult_period,
    infectious_period = x$infectious_period,
    seed_type = x$seed_type,
    replicate = x$replicate,

# Final epidemic size - counts how many farms not susceptible at the end of the simulation = total number of farms infected during the outbreak
    final_infected = sum(
      result$final_states$state != "S"
    ),
# peak epidemic size - finds largest number of infectious farms on any single day
    peak_infectious = max(
      result$history$I
    )
  )
}
```
```{r}
# Run all simulations - loops through every simulation scenario, runs the run_sensitivity() function for each one, stores the results as a list
results <- lapply(
  seq_len(nrow(scenarios)), # creates list from 1 to no. simulation scenarios
  function(i){ # takes each scenario in turn
    run_sensitivity(scenarios[i, ]) # runs 1 simulation using that scenario's parameter values (lapply repeats this for every scenario andstores each result in a list called 'results - so even if the parameter values are identical in all 100 replicates, each run is different because is a stochastic model)
  }
)
```
```{r}
# add a progress bar - replace above block
library(pbapply)

results <- pblapply(
  seq_len(nrow(scenarios)),
  function(i){
    run_sensitivity(scenarios[i, ])
  }
)
```

```{r}
# Combine results - stacks all the one-row dataframes to create one combined data frame
# bind_rows = dplyr
results_RB <- bind_rows(results) 
#results
```

Mean outbreak size by seed type:
  Gives the overall effect across all tested beta/occult/infectious values & stochastic replicates
```{r}
OB_by_seedtype <- results_RB %>%
  group_by(seed_type) %>%
  summarise(
    mean_outbreak_size = mean(final_infected),
    sd_outbreak_size = sd(final_infected),
    mean_peak_infectious = mean(peak_infectious)
  )
OB_by_seedtype
```

Interpret - Across X (check how many run!) simulations, dz introductions into structurally important farms produced larger outbreaks than random introductions, irrespective of parameter changes. 

Robustness across transmission parameter space:
  Shows the proportion of parameter combinations where      structural farms outperform random farms

```{r}
# reshape so each parameter becomes a row and seed types become columns - robustness check

ranking_table3 <- results_RB %>%
  group_by(beta, occult_period, infectious_period, seed_type) %>%
  summarise(
    mean_infected = mean(final_infected),
    .groups = "drop"
  ) %>%
  pivot_wider(
    names_from = seed_type,
    values_from = mean_infected
  )
```

```{r}
ranking_table3 %>%
  summarise(
    out_degree_better = mean(out_degree > random),
    betweenness_better = mean(betweenness > random)
  )
```



The advantage of structurally important farms was maintained across the tested disease parameter space. Outdegree and betweenness farms generated larger outbreaks than random introductions in >90% of parameter combinations

Show effect of changing disease parameters - table grouped by beta because likely to be the biggest driver
```{r}
results_RB %>%
  group_by(beta, seed_type) %>%
  summarise(
    mean_outbreak_size = mean(final_infected),
    .groups = "drop"
  )
```

Across all tested combinations of transmission probability, occult period, and infectious period, structurally important farms consistenly generated larger outbreaks than random introductions. Although absolute outbreaksizes increased with higher transmission probabilitues and longer infectious periods, the relative advantage of structurally important farms (ranking) remained consistent, supporting the robustness of the selected parameter values for subsequent epidemic sims - the relationship of structrally important farms is robust to uncertainty in disease parameters.


Sensitivity analysis conclusion:
  The ranking of high-risk farms was robust to uncertainty in disease parameters: both structural metrics consistently outperformed random farm selection.
  
```{r}
# box plot across all 8100 simulations to visually show the separation between structural and random introductions
ggplot(results_RB, aes(
  x = seed_type,
  y = final_infected,
  fill = seed_type
)) +
  geom_boxplot() +
  labs(
    x = "Seed farm type",
    y = "Final number of infected farms",
    title = "Outbreak size following introduction into structurally important farms"
  ) +
  theme_classic() +
  theme(
    legend.position = "none"
  )
```
For dissertation and sensitivity analysis - final infected farms violin plot (overall outbreak impact) & peak infectious farms violin pot (epidemic intensity)

Use the literature-based baseline parameters for the main simulation results.
Present the sensitivity analysis as evidence that the conclusions are not driven by those specific parameter values. "The baseline parameter values were selected from published estimates of bovine tuberculosis transmission and progression. Sensitivity analyses were performed across plausible ranges reported in the literature."

So - sensitivity analysis has shown robust to changes in parameters, will use:

  beta = 0.005 - middle of test range, avoids assuming very high/very low transmission ans is consistent with bTB being relatively slow to spread compared with acute diseases like FMD
  
  occult period = 120 days - intermediate value; bTB has highly variable latency, 4 month period is a reasonable simplification for a farm-level model
  
  infectious period = 180 days - intermediate value; allows sufficient time for farms to generate secondary infections while avoiding unrealistically long infectiousness - reflects time it could take to identify and impose movement restrictions on an infected farm
  
  


# NEXT STEP

Main research simulations

"An initial unweighted movement model was developed to evaluate whether network structure alone influenced epidemic outcomes. The model was subsequently extended to incorporate movement volume, allowing transmission probability to scale with the number of cattle moved."

NEED TO ADD WEIGHTS - model 3 in https://www.sciencedirect.com/science/article/pii/S1755436510000460#sec1
