---
title: "TEST SIM"
author: "ESINCLAIR"
date: "2026-07-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

https://github.com/digivet-consortium/movenet/blob/main/R/network_creation.R
https://digivet-consortium.github.io/movenet/authors.html#citation

N.B. for SimInf model - graph needs to be in 
A B
A B
A C
A B

Not

A B 3 (movements)
A C 1 etc
```{r}
library(movenet)

# Load an example movement config file:
load_config(system.file("configurations", "ScotEID.yml", package = "movenet"))
#> Successfully loaded config file: C:/Users/cboga/OneDrive - University of Glasgow/Documents/R/win-library/4.1/movenet/configurations/ScotEID.yml
# Read in and reformat an example movement data file:
movement_data <- 
  reformat_data(system.file("extdata", "example_movement_data.csv", package = "movenet"),
                type = "movement")

# Load an example holding config file:
load_config(system.file("configurations", "fakeScotEID_holding.yml", package = "movenet")) 
#> Successfully loaded config file: C:/Users/cboga/OneDrive - University of Glasgow/Documents/R/win-library/4.1/movenet/configurations/fakeScotEID_holding.yml
# Read in and reformat an example holding data file:
holding_data <- 
  reformat_data(system.file("extdata", "example_holding_data.csv", package = "movenet"),
                type = "holding")
```
```{r}
head(movement_data)
```
```{r}
head(holding_data)
```
```{r}
# drop coordinates
library(dplyr)
hold_data_nocoord <- holding_data %>% select(- 'coordinates')
```

```{r}
head(hold_data_nocoord)
```
```{r}
library(SimInf)
time_unit <- 1 #every how much time / how many days to report results for - because in events table - already in weeks - then time step of '1' = 1 week
beta = 0.45 #transmission rate
#Guinat et al 2016: 0.3 between-pen, 0.6 within-pen;
#used in Halasa 2016a as low & high tm levels
epsilon = 0.1 #incubation rate (1/incubation period)
#Zhang et al 2021: 0.1 (0.05-0.25);
#based on Barongo et al 2016, Nielsen et al 2017
gamma = 0.25 #mortality rate
#Zhang et al 2021: 0.25 (0.07-0.5)
#based on Guinat et al 2016
```
```{r}
# build the nodes
nodes <- hold_data_nocoord %>%
  transmute(
    id = cph,
    size = as.integer(herd_size)
  )
```

```{r}
# build the events table - this expands each edge into movement rows (converts igraph into SimInf events)
# transmute is in dpylr package
events <-
  movement_data %>%
  transmute(
    event = "extTrans", #"extTrans" between nodes (or "intTrans","enter","exit") - simif event type for external transfer
    time = as.integer((as.Date(departure_date)- min (as.Date(departure_date))) / 7) +1, # stops havig any week = 0 (0/7 = 0)
    node = match(departure_cph, nodes$id),
    dest = match(dest_cph, nodes$id), # this converts cph number to an integer for simif
    n = qty_pigs, #weight (number pigs)
    proportion = 0, #alternative to n, if want to move proportion (set n as 0)
    select = 2, #col in model select matrix (which compartments to sample from)
    shift = 0) #col in model shift matrix (if want to shift compartment)
```

Set up model
```{r}
n <- length(nodes$id) # number of nodes = number of farms in dataset
# Build the initial conditions = x4 matrix of initial SEIR states
u0 <- data.frame(S = nodes$size, # number of susceptibles = herd size
                 E = rep(0, n),
                 I = rep(0, n),
                 R = rep(0, n))

#Add single infectious individual to a single random node
infected_node<-sample(n,1)
u0[infected_node,1] <- u0[infected_node,1]-1 #take 1 individual out of S
u0[infected_node,3] <- 1 #add 1 individual to I

# Build the model timespan
tspan <- seq(from = min(events$time), #first movement
             to = max(events$time), #last movement
             by = time_unit)

model <- SEIR(u0 = u0, # initial conditions
              tspan = tspan, # timespan
              events = events, #moves between nodes
              beta = beta, #transmission rate
              epsilon = epsilon, #incubation rate (1/incubation period)
              gamma = gamma) #mortality rate

#Change the select_matrix of the model so that the 2nd column indicates
#selection from S, E and I (but not R) compartments
select_matrix(model) <- matrix(c(1, 0, 0, 0, 1, 1, 1, 0), nrow = 4)
```
Run the model

```{r}
#Summary stats of total outbreak size, for 10 simulations
summary(replicate(n = 10, {
  IR_at_end <- tail(trajectory(run(model))[c("I","R")], n)
  #sum(IR_at_end)
  summary(IR_at_end)
}))
```
Most simulations had almost no outbreak.  
Min, 1st quartile, and median are all 1, meaning in at least half of the runs only the single seeded pig was ever infected.

Occasionally the outbreak spread.  
The mean = 27.6 shows that a few larger outbreaks pulled the average up.

Rarely, it spread substantially.  
The maximum = 152 shows the biggest outbreak across the 10 runs.

NB this is pigs not nodes. If want to change to nodes (for dissertation model) - set herd size = 1 for every node; nodes$size <- rep(1, n); 
Then:

S = 1 means node is susceptible

E = 1 means node is exposed

I = 1 means node is infectious

R = 1 means node is removed

# Need to change

  model structure - SEI instead of SEIR?
  the parameters - bTB slow, chrinic, low transmission - this is ASF in pigs
  the initial conditions - farm-level not pig-level infection
  the movement rules
  
  vignette("SimInf")
vignette("trajectory")
vignette("post-processing")

Documentation for:

   view model output
   extract compartment values
   inspect events
   plot trajectories
   analyse node-level results
   
```{r}
browseVignettes("SimInf")
```

N.B. for SimInf model - graph needs to be in 
A B
A B
A C
A B

Not

A B 3 (movements)
A C 1 etc