---
title: "SIMTEST2"
author: "ESINCLAIR"
date: "2026-07-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

https://pmc.ncbi.nlm.nih.gov/articles/PMC3475695/ - SORI model
https://pmc.ncbi.nlm.nih.gov/articles/instance/2600903/bin/rspb20071601s01.pdf 
https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1004038#pcbi-1004038-g001
https://pmc.ncbi.nlm.nih.gov/articles/PMC6122770/#Sec5
https://efsa.onlinelibrary.wiley.com/doi/full/10.2903/j.efsa.2023.8028

https://www.rdocumentation.org/packages/stats/versions/3.5.2/topics/Uniform - R Documentation - runif()

file:///C:/Users/errai/Downloads/978-3-540-78911-6_3.pdf - chapter on maths behind models
Built from work in TEST SIM 2 Rmd

Do farms that are structurally important in the cattle movement network disproportionately contribute to bTB spread over a one-year period?

Hypothesis is:

"Are structurally superspreader farms actually superspreaders when disease dynamics are simulated?"

Basic SEIR model:

not trying to build the most biologically detailed model of bovine TB; trying to compare how different network interventions affect disease spread.

based on:
Green et al. (2006) — farm-level disease transmission through cattle movements. Modelling the initial spread of foot-and-mouth disease through animal movements
Green et al 2008 - Estimates for local and movement-based transmission of bovine tuberculosis in British cattle
Gilbert et al 2005 - This isn't a simulation paper, but it established that cattle movements are a major predictor of bTB occurrence.
BoTMEW  2018 A new bovine tuberculosis model for England and Wales (BoTMEW), explicitly describe replaying movement data
Heath et al. (2008) — preserving the temporal sequence of movement events is important
SimInf (Widgren et al., 2016) — stochastic compartmental models on livestock movement networks.

NETWORK BASED EVENT DRIVEN SEIR MODEL
A stochastic, farm-level, network-based, movement-driven SEIR compartment model.

PSEUDOCODE FOR WRITE UP - for methods chapter - aim to communicte modelling logic independent of code

Initialise farms

↓

Seed infection

↓

For each day

    Update E→I→R

    Get today's movements

    For each movement

        if source == I
            if destination == S

                transmit?

↓

Save daily epidemic

↓

Return results

# UPDATE TO INCLUDE BATCH SIZE in future
Model is farm-level SEIR, not animal level SEIR. each movement is an opportunity for infection to be introduced into the destination herd. more cattle moved = more opportunities for introduction to occur. P(transmission) = 1 - (1-B)^n, where n = batch size.

Interpret Beta as the probability that one infected animal successfully seeds infection in the destination herd. If a batch contains n animals, then the probability of at least one successful transmission becomes P(transmission) = 1-(1-B)^n, where Beta = per animal transmission probabiity; n = number of cattle moved.


SO LOOKING AT - structural characteristics of SS identified in network analysis. Impact of removing these nodes from stochastic simulation (that includes probability of transmission based on 'strength'/weighted batch movement)

PROMPT (ChatGPT)
show me basic complete R code for an SEIR model that tracks btb-like disease spread across nodes (farms). each movement event has a total_cattle batch size. Each farm can send or recieve several batches of cattle per day. I want to be able to seed an infection in a node an view the potential disease spread. I want to be able to remove certain  nodes  identified as superspreaders and compare simulation outcomes. Possibly in the future, I would like to compute probabiltiy of tranmission to be weighted by number of cattle

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
```


```{r}
# This is base model function from gpt prompt - a SEIR compartment stochastic model - farm level rather than animal level (issues with Movenet/DigiVet)

# Everything in this block is inside the run_seir function
# takes input:
  # movement data (movements) and a 
  # starting infected farm (seed_farm)
  # transmission probability (beta) 
  # latent_period - time between becoming infected and infectious - in compartment 'E' - incubating but CANNOT yet infect other farms through  movements - (?make latency move with animal - realistic infection developing further down contact chain?)
  # infectious period - time spent infectious and capable of transmitting infection before moving into R (removed) - model assumption based on time to detect infected farm. During these 80d - movements from the farm can transmit infection according to beta

run_seir <- function(
    movements,
    seed_farm,
    beta = 0.05, 
    latent_period = 60, 
    infectious_period = 180 
){


# Overall timeline if a susceptible farm becomes infected on day 100: 0-99d = susceptible, can't infect others; 100-159d = exposed, can't infect others, 160-339d = infectious, can infect others, 340d onwards = removed, cant infect others

# Build lookup table - pulls every unique farm ID from both the source and destination columns of movements  - combines, removes duplicates, sorts --> (farms) is a clean lookup table of all farms involved in the dataset
  
  farms <- sort(unique(c(
    movements$source,
    movements$destination
  )))
  
# creates a numeric index for the farms, labels each with a number corresponding to the farm ID
  n_farms <- length(farms)

  farm_index <- seq_len(n_farms)
  names(farm_index) <- farms

 
# Convert movements to integer indices #

  movements$source_id <- farm_index[movements$source]
  movements$dest_id   <- farm_index[movements$destination]

  
# Disease state definitions:
  
  # 0 = S = susceptible
  # 1 = E = exposed
  # 2 = I = infectious
  # 3 = R = removed

# set up vectors to track each farms disease status and how long it has been in its current state
  
  state <- integer(n_farms) # creates a vector of zeros - meaning all farms start susceptible

  days <- integer(n_farms) # creates a parallel vector to count how many days each farm has spent in its current state

 
# Seed infection - infect chosen seed farm by setting its state to 2(I)

  state[farm_index[seed_farm]] <- 2


# Storage - prepares storage for the simulation results

  max_day <- max(movements$day) # finds last day in the movement dataset

# History - creates a data frame with one row per day, and columns for the number of farms in each SEIR state, all columns are pre-allocated as integer vectors of length max_day for filling during the simulation
  
  history <- data.frame( 
    day = integer(max_day), 
    S = integer(max_day),
    E = integer(max_day),
    I = integer(max_day),
    R = integer(max_day)
  )
  


# Simulation - advances farms through the E -> I -> R stages as each day of the simulation passes

# the 'for' loop runs one iteration per day

  for(day in seq_len(max_day)){

 
# Progress exposed farms:
 
# the day counters increase, and farms that have spent more than latent_period days in E become infectious (state == 2) and their day counters reset
    
    exposed <- which(state == 1)

    if(length(exposed) > 0){

      days[exposed] <- days[exposed] + 1

      become_I <- exposed[
        days[exposed] >= latent_period
      ]

      state[become_I] <- 2
      days[become_I] <- 0

    }


# Progress infectious farms

    infectious <- which(state == 2)

# Their day counters increase, any farms that have spent more than infectious_period i days in I become removed (state == 3) and their day counters reset    
    
    if(length(infectious) > 0){

      days[infectious] <- days[infectious] + 1

      become_R <- infectious[
        days[infectious] >= infectious_period
      ]

      state[become_R] <- 3
      days[become_R] <- 0

    }


# Today's movements - filters the 'movements' data frame to keep just the rows where 'day' matches the current loop day, giving the set of movements to process for that day.

    todays <- movements[movements$day == day, ]


# Replay movement events - processes each movement and applies infection-spread rules.
# loop over all movements happening today, for each movement, get the source and destination farm indices; skip the movement if the source is not infectious (state != 2); skip the movement if the destination is not susceptible (state != 0). Only movements from the infectious -> susceptible farms proceed to the transmission step that follows
    

    if(nrow(todays) > 0){

      for(i in seq_len(nrow(todays))){

        source <- todays$source_id[i]
        dest   <- todays$dest_id[i]

        # source not infectious
        if(state[source] != 2)
          next

        # destination already infected - i.e. not susceptible
        if(state[dest] != 0)
          next


# Transmission - applies transmission step - an infectious source may infect a susceptible destination. With probability 'beta', the destination farm becomes exposed (state = 1) and is day counter resets to 0 
# https://www.rdocumentation.org/packages/stats/versions/3.5.2/topics/Uniform - runif() - generates random number between 0-1, if number is < beta - infection progresses = Bernoulli
        
        if(runif(1) < beta){

          state[dest] <- 1
          days[dest] <- 0

        }

      }

    }

    
# Save epidemic - logs the SEIR count for the current simulation day - store the day number; count how many farms are S,E,I,R and write those values into the corresponding row of history
    
    history$day[day] <- day
    history$S[day] <- sum(state == 0)
    history$E[day] <- sum(state == 1)
    history$I[day] <- sum(state == 2)
    history$R[day] <- sum(state == 3)

  }


# Return - returns the simulation output as a structured list
  
  list(

    history = history, # stores the full day-by-day SEIR counts

    final_states = data.frame( # data frame listing each farm and its final SEIR state, converts numeric codes (0-3) into labelled factors S,E,I,R

      farm = farms,

      state = factor(
        state,
        levels = 0:3,
        labels = c("S","E","I","R")
      )

    )

  )

}
```

# Provide movement data

```{r}
# read in csv
movements_raw <- read.csv(
    "RQ1 NO EP.csv",
    stringsAsFactors = FALSE
)
# head(movements_raw)
```

```{r}
# check the structure
str(movements_raw)
```
```{r}
# create simulation movement  table
movements <- movements_raw %>%
    select(
        source = org_id,
        destination = dest_id,
        day,
        total_cattle
    )
# head(movements)
```


```{r}
# convert  farm IDs to chr rather than integer to stop e.g. 1 being  treated differently from 001 in future datasets
movements <- movements %>%
    mutate(
        source = as.character(source),
        destination = as.character(destination)
    )
```

```{r}
# check structure
str(movements)
```
```{r}
# check time period
range(movements$day)
```
```{r}
# number of movement events
nrow(movements)
```

```{r}
# check farms that only appear as destinations
farms <- sort(unique(c(
    movements$source,
    movements$destination
)))
```

```{r}
# check movement intensity
movements %>%
    count(day) %>%
    summary()
```
# Run simulation
```{r}
# N.B.this is only running one replicate of the simulation- being stochastic need to run 100-1000 replicates because each will be slightly different
result <- run_seir(
    movements,
    seed_farm = "285" # choose farm to seed infection - needs updated to allow several farms
)

head(result$history)
```

```{r}
# plot the curve - using ggplot

result$history %>%
    pivot_longer(
        cols = c(S,E,I,R),
        names_to = "state",
        values_to = "farms"
    ) %>%
    ggplot(
        aes(
            x = day,
            y = farms,
            colour = state
        )
    ) +
    geom_line(size = 1) +
    theme_minimal()
```
```{r}
result_95 <- run_seir(
    movements,
    seed_farm = "95"
)
```
```{r}
# inspect the outbreak
result_95$final_states %>%
    filter(state != "S")
```
```{r}
tail(result_95$history, 1)
```


This plot shows that infection seeded at farm 7, Susceptible (S) declines from ~300 to ~295.
Exposed (E) rises gradually and peaks at only a few farms.
Infectious (I) remains low (peaking around 5 farms).
Removed (R) accumulates to around 5–6 farms by the end.
So the outbreak seeded at farm 7 caused a small, self-limiting chain of transmission rather than a large epidemic.

That is a plausible outcome for a sparse movement network.

```{r}
# maximum number infected
max(result$history$I)
```

```{r}
tail(result$history, 1)
```


### NEXT THINGS TO DO - INCOPORTATE WEIGHTED BY CATTLE BATCH SIZE, TUNE TRANSMISSION/LATENT ETC TO BTB, tune parameters to reflect bTB
 

Removing superspreaders

```{r}
remove_nodes <- function(movements, nodes){

  movements[
    !(movements$source %in% nodes |
      movements$destination %in% nodes),
  ]

}
```


# Simulaiton outputs:

epidemic size - number of farms infected/proportion of network infected/cumulative incidence

epidemic spread - time to first 10 farms infected/time to peak infected farms/epidemic duration

node level contribution - for each farm: number of secondary infeciton caused/probability of causing a large outbreak/outbreak size conditional on seed infection

Then compare - structural SS farms vs random farms as initial infection seeds.

A strong test - if infection starts in the top 5% central farms, does the resulting epidemic size exceed epidemics in randomly selected farms?

Implement: transmission_probability <- 1 - (1-beta)^nu_cattle_moved
and keep network centrality out of the transmission equation.

Then evaluate whether the independently identified SS (network analysis) nodes actually generate: larger outbreaks/faster outbreaks/more secondary infections/higher outbreak probability - means that simulation is testing the network hypothesis rather than assuming it.


# DEBUGGING
```{r}
result <- run_seir(
    movements,
    seed_farm = "7"
)
```
```{r}
result$final_states %>%
    filter(state != "S")
```